### 2018-09-13 doc/article.tex
A new document to describe results

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{natbib}

%SetFonts

%SetFonts

\begin{document}

\title{Testing Gravity using Type Ia Supernovae Discovered by LSST}
\author[0000-0001-6315-8743]{A.~G.~Kim}
\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}

\author{and others}

%\date{}							% Activate to display a given date or no date

\maketitle
\section{Introduction}
Peculiar velocities provide a measure of $f\sigma_8$, which in turn probes gravity.  Type~Ia supernovae. as a precise distance indicator,
can in turn  provide precise peculiar velocities of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
\citet{2015JCAP...12..033H, 2017JCAP...05..015H}  test and ultimately find consistency between
$\Lambda$CDM and the peculiar velocities of existing SN~Ia samples,

Surveys such as ZTF and LSST are and will discover orders of magnitude more nearby SNe~Ia than currently available.
The motivation of this work is to quantify the probative power of their peculiar velocities in the LSST era.
While there have been a number of articles on the subject,
our analysis brings a higher level of fidelity than performed in previous analyses.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
posteriors using MCMC on reasonable timescales.   We can use our machinery to 
compare different survey parameters, such as redshift depth, total numbers of supernovae,
solid angle/survey geometry, and the implications of different SN~ia follow-up strategies.

Galaxy-count -- peculiar velocity cross-correlation is not yet implemented but is a planned extension of this work.  We would like to
quantify the suppression of sample variance by comparing matter-density and velocities in the same volume \citep{2007PhRvL..99h1301G}.

\section{Simulated Data}

\section{Anaysis}

\section{Results}

\bibliographystyle{aasjournal}
\bibliography{/Users/akim/Documents/alex}

\end{document}

### 2018-09-14 doc/article.tex
continuing on with first draft

@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{C.~Harper}
+\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+\author{C.~Ju}
+\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+\author{D.~Huterer}
+\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
+Arbor, MI 48109, USA }
 
@@
 
-\maketitle
 \section{Introduction}
-Peculiar velocities provide a measure of $f\sigma_8$, which in turn probes gravity.  Type~Ia supernovae. as a precise distance indicator,
-can in turn  provide precise peculiar velocities of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
+Peculiar velocities provide a measure of $f\sigma_8$, which in turn probes gravity.  As precise distance indicators Type~Ia supernovae
+can provide precise peculiar velocities (expressed practically in term of peculiar magnitudes)
+of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
 \citet{2015JCAP...12..033H, 2017JCAP...05..015H}  test and ultimately find consistency between
-$\Lambda$CDM and the peculiar velocities of existing SN~Ia samples,
+$\Lambda$CDM and the peculiar velocities of existing SN~Ia samples.
 
 Surveys such as ZTF and LSST are and will discover orders of magnitude more nearby SNe~Ia than currently available.
-The motivation of this work is to quantify the probative power of their peculiar velocities in the LSST era.
+The motivation of this work is to quantify the probative power of SN~Ia-derived peculiar velocities in the LSST era.
 While there have been a number of articles on the subject,
-our analysis brings a higher level of fidelity than performed in previous analyses.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
+our analysis brings a higher level of fidelity than sought by previous analyses.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
 catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
-posteriors using MCMC on reasonable timescales.   We can use our machinery to 
+posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
 compare different survey parameters, such as redshift depth, total numbers of supernovae,
-solid angle/survey geometry, and the implications of different SN~ia follow-up strategies.
+solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
 
-Galaxy-count -- peculiar velocity cross-correlation is not yet implemented but is a planned extension of this work.  We would like to
-quantify the suppression of sample variance by comparing matter-density and velocities in the same volume \citep{2007PhRvL..99h1301G}.
+The cross-correlation between galaxy-count and peculiar velocity surveys is not yet implemented but is a planned extension of this work.  We would like to
+quantify the suppression of sample variance achieved when considering matter-densities and velocities 
+within the same volume \citep{2007PhRvL..99h1301G}.
 
 \section{Simulated Data}
+The Buzzard (v1.6) galaxy catalog is used.  This is because it is the survey, among those currently available
+in the Generic Catalog Reader of the DESC, with a light cone that covers the largest solid angle of 10,313.24 sq.~deg.  However,
+the survey geometry is different
+from that of LSST.  The catalog is based on a Flat $\Lambda$CDM model with $\Omega_M=0.286$, $\Omega_B=0.047$, and
+$\Omega_\nu=0$,
+The catalog contains140M galaxies with observed $z<0.2$ for which the star formation rate and stellar mass are available; from these  parameters the
+supernova rate of each host is determined using the rates of
+\citet{2012ApJ...755...61S}.  Supernovae are realized for a span of 10 observer-years yielding 17,205 objects.  This total supernova
+production underestimates the expected discovery of 134,574 supernovae in 10 observer-years based on the volumetric
+rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so the sample under consideration in this article
+corresponds to 1.3 years worth of discoveries.
 
-\section{Anaysis}
+\section{Analysis}
+The analysis of this article is identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}.   The peculiar magnitude correlation
+function $\xi_{\delta m \delta m}$ expected from General Relativity, the matter density power spectrum at the CMB assuming the same
+cosmological parameters used for the Buzzard catalog, is calculated using CAMB \citep{Lewis:2002ah}.  Our model for the data covariance
+is
+\begin{equation}
+C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij}.
+\end{equation}
+In this model, we treat the shape of the power spectrum $\xi$ as fixed, and use the parameter $A$ to represent
+deviations from General Relativity $A=1$.  This model is not sensitive to  the shape of
+the power spectrum between gravitational models.  Nevertheless, for
+convenience we make the association that our model gives $f\sigma_8 = A f\sigma_8^{GR}$, where $ f\sigma_8^{GR}$ is the expectation
+from General Relativity.  Special cases of modified gravity could give $A=1$, while inconsistency with $A=1$ is inconsistency with GR.
 
-\section{Results}
+\section{Results of Subsets}
 

### 2018-09-16 doc/article.tex
making progress

@@
 
-\author{and others}
+\author{others}
 
@@
 Peculiar velocities provide a measure of $f\sigma_8$, which in turn probes gravity.  As precise distance indicators Type~Ia supernovae
-can provide precise peculiar velocities (expressed practically in term of peculiar magnitudes)
+can provide precise peculiar velocities (expressed equivalently as peculiar magnitudes)
 of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
@@
 $\Omega_\nu=0$,
-The catalog contains140M galaxies with observed $z<0.2$ for which the star formation rate and stellar mass are available; from these  parameters the
+The catalog contains140M galaxies with observed $z<0.2$.  Each galaxies cosmological redshift, redshift that combines
+which the star formation rate and stellar mass are available; from these  parameters the
 supernova rate of each host is determined using the rates of
@@
 convenience we make the association that our model gives $f\sigma_8 = A f\sigma_8^{GR}$, where $ f\sigma_8^{GR}$ is the expectation
-from General Relativity.  Special cases of modified gravity could give $A=1$, while inconsistency with $A=1$ is inconsistency with GR.
+from General Relativity.  Special cases of modified gravity could give $A=1$, but inconsistency with $A=1$ is is evidence against General Relativity.
 

### 2018-09-16 pv/PVHostGalaxies.py
found out redshift is in the galaxy frame, not the CMB frame.  Get CMB-frame redshifts

@@ -34,7 +34,7 @@ def main(seed, test = True):
     numpy.random.seed(seed)
 
     # catalog quantities necessary to do the analysis
-    quantities = ["redshift", "redshift_true","ra","dec","galaxy_id",
+    quantities = ["redshift_true","ra","dec","galaxy_id",'position_x','position_y','position_z','velocity_x','velocity_y','velocity_z',
                   "truth/COEFFS/0","truth/COEFFS/1","truth/COEFFS/2","truth/COEFFS/3","truth/COEFFS/4"]
 
     # GCRCatalog is not efficient for one-step filtering since derived quantities calculated for everything.  Hence do a first coarse filter.
@@ -50,6 +50,8 @@ def main(seed, test = True):
 
     out = None
 
+    c_light = 299792.458
+
     for data in coadd_cat.get_quantities(quantities, filters = filts, return_iterator=True):
         sfr_met_smass=gdq.get_derived_quantities(numpy.array([data["truth/COEFFS/0"],data["truth/COEFFS/1"],
                                                   data["truth/COEFFS/2"],data["truth/COEFFS/3"],
@@ -63,6 +65,13 @@ def main(seed, test = True):
         data['random_realize'] = []
         for nsn in data['nsne']:
             data['random_realize'].append(numpy.random.normal(scale=1,size=nsn))
+
+        data['redshift']=  data['position_x']*data['velocity_x'] + data['position_y'] * data['velocity_y'] + data['position_z']*data['velocity_z']
+        data['redshift'] = data['redshift']/numpy.sqrt(data['position_x']**2 + data['position_y']**2+data['position_z']**2)
+        data['redshift'] = (1+data['redshift']/c_light)/numpy.sqrt(1-(data['velocity_x']**2 + data['velocity_y']**2+data['velocity_z']**2)/c_light**2)
+        data['redshift'] = (1+data['redshift_true'])*data['redshift'] -1
+
+
         data['mB_true_mn'] = MB + cosmo.distmod(data["redshift_true"]).value
         data['mB_expected'] = MB + cosmo.distmod(data["redshift"]).value
         c = SkyCoord(ra=data["ra"]*u.degree, dec=data["dec"]*u.degree, frame='icrs')

### 2018-09-16 pv/HostGalaxies.py
clean things up for rerun

@@ -112,7 +112,7 @@ if __name__ == "__main__":
     parser.add_argument('--path', dest='path', default='../test/', type = str, required=False)
     args = parser.parse_args()
 
-    hg = HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path)
+    # hg = HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path)
     # sg, sxi = hg.getSubset(decmax=60, bmin=-34, bmax=34)
     # print (hg.xi.shape[0], sxi.shape[0])
-    # HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path).draganFormat()
+    HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path).draganFormat()

### 2018-09-16 pv/buzzard_stellar_mass.py

@@ -73,14 +73,14 @@ class GetDerivedQuantities:
 
         smass = np.dot(self.mass_tot, coeff.T) * (da * 1e6 / 10) ** 2
         ssfr  = np.dot(coeff, self.sfh_tot)
-        met   = np.dot(coeff, self.sfh_tot * self.sfh_met) / ssfr
+        # met   = np.dot(coeff, self.sfh_tot * self.sfh_met) / ssfr
         sfr   = ssfr * smass[:,np.newaxis]
 
         #get the values at z_galaxy
-        met   = met[:,-1]
+        # met   = met[:,-1]
         sfr   = sfr[:,-1]
 
-        return np.vstack((sfr, met, smass))
+        return np.vstack((sfr, np.zeros(len(smass)), smass))
 
     def __init__(self,template_path):
         self.sfh_tot = fitsio.read(template_path, 12)
@@ -140,5 +140,4 @@ if __name__=='__main__':
 
     galaxies  = fitsio.FITS(filename)[-1].read(columns=['COEFFS','Z','DELTAM']) # read relevant info from files
     coeffs    = galaxies['COEFFS']*10**(galaxies['DELTAM'].reshape(-1,1)/2.5)
-    
     sfr, met, smass = get_derived_quantities(kfile,coeffs,galaxies['Z'])

### 2018-09-17 pv/fit.py
add variables for solid angle

@@ -121,6 +121,9 @@ if __name__ == "__main__":
         bmax=34
         fstr='skycut.'
     else:
+        decmax=90
+        bmin=-90
+        bmax=90
         fstr=''
 
     hg = HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path)

### 2018-09-24 doc/article.tex
added cauchy prior to sigmas (which has no effect) and non-linear dispersion to fit (which also has no effect)

@@
 \usepackage{natbib}
+\usepackage{grffile}
 
@@
 of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
-\citet{2015JCAP...12..033H, 2017JCAP...05..015H}  test and ultimately find consistency between
-$\Lambda$CDM and the peculiar velocities of existing SN~Ia samples.
+\citet{2015JCAP...12..033H, 2017JCAP...05..015H}  test and ultimately detect correlations in the peculiar velocities of existing SN~Ia samples.
 
@@
 
-The cross-correlation between galaxy-count and peculiar velocity surveys is not yet implemented but is a planned extension of this work.  We would like to
+The cross-correlation between galaxy-count and peculiar-velocity surveys is not yet implemented but is a planned extension of this work.  We would like to
 quantify the suppression of sample variance achieved when considering matter-densities and velocities 
@@
 \section{Simulated Data}
-The Buzzard (v1.6) galaxy catalog is used.  This is because it is the survey, among those currently available
-in the Generic Catalog Reader of the DESC, with a light cone that covers the largest solid angle of 10,313.24 sq.~deg.  However,
+The Buzzard (v1.6) galaxy catalog is used.  This is because it is the survey, with a light cone that covers 10,313.24 sq.~deg., that covers
+the largest solid angle among those currently available
+in the DESC Generic Catalog Reader.  However,
 the survey geometry is different
-from that of LSST.  The catalog is based on a Flat $\Lambda$CDM model with $\Omega_M=0.286$, $\Omega_B=0.047$, and
-$\Omega_\nu=0$,
-The catalog contains140M galaxies with observed $z<0.2$.  Each galaxies cosmological redshift, redshift that combines
-which the star formation rate and stellar mass are available; from these  parameters the
-supernova rate of each host is determined using the rates of
+from that of LSST.  The catalog is based on a Flat $\Lambda$CDM model with $H_0=70$~km~s$^{-1}$,  $\Omega_M=0.286$, $\Omega_B=0.047$, and
+$\Omega_\nu=0$.
+The catalog contains 140M galaxies with observed redshift $z<0.2$.
+Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
+the radial peculiar velocity is determined, and its
+star formation rate and stellar mass from which the
+supernova rate of each host is determined using 
 \citet{2012ApJ...755...61S}.  Supernovae are realized for a span of 10 observer-years yielding 17,205 objects.  This total supernova
-production underestimates the expected discovery of 134,574 supernovae in 10 observer-years based on the volumetric
+production underestimates the expected discovery of 130,000 supernovae in 10 observer-years based on the volumetric
 rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so the sample under consideration in this article
@@
 
+Each supernova is assigned a magnitude based on the distance modulus of its cosmological
+redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
+intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
+motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
+\citet{2011ApJ...741...67D}, the effects in Eq.~18 are ignored.
+
 \section{Analysis}
-The analysis of this article is identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}.   The peculiar magnitude correlation
-function $\xi_{\delta m \delta m}$ expected from General Relativity, the matter density power spectrum at the CMB assuming the same
-cosmological parameters used for the Buzzard catalog, is calculated using CAMB \citep{Lewis:2002ah}.  Our model for the data covariance
+The analysis of this article is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}.   The peculiar magnitude correlation
+function $\xi_{\delta m \delta m}$ expected from General Relativity starting with the CMB
+matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
+cosmological parameters used for the Buzzard catalog.  Our model for the data covariance
 is
 \begin{equation}
-C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij}.
+C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij}.
 \end{equation}
-In this model, we treat the shape of the power spectrum $\xi$ as fixed, and use the parameter $A$ to represent
+In this model, we treat the shape of the power spectrum $\xi$ as fixed and use the parameter $A$ to represent
 deviations from General Relativity $A=1$.  This model is not sensitive to  the shape of
 the power spectrum between gravitational models.  Nevertheless, for
-convenience we make the association that our model gives $f\sigma_8 = A f\sigma_8^{GR}$, where $ f\sigma_8^{GR}$ is the expectation
+convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
 from General Relativity.  Special cases of modified gravity could give $A=1$, but inconsistency with $A=1$ is is evidence against General Relativity.
+The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
+\begin{equation}
+\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
+\end{equation}
+where $\bar{z}$ is the cosmological redshift of the host galaxy.
+\section{Results of Subsets}
 
-\section{Results of Subsets}
+For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
+reduce errors made in the first-order transformation between velocity and
+magnitude.  This cut removes a  small volume relative to the full survey.
+
+We consider two different scenarios cutting at different redshift depths $z_{max}=0.1$ and  $z_{max}=0.15$.  All the objects in the first
+scenario are in the second scenario.  The results are shown in Fig.~\ref{zmax:fig}.  The points of interest are:
+\begin{itemize}
+\item There is a bias toward low $f\sigma_8$ for low-redshift ($z_{max}=0.1$).  Adding more supernova with increasing redshift
+ decreases the bias  ($z_{max}=0.15$), and $A$ eventually approaches 1 ($z_{max}=0.2$ not shown).
+ \item Based on the analysis of \citet{2017MNRAS.471.3135H}, I added an intrinsic velocity dispersion $\sigma_v$ (km~s$^{-1}$) due to non-linear effects.  This was not
+ in Dragan's model.  My fit finds that this  dispersion is consistent with zero, not well constrained, and not strongly correlated with
+ $A$.  The lack of covariance is consistent with Dragan's own tests.
+ \item The fit intrinsic magnitude dispersion, on the other hand, is larger than the input $\sigma_M=0.08$ mag.  This  indicates that the catalogs
+ contain 
+ extra  velocity dispersion better attributed to intrinsic magnitude dispersion rather than correlated peculiar velocities.
+\end{itemize}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{../out/pvlist.0.08.1234.1.0.1.pkl.2000.png}
+\includegraphics[width=0.5\textwidth]{../out/pvlist.0.08.1234.1.0.15.pkl.2000.png}
+\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
+absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
+to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
+ Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.15$.
+\label{zmax:fig}}
+\end{figure}
+
+\begin{itemize}
+\item Dragan: Is there some way that the formalism in your article breaks down at low-redshift?  I do remove very low-redshift objects at $z<0.01$,
+which I think is conservative.
+ \item Risa/Joe: Is there any reasons for the Buzzard catalog to not capture correlated peculiar velocities on small scales (box size to large?) nor capture
+ non-linear velocities?  (I don't know how to translate a mass resolution of 2.7e10 $M_{\odot}/h$ into a spatial resolution!)
+\end{itemize}
 

### 2018-09-24 pv/HostGalaxies.py

@@ -4,6 +4,7 @@ import copy
 import argparse
 import os.path
 import array
+import astropy.cosmology
 
 
 class HostGalaxies(object):
@@ -19,6 +20,15 @@ class HostGalaxies(object):
         self.catseed=catseed
         self.sigma_mu = sigma_mu
         self.seed=seed
+        # self.nonlinear = 200./2.99792458e5
+
+        # dm/dv = dm/dz dz/dzv dzv/dv
+        #       = dm/dz (1+z)
+        # cosmology = astropy.cosmology.FlatLambdaCDM(70, 0.286)
+        # onepluszv=numpy.sqrt((1-self.nonlinear)/(1+self.nonlinear))
+        # dm = (cosmology.distmod(self.galaxies['redshift_true'])-cosmology.distmod((1+self.galaxies['redshift_true'])*onepluszv-1)).value \
+        #     *(1+self.galaxies['redshift_true'])
+
 
         numpy.random.seed(self.seed)
         rand = []
@@ -26,8 +36,9 @@ class HostGalaxies(object):
             rand.append(numpy.random.normal(scale=sigma_mu,size=nsn))
 
         self.galaxies['mB'] = []
-        for ran , mag in zip(rand, self.galaxies['mB_true_mn']):
-            self.galaxies['mB'].append(ran + mag)
+        for ran , mag, red, red_true in zip(rand, self.galaxies['mB_true_mn'],self.galaxies['redshift'],self.galaxies['redshift_true']):
+            # onepluszspec = (1+red)/(1+red_true)
+            self.galaxies['mB'].append(ran + mag)   # + 10*numpy.log10(onepluszspec))
 
         self.xi = None
 
@@ -45,8 +56,11 @@ class HostGalaxies(object):
                 self.xi[i,j] =v
                 if (i !=j):
                     self.xi[j,i]=v
-            pickle.dump(self.xi,open("{}/pvlist.{}.xi.pkl".format(path,catseed), "wb" ))
+            # pickle.dump(self.xi,open("{}/pvlist.{}.xi.pkl".format(path,catseed), "wb" ))
 
+        # for i in range(ngal):
+        #     print (self.xi[i,i],dm[i]**2)
+        #     self.xi[i,i] = self.xi[i,i]+dm[i]**2
 
     def draganFormat(self, sort=False):
         if sort:
@@ -81,7 +95,7 @@ class HostGalaxies(object):
[…]

### 2018-09-24 pv/fit.py

@@ -11,7 +11,8 @@ import cvxopt.lapack as lapack
 import cvxopt.blas as blas
 from cvxopt import matrix
 from scipy.linalg.lapack import *
-
+import astropy.cosmology
+from scipy.stats import cauchy
 import time
 
 class Fit(object):
@@ -19,14 +20,13 @@ class Fit(object):
     # def __init__(self):
     #     super(Fit, self).__init__()
 
-
     @staticmethod        
-    def lnprob(theta, Deltam, nsne, xi):
-        A, M, sigma = theta
-        if (A <=0 or sigma <0):
+    def lnprob(theta, Deltam, nsne, xi,redshiftterm):
+        A, M, sigma, pv = theta
+        if (A <=0 or sigma <0 or pv<0):
             return -numpy.inf
         C = A*numpy.array(xi)
-        numpy.fill_diagonal(C,C.diagonal()+ sigma**2/nsne)
+        numpy.fill_diagonal(C,C.diagonal()+ sigma**2/nsne + (pv*redshiftterm)**2)
         mterm  = Deltam-M
 
         C = matrix(C)
@@ -37,19 +37,19 @@ class Fit(object):
             return -np.inf
         logdetC= 2*numpy.log(numpy.array(C).diagonal()).sum()
                        
-        lp = -0.5* (logdetC +blas.dot(matrix(mterm), W) )
+        lp = -0.5* (logdetC +blas.dot(matrix(mterm), W) ) + cauchy.logpdf(sigma, loc=0.08, scale=0.5) + cauchy.logpdf(pv, loc=0, scale=600/3e5)
 
         if not numpy.isfinite(lp):
             return -np.inf
         return lp
 
     @staticmethod  
-    def fit(Deltam, nsne, xi, mpi=False, p0=None, nchain=2000, **kwargs):
-        ndim, nwalkers = 3, 8
-        sig = numpy.array([0.1,0.01,0.01])
+    def fit(Deltam, nsne, xi, redshiftterm,mpi=False, p0=None, nchain=2000, **kwargs):
+        ndim, nwalkers = 4, 8
+        sig = numpy.array([0.1,0.01,0.01,50/3e5])
 
         if p0 is None:
-            p0 = [numpy.array([1,0,0.08])+numpy.random.uniform(low = -sig, high=sig) for i in range(nwalkers)]
+            p0 = [numpy.array([1,0,0.08,200/3e5])+numpy.random.uniform(low = -sig, high=sig) for i in range(nwalkers)]
 
         if mpi:
[…]

### 2018-10-26 doc/article.tex
grinding on

@@
 
-\section{Introduction}
-Peculiar velocities provide a measure of $f\sigma_8$, which in turn probes gravity.  As precise distance indicators Type~Ia supernovae
-can provide precise peculiar velocities (expressed equivalently as peculiar magnitudes)
-of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
-\citet{2015JCAP...12..033H, 2017JCAP...05..015H}  test and ultimately detect correlations in the peculiar velocities of existing SN~Ia samples.
+\begin{abstract}
+ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
+their absolute magnitudes, now make them competitive probes of the growth of structure.
+\end{abstract}
 
-Surveys such as ZTF and LSST are and will discover orders of magnitude more nearby SNe~Ia than currently available.
-The motivation of this work is to quantify the probative power of SN~Ia-derived peculiar velocities in the LSST era.
+\section{Connection between SNe~Ia and Gravity}
+
+The growth of structure depends on the expansion history of the universe, the nature and density of its contents, and gravity. 
+It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the baryonic structures
+in the Universe and the peculiar velocities of test masses therein.
+Peculiar velocities are motions, on top of the cosmological expansion, caused by the gravitational attraction
+and repulsion of all sources of inhomogeneity in the Universe.
+
+Baryonic structures and peculiar velocities  can provide a measurement of the combination $fD$, where $D$ is the linear growth factor that
+gives the overall amplitude of  overdensities and
+$$f \equiv \frac{d\ln{D}}{d\ln{a}}$$ is how that amplitude changes with redshift.  General Relativity predicts
+$f=\Omega_M^\gamma$ (with $D$ determined accordingly) where $\gamma=0.55$, whereas other gravity models can be similarly described
+but with different values of $\gamma$
+\citep{2007APh....28..481L}.  Thus growth of structure, through the measurement of $fD$ (or $f\sigma_8$), provides a test of General Relativity.
+We typically refer to $f\sigma_8$ because the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres is often used to normalize the
+overall amplitude of  overdensities.
+
+As distance indicators SNe~Ia
+can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
+of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
+have been used to test and ultimately find spatial correlations in peculiar velocities, may be attributed to the growth of structure
+\citet{2015JCAP...12..033H, 2017JCAP...05..015H}.  However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
+
+Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
+First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\gtrsim 0.12$~mag distance modulus
+dispersion.  However, SNe transmit more information than a light curve shape and single color used in current analysis.
+Recent studies indicate that with the right data, SNe absolute
+magnitudes can be calibrated to $\lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.
+As absolute magnitude uncertainty propagates into fractional velocity uncertainty, improvements to the standard candle lead to
+improvements in their host peculiar velocities.
+Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+hundreds of thousands.
+
+The precision in  $f\sigma_8$ derived from an ten-year LSST-discovered  SN~Ia survey has been projected by \citet{2017ApJ...847..128H},
+from both their  clustering and peculiar velocities.
+The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance uncertainty is assumed for each supernova.
+SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster, and Ly$\alpha$
+RSD measurements are sensitive.
+
+% Requires the booktabs if the memoir class is not being used
+\begin{table}
+   \centering
+   %\topcaption{Table captions are better up top} % requires the topcapt package
+   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
+	\hline
+	Redshift & RSD & RSD + PV & cumulative\\ \hline
+      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
+     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
+     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
+     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
+      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
+     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
+      \hline
+   \end{tabular}
+   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
+   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
+   The cumulative column shows the effective uncertainty from combining the redshift and all other shallower redshift bins.}
+   \label{tab:howlett}
+\end{table}
+
+The projected precisions for LSST-discovered SNe~Ia have a number of interesting features.  Particular velocities contribute increasingly to
+$f\sigma_8$ precision with decreasing redshift.  At redshift 0.3 the differential gain from adding peculiar velocities to the RSD measurement is small.
+The differential gain in the cumulative $f\sigma_8$ constraint is small.  We therefore conclude that the focus should be in $z<0.3$ LSST SN~Ia discoveries.
+After 10 years, the total number density of supernovae is such that both sample and shot noise contribute to the error budget for the volume with $z<0.2$, 
+Galaxy RSD plus SN~Ia peculiar velocities have yet to be calculated.
+
+They perform their analysis in Fourier space, 
+where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
+approximated by
+\begin{equation}
+F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right],
+\end{equation}
+where the he covariance
+\begin{equation}
+C = P_{vv}(k) + \frac{\sigma^2}{n}
+\end{equation}
+has contributions from the power spectrum, noise in the velocity measurement, and the density of velocity probes
+\citep{2017MNRAS.464.2517H}. 
+In the sample variance limit for a sample with fixed depth, the variance in $f\sigma_8$ (and other $\lambda$ parameters)
+is thus inversely proportional to the survey solid-angle $\Omega$, whereas
+in the shot-noise limit the variance is inversely proportional to $\Omega n^2 \propto N^2/\Omega$, where $n$ is the number density,
+and $N$ is the total number of supernovae.
+
 While there have been a number of articles on the subject,
@@
 solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
+
+The motivation of this work is to quantify the probative power of SN~Ia-derived peculiar velocities in the LSST era.
+
+We consider a maximum redshift of $z=0.2$ because over ten years the noise is not shot-noise dominated, meaning that the differential improvement
+in the precision of
+$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
+likely saturate available follow-up resources.
+Our calculations are thus based on the number and solid-angle of $z<0.2$ supernova pre-maximum discoveries 
+for the survey candidates provided by the Project.  Based on these numbers we calculate the Figures of Merit in
+both sample and shot-noise limits, normalized to ``baseline2018a''.
+After 10 years,
+LSST supernovae are at neither extreme; we thus adopt the average of the two FoM's as what we advocate for the survey.
+These averages for a select set of surveys are shown in Figure~\ref{fig:fom}.
 
@@
 
+The 
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.2.pkl.png}
+\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
+absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
+to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
+ Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.15$.
+\label{zmax:fig}}
+\end{figure}
+
+\begin{table}
+   \centering
+   %\topcaption{Table captions are better up top} % requires the topcapt package
+   \begin{tabular}{|cccc|cccc|} % Column formatting, @{} suppresses leading/trailing space
+   \hline
+$z_{max}$ & fraction & $\sigma_{SN}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+\hline
+0.20 &  1.0 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 \\ \hline
+0.20 &  0.2 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
+0.20 &  0.3 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 \\
+0.20 &  0.5 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 \\
+0.10 &  1.0 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
+0.15 &  1.0 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
+0.20 &  1.0 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 \\
+0.20 &  1.0 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 \\
+0.15 &  0.3 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 \\
+0.12 &  0.3 & 0.10 & 746 &   1.62 &   0.70 &  0.085 &  11.28 \\
+0.07 &  1.0 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
+    \hline
+   \end{tabular}
+   \caption{.}
+   \label{tab:booktabs}
+\end{table}
+
 We consider two different scenarios cutting at different redshift depths $z_{max}=0.1$ and  $z_{max}=0.15$.  All the objects in the first
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{../out/pvlist.0.08.1234.1.0.1.pkl.2000.png}
-\includegraphics[width=0.5\textwidth]{../out/pvlist.0.08.1234.1.0.15.pkl.2000.png}
+\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.1.pkl.png}
+\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.2.pkl.png}
+\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
+absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
+to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
+ Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.2$.
+\label{zmax:fig}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
+\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
 \caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova

### 2018-11-01 doc/article.tex
forging on

@@
 RSD measurements are sensitive.
+{\bf Greg do you have to power to add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
 
@@
 The differential gain in the cumulative $f\sigma_8$ constraint is small.  We therefore conclude that the focus should be in $z<0.3$ LSST SN~Ia discoveries.
-After 10 years, the total number density of supernovae is such that both sample and shot noise contribute to the error budget for the volume with $z<0.2$, 
-Galaxy RSD plus SN~Ia peculiar velocities have yet to be calculated.
+After 10 years for $z>0.1$ both sample variance and shot noise contribute to the error budget; adding even more SNe in the volume would improve
+the measurement though not as $\sqrt{N}$.
 
-They perform their analysis in Fourier space, 
+\section{Supernova Peculiar Velocity Survey}
+\citet{2017ApJ...847..128H} perform their analysis in Fourier space, 
 where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
@@
 \end{equation}
-where the he covariance
+where the covariance
 \begin{equation}
@@
 has contributions from the power spectrum, noise in the velocity measurement, and the density of velocity probes
-\citep{2017MNRAS.464.2517H}. 
+\citep{2017MNRAS.464.2517H}.   This equation is for the peculiar-velocity auto-correlation.
+
+The parameters that determine parameter precision are the volume $V$, which
+in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
+magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
 In the sample variance limit for a sample with fixed depth, the variance in $f\sigma_8$ (and other $\lambda$ parameters)
@@
 
+LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active survey area.
+
+The quality of supernova follow-up is an important factor in how well a peculiar-velocity survey can measure $f\sigma_8$.
+LSST light curves of low-redshift SNe, with their relatively poor sampling, plus a spectroscopic redshift
+ can give $\sigma_M \sim 0.15$~mag. Non-LSST follow-up including infrared data or spectrophotometry at peak brightness are projected to give
+$\sigma_M \lesssim 0.08$~mag.
+Uncertainties in $f\sigma_8$ depend on $\sigma^2_M/n$ so the improvement between 0.08 and 0.15~mag dispersions is equivalent to a factor of 3.52
+ in number density, or in survey duration given the expected completeness of $z<0.3$ SN~Ia discoveries.
+
+The variance in $f\sigma_8$ is proportional to solid angle $\Omega$ given a fixed redshift depth and  $\sigma^2_M/n$.  The baseline
+LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<-0.3$ SN~Ia searches, beyond
+the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+coverage of the southern pole, could double
+the sky coverage and halve the variance in $f\sigma_8$.
+
+LSST can be taken to be a SN~Ia discovery plus light-curve builder, or a discovery machine only.  The former comes with a simplified follow-up program of typing
+and/or redshift collection but with higher intrinsic magnitude dispersion.  The latter requires additional follow-up resources to gather more information per supernova to
+lower the intrinsic magnitude dispersion.
+
+\appendix
+\section{Reanalysis in Configuration Space}
+
+This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
 While there have been a number of articles on the subject,
@@
 solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
-
-The motivation of this work is to quantify the probative power of SN~Ia-derived peculiar velocities in the LSST era.
-
-We consider a maximum redshift of $z=0.2$ because over ten years the noise is not shot-noise dominated, meaning that the differential improvement
-in the precision of
-$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
-likely saturate available follow-up resources.
-Our calculations are thus based on the number and solid-angle of $z<0.2$ supernova pre-maximum discoveries 
-for the survey candidates provided by the Project.  Based on these numbers we calculate the Figures of Merit in
-both sample and shot-noise limits, normalized to ``baseline2018a''.
-After 10 years,
-LSST supernovae are at neither extreme; we thus adopt the average of the two FoM's as what we advocate for the survey.
-These averages for a select set of surveys are shown in Figure~\ref{fig:fom}.
-
-The cross-correlation between galaxy-count and peculiar-velocity surveys is not yet implemented but is a planned extension of this work.  We would like to
+We would like to
 quantify the suppression of sample variance achieved when considering matter-densities and velocities 
 within the same volume \citep{2007PhRvL..99h1301G}.
+The cross-correlation between galaxy-count and peculiar-velocity surveys is not yet implemented but is a planned extension of this work.
 
 \section{Simulated Data}
-The Buzzard (v1.6) galaxy catalog is used.  This is because it is the survey, with a light cone that covers 10,313.24 sq.~deg., that covers
-the largest solid angle among those currently available
-in the DESC Generic Catalog Reader.  However,
-the survey geometry is different
-from that of LSST.  The catalog is based on a Flat $\Lambda$CDM model with $H_0=70$~km~s$^{-1}$,  $\Omega_M=0.286$, $\Omega_B=0.047$, and
+The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
+Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
+cosmoDC2 (v1.0) is consistent with GR. }. 
+The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
 $\Omega_\nu=0$.
-The catalog contains 140M galaxies with observed redshift $z<0.2$.
 Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
@@
 supernova rate of each host is determined using 
-\citet{2012ApJ...755...61S}.  Supernovae are realized for a span of 10 observer-years yielding 17,205 objects.  This total supernova
-production underestimates the expected discovery of 130,000 supernovae in 10 observer-years based on the volumetric
-rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so the sample under consideration in this article
-corresponds to 1.3 years worth of discoveries.
+\citet{2012ApJ...755...61S}.   This host-based rate
+underestimates the expected discovery based on the volumetric
+rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
+by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
+does overestimate the total number of pre-maximum supernova discoveries, since most fields cannot continuously monitored throughout the year
+from most ground-based observatories.
 
@@
 \citet{2011ApJ...741...67D}, the effects in Eq.~18 are ignored.
+
+We consider a maximum redshift of $z=0.2$ because over ten years the noise is not shot-noise dominated, meaning that the differential improvement
+in the precision of
+$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
+would likely saturate available follow-up resources.
+Our calculations are thus based on the number and solid-angle of $z<0.2$ supernova pre-maximum discoveries 
+for the survey candidates provided by the Project.
 
@@
    %\topcaption{Table captions are better up top} % requires the topcapt package
-   \begin{tabular}{|cccc|cccc|} % Column formatting, @{} suppresses leading/trailing space
+   \begin{tabular}{|ccc|ccccc|} % Column formatting, @{} suppresses leading/trailing space
    \hline
@@
 \hline
-0.20 &  1.0 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 \\ \hline
-0.20 &  0.2 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
-0.20 &  0.3 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 \\
-0.20 &  0.5 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 \\
-0.10 &  1.0 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
-0.15 &  1.0 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
-0.20 &  1.0 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 \\
-0.20 &  1.0 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 \\
-0.15 &  0.3 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 \\
-0.12 &  0.3 & 0.10 & 746 &   1.62 &   0.70 &  0.085 &  11.28 \\
-0.07 &  1.0 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
+0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
+0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &  0.120 &  12.09 \\
+0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &  0.074 &  12.48 \\
+0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 \\
+0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 \\
+0.20 & 0.30 & 0.10 & 3023 &   1.40 &   0.54 &  0.047 &  12.68 \\
+0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &  0.037 &   9.95 \\
+0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 \\
+0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 \\
+0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 \\
+0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 \\
+0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 \\
+0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 \\
+0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 \\
+0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 \\
+0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &  0.053 &  20.79 \\
+0.20 & 0.65 & 0.10 & 6491 &   1.24 &   0.34 &  0.046 &  17.95 \\
+0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &  0.037 &  14.33 \\
+0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
+0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
+0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
+0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 \\
+0.20 & 1.00 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 \\
+0.20 & 1.00 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 \\
     \hline
    \end{tabular}
-   \caption{.}
-   \label{tab:booktabs}
+   \caption{Signal-to-noise in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
+   the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
+   $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
+   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The signal-to-noise, scaled to an LSST solid angle, is given as 
+    $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
+   \label{tab:booktabs}}
 \end{table}

### 2018-11-01 pv/view.py
handling the different parameters better

@@ -1,19 +1,89 @@
 import pickle
 import numpy
 from chainconsumer import *
+import matplotlib.pyplot as plt
+from astropy.cosmology import FlatLambdaCDM
+import scipy.stats
 
 # names = ['pvlist.0.08.1234.0.2.0.2.pkl','pvlist.0.08.1234.0.3.0.2.pkl','pvlist.0.08.1234.0.5.0.2.pkl','pvlist.0.08.1234.1.0.2.pkl','pvlist.0.08.1234.1.0.1.pkl','pvlist.0.08.1234.1.0.15.pkl']
-nsne = [2021,3023, 5015, 9901,1294,3870, 9901, 9901]
-zmax = [0.2,0.2,0.2,0.2,0.1,0.15,0.2,0.2,0.2]
-frac = [0.2,0.3,0.5,1,1,1,1,1,]
-snsig = [0.08,0.08,0.08,0.08,0.08,0.08,0.1,0.12]
+nsne = [2021,3023, 5015, 9901,1294,3870, 9901, 9901, 1215,746,563,6491, 873,2541,6491,6491, 3023, 3023, 4950]
+zmax = [0.2,0.2,0.2,0.2,0.1,0.15,0.2,0.2, 0.15,0.12,0.07,0.2,0.1, 0.15,0.2, 0.2,0.2, 0.2]
+frac = [0.2,0.3,0.5,1,1,1,1,1,0.3,0.3,1,0.65,0.65, 0.65, 0.65,0.65, 0.3, 0.3, 1]
+snsig = [0.08,0.08,0.08,0.08,0.08,0.08,0.1,0.12,0.12,0.1,0.08,0.08,0.08, 0.08, 0.1, 0.12, 0.1, 0.12, 0.08]
+tag = ["","","","","","","","","","","","","","", "skycut."]
+
+# print(len(frac),len(zmax))
+
+zmax_ = [0.07,0.1,0.15,0.2]
+frac_ = [0.2, 0.3,0.5, 0.65, 1]
+snsig_ = [0.08,0.1,0.12]
+
+names =[]
+zmax=[]
+frac=[]
+snsig=[]
+nsne=[]
+nsne_d = dict()
+nsne_d[(0.2,1)]=9901
+nsne_d[(0.2,0.65)]=6491
+nsne_d[(0.2,0.5)]=5015
[…]

### 2018-11-01 doc/article.tex
continuing on and making sure notation of plots and text are consistent.

@@
 have been used to test and ultimately find spatial correlations in peculiar velocities, may be attributed to the growth of structure
-\citet{2015JCAP...12..033H, 2017JCAP...05..015H}.  However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
+\citet{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.  However, the STON is currently insufficient to perform a meaningful test of GR.
 
@@
 C = P_{vv}(k) + \frac{\sigma^2}{n}
+\label{cov:eq}
 \end{equation}
@@
 While there have been a number of articles on the subject,
-our analysis brings a higher level of fidelity than sought by previous analyses.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
+our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
 catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
@@
 \section{Analysis}
-The analysis of this article is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}.   The peculiar magnitude correlation
-function $\xi_{\delta m \delta m}$ expected from General Relativity starting with the CMB
+The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}.  
+The peculiar magnitude is given by $\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda)$.  For our estimates we introduce one free parameter 
+$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
+
+The peculiar magnitude correlation
+function $\xi_{\delta m \delta m}$ expected from General Relativity acting on the CMB
 matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
-cosmological parameters used for the Buzzard catalog.  Our model for the data covariance
+cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
 is
 \begin{equation}
-C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij}.
+C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
 \end{equation}
+where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
 In this model, we treat the shape of the power spectrum $\xi$ as fixed and use the parameter $A$ to represent
-deviations from General Relativity $A=1$.  This model is not sensitive to  the shape of
-the power spectrum between gravitational models.  Nevertheless, for
+deviations from General Relativity $A=1$.  This model is not sensitive to differences in the shapes
+of  power spectra from different gravitational models.  Nevertheless, for
 convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
@@
 where $\bar{z}$ is the cosmological redshift of the host galaxy.
-\section{Results of Subsets}
+We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
+\citet{2015JCAP...12..033H} find that their results are insensitive to whether these corrections are included or not.
 
@@
 
-The 
+The posterior of the parameters is sampled using MCMC.
+The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
+
+\section{Results of Subsets}
+
+The projected STON in $f\sigma_8$ is calculated for different subsamples of the simulated data.
+For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
+determined.  The STON in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
+$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova-based peculiar velocity in the subset is given by 
+ $\sigma_A N_{gal}^{-0.5}$.
+
+Different parameters describing the input simulated data define the subsamples, 
+which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
+magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
+of sky is not monitored over the full year, we take
+$\phi \sim 0.65$ as an effective number density of discovered supernovae with light-coverage in 10 years.  
+The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
+For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
+is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for different subset analyses.
+
 \begin{figure}
 \centering
-\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.2.pkl.png}
+\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
 \caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
 absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
- Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.15$.
+to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
+part of the simulation.
 \label{zmax:fig}}
@@
    \hline
-$z_{max}$ & fraction & $\sigma_{SN}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+$z_{max}$ & $f$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
 \hline
@@
    \end{tabular}
-   \caption{Signal-to-noise in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
+   \caption{STON in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
    the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
    $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
-   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The signal-to-noise, scaled to an LSST solid angle, is given as 
+   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The STON, scaled to an LSST solid angle, is given as 
     $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
-   \label{tab:booktabs}}
+   \label{tab:subsets}}
 \end{table}
 
-We consider two different scenarios cutting at different redshift depths $z_{max}=0.1$ and  $z_{max}=0.15$.  All the objects in the first
-scenario are in the second scenario.  The results are shown in Fig.~\ref{zmax:fig}.  The points of interest are:
-\begin{itemize}
-\item There is a bias toward low $f\sigma_8$ for low-redshift ($z_{max}=0.1$).  Adding more supernova with increasing redshift
- decreases the bias  ($z_{max}=0.15$), and $A$ eventually approaches 1 ($z_{max}=0.2$ not shown).
- \item Based on the analysis of \citet{2017MNRAS.471.3135H}, I added an intrinsic velocity dispersion $\sigma_v$ (km~s$^{-1}$) due to non-linear effects.  This was not
- in Dragan's model.  My fit finds that this  dispersion is consistent with zero, not well constrained, and not strongly correlated with
- $A$.  The lack of covariance is consistent with Dragan's own tests.
- \item The fit intrinsic magnitude dispersion, on the other hand, is larger than the input $\sigma_M=0.08$ mag.  This  indicates that the catalogs
- contain 
- extra  velocity dispersion better attributed to intrinsic magnitude dispersion rather than correlated peculiar velocities.
-\end{itemize}
-
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.1.pkl.png}
-\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.2.pkl.png}
-\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
-absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
- Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.2$.
-\label{zmax:fig}}
-\end{figure}
+In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
+In Figure~\ref{scaling:fig} the effective LSST STON for $z_{max}=0.2$ is plotted as a function of input $n \sigma^{-2}_M$.
+The fact that the STON is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
+indicates that sample variance is non-negligible.
 
@@
 \includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
-absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
- Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.15$.
-\label{zmax:fig}}
+\caption{Left: Effective LSST STON for $z_{max}=0.2$ is plotted as a function of input  $n \sigma^{-2}_M$.
+In both plots the result of a linear fit is plotted for reference. Right: For $\phi=1$ and $\sigma_M$, the  effective LSST STON as
+a function of $z_{max}$>
+\label{scaling:fig}}
 \end{figure}
 
-\begin{itemize}
-\item Dragan: Is there some way that the formalism in your article breaks down at low-redshift?  I do remove very low-redshift objects at $z<0.01$,
-which I think is conservative.
- \item Risa/Joe: Is there any reasons for the Buzzard catalog to not capture correlated peculiar velocities on small scales (box size to large?) nor capture
- non-linear velocities?  (I don't know how to translate a mass resolution of 2.7e10 $M_{\odot}/h$ into a spatial resolution!)
-\end{itemize}
+A SN peculiar survey benefits from greater redshift depth $z_{max}$ with the increase in volume and numbers of supernovae.  However,
+peculiar velocity uncertainties increase with redshift and the follow-up of fainter high-redshift SNe requires more resources.  Table~\ref{tab:subsets}
+shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
+The Table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ that has 1294 supernovae yields a STON of 15.96, whereas
+a sample with $z_{max}=0.2$ that has (more) 2021 supernovae has a lower STON of 13.61.
+
+%We consider two different scenarios cutting at different redshift depths $z_{max}=0.1$ and  $z_{max}=0.15$.  All the objects in the first
+%scenario are in the second scenario.  The results are shown in Fig.~\ref{zmax:fig}.  The points of interest are:
+%\begin{itemize}
+%\item There is a bias toward low $f\sigma_8$ for low-redshift ($z_{max}=0.1$).  Adding more supernova with increasing redshift
+% decreases the bias  ($z_{max}=0.15$), and $A$ eventually approaches 1 ($z_{max}=0.2$ not shown).
+% \item Based on the analysis of \citet{2017MNRAS.471.3135H}, I added an intrinsic velocity dispersion $\sigma_v$ (km~s$^{-1}$) due to non-linear effects.  This was not
+% in Dragan's model.  My fit finds that this  dispersion is consistent with zero, not well constrained, and not strongly correlated with
+% $A$.  The lack of covariance is consistent with Dragan's own tests.
+% \item The fit intrinsic magnitude dispersion, on the other hand, is larger than the input $\sigma_M=0.08$ mag.  This  indicates that the catalogs
+% contain 
+% extra  velocity dispersion better attributed to intrinsic magnitude dispersion rather than correlated peculiar velocities.
+%\end{itemize}
+
+%\begin{figure}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.1.pkl.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.2.pkl.png}
+%\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
+%absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
+%to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
+% Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.2$.
+%\label{zmax:fig}}
+%\end{figure}
+%
+
+%
+%\begin{itemize}
+%\item Dragan: Is there some way that the formalism in your article breaks down at low-redshift?  I do remove very low-redshift objects at $z<0.01$,
+%which I think is conservative.
+% \item Risa/Joe: Is there any reasons for the Buzzard catalog to not capture correlated peculiar velocities on small scales (box size to large?) nor capture
+% non-linear velocities?  (I don't know how to translate a mass resolution of 2.7e10 $M_{\odot}/h$ into a spatial resolution!)
+%\end{itemize}
 

### 2018-11-01 pv/fit.py

@@ -117,9 +117,9 @@ if __name__ == "__main__":
         args.savef = args.nchain
 
     if args.skycut:
-        decmax=60.
-        bmin=-34.
-        bmax=34
+        decmax=-35.65
+        bmin=-90.
+        bmax=90
         fstr='skycut.'
     else:
         decmax=90
@@ -128,7 +128,7 @@ if __name__ == "__main__":
         fstr=''
 
     hg = HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path)
-    
+
     if args.antecedent is not None:
         chain = pickle.load(open('{}/pvlist.{}.{}.{}.{}.{}pkl.{}'.format(args.path,args.sigma_mu,args.seed,args.frac,args.zmax,fstr,args.antecedent), "rb" ) )
     else:
@@ -141,6 +141,7 @@ if __name__ == "__main__":
     #     print(usexi.shape)
     # elif (args.skycut)
         print (usexi.shape)
+
     else:
         usehg = hg.data
         usexi = hg.xi

### 2018-11-01 pv/view.py

@@ -88,7 +88,7 @@ def plotter(chain):
     c = ChainConsumer()
     dum = numpy.reshape(cutchain,(cutchain.shape[0]*cutchain.shape[1],cutchain.shape[2]))
     dum[:,3] = dum[:,3]*3e5
-    c.add_chain(dum, parameters=["$A$", "$M$","$\sigma_M$","$\sigma_{v}$"])
+    c.add_chain(dum, parameters=["$A$", "$\mathcal{M}$","$\sigma_M$","$\sigma_{v}$"])
     c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
 
 print("{} & {} & {} & {} & {} & {} & {} & {} \\\\".format("$z_{max}$", "fraction", "$\\sigma_{SN}$", "$N_{gal}$", \
@@ -102,7 +102,7 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
     # c = ChainConsumer()
     # dum = numpy.reshape(cutchain,(cutchain.shape[0]*cutchain.shape[1],cutchain.shape[2]))
     # dum[:,3] = dum[:,3]*3e5
-    # c.add_chain(dum, parameters=["$A$", "$M$","$\sigma_M$","$\sigma_{v}$"])
+    # c.add_chain(dum, parameters=["$A$", "$\mathcal{M}$","$\sigma_M$","$\sigma_{v}$"])
     # c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
 
     effston.append(cutchain[:,:,0].mean()/cutchain[:,:,0].std()*numpy.sqrt(18000./760))
@@ -127,8 +127,8 @@ volume = FlatLambdaCDM(71,0.265).comoving_volume(0.2).value
 slope, intercept, r_value, p_value, std_err = scipy.stats.linregress(1e3*nsne[w]/volume/snsig[w]**2,effston[w])
 plt.scatter(1e3*nsne[w]/volume/snsig[w]**2,effston[w])
 plt.plot([0,.7],[intercept,intercept+slope*0.7])
-plt.xlabel(r'$n/\sigma^2$ $\left[10^{-3} mag^{-2} Mpc^{-3}\right]$')
[…]

### 2018-11-02 doc/article.tex
for a version for greg

@@
 \usepackage{grffile}
-
+\usepackage{textcase}
 %SetFonts
@@
 
-\section{Connection between SNe~Ia and Gravity}
+\section{Connection between Type Ia Supernovae Correlations and Gravity}
 
-The growth of structure depends on the expansion history of the universe, the nature and density of its contents, and gravity. 
+The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
 It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the baryonic structures
-in the Universe and the peculiar velocities of test masses therein.
-Peculiar velocities are motions, on top of the cosmological expansion, caused by the gravitational attraction
-and repulsion of all sources of inhomogeneity in the Universe.
+in the Universe and from the peculiar velocities of test masses therein.
+Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
+and repulsion due to inhomogeneities in the Universe.  The peculiar velocity of an object wth a given absolute magnitude
+is determined from its observer magnitude and redshift.  Given a background cosmology, the observed magnitude provides
+an estimate of the cosmological redshift.  Peculiar velocity accounts for the difference between the cosmological and observer redshift.
 
-Baryonic structures and peculiar velocities  can provide a measurement of the combination $fD$, where $D$ is the linear growth factor that
-gives the overall amplitude of  overdensities and
+Baryonic structures and peculiar velocities  provide a measurement of the combination $fD$.  $D$ is the ``linear growth factor'' that
+gives the overall amplitude of  overdensities and the ``linear growth rate''
 $$f \equiv \frac{d\ln{D}}{d\ln{a}}$$ is how that amplitude changes with redshift.  General Relativity predicts
-$f=\Omega_M^\gamma$ (with $D$ determined accordingly) where $\gamma=0.55$, whereas other gravity models can be similarly described
-but with different values of $\gamma$
-\citep{2007APh....28..481L}.  Thus growth of structure, through the measurement of $fD$ (or $f\sigma_8$), provides a test of General Relativity.
-We typically refer to $f\sigma_8$ because the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres is often used to normalize the
-overall amplitude of  overdensities.
+$f \approx \Omega_M^\gamma$ (with $D$ determined accordingly) when $\gamma=0.55$, whereas other gravity models can be similarly described
+with different values of $\gamma$
+\citep{2007APh....28..481L}.  Thus growth of structure, through the measurement of $fD$, provides a test of General Relativity and breaks degeneracies
+between gravitational and dark energy models that explain the accelerating expansion of the Universe.
+The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
+commonly used to normalize the
+overall amplitude of  overdensities instead of $D$, and so the standard parameterization used by the community is $f\sigma_8$.
 
-As distance indicators SNe~Ia
+Baryonic structures are sensitive to $f\sigma_8$ through redshift space distortions (RSD), 
+which gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
+and $\mu$ gives the angular separation.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
+They are independent of the tracer bias because all sources of mass, predominantly dark matter, contribute to the dynamics. As
+mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
+leads to the dependence on $H_0 d_L$.  The cross-correlation between structure and peculiar velocity within the same volume come from the same
+overdensities and hence
+provide independent information unsusceptible to sample variance \citep{2007PhRvL..99h1301G}.
+
+Tracers with precisely known absolute magnitude provide precise cosmological redshifts and hence peculiar velocities.  
+The current generation of peculiar velocity studies use $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
+2016AJ....152...50T}. 
+Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample size by an order of magnitude.
+The galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.
+
+The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
+Nevertheless, their low intrinsic-magnitude uncertainties can 
 can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
 of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
-have been used to test and ultimately find spatial correlations in peculiar velocities, may be attributed to the growth of structure
-\citet{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.  However, the STON is currently insufficient to perform a meaningful test of GR.
+have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
+\citet{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.
+However, the ston is currently insufficient to perform a meaningful test of GR.
 
 Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
-First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\gtrsim 0.12$~mag distance modulus
-dispersion.  However, SNe transmit more information than a light curve shape and single color used in current analysis.
+First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
+dispersion.  However, SNe transmit more information than just the light curve shape and single color used in current analysis.
 Recent studies indicate that with the right data, SNe absolute
-magnitudes can be calibrated to $\lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.
-As absolute magnitude uncertainty propagates into fractional velocity uncertainty, improvements to the standard candle lead to
-improvements in their host peculiar velocities.
+magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.
+One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
 Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands.
+hundreds of thousands.  This is a sample size comparable to those projected by WALLABY and TAIPAN.
 
 The precision in  $f\sigma_8$ derived from an ten-year LSST-discovered  SN~Ia survey has been projected by \citet{2017ApJ...847..128H},
-from both their  clustering and peculiar velocities.
+from both their RSD and peculiar velocities.
 The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance uncertainty is assumed for each supernova.
 SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster, and Ly$\alpha$
-RSD measurements are sensitive.
-{\bf Greg do you have to power to add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
+RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
+{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
 
@@
 
-The projected precisions for LSST-discovered SNe~Ia have a number of interesting features.  Particular velocities contribute increasingly to
+The projected precisions for LSST-discovered SNe~Ia have a number of interesting features.  Relative to
+RSD, particular velocities contribute increasingly 
 $f\sigma_8$ precision with decreasing redshift.  At redshift 0.3 the differential gain from adding peculiar velocities to the RSD measurement is small.
-The differential gain in the cumulative $f\sigma_8$ constraint is small.  We therefore conclude that the focus should be in $z<0.3$ LSST SN~Ia discoveries.
-After 10 years for $z>0.1$ both sample variance and shot noise contribute to the error budget; adding even more SNe in the volume would improve
-the measurement though not as $\sqrt{N}$.
+Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that
+the differential gain in the cumulative  uncertainty is small for surveys with increasing maximum redshift.
+For $z>0.1$, the volume is sufficiently large that we are not yet sample variance limited  given the LSST sample.
 
-\section{Supernova Peculiar Velocity Survey}
+\section{Important Elements in Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
 \citet{2017ApJ...847..128H} perform their analysis in Fourier space, 
@@
 \end{equation}
-where the covariance
+where the covariance for the velocity-velocity correlation
 \begin{equation}
@@
 \end{equation}
-has contributions from the power spectrum, noise in the velocity measurement, and the density of velocity probes
-\citep{2017MNRAS.464.2517H}.   This equation is for the peculiar-velocity auto-correlation.
+is dependent on the power spectrum, noise in the velocity measurement, and the density of velocity probes
+\citep{2017MNRAS.464.2517H}.
 
-The parameters that determine parameter precision are the volume $V$, which
+The survey parameters that determine parameter precision are the volume $V$, which
 in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
 magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
-In the sample variance limit for a sample with fixed depth, the variance in $f\sigma_8$ (and other $\lambda$ parameters)
-is thus inversely proportional to the survey solid-angle $\Omega$, whereas
+For surveys of fixed depth, the variance of $f\sigma_8$ 
+in the sample-variance limit  
+is inversely proportional to the survey solid-angle $\Omega$, whereas
 in the shot-noise limit the variance is inversely proportional to $\Omega n^2 \propto N^2/\Omega$, where $n$ is the number density,
@@
 
-LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active survey area.
+LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
+The nominal observing strategy will obtain photometry every couple of days rotating through the 6 $grizy$ bands.
 
 The quality of supernova follow-up is an important factor in how well a peculiar-velocity survey can measure $f\sigma_8$.
-LSST light curves of low-redshift SNe, with their relatively poor sampling, plus a spectroscopic redshift
- can give $\sigma_M \sim 0.15$~mag. Non-LSST follow-up including infrared data or spectrophotometry at peak brightness are projected to give
+LSST SM~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
+ can give $\sigma_M \sim 0.15$~mag. Non-LSST follow-up including infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
+ \citep{2015ApJ...815...58F} are projected to give
 $\sigma_M \lesssim 0.08$~mag.
 Uncertainties in $f\sigma_8$ depend on $\sigma^2_M/n$ so the improvement between 0.08 and 0.15~mag dispersions is equivalent to a factor of 3.52
- in number density, or in survey duration given the expected completeness of $z<0.3$ SN~Ia discoveries.
+ in number density, or in survey duration given the expected saturation of $z<0.3$ SN~Ia discoveries.
 
@@
 This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
+This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
 While there have been a number of articles on the subject,
@@
 posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
-compare different survey parameters, such as redshift depth, total numbers of supernovae,
+compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
 solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
-We would like to
+Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
 quantify the suppression of sample variance achieved when considering matter-densities and velocities 
 within the same volume \citep{2007PhRvL..99h1301G}.
-The cross-correlation between galaxy-count and peculiar-velocity surveys is not yet implemented but is a planned extension of this work.
 
@@
 by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
-does overestimate the total number of pre-maximum supernova discoveries, since most fields cannot continuously monitored throughout the year
+is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
 from most ground-based observatories.
@@
 motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
-\citet{2011ApJ...741...67D}, the effects in Eq.~18 are ignored.
+\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
+and does not affect our conclusions.
 
-We consider a maximum redshift of $z=0.2$ because over ten years the noise is not shot-noise dominated, meaning that the differential improvement
+We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
 in the precision of
@@
 would likely saturate available follow-up resources.
-Our calculations are thus based on the number and solid-angle of $z<0.2$ supernova pre-maximum discoveries 
-for the survey candidates provided by the Project.
 
 \section{Analysis}
-The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}.  
-The peculiar magnitude is given by $\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda)$.  For our estimates we introduce one free parameter 
+The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
+The peculiar magnitude is given by 
+\begin{equation}
+\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
+\label{data:eqn}
+\end{equation}  We introduce one free parameter 
 $\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
@@
 The peculiar magnitude correlation
-function $\xi_{\delta m \delta m}$ expected from General Relativity acting on the CMB
+function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
 matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
@@
 C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
+\label{cov:eqn}
 \end{equation}
 where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
-In this model, we treat the shape of the power spectrum $\xi$ as fixed and use the parameter $A$ to represent
+In this model, we treat the shape of the power spectrum $\xi$ as fixed and use the amplitude parameter $A$ to represent
 deviations from General Relativity $A=1$.  This model is not sensitive to differences in the shapes
-of  power spectra from different gravitational models.  Nevertheless, for
+of  power spectra from different gravitational models.   Special cases of modified gravity could give $A=1$, but inconsistency with $A=1$ is is evidence against General Relativity.
+Nevertheless, for
 convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
-from General Relativity.  Special cases of modified gravity could give $A=1$, but inconsistency with $A=1$ is is evidence against General Relativity.
+from General Relativity. 
 The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
@@
 We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
-\citet{2015JCAP...12..033H} find that their results are insensitive to whether these corrections are included or not.
+Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
 
@@
 
-The posterior of the parameters is sampled using MCMC.
+The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
+covariance matrix in Eq.~\ref{cov:eqn}.
 The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
@@
 
-The projected STON in $f\sigma_8$ is calculated for different subsamples of the simulated data.
+The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
 For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
-determined.  The STON in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
-$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova-based peculiar velocity in the subset is given by 
- $\sigma_A N_{gal}^{-0.5}$.
+determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
+$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
+ $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
 
@@
 magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
-of sky is not monitored over the full year, we take
-$\phi \sim 0.65$ as an effective number density of discovered supernovae with light-coverage in 10 years.  
+of sky is not monitored over the full year.  
 The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
 For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
-is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for different subset analyses.
+is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
 
@@
 \includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
-\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
+\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
 absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
@@
    \hline
-$z_{max}$ & $f$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
 \hline
-0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
+0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 \\
+0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
+0.07 & 1.00 & 0.10 & 563 &   1.27 &   0.48 &  0.112 &  12.88 \\
+0.07 & 1.00 & 0.12 & 563 &   1.41 &   0.61 &  0.097 &  11.15 \\
+0.07 & 1.00 & 0.15 & 563 &   1.64 &   0.82 &  0.084 &   9.74 \\
 0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &  0.120 &  12.09 \\
+0.10 & 0.50 & 0.08 & 683 &   1.02 &   0.42 &  0.093 &  11.84 \\
+0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 \\
+0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 \\
+0.10 & 0.65 & 0.12 & 873 &   1.11 &   0.57 &  0.066 &   9.48 \\
+0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
+0.10 & 1.00 & 0.10 & 1294 &   0.95 &   0.36 &  0.072 &  12.66 \\
+0.10 & 1.00 & 0.12 & 1294 &   1.04 &   0.43 &  0.067 &  11.71 \\
 0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &  0.074 &  12.48 \\
 0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 \\
+0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 \\
+0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 \\
+0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 \\
+0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 \\
+0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
+0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
 0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 \\
@@
 0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &  0.037 &   9.95 \\
-0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 \\
 0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 \\
-0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 \\
-0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 \\
-0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 \\
-0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 \\
-0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 \\
-0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 \\
 0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &  0.053 &  20.79 \\
@@
 0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &  0.037 &  14.33 \\
-0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
-0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
-0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
 0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 \\
@@
    \end{tabular}
-   \caption{STON in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
+   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
    the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
    $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
-   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The STON, scaled to an LSST solid angle, is given as 
+   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The ston, scaled to an LSST solid angle, is given as 
     $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
@@
 In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
-In Figure~\ref{scaling:fig} the effective LSST STON for $z_{max}=0.2$ is plotted as a function of input $n \sigma^{-2}_M$.
-The fact that the STON is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
-indicates that sample variance is non-negligible.
+In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
+For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
+though the shallowing of the slopes for shallower surveys shows that it does contribute non-negligibly.
+The fact that the ston is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
+indicates that it is non-negligible.
 
 \begin{figure}
+\centering
 \includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
-\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Left: Effective LSST STON for $z_{max}=0.2$ is plotted as a function of input  $n \sigma^{-2}_M$.
-In both plots the result of a linear fit is plotted for reference. Right: For $\phi=1$ and $\sigma_M$, the  effective LSST STON as
-a function of $z_{max}$>
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
 \label{scaling:fig}}
@@
 
-A SN peculiar survey benefits from greater redshift depth $z_{max}$ with the increase in volume and numbers of supernovae.  However,
-peculiar velocity uncertainties increase with redshift and the follow-up of fainter high-redshift SNe requires more resources.  Table~\ref{tab:subsets}
+A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
+peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
+Table~\ref{tab:subsets}
 shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
-The Table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ that has 1294 supernovae yields a STON of 15.96, whereas
-a sample with $z_{max}=0.2$ that has (more) 2021 supernovae has a lower STON of 13.61.
+The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
+a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
+is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.
+
+$\phi \sim 0.65$ is an effective number density of discovered supernovae with light-curve coverage observed in 10 years
 

### 2018-11-02 pv/view.py

@@ -1,3 +1,5 @@
+#!/usr/bin/env python
+
 import pickle
 import numpy
 from chainconsumer import *
@@ -16,7 +18,7 @@ tag = ["","","","","","","","","","","","","","", "skycut."]
 
 zmax_ = [0.07,0.1,0.15,0.2]
 frac_ = [0.2, 0.3,0.5, 0.65, 1]
-snsig_ = [0.08,0.1,0.12]
+snsig_ = [0.08,0.1,0.12,0.15]
 
 names =[]
 zmax=[]
@@ -47,8 +49,8 @@ nsne_d[(0.07, 0.65)]=392
 
 from pathlib import Path
 
-for a2 in frac_:
-    for a1 in zmax_:
+for a1 in zmax_:
+    for a2 in frac_:
         for a3 in snsig_:
             f = Path('/Users/akim/project/PeculiarVelocity/outcosmo/pvlist.{}.1234.{}.{}.pkl'.format(a3,a2,a1))
             print(a1, a2, a3,end='')
@@ -121,25 +123,47 @@ effston = numpy.array(effston)
 # plt.ylabel('ston')
 # plt.savefig('frac_.png')
 # plt.clf()
-
-w = zmax ==0.2
-volume = FlatLambdaCDM(71,0.265).comoving_volume(0.2).value
-slope, intercept, r_value, p_value, std_err = scipy.stats.linregress(1e3*nsne[w]/volume/snsig[w]**2,effston[w])
-plt.scatter(1e3*nsne[w]/volume/snsig[w]**2,effston[w])
-plt.plot([0,.7],[intercept,intercept+slope*0.7])
+volumen = FlatLambdaCDM(71,0.265).comoving_volume(0.01).value
+for usez in numpy.unique(zmax):
+    w = zmax ==usez
+    volume = FlatLambdaCDM(71,0.265).comoving_volume(usez).value-volumen
+    slope, intercept, r_value, p_value, std_err = scipy.stats.linregress(1e3*nsne[w]/volume/snsig[w]**2,effston[w])
+    plt.scatter(1e3*nsne[w]/volume/snsig[w]**2,effston[w],label=r"$z_{{max}}={}$".format(usez))
 plt.xlabel(r'$n \sigma^{-2}_M$ $\left[10^{-3} mag^{-2} Mpc^{-3}\right]$')
 plt.ylabel('Effective LSST STON')
+plt.legend(loc=2)
[…]

### 2018-11-30 doc/article.tex
Respond to Segev comments

@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{S.~BenZvi}
+\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
 \author{C.~Harper}
@@
 
-The precision in  $f\sigma_8$ derived from an ten-year LSST-discovered  SN~Ia survey has been projected by \citet{2017ApJ...847..128H},
+The precision in  $f\sigma_8$ derived from the baseline WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
 from both their RSD and peculiar velocities.
@@
 
+ZTF and Future-ZTF.  Access close to $4\pi$.  LSST does not survey the southern equatorial pole.
+
 \appendix
@@
 where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
-In this model, we treat the shape of the power spectrum $\xi$ as fixed and use the amplitude parameter $A$ to represent
-deviations from General Relativity $A=1$.  This model is not sensitive to differences in the shapes
-of  power spectra from different gravitational models.   Special cases of modified gravity could give $A=1$, but inconsistency with $A=1$ is is evidence against General Relativity.
+The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
+from the linearized GR expectation of the shape of the power spectrum 
+$\xi$.
 Nevertheless, for
@@
 The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
+
+Direct tests of non-GR models for which Poisson's equation or the linear approximation
+does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work.
 

### 2018-11-30 pv/dragan/spline_functions.c
fast evaluation of cov mat

@@ -548,53 +548,75 @@ double Cij_theta_vel_integrand(double log10k, void *p)
     
     double chi1 = r_dimless_tab(z1)/H0_hmpcinv;
     double chi2 = r_dimless_tab(z2)/H0_hmpcinv;
-
+    
+    //v-v covariance fast computation
+    /*
     double prefac0 = pow(5/log(10.0), 2);
     // checked very small diff in results if the starting 1s are zeroed out
     double prefac1 = 1 - (1+z1)/H_dimless(z1, om, w0, wa)/r_dimless_tab(z1);
     double prefac2 = 1 - (1+z2)/H_dimless(z2, om, w0, wa)/r_dimless_tab(z2);
-    if (PRINT_FLAG == 1) printf("%f %f %d\n",  prefac0, prefac1, NSPLINE_PK_CAMB);
+    //if (PRINT_FLAG == 1) printf("%f %f %d\n",  prefac0, prefac1, NSPLINE_PK_CAMB);
     double prefac3 =  1/(2*M_PI*M_PI) *
         log(10.0) * k *                        // conversion: they have dk, I want dlog10(k)
         Pk_CAMB_NR_tab(k) *                       // P(k) from CAMB
         dD_deta_tab(z1) * dD_deta_tab(z2)      // units are [h Mpc^{-1}]^2
         ;
-
-    int lmax;
-
-    /////////////////////////////////////////////////////////////////////////
-    // 16 October 2016: found sufficient conditions on lmax;
-    // drastically smaller lmax in all cases except aligned, high-z objects
-    /////////////////////////////////////////////////////////////////////////
-    //if ((z1 > 0.05) && (z2 > 0.05) && (costh > 0.95)) lmax = LMAX_CIJ_THETA;
-    if (costh > 0.95)  lmax = LMAX_CIJ_THETA;
-    else               lmax = 20.0;
-    //FIX
-    //lmax=10;
     
-    for (ell=0; ell<=lmax; ell++)
-    {
-        if (WINDOW == 3) // no mean of magnitude subtracted
-            fac = gsl_sf_legendre_Pl(ell,costh);
-        if (WINDOW == 2) // mean subtracted with arbitrary window
-        {
-            fac = gsl_sf_legendre_Pl(ell,costh) -
[…]

### 2018-12-12 pv/fit.py
more output

@@ -130,7 +130,7 @@ if __name__ == "__main__":
     hg = HostGalaxies(sigma_mu=args.sigma_mu, catseed=args.seed, path=args.path)
     
     if args.antecedent is not None:
-        chain = pickle.load(open('{}/pvlist.{}.{}.{}.{}.{}pkl.{}'.format(args.path,args.sigma_mu,args.seed,args.frac,args.zmax,fstr,args.antecedent), "rb" ) )
+        chain = pickle.load(open('{}/debug/pvlist.{}.{}.{}.{}.{}pkl.{}'.format(args.path,args.sigma_mu,args.seed,args.frac,args.zmax,fstr,args.antecedent), "rb" ) )
     else:
         chain=None
 
@@ -163,7 +163,7 @@ if __name__ == "__main__":
         else:
             indnm = i+args.antecedent+1
 
-        pickle.dump(chain, open('{}/pvlist.{}.{}.{}.{}.{}pkl.{}'.format(args.path,args.sigma_mu,args.seed,args.frac,args.zmax,fstr,chain.shape[1]), "wb" ) )
+        pickle.dump(chain, open('{}/debug/pvlist.{}.{}.{}.{}.{}pkl.{}'.format(args.path,args.sigma_mu,args.seed,args.frac,args.zmax,fstr,chain.shape[1]), "wb" ) )
 
 #srun -n 1 -c 64 --cpu-bind=sockets python fit.py --path ../out/ --frac 0.5  --nchain 2
 #python fit.py --path ../out/ --nchain 1 frac 0.19

### 2018-12-12 pv/view.py

@@ -3,39 +3,46 @@ import numpy
 from chainconsumer import *
 
 # names = ['pvlist.0.08.1234.0.2.0.2.pkl','pvlist.0.08.1234.0.3.0.2.pkl','pvlist.0.08.1234.0.5.0.2.pkl','pvlist.0.08.1234.1.0.2.pkl','pvlist.0.08.1234.1.0.1.pkl','pvlist.0.08.1234.1.0.15.pkl']
+
 nsne = [2021,3023, 5015, 9901,1294,3870, 9901, 9901]
 zmax = [0.2,0.2,0.2,0.2,0.1,0.15,0.2,0.2,0.2]
-frac = [0.2,0.3,0.5,1,1,1,1,1,]
+#frac = [0.2,0.3,0.5,1,1,1,1,1,]
 snsig = [0.08,0.08,0.08,0.08,0.08,0.08,0.1,0.12]
 
 
 names = []
-for fra,zma,snsi in zip(frac,zmax,snsig):
-    names.append('pvlist.{}.1234.{}.{}.pkl'.format(snsi,fra,zma))
+#for fra,zma,snsi in zip(frac,zmax,snsig):
+#    names.append('pvlist.{}.1234.{}.{}.pkl'.format(snsi,fra,zma))
 
 
-def plotter(chain):
-    c = ChainConsumer()
-    dum = numpy.reshape(cutchain,(cutchain.shape[0]*cutchain.shape[1],cutchain.shape[2]))
-    dum[:,3] = dum[:,3]*3e5
-    c.add_chain(dum, parameters=["$A$", "$M$","$\sigma_M$","$\sigma_{v}$"])
-    c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
+#def plotter(chain):
+    #c = ChainConsumer()
+    #dum = numpy.reshape(cutchain,(cutchain.shape[0]*cutchain.shape[1],cutchain.shape[2]))
+    #dum[:,3] = dum[:,3]*3e5
+    #c.add_chain(dum, parameters=["$A$", "$M$","$\sigma_M$","$\sigma_{v}$"])
+    #c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
 
[…]

### 2018-12-12 doc/article.tex
typo in column header

@@
 \begin{abstract}
-ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.2$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
 hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
-their absolute magnitudes, now make them competitive probes of the growth of structure.
+their absolute magnitudes, now make them competitive probes of the growth of structure.  The peculiar velocity power spectrum derived from SNe~Ia
+is sensitive to the product of the linear growth and amplitude of structure
+$f\sigma_8$.  Cross-correlation with synergistic galaxy surveys further constrains $f \sigma_8$ and the galaxy bias.  Velocities
+SNe~Ia in the next decade will measure the growth of structure  a powerful test of General Relativity in the local $z<0.2$ Universe.
 \end{abstract}
@@
    \hline
-$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
 \hline
-0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 \\
-0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
-0.07 & 1.00 & 0.10 & 563 &   1.27 &   0.48 &  0.112 &  12.88 \\
-0.07 & 1.00 & 0.12 & 563 &   1.41 &   0.61 &  0.097 &  11.15 \\
-0.07 & 1.00 & 0.15 & 563 &   1.64 &   0.82 &  0.084 &   9.74 \\
-0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &  0.120 &  12.09 \\
-0.10 & 0.50 & 0.08 & 683 &   1.02 &   0.42 &  0.093 &  11.84 \\
-0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 \\
-0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 \\
-0.10 & 0.65 & 0.12 & 873 &   1.11 &   0.57 &  0.066 &   9.48 \\
-0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
-0.10 & 1.00 & 0.10 & 1294 &   0.95 &   0.36 &  0.072 &  12.66 \\
-0.10 & 1.00 & 0.12 & 1294 &   1.04 &   0.43 &  0.067 &  11.71 \\
-0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &  0.074 &  12.48 \\
-0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 \\
-0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 \\
-0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 \\
-0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 \\
-0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 \\
-0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
-0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
-0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 \\
-0.20 & 0.30 & 0.10 & 3023 &   1.40 &   0.54 &  0.047 &  12.68 \\
-0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &  0.037 &   9.95 \\
-0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 \\
-0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &  0.053 &  20.79 \\
-0.20 & 0.65 & 0.10 & 6491 &   1.24 &   0.34 &  0.046 &  17.95 \\
-0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &  0.037 &  14.33 \\
-0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 \\
-0.20 & 1.00 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 \\
-0.20 & 1.00 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 \\
+0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &   0.14 &  13.21 \\
+0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &   0.13 &  15.55 \\
+0.07 & 1.00 & 0.10 & 563 &   1.27 &   0.48 &   0.11 &  12.88 \\
+0.07 & 1.00 & 0.12 & 563 &   1.41 &   0.61 &   0.10 &  11.15 \\
+0.07 & 1.00 & 0.15 & 563 &   1.64 &   0.82 &   0.08 &   9.74 \\
+0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &   0.12 &  12.09 \\
+0.10 & 0.50 & 0.08 & 683 &   1.02 &   0.42 &   0.09 &  11.84 \\
+0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &   0.09 &  12.57 \\
+0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &   0.07 &  10.72 \\
+0.10 & 0.65 & 0.12 & 873 &   1.11 &   0.57 &   0.07 &   9.48 \\
+0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &   0.09 &  15.96 \\
+0.10 & 1.00 & 0.10 & 1294 &   0.95 &   0.36 &   0.07 &  12.66 \\
+0.10 & 1.00 & 0.12 & 1294 &   1.04 &   0.43 &   0.07 &  11.71 \\
+0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &   0.07 &  12.48 \\
+0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &   0.06 &  10.36 \\
+0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &   0.07 &  14.93 \\
+0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &   0.07 &  17.05 \\
+0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &   0.06 &  14.43 \\
+0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &   0.05 &  12.42 \\
+0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &   0.07 &  20.00 \\
+0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &   0.06 &  13.61 \\
+0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &   0.06 &  15.16 \\
+0.20 & 0.30 & 0.10 & 3023 &   1.40 &   0.54 &   0.05 &  12.68 \\
+0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &   0.04 &   9.95 \\
+0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &   0.05 &  18.54 \\
+0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &   0.05 &  20.79 \\
+0.20 & 0.65 & 0.10 & 6491 &   1.24 &   0.34 &   0.05 &  17.95 \\
+0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &   0.04 &  14.33 \\
+0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &   0.05 &  24.38 \\
+0.20 & 1.00 & 0.10 & 9901 &   1.06 &   0.25 &   0.04 &  20.33 \\
+0.20 & 1.00 & 0.12 & 9901 &   1.09 &   0.28 &   0.04 &  19.13 \\
     \hline

### 2018-12-12 doc/article.tex
add one more level of precision

--- 2018-11-30 doc/article.tex
@@
    \hline
-$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\sigma_A N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
 \hline

### 2018-12-12 doc/article.tex
fix exponent in table header

@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{D.~Huterer}
+\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
+Arbor, MI 48109, USA }
 \author{C.~Ju}
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
-\author{D.~Huterer}
-\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
-Arbor, MI 48109, USA }
 
@@
 hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
-their absolute magnitudes, now make them competitive probes of the growth of structure.  The peculiar velocity power spectrum derived from SNe~Ia
-is sensitive to the product of the linear growth and amplitude of structure
-$f\sigma_8$.  Cross-correlation with synergistic galaxy surveys further constrains $f \sigma_8$ and the galaxy bias.  Velocities
-SNe~Ia in the next decade will measure the growth of structure  a powerful test of General Relativity in the local $z<0.2$ Universe.
+their absolute magnitudes, now make them competitive probes of the growth of structure.  The peculiar velocity power spectrum
+is sensitive to $f\sigma_8$,
+the product of the linear growth and amplitude of density perturbations.  Cross-correlation with synergistic galaxy surveys further constrains $f \sigma_8$ and the galaxy bias. 
+Thus, in the next decade the peculiar velocities of
+SNe~Ia will provide the growth of structure  in the local $z<0.2$ Universe as a powerful test of General Relativity and other models of gravity.
 \end{abstract}
@@
 Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
-and repulsion due to inhomogeneities in the Universe.  The peculiar velocity of an object wth a given absolute magnitude
-is determined from its observer magnitude and redshift.  Given a background cosmology, the observed magnitude provides
-an estimate of the cosmological redshift.  Peculiar velocity accounts for the difference between the cosmological and observer redshift.
+and repulsion of densitiy inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
+is determined from its observer magnitude and redshift. For a given a background cosmology, the observed magnitude provides
+an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and observer redshifts.
 
 Baryonic structures and peculiar velocities  provide a measurement of the combination $fD$.  $D$ is the ``linear growth factor'' that
-gives the overall amplitude of  overdensities and the ``linear growth rate''
+gives the overall amplitude of  overdensities, and the ``linear growth rate''
 $$f \equiv \frac{d\ln{D}}{d\ln{a}}$$ is how that amplitude changes with redshift.  General Relativity predicts
-$f \approx \Omega_M^\gamma$ (with $D$ determined accordingly) when $\gamma=0.55$, whereas other gravity models can be similarly described
+$f \approx \Omega_M^\gamma$ (with $D$ determined accordingly) with $\gamma=0.55$, whereas other gravity models can be similarly described
 with different values of $\gamma$
-\citep{2007APh....28..481L}.  Thus growth of structure, through the measurement of $fD$, provides a test of General Relativity and breaks degeneracies
+\citep{2007APh....28..481L}.  The growth of structure, through the measurement of $fD$, provides a test of General Relativity and breaks degeneracies
 between gravitational and dark energy models that explain the accelerating expansion of the Universe.
 The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
-commonly used to normalize the
-overall amplitude of  overdensities instead of $D$, and so the standard parameterization used by the community is $f\sigma_8$.
+commonly used in place of $D$ to normalize the
+overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.
 
@@
 and $\mu$ gives the angular separation.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
-They are independent of the tracer bias because all sources of mass, predominantly dark matter, contribute to the dynamics. As
+Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and are so bias-free.
+As
 mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
-leads to the dependence on $H_0 d_L$.  The cross-correlation between structure and peculiar velocity within the same volume come from the same
-overdensities and hence
-provide independent information unsusceptible to sample variance \citep{2007PhRvL..99h1301G}.
+leads to the dependence on $H_0 d_L$. 
+Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
+insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
+than their naive sum.
 
-Tracers with precisely known absolute magnitude provide precise cosmological redshifts and hence peculiar velocities.  
+The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
 The current generation of peculiar velocity studies use $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
@@
 Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample size by an order of magnitude.
-The galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.
+These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.
 
@@
    \hline
-$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
 \hline

### 2018-12-12 doc/article.tex
before sending to Satya

@@
 \begin{abstract}
-ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.2$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
 hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
@@
 Thus, in the next decade the peculiar velocities of
-SNe~Ia will provide the growth of structure  in the local $z<0.2$ Universe as a powerful test of General Relativity and other models of gravity.
+SNe~Ia will provide the growth of structure  in the local $z<0.3$ Universe as a powerful test of General Relativity and other models of gravity.
 \end{abstract}
@@
 The current generation of peculiar velocity studies use $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
-2016AJ....152...50T}. 
-Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample size by an order of magnitude.
-These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.
+2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
+Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample size by an order of magnitude
+over half the sky to a depth of $z=0.1$.
 
@@
 have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
-\citet{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.
-However, the ston is currently insufficient to perform a meaningful test of GR.
+\citep{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.
+However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
 
@@
 First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
-dispersion.  However, SNe transmit more information than just the light curve shape and single color used in current analysis.
+dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
 Recent studies indicate that with the right data, SNe absolute
@@
 Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands.  This is a sample size comparable to those projected by WALLABY and TAIPAN.
+hundreds of thousands.  This is a sample size comparable to the number of galaxies projected by WALLABY and TAIPAN.
 
@@
 
-The projected precisions for LSST-discovered SNe~Ia have a number of interesting features.  Relative to
-RSD, particular velocities contribute increasingly 
-$f\sigma_8$ precision with decreasing redshift.  At redshift 0.3 the differential gain from adding peculiar velocities to the RSD measurement is small.
+The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
 Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
-the differential gain in the cumulative  uncertainty is small for surveys with increasing maximum redshift.
-For $z>0.1$, the volume is sufficiently large that we are not yet sample variance limited  given the LSST sample.
+there is little benefit in going beyond
+$z=0.3$.
+The constraining power on  $f\sigma_8$ 
+comes primarily from peculiar velocities
+at $z<0.2$, but are dominated by RSD by $z=0.3$.
+It turns out that for $z>0.1$, the volume is sufficiently large that the LSST sample is not yet sample variance limited.
 
@@
 where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
-approximated by
 \begin{equation}

### 2018-12-13 doc/article.tex
Table only had STON in A.  Now provide it for fsigma8

@@
 
+The precision in the determination of the combination $\lambda= f\sigma_8$ is estimated using the Fisher matrix.  To first order, 
+$P_{vv} \propto (f\sigma_8)^2$.
+
 The survey parameters that determine parameter precision are the volume $V$, which
@@
 LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-The nominal observing strategy will obtain photometry every couple of days rotating through the 6 $grizy$ bands.
+Equating LSST discovery with having a meaningful peculiar velocity, $n$ is thus a non-tunable parameter.  This leaves
+solid angle, redshift depth, and intrinsic magnitude dispersion as the important tunable parameters in the design of a peculiar velocity survey.
 
+\subsection{Solid Angle}
+
+The variance in $f\sigma_8$ is proportional to solid angle $\Omega$ given a fixed redshift depth and  $\sigma^2_M/n$.  The baseline
+LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
+the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+coverage of the southern equatorial pole, could double
+the sky coverage and halve the variance in $f\sigma_8$.
+
+\subsection{Intrinsic Magnitude Dispersion}
 The quality of supernova follow-up is an important factor in how well a peculiar-velocity survey can measure $f\sigma_8$.
-LSST SM~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
- can give $\sigma_M \sim 0.15$~mag. Non-LSST follow-up including infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
+Supplemental non-LSST data can access SN models that yield lower intrinsic-magnitude dispersion and systematic errors.
+LSST SN~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
+ can give $\sigma_M \sim 0.15$~mag.
+ Non-LSST follow-up including infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
  \citep{2015ApJ...815...58F} are projected to give
@@
 Uncertainties in $f\sigma_8$ depend on $\sigma^2_M/n$ so the improvement between 0.08 and 0.15~mag dispersions is equivalent to a factor of 3.52
- in number density, or in survey duration given the expected saturation of $z<0.3$ SN~Ia discoveries.
-
-The variance in $f\sigma_8$ is proportional to solid angle $\Omega$ given a fixed redshift depth and  $\sigma^2_M/n$.  The baseline
-LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<-0.3$ SN~Ia searches, beyond
-the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-coverage of the southern pole, could double
-the sky coverage and halve the variance in $f\sigma_8$.
+ in number density, or in survey duration. 
+While LSST data can be used to obtain photometric redshifts from host galaxies and supernova light curves, the spectroscopic
+redshifts needed for precise host-galaxy peculiar velocities require other resources.
 
@@
 
-ZTF and Future-ZTF.  Access close to $4\pi$.  LSST does not survey the southern equatorial pole.
+The nominal observing strategy will obtain photometry every couple of days rotating through the 6 $grizy$ bands.  Although SNe have good
+overall sampling, the per-band light curves are poorly sampled.
 
@@
    %\topcaption{Table captions are better up top} % requires the topcapt package
-   \begin{tabular}{|ccc|ccccc|} % Column formatting, @{} suppresses leading/trailing space
+   \begin{tabular}{|ccc|ccccc|c|} % Column formatting, @{} suppresses leading/trailing space
    \hline
-$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$ \\
+$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$&  STON $f\sigma_8$ \\
 \hline
-0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 \\
-0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 \\
-0.07 & 1.00 & 0.10 & 563 &   1.27 &   0.48 &  0.112 &  12.88 \\
-0.07 & 1.00 & 0.12 & 563 &   1.41 &   0.61 &  0.097 &  11.15 \\
-0.07 & 1.00 & 0.15 & 563 &   1.64 &   0.82 &  0.084 &   9.74 \\
-0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &  0.120 &  12.09 \\
-0.10 & 0.50 & 0.08 & 683 &   1.02 &   0.42 &  0.093 &  11.84 \\
-0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 \\
-0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 \\
-0.10 & 0.65 & 0.12 & 873 &   1.11 &   0.57 &  0.066 &   9.48 \\
-0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 \\
-0.10 & 1.00 & 0.10 & 1294 &   0.95 &   0.36 &  0.072 &  12.66 \\
-0.10 & 1.00 & 0.12 & 1294 &   1.04 &   0.43 &  0.067 &  11.71 \\
-0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &  0.074 &  12.48 \\
-0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 \\
-0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 \\
-0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 \\
-0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 \\
-0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 \\
-0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 \\
-0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 \\
-0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 \\
-0.20 & 0.30 & 0.10 & 3023 &   1.40 &   0.54 &  0.047 &  12.68 \\
-0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &  0.037 &   9.95 \\
-0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 \\
-0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &  0.053 &  20.79 \\
-0.20 & 0.65 & 0.10 & 6491 &   1.24 &   0.34 &  0.046 &  17.95 \\
-0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &  0.037 &  14.33 \\
-0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 \\
-0.20 & 1.00 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 \\
-0.20 & 1.00 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 \\
+0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 &  26.42\\
+0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 &  31.11\\
+0.07 & 1.00 & 0.10 & 563 &   1.27 &   0.48 &  0.112 &  12.88 &  25.77\\
+0.07 & 1.00 & 0.12 & 563 &   1.41 &   0.61 &  0.097 &  11.15 &  22.30\\
+0.07 & 1.00 & 0.15 & 563 &   1.64 &   0.82 &  0.084 &   9.74 &  19.48\\
+0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &  0.120 &  12.09 &  24.18\\
+0.10 & 0.50 & 0.08 & 683 &   1.02 &   0.42 &  0.093 &  11.84 &  23.67\\
+0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 &  25.14\\
+0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 &  21.43\\
+0.10 & 0.65 & 0.12 & 873 &   1.11 &   0.57 &  0.066 &   9.48 &  18.96\\
+0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 &  31.91\\
+0.10 & 1.00 & 0.10 & 1294 &   0.95 &   0.36 &  0.072 &  12.66 &  25.33\\
+0.10 & 1.00 & 0.12 & 1294 &   1.04 &   0.43 &  0.067 &  11.71 &  23.41\\
+0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &  0.074 &  12.48 &  24.96\\
+0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 &  20.73\\
+0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 &  29.86\\
+0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 &  34.11\\
+0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 &  28.87\\
+0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 &  24.83\\
+0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 &  40.00\\
+0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 &  27.21\\
+0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 &  30.32\\
+0.20 & 0.30 & 0.10 & 3023 &   1.40 &   0.54 &  0.047 &  12.68 &  25.37\\
+0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &  0.037 &   9.95 &  19.89\\
+0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 &  37.07\\
+0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &  0.053 &  20.79 &  41.58\\
+0.20 & 0.65 & 0.10 & 6491 &   1.24 &   0.34 &  0.046 &  17.95 &  35.89\\
+0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &  0.037 &  14.33 &  28.65\\
+0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 &  48.77\\
+0.20 & 1.00 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 &  40.66\\
+0.20 & 1.00 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 &  38.26\\
     \hline

### 2018-12-13 pv/view.py

@@ -108,10 +108,10 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
     # c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
 
     effston.append(cutchain[:,:,0].mean()/cutchain[:,:,0].std()*numpy.sqrt(18000./760))
-    print("{:4.2f} & {:4.2f} & {:4.2f} & {} & {:6.2f} & {:6.2f} & {:6.3f} & {:6.2f} \\\\".format( \
+    print("{:4.2f} & {:4.2f} & {:4.2f} & {} & {:6.2f} & {:6.2f} & {:6.3f} & {:6.2f} & {:6.2f}\\\\".format( \
         zma,fra,snsi,nsn,cutchain[:,:,0].mean(), cutchain[:,:,0].std(), \
         cutchain[:,:,0].mean()/cutchain[:,:,0].std()/numpy.sqrt(nsn), \
-        effston[-1]))
+        effston[-1],effston[-1]*2))
 
 effston = numpy.array(effston)
 

### 2018-12-13 doc/article.tex
forging on

@@
 \citep{2017MNRAS.464.2517H}.  
-
-The precision in the determination of the combination $\lambda= f\sigma_8$ is estimated using the Fisher matrix.  To first order, 
-$P_{vv} \propto (f\sigma_8)^2$.
+The primary dependence on the growth of structure is through
+\begin{equation}
+\frac{\partial C}{\partial \lambda} = \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
+\end{equation}
+for the parameter choice
+$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.
 
@@
 magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
+LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
+Equating LSST discovery with having a meaningful peculiar velocity, $n$ is thus a non-tunable parameter.
+
 For surveys of fixed depth, the variance of $f\sigma_8$ 
@@
 and $N$ is the total number of supernovae.
-
-LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-Equating LSST discovery with having a meaningful peculiar velocity, $n$ is thus a non-tunable parameter.  This leaves
-solid angle, redshift depth, and intrinsic magnitude dispersion as the important tunable parameters in the design of a peculiar velocity survey.
 

### 2018-12-13 pv/view.py
new plot to show that not in shot-noise limit

@@ -106,14 +106,37 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
     # dum[:,3] = dum[:,3]*3e5
     # c.add_chain(dum, parameters=["$A$", "$\mathcal{M}$","$\sigma_M$","$\sigma_{v}$"])
     # c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
-
-    effston.append(cutchain[:,:,0].mean()/cutchain[:,:,0].std()*numpy.sqrt(18000./760))
-    print("{:4.2f} & {:4.2f} & {:4.2f} & {} & {:6.2f} & {:6.2f} & {:6.3f} & {:6.2f} & {:6.2f}\\\\".format( \
-        zma,fra,snsi,nsn,cutchain[:,:,0].mean(), cutchain[:,:,0].std(), \
-        cutchain[:,:,0].mean()/cutchain[:,:,0].std()/numpy.sqrt(nsn), \
-        effston[-1],effston[-1]*2))
+    fsigma8 = numpy.sqrt(cutchain[:,:,0])
+    effston.append(fsigma8.mean()/fsigma8.std()*numpy.sqrt(18000./760))
+    # print("{:4.2f} & {:4.2f} & {:4.2f} & {} & {:6.2f} & {:6.2f} & {:6.3f} & {:6.2f} \\\\".format( \
+    #     zma,fra,snsi,nsn,fsigma8 .mean(), fsigma8 .std(), \
+    #     fsigma8.mean()/fsigma8.std()/numpy.sqrt(nsn), \
+    #     effston[-1]))
+    print("{:4.2f} & {:4.2f} & {:4.2f}  & {:6.3f} & {:6.2f} \\\\".format( \
+        zma,fra,snsi, effston[-1], \
+        effston[-1]/numpy.sqrt(18000./760*nsn)\
+        ))
 
 effston = numpy.array(effston)
+
+for z_ in zmax_:
+    fs = []
+    ef = []
+    for f_ in frac_:
+        for sn_ in snsig_:
+            w = numpy.logical_and.reduce((frac==f_ ,zmax==z_, snsig==sn_))
+            if w.sum() != 0:
+                fs.append(sn_**2/f_)
+                ef.append(effston[w])
+    fs = numpy.array(fs)*0.65/0.1**2
[…]

### 2018-12-14 doc/article.tex
forging on in edits

@@
 Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands.  This is a sample size comparable to the number of galaxies projected by WALLABY and TAIPAN.
+hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia for which good light curves can be measured.
+This is a sample size comparable to the number of galaxies projected by WALLABY and TAIPAN.
 
@@
 
-\section{Important Elements in Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
+\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
+\subsection{Parameters that Determine Scientific Performance}
+
 \citet{2017ApJ...847..128H} perform their analysis in Fourier space, 
@@
 \end{equation}
-is dependent on the power spectrum, noise in the velocity measurement, and the density of velocity probes
+is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
 \citep{2017MNRAS.464.2517H}.  
@@
 magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
+
 LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-Equating LSST discovery with having a meaningful peculiar velocity, $n$ is thus a non-tunable parameter.
+Equating LSST discovery with having a meaningful peculiar velocity, the number density is$n$ is thus a non-tunable parameter, which only increases with survey duration.
+The remaining free parameters are then $\Omega$, $z_{max}$, and $\sigma_M$.
 
-For surveys of fixed depth, the variance of $f\sigma_8$ 
-in the sample-variance limit  
-is inversely proportional to the survey solid-angle $\Omega$, whereas
-in the shot-noise limit the variance is inversely proportional to $\Omega n^2 \propto N^2/\Omega$, where $n$ is the number density,
-and $N$ is the total number of supernovae.
+While LSST data can be used to obtain photometric redshifts from host galaxies and supernova light curves, the spectroscopic
+redshifts needed for precise host-galaxy peculiar velocities require other resources.
 
-\subsection{Solid Angle}
-
-The variance in $f\sigma_8$ is proportional to solid angle $\Omega$ given a fixed redshift depth and  $\sigma^2_M/n$.  The baseline
+\begin{itemize}
+\item Solid Angle $\Omega$:
+The variance in $f\sigma_8$ is proportional to solid angle $\Omega$.  The baseline
 LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
@@
 the sky coverage and halve the variance in $f\sigma_8$.
+\item Redshift Depth $z_{max}$:
+For the number densities and intrinsic magnitude dispersions relevant to LSST, the noise is not sample-variance
+limited.  We have run simulations of LSST SN surveys
+and calculated the signal-to-noise (STON) of $f\sigma_8$ from peculiar velocities. These simulations show, as seen in the left plot
+of  Figure~\ref{notshot:fig}, that an increase in the number the number density of tracers leads to significant improvements in STON.
+It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of  STON on $\sigma_M^2/n$.
 
-\subsection{Intrinsic Magnitude Dispersion}
-The quality of supernova follow-up is an important factor in how well a peculiar-velocity survey can measure $f\sigma_8$.
-Supplemental non-LSST data can access SN models that yield lower intrinsic-magnitude dispersion and systematic errors.
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
+\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
+\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities.  For several limiting redshift depths, the STON is plotted as a
+function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
+Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
+\label{notshot:fig}}
+\end{figure}
+
+The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers as shown in the
+right plot of   Figure~\ref{notshot:fig}, shows that LSST is in the regime where adding one tracer at lower redshift better improves the 
+STON relative to the addition of one at higher redshift.
+
+Scientific leverage dictates that low-redshift sources are most valuable.
+This aligns with observational considerations, for which closer and hence brighter objects require more modest follow-up resources.
+Therefore, $z_{max}$ depends on the follow-up program as the redshift where complete follow-up saturates resources.
+
+\item Intrinsic Magnitude Dispersion $\sigma_M$:
+Just as an increase in tracer number density beyond that of a ten-year LSST survey improves the STON in the measurement
+of $f\sigma_8$, as noted in the discussion on $z_{max}$, so too does a decrease in the supernova intrinsic magnitude dispersion.
+
+The nominal LSST WFD observing plan is to have
+each piece of sky observed every couple of days rotating through the 6 $grizy$ bands. The data is sufficient to have effectively
+complete $z<0.3$ SN~Ia discoveries.  However,  the per-band light curves are poorly sampled, which limits the accuracy with which distances
+can be derived.
 LSST SN~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
  can give $\sigma_M \sim 0.15$~mag.
- Non-LSST follow-up including infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
+
+Supplemental non-LSST follow-up data in the form or improved temporal light-curve sampling (accurate rise and decline times), 
+expanded (UV, NIR) wavelength coverage, and spectral features
+can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
+For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
  \citep{2015ApJ...815...58F} are projected to give
@@
  in number density, or in survey duration. 
-While LSST data can be used to obtain photometric redshifts from host galaxies and supernova light curves, the spectroscopic
-redshifts needed for precise host-galaxy peculiar velocities require other resources.
+\end{itemize}
 
-LSST can be taken to be a SN~Ia discovery plus light-curve builder, or a discovery machine only.  The former comes with a simplified follow-up program of typing
-and/or redshift collection but with higher intrinsic magnitude dispersion.  The latter requires additional follow-up resources to gather more information per supernova to
-lower the intrinsic magnitude dispersion.
+\subsection{Designing an SN~Ia Peculiar Velocity Survey}
+There are different options to consider in the design of a SN~Ia peculiar velocity survey, though at this point it is difficult to assess which is optimal
+or most cost effective.  While there are a variety
+of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
+there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
+Constructing such a model is a parallel
+endeavor that will benefit peculiar-velocity  and other SN~Ia science.  Instead of making specific recommendations, we make several general conclusions
+based upon the above findings.
 
-The nominal observing strategy will obtain photometry every couple of days rotating through the 6 $grizy$ bands.  Although SNe have good
-overall sampling, the per-band light curves are poorly sampled.
-
-\appendix
-\section{Reanalysis in Configuration Space}
-
-This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
-This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
-While there have been a number of articles on the subject,
-our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
-catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
-posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
-compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
-solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
-Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
-quantify the suppression of sample variance achieved when considering matter-densities and velocities 
-within the same volume \citep{2007PhRvL..99h1301G}.
-
-\section{Simulated Data}
-The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
-Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
-cosmoDC2 (v1.0) is consistent with GR. }. 
-The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
-$\Omega_\nu=0$.
-Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
-the radial peculiar velocity is determined, and its
-star formation rate and stellar mass from which the
-supernova rate of each host is determined using 
-\citet{2012ApJ...755...61S}.   This host-based rate
-underestimates the expected discovery based on the volumetric
-rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
-by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
-is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
-from most ground-based observatories.
-
-Each supernova is assigned a magnitude based on the distance modulus of its cosmological
-redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
-intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
-motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
-\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
-and does not affect our conclusions.
-
-We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
-in the precision of
-$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
-would likely saturate available follow-up resources.
-
-\section{Analysis}
-The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
-The peculiar magnitude is given by 
-\begin{equation}
-\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
-\label{data:eqn}
-\end{equation}  We introduce one free parameter 
-$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
-
-The peculiar magnitude correlation
-function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
-matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
-cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
-is
-\begin{equation}
-C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
-\label{cov:eqn}
-\end{equation}
-where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
-The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
-from the linearized GR expectation of the shape of the power spectrum 
-$\xi$.
-Nevertheless, for
-convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
-from General Relativity. 
-The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
-\begin{equation}
-\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
-\end{equation}
-where $\bar{z}$ is the cosmological redshift of the host galaxy.
-We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
-Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
-
-For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
-reduce errors made in the first-order transformation between velocity and
-magnitude.  This cut removes a  small volume relative to the full survey.
-
-The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
-covariance matrix in Eq.~\ref{cov:eqn}.
-The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
-
-Direct tests of non-GR models for which Poisson's equation or the linear approximation
-does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work.
-
-\section{Results of Subsets}
-
-The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
-For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
-determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
-$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
- $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
-
-Different parameters describing the input simulated data define the subsamples, 
-which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
-magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
-of sky is not monitored over the full year.  
-The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
-For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
-is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
-\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
-absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
-part of the simulation.
-\label{zmax:fig}}
-\end{figure}
-
-\begin{table}
-   \centering
-   %\topcaption{Table captions are better up top} % requires the topcapt package
-   \begin{tabular}{|ccc|ccccc|c|} % Column formatting, @{} suppresses leading/trailing space
-   \hline
-$z_{max}$ & $\phi$ & $\sigma_{M}$ & $N_{gal}$ & $\bar{A}$ & $\sigma_A$ & $\bar{A} \sigma_A^{-1} N_{gal}^{-0.5}$ & $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$&  STON $f\sigma_8$ \\
-\hline
-0.07 & 0.65 & 0.08 & 392 &   1.34 &   0.49 &  0.137 &  13.21 &  26.42\\
-0.07 & 1.00 & 0.08 & 563 &   1.16 &   0.36 &  0.135 &  15.55 &  31.11\\
-0.07 & 1.00 & 0.10 & 563 &   1.27 &   0.48 &  0.112 &  12.88 &  25.77\\
-0.07 & 1.00 & 0.12 & 563 &   1.41 &   0.61 &  0.097 &  11.15 &  22.30\\
-0.07 & 1.00 & 0.15 & 563 &   1.64 &   0.82 &  0.084 &   9.74 &  19.48\\
-0.10 & 0.30 & 0.08 & 429 &   1.26 &   0.51 &  0.120 &  12.09 &  24.18\\
-0.10 & 0.50 & 0.08 & 683 &   1.02 &   0.42 &  0.093 &  11.84 &  23.67\\
-0.10 & 0.65 & 0.08 & 873 &   0.89 &   0.35 &  0.087 &  12.57 &  25.14\\
-0.10 & 0.65 & 0.10 & 873 &   1.01 &   0.46 &  0.075 &  10.72 &  21.43\\
-0.10 & 0.65 & 0.12 & 873 &   1.11 &   0.57 &  0.066 &   9.48 &  18.96\\
-0.10 & 1.00 & 0.08 & 1294 &   0.91 &   0.28 &  0.091 &  15.96 &  31.91\\
-0.10 & 1.00 & 0.10 & 1294 &   0.95 &   0.36 &  0.072 &  12.66 &  25.33\\
-0.10 & 1.00 & 0.12 & 1294 &   1.04 &   0.43 &  0.067 &  11.71 &  23.41\\
-0.15 & 0.30 & 0.08 & 1215 &   1.42 &   0.55 &  0.074 &  12.48 &  24.96\\
-0.15 & 0.30 & 0.12 & 1215 &   1.97 &   0.92 &  0.061 &  10.36 &  20.73\\
-0.15 & 0.50 & 0.08 & 1982 &   1.12 &   0.37 &  0.069 &  14.93 &  29.86\\
-0.15 & 0.65 & 0.08 & 2541 &   1.15 &   0.33 &  0.070 &  17.05 &  34.11\\
-0.15 & 0.65 & 0.10 & 2541 &   1.15 &   0.39 &  0.059 &  14.43 &  28.87\\
-0.15 & 0.65 & 0.12 & 2541 &   1.30 &   0.51 &  0.051 &  12.42 &  24.83\\
-0.15 & 1.00 & 0.08 & 3870 &   1.00 &   0.24 &  0.066 &  20.00 &  40.00\\
-0.20 & 0.20 & 0.08 & 2021 &   1.66 &   0.59 &  0.062 &  13.61 &  27.21\\
-0.20 & 0.30 & 0.08 & 3023 &   1.34 &   0.43 &  0.057 &  15.16 &  30.32\\
-0.20 & 0.30 & 0.10 & 3023 &   1.40 &   0.54 &  0.047 &  12.68 &  25.37\\
-0.20 & 0.30 & 0.12 & 3023 &   1.48 &   0.72 &  0.037 &   9.95 &  19.89\\
-0.20 & 0.50 & 0.08 & 5015 &   1.13 &   0.30 &  0.054 &  18.54 &  37.07\\
-0.20 & 0.65 & 0.08 & 6491 &   1.28 &   0.30 &  0.053 &  20.79 &  41.58\\
-0.20 & 0.65 & 0.10 & 6491 &   1.24 &   0.34 &  0.046 &  17.95 &  35.89\\
-0.20 & 0.65 & 0.12 & 6491 &   1.33 &   0.45 &  0.037 &  14.33 &  28.65\\
-0.20 & 1.00 & 0.08 & 9901 &   1.18 &   0.23 &  0.050 &  24.38 &  48.77\\
-0.20 & 1.00 & 0.10 & 9901 &   1.06 &   0.25 &  0.042 &  20.33 &  40.66\\
-0.20 & 1.00 & 0.12 & 9901 &   1.09 &   0.28 &  0.040 &  19.13 &  38.26\\
-    \hline
-   \end{tabular}
-   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
-   the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
-   $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
-   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The ston, scaled to an LSST solid angle, is given as 
-    $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
-   \label{tab:subsets}}
-\end{table}
-
-In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
-In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
-For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
-though the shallowing of the slopes for shallower surveys shows that it does contribute non-negligibly.
-The fact that the ston is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
-indicates that it is non-negligible.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
-\label{scaling:fig}}
-\end{figure}
-
-A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
-peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
-Table~\ref{tab:subsets}
-shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
-The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
-a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
-is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.
-
-$\phi \sim 0.65$ is an effective number density of discovered supernovae with light-curve coverage observed in 10 years
-
-%We consider two different scenarios cutting at different redshift depths $z_{max}=0.1$ and  $z_{max}=0.15$.  All the objects in the first
-%scenario are in the second scenario.  The results are shown in Fig.~\ref{zmax:fig}.  The points of interest are:
-%\begin{itemize}
-%\item There is a bias toward low $f\sigma_8$ for low-redshift ($z_{max}=0.1$).  Adding more supernova with increasing redshift
-% decreases the bias  ($z_{max}=0.15$), and $A$ eventually approaches 1 ($z_{max}=0.2$ not shown).
-% \item Based on the analysis of \citet{2017MNRAS.471.3135H}, I added an intrinsic velocity dispersion $\sigma_v$ (km~s$^{-1}$) due to non-linear effects.  This was not
-% in Dragan's model.  My fit finds that this  dispersion is consistent with zero, not well constrained, and not strongly correlated with
-% $A$.  The lack of covariance is consistent with Dragan's own tests.
-% \item The fit intrinsic magnitude dispersion, on the other hand, is larger than the input $\sigma_M=0.08$ mag.  This  indicates that the catalogs
-% contain 
-% extra  velocity dispersion better attributed to intrinsic magnitude dispersion rather than correlated peculiar velocities.
-%\end{itemize}
-
+\begin{itemize}
+\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
+to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.
+\item Role of LSST:
+LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence of 
+ produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone.  Leaving the
+ distance determination for other telescope, LSST should go for $\Omega$.
+\item Follow-up resources.  21 mag.  2--4m-class telescopes.
+\end{itemize}
+\section{Conclusions}
+%
+%
+%\appendix
+%\section{Reanalysis in Configuration Space}
+%
+%This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
+%This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
+%While there have been a number of articles on the subject,
+%our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
+%catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
+%posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
+%compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
+%solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
+%Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
+%quantify the suppression of sample variance achieved when considering matter-densities and velocities 
+%within the same volume \citep{2007PhRvL..99h1301G}.
+%
+%\section{Simulated Data}
+%The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
+%Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
+%cosmoDC2 (v1.0) is consistent with GR. }. 
+%The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
+%$\Omega_\nu=0$.
+%Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
+%the radial peculiar velocity is determined, and its
+%star formation rate and stellar mass from which the
+%supernova rate of each host is determined using 
+%\citet{2012ApJ...755...61S}.   This host-based rate
+%underestimates the expected discovery based on the volumetric
+%rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
+%by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
+%is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
+%from most ground-based observatories.
+%
+%
+%Each supernova is assigned a magnitude based on the distance modulus of its cosmological
+%redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
+%intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
+%motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
+%\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
+%and does not affect our conclusions.
+%
+%We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
+%in the precision of
+%$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
+%would likely saturate available follow-up resources.
+%
+%
+%\section{Analysis}
+%The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
+%The peculiar magnitude is given by 
+%\begin{equation}
+%\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
+%\label{data:eqn}
+%\end{equation}  We introduce one free parameter 
+%$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
+%
+%
+%The peculiar magnitude correlation
+%function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
+%matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
+%cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
+%is
+%\begin{equation}
+%C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
+%\label{cov:eqn}
+%\end{equation}
+%where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
+%The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
+%from the linearized GR expectation of the shape of the power spectrum 
+%$\xi$.
+%Nevertheless, for
+%convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
+%from General Relativity. 
+%The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
+%\begin{equation}
+%\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
+%\end{equation}
+%where $\bar{z}$ is the cosmological redshift of the host galaxy.
+%We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
+%Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
+%
+%For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
+%reduce errors made in the first-order transformation between velocity and
+%magnitude.  This cut removes a  small volume relative to the full survey.
+%
+%The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
+%covariance matrix in Eq.~\ref{cov:eqn}.
+%The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
+%
+%Direct tests of non-GR models for which Poisson's equation or the linear approximation
+%does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work. 
+%
+%\section{Results of Subsets}
+%
+%The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
+%For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
+%determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
+%$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
+% $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
+%
+%Different parameters describing the input simulated data define the subsamples, 
+%which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
+%magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
+%of sky is not monitored over the full year.  
+%The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
+%For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
+%is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
+%
 %\begin{figure}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.1.pkl.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.1.0.2.pkl.png}
-%\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $M$ is the supernova
+%\centering
+%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
+%\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
 %absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-%to peculiar velocity.   The parameters for which I controlled the input, $M=0$ and $\sigma_M=0.08$~mag, are shown in dotted lines.
-% Left: A survey truncated at $z_{max}=0.1$.  Right: A survey truncated at $z_{max}=0.2$.
+%to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
+%part of the simulation.
 %\label{zmax:fig}}
@@
 %
-
+%\begin{table}
+%   \centering
+%   %\topcaption{Table captions are better up top} % requires the topcapt package
+%   \begin{tabular}{|ccc|cc|} % Column formatting, @{} suppresses leading/trailing space
+%   \hline
+%$z_{max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
+%\hline
+%0.07 & 0.65 & 0.08  & 25.795 &   0.27 \\
+%0.07 & 1.00 & 0.08  & 31.163 &   0.27 \\
+%0.07 & 1.00 & 0.10  & 25.421 &   0.22 \\
+%0.07 & 1.00 & 0.12  & 22.173 &   0.19 \\
+%0.07 & 1.00 & 0.15  & 18.831 &   0.16 \\
+%0.10 & 0.30 & 0.08  & 23.784 &   0.24 \\
+%0.10 & 0.50 & 0.08  & 23.501 &   0.18 \\
+%0.10 & 0.65 & 0.08  & 24.994 &   0.17 \\
+%0.10 & 0.65 & 0.10  & 20.866 &   0.15 \\
+%0.10 & 0.65 & 0.12  & 18.778 &   0.13 \\
+%0.10 & 1.00 & 0.08  & 31.704 &   0.18 \\
+%0.10 & 1.00 & 0.10  & 25.687 &   0.15 \\
+%0.10 & 1.00 & 0.12  & 22.982 &   0.13 \\
+%0.15 & 0.30 & 0.08  & 24.799 &   0.15 \\
+%0.15 & 0.30 & 0.12  & 19.527 &   0.12 \\
+%0.15 & 0.50 & 0.08  & 29.550 &   0.14 \\
+%0.15 & 0.65 & 0.08  & 33.921 &   0.14 \\
+%0.15 & 0.65 & 0.10  & 28.655 &   0.12 \\
+%0.15 & 0.65 & 0.12  & 24.111 &   0.10 \\
+%0.15 & 1.00 & 0.08  & 40.158 &   0.13 \\
+%0.20 & 0.20 & 0.08  & 26.487 &   0.12 \\
+%0.20 & 0.30 & 0.08  & 29.828 &   0.11 \\
+%0.20 & 0.30 & 0.10  & 24.019 &   0.09 \\
+%0.20 & 0.30 & 0.12  & 19.181 &   0.07 \\
+%0.20 & 0.50 & 0.08  & 36.680 &   0.11 \\
+%0.20 & 0.65 & 0.08  & 41.591 &   0.11 \\
+%0.20 & 0.65 & 0.10  & 35.196 &   0.09 \\
+%0.20 & 0.65 & 0.12  & 27.987 &   0.07 \\
+%0.20 & 1.00 & 0.08  & 48.368 &   0.10 \\
+%0.20 & 1.00 & 0.10  & 40.835 &   0.08 \\
+%0.20 & 1.00 & 0.12  & 37.875 &   0.08 \\
+%    \hline
+%   \end{tabular}
+%   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
+%   the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
+%   $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
+%   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The ston, scaled to an LSST solid angle, is given as 
+%    $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
+%   \label{tab:subsets}}
+%\end{table}
 %
-%\begin{itemize}
-%\item Dragan: Is there some way that the formalism in your article breaks down at low-redshift?  I do remove very low-redshift objects at $z<0.01$,
-%which I think is conservative.
-% \item Risa/Joe: Is there any reasons for the Buzzard catalog to not capture correlated peculiar velocities on small scales (box size to large?) nor capture
-% non-linear velocities?  (I don't know how to translate a mass resolution of 2.7e10 $M_{\odot}/h$ into a spatial resolution!)
-%\end{itemize}
+%In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
+%In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
+%For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
+%though the shallowing of the slopes for shallower surveys shows that it does contribute non-negligibly.
+%The fact that the ston is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
+%indicates that it is non-negligible.
+%
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
+%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+%\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
+%\label{scaling:fig}}
+%\end{figure}
+%
+%A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
+%peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
+%Table~\ref{tab:subsets}
+%shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
+%The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
+%a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
+%is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.
+%
+%
+%$\phi \sim 0.65$ is an effective number density of discovered supernovae with light-curve coverage observed in 10 years
+%
 

### 2018-12-14 pv/view.py

@@ -112,8 +112,8 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
     #     zma,fra,snsi,nsn,fsigma8 .mean(), fsigma8 .std(), \
     #     fsigma8.mean()/fsigma8.std()/numpy.sqrt(nsn), \
     #     effston[-1]))
-    print("{:4.2f} & {:4.2f} & {:4.2f}  & {:6.3f} & {:6.2f} \\\\".format( \
-        zma,fra,snsi, effston[-1], \
+    print("{:4.2f} & {:4.2f} & {:4.2f} & {} & {:6.3f} & {:6.2f} \\\\".format( \
+        zma,fra,snsi, nsn*18000./760, effston[-1], \
         effston[-1]/numpy.sqrt(18000./760*nsn)\
         ))
 
@@ -131,11 +131,30 @@ for z_ in zmax_:
     fs = numpy.array(fs)*0.65/0.1**2
     plt.scatter(fs,ef,label=r"$z_{{max}}={}$".format(z_))
 
-plt.xlabel(r"$\frac{\sigma^2}{n} \left[\frac{n(10\ yr)}{(0.1\ mag)^2}\right]$",fontsize=14)
+plt.xlabel(r"$\frac{\sigma_M^2}{n} \left[\frac{n(10\ \mathrm{yr})}{(0.1\ \mathrm{mag})^2}\right]$",fontsize=14)
 plt.ylabel(r"STON",fontsize=14)
 plt.legend()
 plt.tight_layout()
 plt.savefig('notshot.png')
+plt.clf()
+
+for z_ in zmax_:
+    fs = []
+    ef = []
+    for f_ in frac_:
+        for sn_ in snsig_:
+            w = numpy.logical_and.reduce((frac==f_ ,zmax==z_, snsig==sn_))
+            if w.sum() != 0:
+                fs.append(sn_**2/f_)
+                ef.append(effston[w]/numpy.sqrt(18000./760*nsne[w]))
+    fs = numpy.array(fs)*0.65/0.1**2
+    plt.scatter(fs,ef,label=r"$z_{{max}}={}$".format(z_))
+
+plt.xlabel(r"$\frac{\sigma_M^2}{n} \left[\frac{n(10\ \mathrm{yr})}{(0.1\ \mathrm{mag})^2}\right]$",fontsize=14)
+plt.ylabel(r"STON per SN",fontsize=14)
[…]

### 2018-12-14 doc/article.tex
A version, not quite smooth, to give to Eric.

@@
 RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
+SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<-.3$  was well as galaxy surveys will at $z \sim 0.6$ using RSD. 
 {\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
-
 % Requires the booktabs if the memoir class is not being used
@@
 \section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
-\subsection{Parameters that Determine Scientific Performance}
-
-\citet{2017ApJ...847..128H} perform their analysis in Fourier space, 
+\subsection{Survey Parameters that Determine Scientific Performance}
+\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
+They perform their analysis in Fourier space, 
 where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
@@
 \citep{2017MNRAS.464.2517H}.  
-The primary dependence on the growth of structure is through
+The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 
 \begin{equation}
-\frac{\partial C}{\partial \lambda} = \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
+ \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
 \end{equation}
@@
 
-The survey parameters that determine parameter precision are the volume $V$, which
+The survey parameters that  enter the Fisher matrix are the volume $V$, which
 in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
 magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
-
 LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-Equating LSST discovery with having a meaningful peculiar velocity, the number density is$n$ is thus a non-tunable parameter, which only increases with survey duration.
-The remaining free parameters are then $\Omega$, $z_{max}$, and $\sigma_M$.
-
-While LSST data can be used to obtain photometric redshifts from host galaxies and supernova light curves, the spectroscopic
-redshifts needed for precise host-galaxy peculiar velocities require other resources.
+Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a non-tunable parameter, which only increases with survey duration.
+Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.
 
@@
 \item Redshift Depth $z_{max}$:
-For the number densities and intrinsic magnitude dispersions relevant to LSST, the noise is not sample-variance
+For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
 limited.  We have run simulations of LSST SN surveys
-and calculated the signal-to-noise (STON) of $f\sigma_8$ from peculiar velocities. These simulations show, as seen in the left plot
-of  Figure~\ref{notshot:fig}, that an increase in the number the number density of tracers leads to significant improvements in STON.
-It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of  STON on $\sigma_M^2/n$.
+and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
+of  Figure~\ref{notshot:fig}, that an increase in the number the number density of tracers leads to a significant improvement in STON.
+It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$.
 
@@
 
-The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers as shown in the
-right plot of   Figure~\ref{notshot:fig}, shows that LSST is in the regime where adding one tracer at lower redshift better improves the 
-STON relative to the addition of one at higher redshift.
+The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
+$z_{max}$.  This is shown in the
+right plot of   Figure~\ref{notshot:fig}.
 
-Scientific leverage dictates that low-redshift sources are most valuable.
-This aligns with observational considerations, for which closer and hence brighter objects require more modest follow-up resources.
-Therefore, $z_{max}$ depends on the follow-up program as the redshift where complete follow-up saturates resources.
+A low-redshift tracer is more valuable than one at higher redshift.
 
 \item Intrinsic Magnitude Dispersion $\sigma_M$:
-Just as an increase in tracer number density beyond that of a ten-year LSST survey improves the STON in the measurement
-of $f\sigma_8$, as noted in the discussion on $z_{max}$, so too does a decrease in the supernova intrinsic magnitude dispersion.
+A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$,
+as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
+limited so the $\sigma^2_M/n$-term
+in the Fisher matrix is important.
+Survey follow-up stratey,
+which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
+This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
+ in number density, or equivalently in survey duration.
 
-The nominal LSST WFD observing plan is to have
-each piece of sky observed every couple of days rotating through the 6 $grizy$ bands. The data is sufficient to have effectively
-complete $z<0.3$ SN~Ia discoveries.  However,  the per-band light curves are poorly sampled, which limits the accuracy with which distances
-can be derived.
-LSST SN~Ia light curves from the WFD survey, with their relatively poor sampling, plus a spectroscopic redshift
- can give $\sigma_M \sim 0.15$~mag.
-
-Supplemental non-LSST follow-up data in the form or improved temporal light-curve sampling (accurate rise and decline times), 
-expanded (UV, NIR) wavelength coverage, and spectral features
-can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
-For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
- \citep{2015ApJ...815...58F} are projected to give
-$\sigma_M \lesssim 0.08$~mag.
-Uncertainties in $f\sigma_8$ depend on $\sigma^2_M/n$ so the improvement between 0.08 and 0.15~mag dispersions is equivalent to a factor of 3.52
- in number density, or in survey duration. 
+The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively we consider
+a pure SN~Ia sample obtained through
+spectroscopic classification.  The feasibility of using photometric classification is an important subject of research
+that can have implications for survey planning.
 \end{itemize}
@@
 there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
-Constructing such a model is a parallel
-endeavor that will benefit peculiar-velocity  and other SN~Ia science.  Instead of making specific recommendations, we make several general conclusions
-based upon the above findings.
+Here we  make several general conclusions
+based upon the above findings, while noting that 
+constructing such a model is a parallel
+endeavor that will benefit peculiar-velocity  and other SN~Ia science.
 
 \begin{itemize}
+\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
 \item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
-to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.
+to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
+available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
 \item Role of LSST:
 LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence of 
- produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone.  Leaving the
- distance determination for other telescope, LSST should go for $\Omega$.
-\item Follow-up resources.  21 mag.  2--4m-class telescopes.
+ produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
+The wide-field of LSST is
+ well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
+$z<0.3$ SNe~Ia.  LSST is  suited for maximizing $\Omega$ while other (cheaper) resources can improve $\sigma_M$.
+\item Follow-up resources to determine per-SN distances:
+Supplemental non-LSST follow-up data in the form or improved temporal light-curve sampling (accurate rise and decline times), 
+expanded (UV, NIR) wavelength coverage, and spectral features
+can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
+For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
+ \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.
+
+Scientific leverage dictates that low-redshift sources are most valuable.
+This aligns with observational considerations, for which closer and hence brighter objects require more modest follow-up resources.
+Therefore, $z_{max}$ depends on the follow-up program as the redshift where complete follow-up saturates resources.
+
+A follow-up survey with $z_{max}=0.2$ would have $\sim 150,000$ targets over 10 years.  These sources will be observed when
+brighter than $r \sim 20.5$~mag and so will be accessible to 2--4m-class telescopes.  
 \end{itemize}
+
 \section{Conclusions}
+SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
+high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
+of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ as well
+as galaxy surveys, but at lower redshifts that provide better leverage to test gravity models.
+While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
+their probative power.  The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
+access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
+are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should already provide leading measurements of
+$f\sigma_8$ at $z\sim 0.1$.
+
 %

### 2018-12-18 doc/article.tex
draft for larger group

@@
 \author[0000-0001-6315-8743]{A.~G.~Kim}
-\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
     1 Cyclotron Road, Berkeley, CA, 94720}
@@
 \author{C.~Harper}
-\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
     1 Cyclotron Road, Berkeley, CA, 94720}
@@
 \author{C.~Ju}
-\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+\author{E.~V.~Linder}
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
     1 Cyclotron Road, Berkeley, CA, 94720}
@@
 
+In the late 1990's, Type~Ia supernovae (SNe~Ia) were used as distance probes to measure the homogeneous expansion history of the Universe.  The remarkable discovery
+that the expansion is accelerating  has called into question our basic understanding of the gravitational forces within the Universe.  Either it
+is dominated by a ``dark energy'' that is gravitationally repulsive, or General Relativity is inadequate and needs to be replaced by a modified theory of
+gravity.
+It is only appropriate that in the upcoming decade, with their sheer numbers and improved distance precisions, SNe~Ia will provide measurements of the {\it inhomogeneous} motions of structures in the Universe
+that will provide an unmatched test of whether dark energy or modified gravity is responsible for the accelerating expansion of the Universe.
+
 The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
-It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the baryonic structures
+It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the light-emittting structures
 in the Universe and from the peculiar velocities of test masses therein.
 Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
-and repulsion of densitiy inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
-is determined from its observer magnitude and redshift. For a given a background cosmology, the observed magnitude provides
-an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and observer redshifts.
+and repulsion of density inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
+can be determined from its observer-frame magnitude and redshift. For a given background cosmology, the observed magnitude provides
+an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and CMB-frame redshifts.
 
@@
 which gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
-and $\mu$ gives the angular separation.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
-Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and are so bias-free.
+and $\mu$ gives the angular separation of Fourier modes to the line-of-sight.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
+Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and so are bias-free.
 As
@@
 The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
-The current generation of peculiar velocity studies use $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
+The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
 2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
-Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample size by an order of magnitude
+Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample sizes by an order of magnitude
 over half the sky to a depth of $z=0.1$.
@@
 The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
-Nevertheless, their low intrinsic-magnitude uncertainties can 
+Nevertheless, their low intrinsic-magnitude uncertainties 
 can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
@@
 from both their RSD and peculiar velocities.
-The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance uncertainty is assumed for each supernova.
-SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster, and Ly$\alpha$
+The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
+SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster (through the
+kinematic S-Z effect), and Ly$\alpha$
 RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
-SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<-.3$  was well as galaxy surveys will at $z \sim 0.6$ using RSD. 
-{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
+SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
+This can be compared to the current 15\% uncertainty from 6dF \citep{2017MNRAS.471..839A}, 3\% uncertainty projected
+for the combined TAIPAN and WALLABY+WNSHS surveys \citep{2017MNRAS.464.2517H} for low-redshift measurements of $f\sigma_8$.
+At larger $z>0.3$ redshifts DESI projects 10\% accuracies.
+
+%{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
 % Requires the booktabs if the memoir class is not being used
@@
    \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
-   The cumulative column shows the effective uncertainty from combining the redshift and all other shallower redshift bins.}
+   The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.}
    \label{tab:howlett}
@@
 They perform their analysis in Fourier space, 
-where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
+where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C(k)$ is
 \begin{equation}
@@
 in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
-magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
+magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
 LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a non-tunable parameter, which only increases with survey duration.
+Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
 Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.
@@
 \item Solid Angle $\Omega$:
-The variance in $f\sigma_8$ is proportional to solid angle $\Omega$.  The baseline
+The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
 LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
@@
 and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
-of  Figure~\ref{notshot:fig}, that an increase in the number the number density of tracers leads to a significant improvement in STON.
+of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
 It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$.
@@
 \includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
-\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities.  For several limiting redshift depths, the STON is plotted as a
+\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  For several limiting redshift depths, the STON is plotted as a
 function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
@@
 in the Fisher matrix is important.
-Survey follow-up stratey,
+Survey follow-up strategy,
 which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
@@
 a pure SN~Ia sample obtained through
-spectroscopic classification.  The feasibility of using photometric classification is an important subject of research
-that can have implications for survey planning.
+spectroscopic classification, as the requisite follow-up resources for a  complete $z<0.3$ sample is within reason. 
+The feasibility of using photometric classification is an important subject of research
+that can have implications for survey planning, particularly in extending the redshift range.
 \end{itemize}
@@
 \item Role of LSST:
-LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence of 
+LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence  
  produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
@@
 \item Follow-up resources to determine per-SN distances:
-Supplemental non-LSST follow-up data in the form or improved temporal light-curve sampling (accurate rise and decline times), 
+Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
 expanded (UV, NIR) wavelength coverage, and spectral features
@@
 high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
-of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ as well
-as galaxy surveys, but at lower redshifts that provide better leverage to test gravity models.
+of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ significantly better
+than galaxy surveys, at lower redshifts that provide better leverage to test gravity models.
 While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
@@
 access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
-are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should already provide leading measurements of
+are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
 $f\sigma_8$ at $z\sim 0.1$.

### 2018-12-18 pv/view.py

@@ -118,8 +118,8 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
         ))
 
 effston = numpy.array(effston)
-
-for z_ in zmax_:
+mkrs=['o','s','*','D']
+for z_,m in zip(zmax_,mkrs):
     fs = []
     ef = []
     for f_ in frac_:
@@ -129,16 +129,17 @@ for z_ in zmax_:
                 fs.append(sn_**2/f_)
                 ef.append(effston[w])
     fs = numpy.array(fs)*0.65/0.1**2
-    plt.scatter(fs,ef,label=r"$z_{{max}}={}$".format(z_))
+    plt.scatter(fs,ef,label=r"$z_{{max}}={}$".format(z_),marker=m)
 
-plt.xlabel(r"$\frac{\sigma_M^2}{n} \left[\frac{n(10\ \mathrm{yr})}{(0.1\ \mathrm{mag})^2}\right]$",fontsize=14)
+plt.xlabel(r"$\frac{\sigma_M^2}{n} \left[\frac{n(10\ \mathrm{yr})}{(0.1\ \mathrm{mag})^2}\right]$",fontsize=16)
 plt.ylabel(r"STON",fontsize=14)
-plt.legend()
+plt.legend(fontsize=14)
 plt.tight_layout()
 plt.savefig('notshot.png')
 plt.clf()
 
-for z_ in zmax_:
+
+for z_,m in zip(zmax_,mkrs):
     fs = []
     ef = []
     for f_ in frac_:
@@ -148,11 +149,11 @@ for z_ in zmax_:
                 fs.append(sn_**2/f_)
                 ef.append(effston[w]/numpy.sqrt(18000./760*nsne[w]))
     fs = numpy.array(fs)*0.65/0.1**2
-    plt.scatter(fs,ef,label=r"$z_{{max}}={}$".format(z_))
+    plt.scatter(fs,ef,label=r"$z_{{max}}={}$".format(z_), marker=m)
 
-plt.xlabel(r"$\frac{\sigma_M^2}{n} \left[\frac{n(10\ \mathrm{yr})}{(0.1\ \mathrm{mag})^2}\right]$",fontsize=14)
+plt.xlabel(r"$\frac{\sigma_M^2}{n} \left[\frac{n(10\ \mathrm{yr})}{(0.1\ \mathrm{mag})^2}\right]$",fontsize=16)
[…]

### 2019-01-03 doc/article.tex
use Chao's results

@@
 \documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
+\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
 								% TeX will automatically convert eps --> pdf in pdflatex		
@@
 \usepackage{textcase}
+\usepackage{hyperref}
 %SetFonts
@@
 
-\title{Testing Gravity using Type Ia Supernovae Discovered by LSST}
+\thispagestyle{empty}
+\pagenumbering{gobble} 
+\begin{description}
+\item[Title]Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Survey
+\item[Thematic Science Area] Cosmology and Fundamental Physics
+\item[Authors] and a list of authors
+\item[Lead Author Contact Information]Alex Kim; Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720;
+1-510-486-4621;
+\href{mailto:agkim@lbl.gov}{agkim@lbl.gov}
+\end{description}
+
+\clearpage
+\newpage
+\setcounter{page}{0}
+\pagenumbering{arabic}
+\setcounter{page}{1}
+\title{Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys}
 \author[0000-0001-6315-8743]{A.~G.~Kim}
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{P.~Antilogus}
+\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
 \author{S.~BenZvi}
 \affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+\author{S.~Gontcho A Gontcho}
+\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+\author{R.~Graziani}
+\affiliation{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
+Physique de Clermont, F-63000 Clermont-Ferrand, France}    
 \author{C.~Harper}
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{C.~Howlett}
+\affiliation{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
 \author{D.~Huterer}
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{P.-F.~Leget}
+\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
 \author{E.~V.~Linder}
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
-
+\author{P.~McDonald}
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+\author{J.~Nordin}
+\affiliation{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
+\author{S.~Perlmutter}
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+\affiliation{Department of Physics, University of California Berkeley, 366 LeConte Hall MC 7300, Berkeley, CA, 94720-7300}
+\author{N.~Regnault}
+\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+\author{M.~Rigault}
+\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
+1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+Lyon, France}    
 \author{others}
@@
 with different values of $\gamma$
-\citep{2007APh....28..481L}.  The growth of structure, through the measurement of $fD$, provides a test of General Relativity and breaks degeneracies
+\citep{2007APh....28..481L}.  
+The growth of structure, through the measurement of $fD$, provides a test of General Relativity and breaks degeneracies
 between gravitational and dark energy models that explain the accelerating expansion of the Universe.
+Non-GR models may  also predict a change in the redshift- and scale-dependence of the growth, such observations provide additional leverage in probing gravity.
 The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
@@
 Baryonic structures are sensitive to $f\sigma_8$ through redshift space distortions (RSD), 
-which gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
+which on large scales gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
 and $\mu$ gives the angular separation of Fourier modes to the line-of-sight.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
@@
 mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
-leads to the dependence on $H_0 d_L$. 
+leads to the dependence on $H_0 d_L$. (Note that the combination $H_0 d_L$ is independent of $H_0$.)
 Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
@@
 Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample sizes by an order of magnitude
-over half the sky to a depth of $z=0.1$.
+over 75\% of the sky to a maximum depth of $z=0.1$.
 
@@
 have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
-\citep{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.
+\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
 However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
@@
 Recent studies indicate that with the right data, SNe absolute
-magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.
+magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
+established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
+that is observed using current calibrations.
 One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
@@
 hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia for which good light curves can be measured.
-This is a sample size comparable to the number of galaxies projected by WALLABY and TAIPAN.
+This is a sample size with more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
+Measurement of the velocity field using LSST-discovered
+SNe~Ia has been previously quantified by \citet{2011PhRvD..83d3004B,2017JCAP...01..060O}.
 
@@
 The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
-SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster (through the
-kinematic S-Z effect), and Ly$\alpha$
+SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
+kinematic S-Z effect)
 RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
@@
 	\hline
-	Redshift & RSD & RSD + PV & cumulative\\ \hline
+	& RSD & RSD + PV & RSD+PV\\
+	Redshift 	 & ($z$-bin) & ($z$-bin)& (cumulative)\\  \hline
       $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
@@
 $z=0.3$.
-The constraining power on  $f\sigma_8$ 
+At $z<0.2$ the constraining power on  $f\sigma_8$ 
 comes primarily from peculiar velocities
-at $z<0.2$, but are dominated by RSD by $z=0.3$.
+but by $z=0.3$ RSD is more important.
 It turns out that for $z>0.1$, the volume is sufficiently large that the LSST sample is not yet sample variance limited.
@@
 the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-coverage of the southern equatorial pole, could double
-the sky coverage and halve the variance in $f\sigma_8$.
+coverage of the southern equatorial pole, could significantly increase
+the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
 \item Redshift Depth $z_{max}$:
@@
 \includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
-\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  For several limiting redshift depths, the STON is plotted as a
+\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
+realizations made in simulating the surveys.
+For several limiting redshift depths, the STON is plotted as a
 function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
@@
 
-The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively we consider
+The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
 a pure SN~Ia sample obtained through
-spectroscopic classification, as the requisite follow-up resources for a  complete $z<0.3$ sample is within reason. 
+spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
 The feasibility of using photometric classification is an important subject of research
@@
  well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
-$z<0.3$ SNe~Ia.  LSST is  suited for maximizing $\Omega$ while other (cheaper) resources can improve $\sigma_M$.
+$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
+The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
+reach a targeted $\sigma_M$.
 \item Follow-up resources to determine per-SN distances:

### 2019-01-13 doc/article.tex
rewrite second half

@@
 								% TeX will automatically convert eps --> pdf in pdflatex		
-\usepackage{amssymb}
+\usepackage{amssymb,amsmath}
 \usepackage{natbib}
@@
 The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
+Their Figure~4b) shows how well these measurements can distinguish between $\gamma=0.42, 0.55, 0.68$
+\citep[$f(R)$, GR,  and DGP gravities,][]{2007APh....28..481L} gravities respectively.
 SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
@@
 \subsection{Survey Parameters that Determine Scientific Performance}
-\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
+
+The expected peculiar velocity signal  is described by its power spectrum $P_{vv}$.
+The range $k$'s  used in an analysis depends on the depth of the survey $k_{min} \approx (2 z_{max})^{-1}$, and the smallest
+scales where the theoretical predictions are reliable, say $k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$.
+The top panel of Figure~\ref{detC:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
+at $z=0$ assuming General Relativity and linear theory.  The panel shows available $k$ ranges for several surveys with different
+maximum redshift: the $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes contribute the bulk of the peculiar velocity signal.
+The sensitivity to gravity primarily resides in the normalization of the power spectrum, as  $P_{vv} \propto (f\sigma_8)^2$.
+Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$  in Figure~\ref{power:fig}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{src/power.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{ 
+\label{power:fig}}
+\end{figure}
+
+\citet{2017ApJ...847..128H} project how complementary galaxy and peculiar velocity surveys in a common volume can measure $f\sigma_8$. 
 They perform their analysis in Fourier space, 
-where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C(k)$ is
+where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C$ is
 \begin{equation}
 F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-\frac{\partial C}{\partial \lambda_j} \right],
+\frac{\partial C}{\partial \lambda_j} \right]
+\label{fisher:eqn}
 \end{equation}
-where the covariance for the velocity-velocity correlation
+and the covariance is 
 \begin{equation}
-C = P_{vv}(k) + \frac{\sigma^2}{n}
+C(k,\mu)  =
+  \begin{bmatrix}
+   P_{gg}(k,\mu) + \frac{1}{n_g} &
+   P_{vg}(k,\mu)  \\
+   P_{vg}(k,\mu)  &
+  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+   \end{bmatrix}.
 \label{cov:eq}
 \end{equation}
-is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
-\citep{2017MNRAS.464.2517H}.  
-The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 
+The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
+\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the bias parameter.
+The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the probe velocity dispersion
+is $\sigma$. 
+ The Fisher matrix in  Eq.~\ref{fisher:eqn} is evaluated by \citet{2017MNRAS.464.2517H}  to generate the $f \sigma_8$ uncertainties in Table~\ref{tab:howlett}.
+
+The integrand
 \begin{equation}
- \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
+I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
+\frac{\partial C}{\partial  (f\sigma_8)} \right]
 \end{equation}
-for the parameter choice
-$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.
+in Eq.~\ref{fisher:eqn}  represents how much a differential Fourier
+volume element with wavenumber $k$ and line of sight $\mu$ contributes to $F_{f\sigma_{8}\,f\sigma_{8}}$;
+it is therefore  useful in evaluating the design of peculiar velocity surveys. 
+We evaluate these surveys relative to the $I_{max}$ of a sample-variance limited
+peculiar velocity survey with\ $\sigma/n =0$.
+In the following we consider as given a  TAIPAN-like
+galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ that overlaps with the peculiar velocity surveys.
+
+First, we show the benefit of having the  galaxy and velocity surveys in a common volume by considering the degradation that occurs when 
+they are treated as being
+independent.  For this exercise the velocity survey is taken to be sample-variance limited.
+The resulting $I/I_{max}$  is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig},
+and is found to lie between $0.1-0.2$.
+Taking advantage of the fact that the same dark-matter fluctuations produce the galaxies and velocities 
+in the common volume decreases
+the variance in $f\sigma_8$ by factors of 5 to 10.
+
+Realistic surveys have shot noise. We consider two- and ten-year
+supernova densities $n$  based on SN~Ia rates 
+with 0.65 sample efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
+assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
+The $I/I_{max}$ of these survey scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
+Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ meaning that they are sample-noise limited.
+Individual $k>0.01$ modes have  $I/I_{max}<1$ showing that shot noise
+is significant.  Recall that collectively,  larger $k$ modes are the important contributors to the peculiar-velocity signal;
+basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$..
+The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
 
@@
 It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
-
-\begin{figure}
-\centering
-\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
-\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
-\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
-realizations made in simulating the surveys.
-For several limiting redshift depths, the STON is plotted as a
-function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
-Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
-\label{notshot:fig}}
-\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
+%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
+%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
+%realizations made in simulating the surveys.
+%For several limiting redshift depths, the STON is plotted as a
+%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
+%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
+%\label{notshot:fig}}
+%\end{figure}
 
@@
 
+A low-redshift supernova survey is insensitive to two important uncertainties
+associated with the SN~Ia Hubble diagram. The measurement is relatively insensitive to absolute color calibration, due to the limited
+range of {\it observer} wavelengths used to standardize restframe distance moduli.
+Depending on which cameras are used to derive distances, instrumental (rather than absolute)
+calibration may be sufficient.  Progenitor population  is not expected to evolve significantly in the $0.01<z<0.2$ Universe.
+
 \section{Conclusions}
@@
 
+\appendix
+\section{Reanalysis in Configuration Space}
+
+This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
+This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
+While there have been a number of articles on the subject,
+our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
+catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
+posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
+compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
+solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
+Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
+quantify the suppression of sample variance achieved when considering matter-densities and velocities 
+within the same volume \citep{2007PhRvL..99h1301G}.
+
+\section{Simulated Data}
+The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
+Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
+cosmoDC2 (v1.0) is consistent with GR. }. 
+The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
+$\Omega_\nu=0$.
+Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
+the radial peculiar velocity is determined, and its
+star formation rate and stellar mass from which the
+supernova rate of each host is determined using 
+\citet{2012ApJ...755...61S}.   This host-based rate
+underestimates the expected discovery based on the volumetric
+rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
+by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
+is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
+from most ground-based observatories.
+
+Each supernova is assigned a magnitude based on the distance modulus of its cosmological
+redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
+intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
+motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
+\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
+and does not affect our conclusions.
+
+We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
+in the precision of
+$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
+would likely saturate available follow-up resources.
+
+\section{Analysis}
+The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
+The peculiar magnitude is given by 
+\begin{equation}
+\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
+\label{data:eqn}
+\end{equation}  We introduce one free parameter 
+$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
+
+The peculiar magnitude correlation
+function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
+matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
+cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
+is
+\begin{equation}
+C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
+\label{cov:eqn}
+\end{equation}
+where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
+The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
+from the linearized GR expectation of the shape of the power spectrum 
+$\xi$.
+Nevertheless, for
+convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
+from General Relativity. 
+The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
+\begin{equation}
+\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
+\end{equation}
+where $\bar{z}$ is the cosmological redshift of the host galaxy.
+We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
+Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
+
+For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
+reduce errors made in the first-order transformation between velocity and
+magnitude.  This cut removes a  small volume relative to the full survey.
+
+The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
+covariance matrix in Eq.~\ref{cov:eqn}.
+The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
+
+Direct tests of non-GR models for which Poisson's equation or the linear approximation
+does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work.
+
+\section{Results of Subsets}
+
+The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
+For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
+determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
+$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
+ $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
+
+Different parameters describing the input simulated data define the subsamples, 
+which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
+magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
+of sky is not monitored over the full year.  
+The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
+For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
+is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
+\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
+absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
+to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
+part of the simulation.
+\label{zmax:fig}}
+\end{figure}
+
+\begin{table}
+   \centering
+   %\topcaption{Table captions are better up top} % requires the topcapt package
+   \begin{tabular}{|ccc|cc|} % Column formatting, @{} suppresses leading/trailing space
+   \hline
+$z_{max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
+\hline
+0.07 & 0.65 & 0.08  & 25.795 &   0.27 \\
+0.07 & 1.00 & 0.08  & 31.163 &   0.27 \\
+0.07 & 1.00 & 0.10  & 25.421 &   0.22 \\
+0.07 & 1.00 & 0.12  & 22.173 &   0.19 \\
+0.07 & 1.00 & 0.15  & 18.831 &   0.16 \\
+0.10 & 0.30 & 0.08  & 23.784 &   0.24 \\
+0.10 & 0.50 & 0.08  & 23.501 &   0.18 \\
+0.10 & 0.65 & 0.08  & 24.994 &   0.17 \\
+0.10 & 0.65 & 0.10  & 20.866 &   0.15 \\
+0.10 & 0.65 & 0.12  & 18.778 &   0.13 \\
+0.10 & 1.00 & 0.08  & 31.704 &   0.18 \\
+0.10 & 1.00 & 0.10  & 25.687 &   0.15 \\
+0.10 & 1.00 & 0.12  & 22.982 &   0.13 \\
+0.15 & 0.30 & 0.08  & 24.799 &   0.15 \\
+0.15 & 0.30 & 0.12  & 19.527 &   0.12 \\
+0.15 & 0.50 & 0.08  & 29.550 &   0.14 \\
+0.15 & 0.65 & 0.08  & 33.921 &   0.14 \\
+0.15 & 0.65 & 0.10  & 28.655 &   0.12 \\
+0.15 & 0.65 & 0.12  & 24.111 &   0.10 \\
+0.15 & 1.00 & 0.08  & 40.158 &   0.13 \\
+0.20 & 0.20 & 0.08  & 26.487 &   0.12 \\
+0.20 & 0.30 & 0.08  & 29.828 &   0.11 \\
+0.20 & 0.30 & 0.10  & 24.019 &   0.09 \\
+0.20 & 0.30 & 0.12  & 19.181 &   0.07 \\
+0.20 & 0.50 & 0.08  & 36.680 &   0.11 \\
+0.20 & 0.65 & 0.08  & 41.591 &   0.11 \\
+0.20 & 0.65 & 0.10  & 35.196 &   0.09 \\
+0.20 & 0.65 & 0.12  & 27.987 &   0.07 \\
+0.20 & 1.00 & 0.08  & 48.368 &   0.10 \\
+0.20 & 1.00 & 0.10  & 40.835 &   0.08 \\
+0.20 & 1.00 & 0.12  & 37.875 &   0.08 \\
+    \hline
+   \end{tabular}
+   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
+   the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
+   $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
+   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The ston, scaled to an LSST solid angle, is given as 
+    $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
+   \label{tab:subsets}}
+\end{table}
+
+In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
+In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
+For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
+though the shallowing of the slopes for shallower surveys shows that it does contribute non-negligibly.
+The fact that the ston is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
+indicates that it is non-negligible.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
+\label{scaling:fig}}
+\end{figure}
+
+A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
+peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
+Table~\ref{tab:subsets}
+shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
+The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
+a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
+is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.
+
+$\phi \sim 0.65$ is an effective number density of discovered supernovae with light-curve coverage observed in 10 years
+
 %
 %
-%\appendix
-%\section{Reanalysis in Configuration Space}
-%
-%This Appendix describes a new analysis used to project precisions on $f\sigma_8$ using peculiar velocities derived from SNe~Ia.
-%This analysis may or may not be useful for the Decadal Survey White Paper, but will be helpful for LSST planning.
-%While there have been a number of articles on the subject,
-%our analysis brings a higher level of fidelity than sought previously.  We simulate SNe~Ia hosted by galaxies in a mock galaxy
-%catalog. The numbers of SNe are sufficiently small to allow fast evaluations of the likelihood, which enable the determination of parameter
-%posteriors using MCMC on reasonable computing timescales.   We can use our machinery to 
-%compare different survey parameters, such as redshift depth, total numbers of supernovae, non-uniform spatial density,
-%solid angle/survey geometry, and SN~Ia intrinsic magnitude dispersion.
-%Adding the cross-correlation between galaxy-count and peculiar-velocity surveys is active work, so we can
-%quantify the suppression of sample variance achieved when considering matter-densities and velocities 
-%within the same volume \citep{2007PhRvL..99h1301G}.
-%
-%\section{Simulated Data}
-%The cosmoDC2 (v1.0) covering 706 sq.~deg is used\footnote{We previously used
-%Buzzard (v1.6) galaxy catalog is used, but it was found to give peculiar velocities inconsistent with general relativity.  As we will see,
-%cosmoDC2 (v1.0) is consistent with GR. }. 
-%The catalog is based on a Flat $\Lambda$CDM model with $H_0=71$~km~s$^{-1}$,  $\Omega_M=0.265$, $\Omega_B=0.0448$, and
-%$\Omega_\nu=0$.
-%Each galaxy has its cosmological redshift, the $x$-, $y$-, $z$-components of its peculiar velocity from which
-%the radial peculiar velocity is determined, and its
-%star formation rate and stellar mass from which the
-%supernova rate of each host is determined using 
-%\citet{2012ApJ...755...61S}.   This host-based rate
-%underestimates the expected discovery based on the volumetric
-%rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
-%by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
-%is greater than the total number of possibly useful supernova discoveries, since most fields cannot be continuously monitored throughout the year
-%from most ground-based observatories.
-%
-%
-%Each supernova is assigned a magnitude based on the distance modulus of its cosmological
-%redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
-%intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric
-%motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
-%\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
-%and does not affect our conclusions.
-%
-%We consider a maximum redshift of $z=0.2$ because over ten years the noise is not sample-variance dominated, meaning that the differential improvement
-%in the precision of
-%$f\sigma_8$ due to a lower-redshift supernova is greater than that due to a high-redshift supernova.  Following all LSST SNe~Ia below this redshift 
-%would likely saturate available follow-up resources.
-%
-%
-%\section{Analysis}
-%The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
-%The peculiar magnitude is given by 
+%\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
+%They perform their analysis in Fourier space, 
+%where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C(k)$ is
 %\begin{equation}
-%\delta_m=(m - M) - \mu(z;H_0, \Omega_M, \Omega_\Lambda).
-%\label{data:eqn}
-%\end{equation}  We introduce one free parameter 
-%$\mathcal{M} = M - 5\log{H_0} + \text{const}$ and do not consider the other cosmological parameters, whose effects are relatively weak at low-redshift.
-%
-%
-%The peculiar magnitude correlation
-%function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
-%matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
-%cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
-%is
+%F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+%\frac{\partial C}{\partial \lambda_j} \right],
+%\end{equation}
+%where the covariance for the velocity-velocity correlation
 %\begin{equation}
-%C_{ij} = A\xi_{\delta m \delta m}(\mathbf{r_i},\mathbf{r_j}) + \frac{\sigma_M^2}{N_i} \delta_{ij} + \sigma^2_{NL}(z_i;\sigma_{v})\delta_{ij},
-%\label{cov:eqn}
+%C = P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+%\label{cov:eq}
 %\end{equation}
-%where $N_i$ is the number of SNe~Ia in galaxy $i$ and $\sigma_M$ is the intrinsic SN magnitude dispersion.
-%The model is used to test General Relativity through deviations from $A=1$ but is not sensitive to deviations
-%from the linearized GR expectation of the shape of the power spectrum 
-%$\xi$.
-%Nevertheless, for
-%convenience we say that our model gives $f\sigma_8 = A (f\sigma_8)^{GR}$, where $(f\sigma_8)^{GR}$ is the expectation
-%from General Relativity. 
-%The final term includes extra magnitude dispersion produced by non-linear effects on velocity, parameterized with $\sigma_{v}$, such that
+%is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
+%\citep{2017MNRAS.464.2517H}.  
+%The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 
 %\begin{equation}
-%\sigma_{NL} = \frac{5}{\ln{10}} \frac{1+\bar{z}}{\bar{z}} \sigma_v,
+% \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
 %\end{equation}
-%where $\bar{z}$ is the cosmological redshift of the host galaxy.
-%We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
-%Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
-%
-%For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
-%reduce errors made in the first-order transformation between velocity and
-%magnitude.  This cut removes a  small volume relative to the full survey.
-%
-%The posterior of the parameters is sampled using emcee.  The likelihood is the Normal distribution with the data in Eq.~\ref{data:eqn} and
-%covariance matrix in Eq.~\ref{cov:eqn}.
-%The priors are positive and flat for $A$, flat for $M$, positive Cauchy(0.08, 0.5) for $\sigma_M$, and  positive Cauchy(0, 600 km~s$^{-1}$) for $\sigma_{v}$.
-%
-%Direct tests of non-GR models for which Poisson's equation or the linear approximation
-%does not apply requires a model-dependent $\xi$.  We defer the analysis of such higher-fidelity models to later work. 
-%
-%\section{Results of Subsets}
-%
-%The  ston in $f\sigma_8$ is calculated for different subsamples of the simulated data.
-%For each subsample  the mean  and standard deviation, $(\bar{A}, \sigma_A$) of the amplitude parameter is
-%determined.  The ston in $f\sigma_8$ for the subsample is thus  $\bar{A} \sigma^{-1}_A$, and that for a 18,000 sq.~deg. survey  is approximated by
-%$\sqrt{18000/760}\bar{A}\sigma^{-1}_A$.  The effective constraining power of a single supernova in the subset is chracterized by 
-% $\bar{A} \sigma^{-1}_A N_{gal}^{-0.5}$.
-%
-%Different parameters describing the input simulated data define the subsamples, 
-%which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
-%magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
-%of sky is not monitored over the full year.  
-%The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
-%For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
-%is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\textwidth]{../outcosmo/pvlist.0.08.1234.0.65.0.2.pkl.png}
-%\caption{Confidence regions for the model parameters.  $A=(f\sigma_8)/(f\sigma_8)_{GR}$, $\mathcal{M}$ is the supernova
-%absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-%to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
-%part of the simulation.
-%\label{zmax:fig}}
-%\end{figure}
-%
-%\begin{table}
-%   \centering
-%   %\topcaption{Table captions are better up top} % requires the topcapt package
-%   \begin{tabular}{|ccc|cc|} % Column formatting, @{} suppresses leading/trailing space
-%   \hline
-%$z_{max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
-%\hline
-%0.07 & 0.65 & 0.08  & 25.795 &   0.27 \\
-%0.07 & 1.00 & 0.08  & 31.163 &   0.27 \\
-%0.07 & 1.00 & 0.10  & 25.421 &   0.22 \\
-%0.07 & 1.00 & 0.12  & 22.173 &   0.19 \\
-%0.07 & 1.00 & 0.15  & 18.831 &   0.16 \\
-%0.10 & 0.30 & 0.08  & 23.784 &   0.24 \\
-%0.10 & 0.50 & 0.08  & 23.501 &   0.18 \\
-%0.10 & 0.65 & 0.08  & 24.994 &   0.17 \\
-%0.10 & 0.65 & 0.10  & 20.866 &   0.15 \\
-%0.10 & 0.65 & 0.12  & 18.778 &   0.13 \\
-%0.10 & 1.00 & 0.08  & 31.704 &   0.18 \\
-%0.10 & 1.00 & 0.10  & 25.687 &   0.15 \\
-%0.10 & 1.00 & 0.12  & 22.982 &   0.13 \\
-%0.15 & 0.30 & 0.08  & 24.799 &   0.15 \\
-%0.15 & 0.30 & 0.12  & 19.527 &   0.12 \\
-%0.15 & 0.50 & 0.08  & 29.550 &   0.14 \\
-%0.15 & 0.65 & 0.08  & 33.921 &   0.14 \\
-%0.15 & 0.65 & 0.10  & 28.655 &   0.12 \\
-%0.15 & 0.65 & 0.12  & 24.111 &   0.10 \\
-%0.15 & 1.00 & 0.08  & 40.158 &   0.13 \\
-%0.20 & 0.20 & 0.08  & 26.487 &   0.12 \\
-%0.20 & 0.30 & 0.08  & 29.828 &   0.11 \\
-%0.20 & 0.30 & 0.10  & 24.019 &   0.09 \\
-%0.20 & 0.30 & 0.12  & 19.181 &   0.07 \\
-%0.20 & 0.50 & 0.08  & 36.680 &   0.11 \\
-%0.20 & 0.65 & 0.08  & 41.591 &   0.11 \\
-%0.20 & 0.65 & 0.10  & 35.196 &   0.09 \\
-%0.20 & 0.65 & 0.12  & 27.987 &   0.07 \\
-%0.20 & 1.00 & 0.08  & 48.368 &   0.10 \\
-%0.20 & 1.00 & 0.10  & 40.835 &   0.08 \\
-%0.20 & 1.00 & 0.12  & 37.875 &   0.08 \\
-%    \hline
-%   \end{tabular}
-%   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
-%   the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
-%   $N_{gal}$, the signal $\bar{A}$ and its uncertainty $\sigma_A$ are for the cosmoDC2 (v1.0) 760 sq.~deg. volume only.
-%   The uncertainty ``per'' supernova is given by  $\sigma_A N_{gal}^{-0.5}$.  The ston, scaled to an LSST solid angle, is given as 
-%    $\bar{A} \sigma_A^{-1} (18000/760)^{0.5}$.
-%   \label{tab:subsets}}
-%\end{table}
-%
-%In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
-%In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
-%For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
-%though the shallowing of the slopes for shallower surveys shows that it does contribute non-negligibly.
-%The fact that the ston is not constant indicates that not of the subsamples is sample-variance limited, though its non-linearity
-%indicates that it is non-negligible.
-%
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\textwidth]{../outcosmo/fracsnsig2_.png}
-%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-%\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
-%\label{scaling:fig}}
-%\end{figure}
-%
-%A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
-%peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
-%Table~\ref{tab:subsets}
-%shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
-%The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
-%a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
-%is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.
-%
-%
-%$\phi \sim 0.65$ is an effective number density of discovered supernovae with light-curve coverage observed in 10 years
+%for the parameter choice
+%$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.
 %

### 2019-01-13 doc/src/cov.py

@@ -1,0 +1,268 @@
+#!/usr/bin/env python
+
+import numpy
+import matplotlib.pyplot as plt
+import matplotlib
+
+
+matter = numpy.loadtxt("../../pv/dragan/matterpower.dat")
+
+# vv
+f=0.28**0.55
+
+def Pvv(mu):
+    return (f*mu*100)**2*matter[:,1]/matter[:,0]**2
+
+def Pvv_l(mu):
+    return 2*f*(mu*100)**2*matter[:,1]/matter[:,0]**2
+
+
+# gg
+b= 1.2
+def Pgg(mu):
+    return (b+f*mu**2)**2*matter[:,1]
+
+def Pgg_l(mu):
+    return 2*mu**2*(b+f*mu**2)*matter[:,1]
+
+# gv
+def Pgv(mu):
+    return (b+f*mu**2)*(f*mu*100)*matter[:,1]/matter[:,0]
+
+def Pgv_l(mu):
+    return (mu**2*(f*mu*100)+(b+f*mu**2)*(mu*100))*matter[:,1]/matter[:,0]
+
+# SN rates
+restrate =  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
+twoyeardensity = 0.65*2*restrate
+tenyeardensity = 0.65*10*restrate
+
+sigm = 0.08
+
+
+# galaxy density (TAIPAN Howlett et al.)
+galdensity = 1e-3
+
+
+mus=[1,0.5]
+lss=[':','-']
+zs = numpy.array([0.05,0.1,.2])
+colors=['blue','green','red']
+sigv_factor = numpy.log(10)/5*zs/(1+zs)
+sigvs =sigm * sigv_factor *3e5
+
+fig, axs = plt.subplots(3,1,sharex=True,figsize=(5,7))
+
+for mu,ax in zip(mus,axs[1:]):
+    shotless=[]
+    nocov=[]
+    pvv = Pvv(mu)
+    pgg = Pgg(mu)
+    pgv = Pgv(mu)
+
+    pvv_l = Pvv_l(mu)
+    pgg_l = Pgg_l(mu)
+    pgv_l = Pgv_l(mu)    
+    for p00, p01, p11,p00_l, p01_l, p11_l in zip(pgg,pgv,pvv,pgg_l,pgv_l,pvv_l):
+
+        C_l = numpy.array([[p00_l,p01_l],[p01_l,p11_l]])
[…]

### 2019-01-15 doc/article.tex
finish draft of article

@@
 \usepackage{amssymb,amsmath}
-\usepackage{natbib}
+\usepackage{natbib,verbatim}
 \usepackage{grffile}
@@
 but by $z=0.3$ RSD is more important.
-It turns out that for $z>0.1$, the volume is sufficiently large that the LSST sample is not yet sample variance limited.
 
 \section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
-\subsection{Survey Parameters that Determine Scientific Performance}
+\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
 
-The expected peculiar velocity signal  is described by its power spectrum $P_{vv}$.
-The range $k$'s  used in an analysis depends on the depth of the survey $k_{min} \approx (2 z_{max})^{-1}$, and the smallest
-scales where the theoretical predictions are reliable, say $k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$.
-The top panel of Figure~\ref{detC:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
-at $z=0$ assuming General Relativity and linear theory.  The panel shows available $k$ ranges for several surveys with different
-maximum redshift: the $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes contribute the bulk of the peculiar velocity signal.
+The  peculiar velocity signal  can be described by its power spectrum $P_{vv}$.
 The sensitivity to gravity primarily resides in the normalization of the power spectrum, as  $P_{vv} \propto (f\sigma_8)^2$.
-Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$  in Figure~\ref{power:fig}.
+Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
+though this detail is not considered here.
+The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
+at $z=0$ expected from General Relativity and linear theory. Available $k$ ranges for surveys with different
+maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
+$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
+scale where  theory predictions are reliable.
+Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{src/power.png}
+\includegraphics[width=0.8\textwidth]{src/power.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{ 
+\caption{Top panel: Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
+Lower two panels: $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
+the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
+independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
+of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
+redshift depths $z_{max}$.
 \label{power:fig}}
@@
 
-\citet{2017ApJ...847..128H} project how complementary galaxy and peculiar velocity surveys in a common volume can measure $f\sigma_8$. 
-They perform their analysis in Fourier space, 
-where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C$ is
+\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
+ obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
+Their analysis is performed in Fourier space, 
+where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
 \begin{equation}
@@
 \end{equation}
-and the covariance is 
+ and covariance
 \begin{equation}
@@
 The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
-\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the bias parameter.
-The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the probe velocity dispersion
+\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
+The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
 is $\sigma$. 
- The Fisher matrix in  Eq.~\ref{fisher:eqn} is evaluated by \citet{2017MNRAS.464.2517H}  to generate the $f \sigma_8$ uncertainties in Table~\ref{tab:howlett}.
-
-The integrand
+The integrand in the Fisher-element calculation
 \begin{equation}
@@
 \end{equation}
-in Eq.~\ref{fisher:eqn}  represents how much a differential Fourier
-volume element with wavenumber $k$ and line of sight $\mu$ contributes to $F_{f\sigma_{8}\,f\sigma_{8}}$;
-it is therefore  useful in evaluating the design of peculiar velocity surveys. 
-We evaluate these surveys relative to the $I_{max}$ of a sample-variance limited
-peculiar velocity survey with\ $\sigma/n =0$.
-In the following we consider as given a  TAIPAN-like
+represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
+differential Fourier
+volume element.
+
+\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
+The performance of a peculiar velocity survey depends on its overlap with galaxy surveys, solid-angle coverage, redshift depth, and intrinsic magnitude dispersion.
+In the following, we consider as given a  TAIPAN-like
 galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ that overlaps with the peculiar velocity surveys.
+While the expected variance in $f\sigma_8$ is $\left(F^{-1}\right)_{{f\sigma_{8}\,f\sigma_{8}}}$, we use $I$ as a proxy to help understand
+how survey densities and shot noise affect science performance.
+Surveys are compared relative to the $I_{max}$ of an idealized sample-variance limited peculiar-velocity survey with\ $\sigma/n =0$.
 
-First, we show the benefit of having the  galaxy and velocity surveys in a common volume by considering the degradation that occurs when 
-they are treated as being
-independent.  For this exercise the velocity survey is taken to be sample-variance limited.
-The resulting $I/I_{max}$  is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig},
-and is found to lie between $0.1-0.2$.
-Taking advantage of the fact that the same dark-matter fluctuations produce the galaxies and velocities 
-in the common volume decreases
-the variance in $f\sigma_8$ by factors of 5 to 10.
+First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
+Consider two- and ten-year
+SN~Ia densities $n$  based on SN~Ia rates 
+with 0.65 efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
+assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
+The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
+Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
+Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
+is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
+Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
+%The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
 
-Realistic surveys have shot noise. We consider two- and ten-year
-supernova densities $n$  based on SN~Ia rates 
-with 0.65 sample efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
-assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
-The $I/I_{max}$ of these survey scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
-Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ meaning that they are sample-noise limited.
-Individual $k>0.01$ modes have  $I/I_{max}<1$ showing that shot noise
-is significant.  Recall that collectively,  larger $k$ modes are the important contributors to the peculiar-velocity signal;
-basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$..
-The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
+{\bf Solid Angle $\Omega$:}
+$F_{f\sigma_{8}\,f\sigma_{8}}$ is  proportional to volume and hence the survey solid angle $\Omega$.  The baseline
+LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
+the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+coverage of the southern equatorial pole could significantly increase
+the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
 
-The survey parameters that  enter the Fisher matrix are the volume $V$, which
-in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
-magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
-LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
-Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.
+%s
+% and so helps illustrate 
+%how how the density, $n$, and intrinsic SN magnitude dispersion of peculiar velocity probes, as
+%$\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, affect survey precision.
 
-\begin{itemize}
-\item Solid Angle $\Omega$:
-The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
-LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
-the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-coverage of the southern equatorial pole, could significantly increase
-the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
-\item Redshift Depth $z_{max}$:
-For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
-limited.  We have run simulations of LSST SN surveys
-and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
-of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
-It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
-%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
-%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
-%realizations made in simulating the surveys.
-%For several limiting redshift depths, the STON is plotted as a
-%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
-%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
-%\label{notshot:fig}}
-%\end{figure}
+{\bf Common-Volume Galaxy  Survey:}
+Galaxy and velocity surveys in different volumes do not have the off-diagonal cross-correlation terms in Eq.~\ref{cov:eq}.
+The degradation in probative power   $I/I_{max}$, when sample-variance limited,
+is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig}.
+The curves lie between $0.1-0.2$, showing
+that accounting for the common dark-matter fluctuations that give rise to galaxy densities and peculiar velocities
+reduces the variance in $f\sigma_8$ by factors of 5 to 10.
 
-The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
-$z_{max}$.  This is shown in the
-right plot of   Figure~\ref{notshot:fig}.
+{\bf Redshift Depth $z_{max}$:}
+Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is the expense of additional depth worth neglecting lower redshift
+SNe? The answer is found in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
+strongest probative power and are sensitive more to peculiar velocities rather than galaxy overdensities.  Despite
+the differential volume increase with redshift depth (Eq,~\ref{fisher:eqn}), the degradation in intrinsic peculiar velocity $\propto z/(1+z)$
+and being shot-noise-limited  makes a low-redshift tracer more valuable than one at higher redshift.
 
-A low-redshift tracer is more valuable than one at higher redshift.
-
-\item Intrinsic Magnitude Dispersion $\sigma_M$:
-A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$,
-as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
-limited so the $\sigma^2_M/n$-term
-in the Fisher matrix is important.
+{\bf Intrinsic Magnitude Dispersion $\sigma_M$}:
+Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
+Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
+%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
+%limited so the $\sigma^2_M/n$-term
+%in the Fisher matrix is important.
 Survey follow-up strategy,
@@
 
-The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
-a pure SN~Ia sample obtained through
-spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
-The feasibility of using photometric classification is an important subject of research
-that can have implications for survey planning, particularly in extending the redshift range.
-\end{itemize}
+%\subsection{Survey Parameters that Determine Scientific Performance}
+
+%The survey parameters that  enter the Fisher matrix are the volume $V$, which
+%in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
+%magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
+
+%
+%
+%\begin{itemize}
+%\item Solid Angle $\Omega$:
+%The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
+%LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
+%the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+%coverage of the southern equatorial pole, could significantly increase
+%the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
+%\item Redshift Depth $z_{max}$:
+%Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is it worth doing so at the expense of neglecting lower redshift
+%SNe? This is answered in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
+%stronger probative power and benefit most from peculiar velocities.  Although
+%a differential increase in redshift depth increases the volume in Eq,~\ref{fisher:eqn}, the degradation in  intrinsic peculiar velocity $\propto z/(1+z)$
+%and the fact that shot noise is still important leads to the conclusion that a low-redshift tracer is more valuable than one at higher redshift.
+%
+%%For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
+%%limited.  We have run simulations of LSST SN surveys
+%%and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
+%%of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
+%%It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
+%%
+%%\begin{figure}
+%%\centering
+%%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
+%%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
+%%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
+%%realizations made in simulating the surveys.
+%%For several limiting redshift depths, the STON is plotted as a
+%%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
+%%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
+%%\label{notshot:fig}}
+%%\end{figure}
+%%
+%%The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
+%%$z_{max}$.  This is shown in the
+%%right plot of   Figure~\ref{notshot:fig}.
+%%
+%
+%
+%\item Intrinsic Magnitude Dispersion $\sigma_M$:
+%Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
+%A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
+%%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
+%%limited so the $\sigma^2_M/n$-term
+%%in the Fisher matrix is important.
+%Survey follow-up strategy,
+%which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
+%This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
+% in number density, or equivalently in survey duration. 
+%\end{itemize}
 
@@
 
+%Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
+%Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.
+
 \begin{itemize}
 \item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
+The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
+a pure SN~Ia sample obtained through
+spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
+The feasibility of using photometric classification is an important subject of research
+that can have implications for survey planning, particularly in extending the redshift range.
 \item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
@@
 \item Role of LSST:
-LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence  
+LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.
+LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
+The nominal cadence  
  produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
@@
 $f\sigma_8$ at $z\sim 0.1$.
+
+\begin{comment}
 
@@
 
+\end{comment}
+
 \bibliographystyle{aasjournal}

### 2019-01-15 doc/src/cov.py

@@ -51,7 +51,7 @@ colors=['blue','green','red']
 sigv_factor = numpy.log(10)/5*zs/(1+zs)
 sigvs =sigm * sigv_factor *3e5
 
-fig, axs = plt.subplots(3,1,sharex=True,figsize=(5,7))
+fig, axs = plt.subplots(3,1,sharex=True,figsize=(8,5))
 
 for mu,ax in zip(mus,axs[1:]):
     shotless=[]
@@ -110,12 +110,13 @@ for mu,ax in zip(mus,axs[1:]):
 # plt.plot(matter[:,0],shotless,label='Sample Variance Limit')
 # plt.plot(matter[:,0],twoyear,label='Two Year')
 # plt.plot(matter[:,0],tenyear,label='Ten Year')
-    ax.set_xlim((5e-4,.5))
+    ax.set_xlim((6e-4,.2))
     ax.set_ylim((.5e-3,1))
     ax.set_yscale("log", nonposx='clip')
     ax.set_xscale("log", nonposx='clip')
-    ax.set_title(r'$\mu={}$'.format(mu))
-    ax.set_ylabel(r'Tr$\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$')
+    ax.text(0.2,0.4,r'$\mu={}$'.format(mu),transform=ax.transAxes,fontsize=14)
+    # ax.set_ylabel(r'Tr$\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$')
+    ax.set_ylabel(r'$I/I_{max}$')
 # plt.legend()
 axs[1].legend(prop={'size': 6})
 axs[2].set_xlabel(r'$k$ [$h$ Mpc$^{-1}$]')
@@ -130,13 +131,14 @@ axs[0].plot([1./(zs[2]*3e3*2),0.1],[numpy.exp(dum*10),numpy.exp(dum*10)],color=c
 axs[0].text(0.1, numpy.exp(dum*10.6),r'$z_{{max}}={}$'.format(zs[0]),color=colors[0])
 axs[0].text(0.1, numpy.exp(dum*10.3),r'$z_{{max}}={}$'.format(zs[1]),color=colors[1])
 axs[0].text(0.1, numpy.exp(dum*10),r'$z_{{max}}={}$'.format(zs[2]),color=colors[2])
-axs[0].set_xlim((5e-4,.5))
[…]

### 2019-01-16 doc/src/cov.py
add new plot

@@ -16,7 +16,6 @@ def Pvv(mu):
 def Pvv_l(mu):
     return 2*f*(mu*100)**2*matter[:,1]/matter[:,0]**2
 
-
 # gg
 b= 1.2
 def Pgg(mu):
@@ -25,12 +24,19 @@ def Pgg(mu):
 def Pgg_l(mu):
     return 2*mu**2*(b+f*mu**2)*matter[:,1]
 
+def Pgg_b(mu):
+    return 2*(b+f*mu**2)*matter[:,1]
+
 # gv
 def Pgv(mu):
     return (b+f*mu**2)*(f*mu*100)*matter[:,1]/matter[:,0]
 
 def Pgv_l(mu):
     return (mu**2*(f*mu*100)+(b+f*mu**2)*(mu*100))*matter[:,1]/matter[:,0]
+
+def Pgv_b(mu):
+    return (f*mu*100)*matter[:,1]/matter[:,0]
+
 
 # SN rates
 restrate =  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
@@ -39,18 +45,244 @@ tenyeardensity = 0.65*10*restrate
 
 sigm = 0.08
 
-
 # galaxy density (TAIPAN Howlett et al.)
-galdensity = 1e-3
-
+ng = 1e-3
 
 mus=[1,0.5]
-lss=[':','-']
-zs = numpy.array([0.05,0.1,.2])
-colors=['blue','green','red']
-sigv_factor = numpy.log(10)/5*zs/(1+zs)
-sigvs =sigm * sigv_factor *3e5
+#lss=[':','-']
 
+# sigv_factor = numpy.log(10)/5*zs/(1+zs)
+# sigvs =sigm * sigv_factor *3e5
+
+# the four traces in an array.  trace is a function of (z)
+def trace(z, mu, sigm, n,ng,independent=False):
+
+    pvv = Pvv(mu)
+    pgg = Pgg(mu)
+    pvg = Pgv(mu)
+
+    pvv_l = Pvv_l(mu)
+    pgg_l = Pgg_l(mu)
+    pvg_l = Pgv_l(mu)  
+
+    pgg_b = Pgg_b(mu)
+    pvg_b = Pgv_b(mu)
+
+    sigv_factor = numpy.log(10)/5*z/(1+z)
+    sigv =sigm * sigv_factor *3e5
+    shot = sigv**2/n
+    trace_ll = []
+    trace_lb = []
+    trace_bl = []
+    trace_bb = []
[…]

### 2019-01-17 doc/src/partials.py
new plots

@@ -1,0 +1,252 @@
+#!/usr/bin/env python
+
+import numpy
+import matplotlib.pyplot as plt
+import matplotlib
+
+
+matter = numpy.loadtxt("../../pv/dragan/matterpower.dat")
+
+# vv
+f=0.28**0.55
+
+# SN rates
+restrate = 0.65*  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
+
+
+sigm = 0.08
+
+# galaxy density (TAIPAN Howlett et al.)
+ng = 1e-3
+
+
+def Pvv(mu):
+    return (f*mu*100)**2*matter[:,1]/matter[:,0]**2
+
+def Pvv_l(mu):
+    return 2*f*(mu*100)**2*matter[:,1]/matter[:,0]**2
+
+# gg
+b= 1.2
+def Pgg(mu):
+    return (b+f*mu**2)**2*matter[:,1]
+
+def Pgg_l(mu):
+    return 2*mu**2*(b+f*mu**2)*matter[:,1]
+
+def Pgg_b(mu):
+    return 2*(b+f*mu**2)*matter[:,1]
+
+# gv
+def Pgv(mu):
+    return (b+f*mu**2)*(f*mu*100)*matter[:,1]/matter[:,0]
+
+def Pgv_l(mu):
+    return (mu**2*(f*mu*100)+(b+f*mu**2)*(mu*100))*matter[:,1]/matter[:,0]
+
+def Pgv_b(mu):
+    return (f*mu*100)*matter[:,1]/matter[:,0]
+
+def finvp(f00,f11,f01,f00p,f11p,f01p):
+    den = f00*f11 -f01**2
+    return f11p /den - f11*(f00p*f11 + f00*f11p - 2*f01*f01p)/den**2
+
+def Cmatrices(z,mu,ng,duration,sigm):
+    n = duration*restrate
+    zz = z * 100./3e5
+    sigv_factor = numpy.log(10)/5*zz/(1+zz)
+    sigv =sigm * sigv_factor *3e5
+    ninv = sigv**2/n
+
+    dsdn = sigv**2/n**2
+
+    nginv = 1./ng
+
+    pggs = Pgg(mu)
+    pgvs = Pgv(mu)
+    pvvs = Pvv(mu)
+
+    pggs_l = Pgg_l(mu)
+    pgvs_l = Pgv_l(mu)
[…]

### 2019-01-18 doc/src/partials.py
add fidelity to this

@@ -3,22 +3,55 @@
 import numpy
 import matplotlib.pyplot as plt
 import matplotlib
+import scipy.integrate as integrate
+from astropy.cosmology import FlatLambdaCDM
 
+# cosmology
+OmegaM0 = 0.28
+cosmo = FlatLambdaCDM(H0=100, Om0=OmegaM0)
 
+# power spectrum from CAMB
 matter = numpy.loadtxt("../../pv/dragan/matterpower.dat")
 
 # vv
-f=0.28**0.55
+f=OmegaM0**0.55
 
-# SN rates
+# SN properties
 restrate = 0.65*  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
-
-
 sigm = 0.08
 
 # galaxy density (TAIPAN Howlett et al.)
 ng = 1e-3
 
+
+
+def OmegaM(a):
+    return OmegaM0/(OmegaM0 + 0.72*a**3)
+
+def integrand_dlnfs(a):
+    return numpy.log(OmegaM(a))* OmegaM(a)**0.55 / a
+
+def dlnfs8dg(a):
+    return numpy.log(OmegaM(a)) + integrate.quad(integrand_dlnfs, 1e-8, a)[0]
+
+def integrand_D(a):
+    return  OmegaM(a)**0.55 / a
+
+# normalized to a=1
+def D(a):
+    return numpy.exp(-integrate.quad(integrand_D, a,1)[0])
+
+# a=1
+# print(numpy.log(OmegaM(a)),integrate.quad(integrand, 1e-8, a)[0],numpy.log(OmegaM(a)) + integrate.quad(integrand, 1e-8, a)[0])
+# a=1/(1+0.1)
+# print(numpy.log(OmegaM(a)),integrate.quad(integrand, 1e-8, a)[0],numpy.log(OmegaM(a)) + integrate.quad(integrand, 1e-8, a)[0])
+# a=0.5
+# print(numpy.log(OmegaM(a)),integrate.quad(integrand, 1e-8, a)[0],numpy.log(OmegaM(a)) + integrate.quad(integrand, 1e-8, a)[0])
+# a=1/(1+0.5)
[…]

### 2019-01-19 doc/src/partials.py
use SNe galaxies as tracers

@@ -86,15 +86,16 @@ def finvp(f00,f11,f01,f00p,f11p,f01p):
 
 def Cmatrices(z,mu,ng,duration,sigm):
     a = 1/(1.+z)
+    OmegaM_a = OmegaM(a)
     n = duration*restrate
     sigv_factor = numpy.log(10)/5*z/(1+z)
     sigv =sigm * sigv_factor *3e5
     lnfgfactor = dlnfs8dg(a)
     fDfactor = D(a)*(OmegaM(a)/OmegaM0)**0.55
 
-    ninv = sigv**2/n
 
-    dsdn = sigv**2/n**2
+    sigv2 = sigv**2
+    ninv = sigv2/n
 
     nginv = 1./ng
 
@@ -113,16 +114,20 @@ def Cmatrices(z,mu,ng,duration,sigm):
     Cinv = []
     dCdl = []
     dCdb = []
-    Cinvs = []
+    # Cinvs = []
+    Cinvn = []
     for pgg, pgv, pvv, pgg_l, pgv_l, pvv_l, pgg_b, pgv_b  in zip(pggs,pgvs,pvvs,pggs_l,pgvs_l,pvvs_l,pggs_b,pgvs_b):
-        C.append(numpy.array([[pgg+nginv,pgv],[pgv,pvv+ninv]]))
+        C.append(numpy.array([[pgg+ninv,pgv],[pgv,pvv+ninv]]))
         Cinv.append(numpy.linalg.inv(C[-1]))
-        dCdl.append(numpy.array([[pgg_l,pgv_l],[pgv_l,pvv_l]])*OmegaM(a)**0.55*lnfgfactor) # second term converts from fsigma8 to gamma
+        dCdl.append(numpy.array([[pgg_l,pgv_l],[pgv_l,pvv_l]])*OmegaM_a**0.55*lnfgfactor) # second term converts from fsigma8 to gamma
         dCdb.append(numpy.array([[pgg_b,pgv_b],[pgv_b,0]]))
-        den = (pgg+nginv)*(pvv+ninv)-pgv**2
-        Cinvs.append(1./den * numpy.array([[1,0],[0,0]]) - (pgg+nginv)/den**2 * numpy.array([[pvv+ninv,-pgv],[-pgv,pgg+nginv]]))
[…]

### 2019-01-19 doc/article.tex
updating

@@
 \usepackage{hyperref}
+\usepackage[overload]{textcase}
+
 %SetFonts
@@
 \author[0000-0001-6315-8743]{A.~G.~Kim}
+\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+\author{G.~Aldering}
 \affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
@@
 \begin{abstract}
-ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+Cadence wide-field imaging surveys 
+in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
 hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
@@
 
-\section{Connection between Type Ia Supernovae Correlations and Gravity}
+\section{}
 
@@
 
-The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
-It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the light-emittting structures
-in the Universe and from the peculiar velocities of test masses therein.
-Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
-and repulsion of density inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
-can be determined from its observer-frame magnitude and redshift. For a given background cosmology, the observed magnitude provides
-an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and CMB-frame redshifts.
+In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure the the influence of gravity on structure formation within the Universe.
+Peculiar velocities are sensitive to the growth of structure through the combination $fD$, where $D$ is  the linear growth factor and
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The growth of structure depends on gravity, indeed 
+\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
+give
+$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization for gravity, peculiar velocity
+surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$ \citep[see Figure~2 of][to see
+$fD$ plotted as a function of redshift for $\Lambda$CDM]{1475-7516-2013-04-031}.
 
-Baryonic structures and peculiar velocities  provide a measurement of the combination $fD$.  $D$ is the ``linear growth factor'' that
-gives the overall amplitude of  overdensities, and the ``linear growth rate''
-$$f \equiv \frac{d\ln{D}}{d\ln{a}}$$ is how that amplitude changes with redshift.  General Relativity predicts
-$f \approx \Omega_M^\gamma$ (with $D$ determined accordingly) with $\gamma=0.55$, whereas other gravity models can be similarly described
-with different values of $\gamma$
-\citep{2007APh....28..481L}.  
-The growth of structure, through the measurement of $fD$, provides a test of General Relativity and breaks degeneracies
-between gravitational and dark energy models that explain the accelerating expansion of the Universe.
-Non-GR models may  also predict a change in the redshift- and scale-dependence of the growth, such observations provide additional leverage in probing gravity.
+Peculiar velocity surveys have already been  used to measure $fD$, though not at a precision to distinguish gravity models.
+ \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane method and get distances to
+get a 15\% uncertainty in $fD$ at $z\approx 0$ using density and velocity cross-correlations.
+The upcoming 
+TAIPAN survey will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
+and the WALLABY+WNSHS surveys will obtain Tully-Fisher galaxies at densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
+$z=0-0.1$.
+Covering 75\% of the sky to a maximum depth of $z=0.1$, these surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
+Existing SN~Ia samples
+have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
+\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
+For reference, t $z>0.3$ redshifts DESI projects 10\% precision of $fD$ using RSD alone.
+%Measurement of the velocity field using LSST-discovered
+%SNe~Ia has been quantified by \citet{2011PhRvD..83d3004B,2017JCAP...01..060O}
+
+Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
+First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
+dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
+Recent studies indicate that with the right data, SNe absolute
+magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
+established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
+that is observed using current calibrations.
+One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
+Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
+  will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia for which good light curves can be measured.
+This is a sample size with more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
+
+A 10-year LSST SN survey can produce interesting $fD$ measurements in redshift bins, as shown in Table~1 of 
+\citet{2017ApJ...847..128H}.
+Projected percent uncertainties in $f\sigma_8$ from 
+\citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
+The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.
+
+%      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
+%     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
+%     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
+%     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
+%      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
+%     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
+
+The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
+Nevertheless, their low intrinsic-magnitude uncertainties 
+can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
+of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
+
+\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
+In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
+The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
+The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
+The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.
+
+It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
+In linear theory, the overdensity field can be expressed as
+$
+\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
+where $D(a)$ is  the linear growth factor.
+The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
+where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
+Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
+from the CMB, provide a measure of $fD$
+(or  equivalently $f\sigma_8$\footnote{
 The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
 commonly used in place of $D$ to normalize the
-overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.
+overall amplitude of  overdensities.}).
 
-Baryonic structures are sensitive to $f\sigma_8$ through redshift space distortions (RSD), 
-which on large scales gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
-and $\mu$ gives the angular separation of Fourier modes to the line-of-sight.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
-Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and so are bias-free.
-As
-mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
-leads to the dependence on $H_0 d_L$. (Note that the combination $H_0 d_L$ is independent of $H_0$.)
-Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
-insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
-than their naive sum.
+The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
+Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
+though this detail is not considered here.
+The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
+at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
+maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
+$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
+scale where  theory predictions are reliable.
+Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{src/new2.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
+The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
+redshift depths $z_{max}$.
+\label{power:fig}}
+\end{figure}
 
@@
 2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.
-Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample sizes by an order of magnitude
-over 75\% of the sky to a maximum depth of $z=0.1$.
 
@@
 
-Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
-First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
-dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
-Recent studies indicate that with the right data, SNe absolute
-magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
-established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
-that is observed using current calibrations.
-One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
-Secondly,  ZTF today and LSST in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia for which good light curves can be measured.
-This is a sample size with more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
-Measurement of the velocity field using LSST-discovered
-SNe~Ia has been previously quantified by \citet{2011PhRvD..83d3004B,2017JCAP...01..060O}.
+%The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
+%Their Figure~4b) shows how well these measurements can distinguish between $\gamma=0.42, 0.55, 0.68$
+%\citep[$f(R)$, GR,  and DGP gravities,][]{2007APh....28..481L} gravities respectively.
+%SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
+%kinematic S-Z effect)
+%RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
+%SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
+%
+%%{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
+%% Requires the booktabs if the memoir class is not being used
+%\begin{table}
+%   \centering
+%   %\topcaption{Table captions are better up top} % requires the topcapt package
+%   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
+%	\hline
+%	& RSD & RSD + PV & RSD+PV\\
+%	Redshift 	 & ($z$-bin) & ($z$-bin)& (cumulative)\\  \hline
+%      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
+%     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
+%     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
+%     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
+%      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
+%     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
+%      \hline
+%   \end{tabular}
+%   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
+%   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
+%   The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.}
+%   \label{tab:howlett}
+%\end{table}
 
-The precision in  $f\sigma_8$ derived from the baseline WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
-from both their RSD and peculiar velocities.
-The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
-Their Figure~4b) shows how well these measurements can distinguish between $\gamma=0.42, 0.55, 0.68$
-\citep[$f(R)$, GR,  and DGP gravities,][]{2007APh....28..481L} gravities respectively.
-SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
-kinematic S-Z effect)
-RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
-SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
-This can be compared to the current 15\% uncertainty from 6dF \citep{2017MNRAS.471..839A}, 3\% uncertainty projected
-for the combined TAIPAN and WALLABY+WNSHS surveys \citep{2017MNRAS.464.2517H} for low-redshift measurements of $f\sigma_8$.
-At larger $z>0.3$ redshifts DESI projects 10\% accuracies.
+Baryonic structures are sensitive to $fD$ through redshift space distortions (RSD), 
+which on large scales gives the combination $(b + f \mu^2)D$ where $b$ is the bias between the tracer and dark matter
+and $\mu$ gives the angular separation of Fourier modes to the line-of-sight. 
+The precision in  $fD$ derived from the baseline LSST WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
+from RSD of supernovae and peculiar velocities.
 
-%{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
-% Requires the booktabs if the memoir class is not being used
-\begin{table}
-   \centering
-   %\topcaption{Table captions are better up top} % requires the topcapt package
-   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
-	\hline
-	& RSD & RSD + PV & RSD+PV\\
-	Redshift 	 & ($z$-bin) & ($z$-bin)& (cumulative)\\  \hline
-      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
-     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
-     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
-     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
-      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
-     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
-      \hline
-   \end{tabular}
-   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
-   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
-   The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.}
-   \label{tab:howlett}
-\end{table}
+%Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
+%insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
+%than their naive sum.
 
-The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
-Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
-there is little benefit in going beyond
-$z=0.3$.
-At $z<0.2$ the constraining power on  $f\sigma_8$ 
-comes primarily from peculiar velocities
-but by $z=0.3$ RSD is more important.
+\section{Probing Gravity}
+Our scientific interest is in probing gravity.
+The combination $fD$ is sensitive to gravity. The growth rate $f$ describes how fast overdensities increase, which
+depends on how fast gravity makes mass move.
+General Relativity predicts
+$f \approx \Omega_M^\gamma$ with $\gamma=0.55$, whereas other gravity models can be similarly described
+with different values of
+$\gamma=0.42, 0.68$ for $f(R)$,  and DGP gravities respectively 
+\citep{2007APh....28..481L}.
 
-\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
-\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
+As a probe of gravity, peculiar velocity surveys are sensitive to $fD=\Omega_M^{\gamma} a \exp{\left(\left(\int_0^a \Omega_M^{\gamma} -1\right)  d\ln(a)\right)}$.
+Note that $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega} d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ in $\Lambda$CDM,
+meaning that surveys at low redshift better measure $\gamma$ given a fixed precent uncertainty in $fD$.
+Non-GR models may  also predict a change in the scale-dependence of the growth, such observations provide additional leverage in probing gravity
+but are not considered here.
 
-The  peculiar velocity signal  can be described by its power spectrum $P_{vv}$.
-The sensitivity to gravity primarily resides in the normalization of the power spectrum, as  $P_{vv} \propto (f\sigma_8)^2$.
-Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
-though this detail is not considered here.
-The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
-at $z=0$ expected from General Relativity and linear theory. Available $k$ ranges for surveys with different
-maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
-$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
-scale where  theory predictions are reliable.
-Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{src/power.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Top panel: Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
-Lower two panels: $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
-the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
-independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
-of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
-redshift depths $z_{max}$.
-\label{power:fig}}
-\end{figure}
-
-\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
- obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
-Their analysis is performed in Fourier space, 
-where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
-\begin{equation}
-F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-\frac{\partial C}{\partial \lambda_j} \right]
+The uncertainties in $\gamma$ for  RSD and peculiar velocity surveys are estimated using a
+Fisher matrix analysis
+similar to but with lower fidelity to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
+The analysis is performed in Fourier space, 
+where the Fisher information matrix of parameters $\lambda \in \{\gamma,b\}$ from random Gaussian field with mean zero is given by
+\begin{align}
+F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right]\\
+& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
 \label{fisher:eqn}
-\end{equation}
+\end{align}
  and covariance
@@
 \end{equation}
+The uncertainty in $\gamma$ is $\sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
+
+The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic magnitude dispersion 
+through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, and the survey duration $t$ through $n=\epsilon rt$ where $r$ is the observer SN~Ia rate
+and $\epsilon$ is the sample-selection efficiency.
+An analysis in which the RSD and velocity surveys are treated independently (e.g.\ they do not cover the same volume) is considered by
+setting the off-diagonal elements of $C$ to zero.
+
+We consider fiducial peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations or $t=2$ and 10~years, and with or without
+cross-correlations accounted for.
+The other survey parameters
+$\Omega=3$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
+A TAIPAN-like
+galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ is used for the RSD measurement.
+The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}, which shows a number of interesting features:
+\begin{itemize}
+\item Even shallow, two-year surveys give gammas for 
+\end{itemize}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{src/var.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{
+\label{var:fig}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.47\textwidth]{src/dvardz.png}
+\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{
+\label{dvar:fig}}
+\end{figure}
+
+%The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
+%It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the light-emittting structures
+%in the Universe and from the peculiar velocities of test masses therein.
+%Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
+%and repulsion of density inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
+%can be determined from its observer-frame magnitude and redshift. For a given background cosmology, the observed magnitude provides
+%an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and CMB-frame redshifts.
+
+%Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
+%Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and so are bias-free.
+%As
+%mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
+%leads to the dependence on $H_0 d_L$. (Note that the combination $H_0 d_L$ is independent of $H_0$.)
+
+%
+%The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
+%Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
+%there is little benefit in going beyond
+%$z=0.3$.
+%At $z<0.2$ the constraining power on  $f\sigma_8$ 
+%comes primarily from peculiar velocities
+%but by $z=0.3$ RSD is more important.
+
+\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
+\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
+
+\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
+ obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
+Their analysis is performed in Fourier space, 
+where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
+\begin{align}
+F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right]\\
+& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
+\label{fisher:eqn}
+\end{align}
+ and covariance
+\begin{equation}
+C(k,\mu)  =
+  \begin{bmatrix}
+   P_{gg}(k,\mu) + \frac{1}{n_g} &
+   P_{vg}(k,\mu)  \\
+   P_{vg}(k,\mu)  &
+  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+   \end{bmatrix}.
+\label{cov:eq}
+\end{equation}
+
 The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
@@
 
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\textwidth]{src/new.png}
+%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
+%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
+%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
+%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
+%redshift depths $z_{max}$.
+%\label{shot:fig}}
+%\end{figure}
+
 \subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
-The performance of a peculiar velocity survey depends on its overlap with galaxy surveys, solid-angle coverage, redshift depth, and intrinsic magnitude dispersion.
-In the following, we consider as given a  TAIPAN-like
-galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ that overlaps with the peculiar velocity surveys.
-While the expected variance in $f\sigma_8$ is $\left(F^{-1}\right)_{{f\sigma_{8}\,f\sigma_{8}}}$, we use $I$ as a proxy to help understand
-how survey densities and shot noise affect science performance.
-Surveys are compared relative to the $I_{max}$ of an idealized sample-variance limited peculiar-velocity survey with\ $\sigma/n =0$.
+First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets.
 
-First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
-Consider two- and ten-year
-SN~Ia densities $n$  based on SN~Ia rates 
-with 0.65 efficiency, and velocity dispersions $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$
-assuming a supernova magnitude dispersion of $\sigma_M=0.08$ and inserting values of $z=0.05,0.1, 0.2$.
 The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
@@
 $f\sigma_8$ at $z\sim 0.1$.
+
+\newpage
+We are interested in
+$$F^{-1}_{00,\alpha} = \frac{F_{11,\alpha}}{F_{00}F_{11}-F_{01}^2} - \frac{F_{11}}{(F_{00}F_{11}-F_{01}^2)^2}\left(F_{00,\alpha}F_{11}+
+F_{00}F_{11,\alpha} - 2 F_{01}F_{01,\alpha}\right).$$
+
+\begin{align}
+F_{ij} &  = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
+\end{align}
+
+$$F_{ij,z} =  \frac{\Omega}{8\pi^2} \frac{dr}{dz} r^2  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1}  k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk$$
+
+$$F_{ij,s} =  \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[
+C^{-1}_{,s} \frac{\partial C}{\partial \lambda_i} C^{-1}\frac{\partial C}{\partial \lambda_j} +
+C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}_{,s}\frac{\partial C}{\partial \lambda_j} 
+\right] d\mu\,dk\,dr$$
+
+\begin{equation}
+C =
+  \begin{bmatrix}
+   P_{gg}(k,\mu) + \frac{1}{n} &
+   P_{vg}(k,\mu)  \\
+   P_{vg}(k,\mu)  &
+  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+   \end{bmatrix}.
+\end{equation}
+
+\begin{align}
+C^{-1}_{,n}  =   &
+ \frac{n^{-2}( (P_{gg}(k,\mu) + \frac{1}{n})\sigma^2  +(P_{vv}(k,\mu) + \frac{\sigma^2}{n})) }
+ {\left((P_{gg}(k,\mu) + \frac{1}{n})(P_{vv}(k,\mu) + \frac{\sigma^2}{n})- P_{vg}(k,\mu)^2\right)^2} \\
+& \times 
+     \begin{bmatrix}
+     P_{vv}(k,\mu) + \frac{\sigma^2}{n} &
+   -P_{vg}(k,\mu)  \\
+  - P_{vg}(k,\mu)  &
+  P_{gg}(k,\mu) + \frac{1}{n} 
+   \end{bmatrix}\\
+   & -  \frac{n^{-2}}{(P_{gg}(k,\mu) + \frac{1}{n})(P_{vv}(k,\mu) + \frac{\sigma^2}{n})- P_{vg}(k,\mu)^2}
+ \begin{bmatrix}
+   \sigma^2 &
+   0\\
+   0 &
+  1
+   \end{bmatrix}.
+ \end{align}
+
+%
+%
+%\begin{align}
+%C^{-1}_{,s}  = &  \frac{1}{(P_{gg}(k,\mu) + \frac{1}{n_g})(P_{vv}(k,\mu) + s)- P_{vg}(k,\mu)^2}
+% \begin{bmatrix}
+%   1 &
+%   0\\
+%   0 &
+%  0
+%   \end{bmatrix}.\\
+%  & -
+% \frac{ P_{gg}(k,\mu) + \frac{1}{n_g} }{\left((P_{gg}(k,\mu) + \frac{1}{n_g})(P_{vv}(k,\mu) + s)- P_{vg}(k,\mu)^2\right)^2} \\
+%& \times 
+%     \begin{bmatrix}
+%     P_{vv}(k,\mu) + s &
+%   -P_{vg}(k,\mu)  \\
+%  - P_{vg}(k,\mu)  &
+%  P_{gg}(k,\mu) + \frac{1}{n_g} 
+%   \end{bmatrix}.
+% \end{align}
 

### 2019-01-19 doc/src/partials.py
fix bug in renormalization in power spectra

@@ -53,32 +53,32 @@ def D(a):
 # print(numpy.log(OmegaM(a)),integrate.quad(integrand, 1e-8, a)[0],numpy.log(OmegaM(a)) + integrate.quad(integrand, 1e-8, a)[0])
 # wefwe
 
-def Pvv(mu):
-    return (f*mu*100)**2*matter[:,1]/matter[:,0]**2
+def Pvv(mu,f,D):
+    return (f*D*mu*100)**2*matter[:,1]/matter[:,0]**2
 
-def Pvv_l(mu):
-    return 2*f*(mu*100)**2*matter[:,1]/matter[:,0]**2
+def Pvv_l(mu,f,D):
+    return 2*f*D*(mu*100)**2*matter[:,1]/matter[:,0]**2
 
 # gg
 b= 1.2
-def Pgg(mu):
-    return (b+f*mu**2)**2*matter[:,1]
+def Pgg(mu,f,D):
+    return (b*D+f*D*mu**2)**2*matter[:,1]
 
-def Pgg_l(mu):
-    return 2*mu**2*(b+f*mu**2)*matter[:,1]
+def Pgg_l(mu,f,D):
+    return 2*mu**2*(b*D+f*D*mu**2)*matter[:,1]
 
-def Pgg_b(mu):
-    return 2*(b+f*mu**2)*matter[:,1]
+def Pgg_b(mu,f,D):
+    return 2*(b*D+f*D*mu**2)*matter[:,1]
 
 # gv
-def Pgv(mu):
-    return (b+f*mu**2)*(f*mu*100)*matter[:,1]/matter[:,0]
+def Pgv(mu,f,D):
+    return (b*D+f*D*mu**2)*(f*D*mu*100)*matter[:,1]/matter[:,0]
 
-def Pgv_l(mu):
-    return (mu**2*(f*mu*100)+(b+f*mu**2)*(mu*100))*matter[:,1]/matter[:,0]
+def Pgv_l(mu,f,D):
+    return (mu**2*(f*D*mu*100)+(b*D+f*D*mu**2)*(mu*100))*matter[:,1]/matter[:,0]
 
-def Pgv_b(mu):
-    return (f*mu*100)*matter[:,1]/matter[:,0]
+def Pgv_b(mu,f,D):
+    return (f*D*mu*100)*matter[:,1]/matter[:,0]
 
 def finvp(f00,f11,f01,f00p,f11p,f01p):
     den = f00*f11 -f01**2
[…]

### 2019-01-19 doc/article.tex
plowing on

@@
 In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure the the influence of gravity on structure formation within the Universe.
-Peculiar velocities are sensitive to the growth of structure through the combination $fD$, where $D$ is  the linear growth factor and
-$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The growth of structure depends on gravity, indeed 
+The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
+The $\Lambda$CDM prediction for the $z=0$ velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
 \citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
@@
 $fD$ plotted as a function of redshift for $\Lambda$CDM]{1475-7516-2013-04-031}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{src/new2.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
+The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
+redshift depths $z_{max}$.
+\label{power:fig}}
+\end{figure}
 
@@
 \citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
-For reference, t $z>0.3$ redshifts DESI projects 10\% precision of $fD$ using RSD alone.
+For reference, DESI projects 10\% precision of $fD$ at $z>0.3$ redshifts  using RSD alone.
 %Measurement of the velocity field using LSST-discovered
@@
 
-Two advances in the upcoming decade will make SNe~Ia  important probes of $f\sigma_8$.
+Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
 First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
@@
   will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia for which good light curves can be measured.
-This is a sample size with more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
+hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
+with number density  $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$
+ for which good light curves can be measured. This sample has comparable
+ number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
 
-A 10-year LSST SN survey can produce interesting $fD$ measurements in redshift bins, as shown in Table~1 of 
-\citet{2017ApJ...847..128H}.
-Projected percent uncertainties in $f\sigma_8$ from 
-\citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
-The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.
+For example, over the course of a decade an LSST SN survey can produce interesting $fD$ measurements in several
+redshift bins.   
+\citet{2017ApJ...847..128H} find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, where at
+$0<z<0.2$ most of the signal comes from peculiar velocities and at higher redshifts $RSD$ is more constraining.
 
-%      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
-%     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
-%     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
-%     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
-%      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
-%     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
+Our scientific interest is in probing gravity, so while most of the literature cast peculiar velocity results in terms of 
+$fD$, the interesting parameter is $\gamma$.  Survey design should be tailored toward minimizing the uncertainty
+in 
+$\sigma_\gamma$ rather than $\sigma_{\ln{fD}}$.  To illustrate the distinction,
+ $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega} d\ln{a}$, which for $\Lambda$CDM
+ $\approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively; given two surveys with the same percentage precision  in $fD$, the one at lower redshift
+better constrains $\gamma$.  We thus thus relate surveys to $\sigma_\gamma$, where we assume a constant $\gamma$
+over the redshift range of the survey.
 
-The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
-Nevertheless, their low intrinsic-magnitude uncertainties 
-can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
-of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
+%
+%The combination $fD$ is sensitive to gravity. The growth rate $f$ describes how fast overdensities increase, which
+%depends on how fast gravity makes mass move.
+%General Relativity predicts
+%$f \approx \Omega_M^\gamma$ with $\gamma=0.55$, whereas other gravity models can be similarly described
+%with different values of
+%$\gamma=0.42, 0.68$ for $f(R)$,  and DGP gravities respectively 
+%\citep{2007APh....28..481L}.
 
+The uncertainties in $\gamma$ are estimated using a
+Fisher matrix analysis
+similar  to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
+The analysis is performed in Fourier space, 
+where the Fisher information matrix of parameters $\lambda \in \{\gamma,bD\}$ with
+the  tracer density and velocity power spectra used as data is
+\begin{align}
+F_{ij} 
+%& = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+%\frac{\partial C}{\partial \lambda_j} \right]\\
+& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
+\label{fisher:eqn}
+\end{align}
+where
+\begin{equation}
+C(k,\mu)  =
+  \begin{bmatrix}
+   P_{\delta \delta}(k,\mu) + \frac{1}{n} &
+   P_{v\delta}(k,\mu)  \\
+   P_{v\delta}(k,\mu)  &
+  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+   \end{bmatrix}.
+\label{cov:eq}
+\end{equation}
+The parameter dependence comes in through $P_{vv}\propto ((fD)(\gamma)\mu)^2$, the SN~Ia power spectrum $P_{\delta \delta }\propto ((bD) + (fD)(\gamma)\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
+\propto  ((bD) + (fD)\mu^2)f\sigma_8$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
+and the line of sight.
+
+The uncertainty in $\gamma$ is $\sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
+Non-GR models may  also predict a change in the scale-dependence of the growth or deviations
+from $f=\Omega_M^{\gamma}$, such observations provide additional leverage in probing gravity
+but are not considered here.
+
+The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic magnitude dispersion 
+through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, and the survey duration $t$ through $n=\epsilon rt$ where $r$ is the observer SN~Ia rate
+and $\epsilon$ is the sample-selection efficiency.
+An analysis in which the RSD and velocity surveys are treated independently (e.g.\ they do not cover the same volume) is considered by
+setting the off-diagonal elements of $C$ to zero.
+
+We consider fiducial peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations or $t=2$ and 10~years, and with or without
+cross-correlations accounted for.
+The other survey parameters
+$\Omega=3$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
+A TAIPAN-like
+galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ is used for the RSD measurement.
+The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}, which shows a number of interesting features:
+\begin{itemize}
+\item Even shallow, two-year surveys give gammas for 
+\end{itemize}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{src/var.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{
+\label{var:fig}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.47\textwidth]{src/dvardz.png}
+\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{
+\label{dvar:fig}}
+\end{figure}
+
+%
+%
+%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
+%Nevertheless, their low intrinsic-magnitude uncertainties 
+%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
+%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
+%
 \section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
@@
 Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{src/new2.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
-The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
-redshift depths $z_{max}$.
-\label{power:fig}}
-\end{figure}
 
@@
 \section{Probing Gravity}
-Our scientific interest is in probing gravity.
-The combination $fD$ is sensitive to gravity. The growth rate $f$ describes how fast overdensities increase, which
-depends on how fast gravity makes mass move.
-General Relativity predicts
-$f \approx \Omega_M^\gamma$ with $\gamma=0.55$, whereas other gravity models can be similarly described
-with different values of
-$\gamma=0.42, 0.68$ for $f(R)$,  and DGP gravities respectively 
-\citep{2007APh....28..481L}.
-
-As a probe of gravity, peculiar velocity surveys are sensitive to $fD=\Omega_M^{\gamma} a \exp{\left(\left(\int_0^a \Omega_M^{\gamma} -1\right)  d\ln(a)\right)}$.
-Note that $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega} d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ in $\Lambda$CDM,
-meaning that surveys at low redshift better measure $\gamma$ given a fixed precent uncertainty in $fD$.
-Non-GR models may  also predict a change in the scale-dependence of the growth, such observations provide additional leverage in probing gravity
-but are not considered here.
-
-The uncertainties in $\gamma$ for  RSD and peculiar velocity surveys are estimated using a
-Fisher matrix analysis
-similar to but with lower fidelity to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
-The analysis is performed in Fourier space, 
-where the Fisher information matrix of parameters $\lambda \in \{\gamma,b\}$ from random Gaussian field with mean zero is given by
-\begin{align}
-F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-\frac{\partial C}{\partial \lambda_j} \right]\\
-& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
-\label{fisher:eqn}
-\end{align}
- and covariance
-\begin{equation}
-C(k,\mu)  =
-  \begin{bmatrix}
-   P_{gg}(k,\mu) + \frac{1}{n_g} &
-   P_{vg}(k,\mu)  \\
-   P_{vg}(k,\mu)  &
-  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
-   \end{bmatrix}.
-\label{cov:eq}
-\end{equation}
-The uncertainty in $\gamma$ is $\sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
-
-The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic magnitude dispersion 
-through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, and the survey duration $t$ through $n=\epsilon rt$ where $r$ is the observer SN~Ia rate
-and $\epsilon$ is the sample-selection efficiency.
-An analysis in which the RSD and velocity surveys are treated independently (e.g.\ they do not cover the same volume) is considered by
-setting the off-diagonal elements of $C$ to zero.
-
-We consider fiducial peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations or $t=2$ and 10~years, and with or without
-cross-correlations accounted for.
-The other survey parameters
-$\Omega=3$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
-A TAIPAN-like
-galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ is used for the RSD measurement.
-The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}, which shows a number of interesting features:
-\begin{itemize}
-\item Even shallow, two-year surveys give gammas for 
-\end{itemize}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{src/var.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{
-\label{var:fig}}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.47\textwidth]{src/dvardz.png}
-\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{
-\label{dvar:fig}}
-\end{figure}
-
 %The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 

### 2019-01-24 doc/article.tex
forging on with edits

@@
 
-In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure the the influence of gravity on structure formation within the Universe.
+In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
 The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
-$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
+To be precise, the peculiar
+velocity power spectrum has dependence $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
+$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
+To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
+at low redshift.   The $H$-dependence of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
+}
 The $\Lambda$CDM prediction for the $z=0$ velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
 \citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
-give
-$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization for gravity, peculiar velocity
-surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$ \citep[see Figure~2 of][to see
-$fD$ plotted as a function of redshift for $\Lambda$CDM]{1475-7516-2013-04-031}.
+follow the relation
+$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization to model gravity, peculiar velocity
+surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose $\Lambda$CDM prediction is plotted 
+in Figure~2 of  \citet{1475-7516-2013-04-031}.
 
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Volume-weighted velocity power spectrum $k^3P_{vv}$ at $z=0$ for $\mu=1, 0,5$ as predicted for General Relativity in the linear regime.
+\caption{Volume-weighted velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
+Volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$, expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions.
 The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
@@
 
-Peculiar velocity surveys have already been  used to measure $fD$, though not at a precision to distinguish gravity models.
- \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane method and get distances to
+Peculiar velocity surveys have already been  used to measure $fD$, though not to a level where  gravity models can be precisely distinguished.
+ \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane method to
 get a 15\% uncertainty in $fD$ at $z\approx 0$ using density and velocity cross-correlations.
@@
 Covering 75\% of the sky to a maximum depth of $z=0.1$, these surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
+(For reference, DESI projects 10\% precision of $fD$ at $z>0.3$ redshifts  using RSD alone.)
 Existing SN~Ia samples
@@
 \citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
-For reference, DESI projects 10\% precision of $fD$ at $z>0.3$ redshifts  using RSD alone.
 %Measurement of the velocity field using LSST-discovered
@@
 hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
-with number density  $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$
- for which good light curves can be measured. This sample has comparable
+ for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$
+ (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)
+  This sample has comparable
  number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
 
-For example, over the course of a decade an LSST SN survey can produce interesting $fD$ measurements in several
+Given these two advances, over the course of a decade an LSST SN survey can produce interesting $fD$ measurements in several
 redshift bins.   
@@
 Our scientific interest is in probing gravity, so while most of the literature cast peculiar velocity results in terms of 
-$fD$, the interesting parameter is $\gamma$.  Survey design should be tailored toward minimizing the uncertainty
-in 
+$fD$, we now focus on $\gamma$.  Survey design should be tailored toward minimizing the uncertainty in $\gamma$
 $\sigma_\gamma$ rather than $\sigma_{\ln{fD}}$.  To illustrate the distinction,
@@
  $\approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively; given two surveys with the same percentage precision  in $fD$, the one at lower redshift
-better constrains $\gamma$.  We thus thus relate surveys to $\sigma_\gamma$, where we assume a constant $\gamma$
-over the redshift range of the survey.
-
-%
-%The combination $fD$ is sensitive to gravity. The growth rate $f$ describes how fast overdensities increase, which
-%depends on how fast gravity makes mass move.
-%General Relativity predicts
-%$f \approx \Omega_M^\gamma$ with $\gamma=0.55$, whereas other gravity models can be similarly described
-%with different values of
-%$\gamma=0.42, 0.68$ for $f(R)$,  and DGP gravities respectively 
-%\citep{2007APh....28..481L}.
-
+better constrains $\gamma$. 
 The uncertainties in $\gamma$ are estimated using a
@@
 The analysis is performed in Fourier space, 
-where the Fisher information matrix of parameters $\lambda \in \{\gamma,bD\}$ with
-the  tracer density and velocity power spectra used as data is
+where the Fisher information matrix of parameters $\lambda \in \{\gamma,bD\}$ is
 \begin{align}
@@
 \end{equation}
-The parameter dependence comes in through $P_{vv}\propto ((fD)(\gamma)\mu)^2$, the SN~Ia power spectrum $P_{\delta \delta }\propto ((bD) + (fD)(\gamma)\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
-\propto  ((bD) + (fD)\mu^2)f\sigma_8$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
-and the line of sight.
-
-The uncertainty in $\gamma$ is $\sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
-Non-GR models may  also predict a change in the scale-dependence of the growth or deviations
-from $f=\Omega_M^{\gamma}$, such observations provide additional leverage in probing gravity
+The parameter dependence enters through $fD$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia overdensity
+power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
+\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
+and the line of sight.  Note that while the $bD$ term does contain information on $\gamma$, we do not use it in this analysis.
+The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
+Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
 but are not considered here.
 
-The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic magnitude dispersion 
+The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic SN~Ia magnitude dispersion 
 through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, and the survey duration $t$ through $n=\epsilon rt$ where $r$ is the observer SN~Ia rate
 and $\epsilon$ is the sample-selection efficiency.
-An analysis in which the RSD and velocity surveys are treated independently (e.g.\ they do not cover the same volume) is considered by
+An analysis in which the RSD and velocity surveys are treated independently is considered by
 setting the off-diagonal elements of $C$ to zero.
 
-We consider fiducial peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations or $t=2$ and 10~years, and with or without
-cross-correlations accounted for.
+We calculate  $\sigma_\gamma$ of  peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations of $t=2$ and 10~years, covering $3\pi$  of sky, and with 
+analyses accounting for or neglecting
+cross-correlations.
 The other survey parameters
 $\Omega=3$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
-A TAIPAN-like
-galaxy survey with $b=1.2$ and number density $n_g=10^{-3}h^3$\,Mpc$^{-3}$ is used for the RSD measurement.
-The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}, which shows a number of interesting features:
-\begin{itemize}
-\item Even shallow, two-year surveys give gammas for 
-\end{itemize}
+The bias of SN~Ia hosts is taken to be $b=1.2$.
+The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}.
 
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{
+\caption{Projected $\sigma_\gamma$ of  peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations of $t=2$ and 10~years, and with 
+analyses accounting for or neglecting
+cross-correlations.
 \label{var:fig}}
@@
 
+Shallow, short-duration peculiar-velocity surveys already achieve meaningful constraints on gravity.   A 2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
+which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $2\sigma$ level.
+Using cross-correlation information decreases $\sigma_\gamma$ by  factors of $\sim 0.8$.   
+ This modest improvement 
+comes from the suppression of  sample variance.
+It is the density measurement that is sample-variance
+limited,  the peculiar velocity measurement is shot-noise limited.
+
+Having established that peculiar velocity surveys can produce a precise test of gravity, we now consider survey optimization through the 
+survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also show the relative importance of different
+redshifts with a given survey.  An important survey parameter that is not considered is cost.
+\begin{description}
+\item[Solid Angle $\Omega$]
+The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$.  The baseline
+LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
+the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+coverage of the southern equatorial pole could significantly increase
+the sky coverage with a corresponding decrease in the variance of $\gamma$.
+\item[Differential Redshift Bin $z$] Certain differential redshift bins  constrain $\gamma$ more strongly than others.  
+The left panel
+of Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
+In both cases the shape of the curve is the same, with the amplitude monotonically decreasing from $z=0.01$ out to $z=0.2$.  If we had
+to sculpt the distribution (say due to limited follow-up resources), the preference would be to sculpt out the highest redshift bins
+resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the SN-discovery distribution and the sample is unambiguously
+parameterized by $z_{max}$.  Note that this result is based on constraining a constant redshift-independent $\gamma$.
 \begin{figure}
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{
+\caption{Left: $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.  Right: $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
+ $z_{max}$ for two- and ten-year surveys.
 \label{dvar:fig}}
 \end{figure}
+\item[Redshift Depth $z_{max}$]
+Increasing the survey redshift depth increases the $\gamma$ precision though that precision asymptotes
+$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
+ \item[Survey duration $t$] There is a significant decrease on $\sigma_\gamma$ when going from a 2- to ten-year survey, as shown
+ in  Figure~\ref{var:fig}.   The right panel of Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
+ $z_{max}$ for two- and ten-year surveys.  The non-zero amplitudes show that shot-noise affects both two- and ten-year surveys, but
+ much less so for the latter more so at low redshifts.  After 10-years, the effect of sample-variance important for shallower surveys.
+\item[Intrinsic Magnitude Dispersion $\sigma_M$]
+The relationship between intrinsic magnitude dispersion and precision in $\gamma$ can be seen through the 
+previous result, as $\partial \sigma_\gamma/\partial \sigma_M = \partial \sigma_\gamma / \partial \ln{t}$.
+Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
+Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
+%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
+%limited so the $\sigma^2_M/n$-term
+%in the Fisher matrix is important.
+Survey follow-up strategy,
+which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
+This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
+ in number density, or equivalently in survey duration. 
+ \end{description}
+
+Shot-noise limit touched upon several times in the preceding discussion.  Go back to Figure.
+
+There are a number of interesting features to note:
+\begin{itemize}
+\item Ten-year survey has lower  $\sigma_\gamma$  relative to  two-year surveys with factors that are redshift-dependent.  For $z_{max}=0.3$
+the uncertainties are reduced by a factor $\sim 0.5$ whereas at  $z_{max}=0.7$ the factor is $\sim 0.7$
+\end{itemize}
+
+ Shot noise
+is important  Figure~\ref{power:fig}.
 
@@
 %The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
-
-{\bf Solid Angle $\Omega$:}
-$F_{f\sigma_{8}\,f\sigma_{8}}$ is  proportional to volume and hence the survey solid angle $\Omega$.  The baseline
-LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
-the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-coverage of the southern equatorial pole could significantly increase
-the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
-
-%s
-% and so helps illustrate 
-%how how the density, $n$, and intrinsic SN magnitude dispersion of peculiar velocity probes, as
-%$\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, affect survey precision.
-
-{\bf Common-Volume Galaxy  Survey:}
-Galaxy and velocity surveys in different volumes do not have the off-diagonal cross-correlation terms in Eq.~\ref{cov:eq}.
-The degradation in probative power   $I/I_{max}$, when sample-variance limited,
-is shown as the black dashed curves in the bottom two panels of Figure~\ref{power:fig}.
-The curves lie between $0.1-0.2$, showing
-that accounting for the common dark-matter fluctuations that give rise to galaxy densities and peculiar velocities
-reduces the variance in $f\sigma_8$ by factors of 5 to 10.
-
-{\bf Redshift Depth $z_{max}$:}
-Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is the expense of additional depth worth neglecting lower redshift
-SNe? The answer is found in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
-strongest probative power and are sensitive more to peculiar velocities rather than galaxy overdensities.  Despite
-the differential volume increase with redshift depth (Eq,~\ref{fisher:eqn}), the degradation in intrinsic peculiar velocity $\propto z/(1+z)$
-and being shot-noise-limited  makes a low-redshift tracer more valuable than one at higher redshift.
-
-{\bf Intrinsic Magnitude Dispersion $\sigma_M$}:
-Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
-Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
-%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
-%limited so the $\sigma^2_M/n$-term
-%in the Fisher matrix is important.
-Survey follow-up strategy,
-which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
-This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
- in number density, or equivalently in survey duration.
-
 %\subsection{Survey Parameters that Determine Scientific Performance}
@@
 
+\begin{align}
+C^{-1}_{,\sigma_M}  =   & \frac{\ln{10}}{5} \frac{z}{1+z} \frac{2\sigma}{n}
+( \frac{1}{(P_{gg}(k,\mu) + \frac{1}{n})(P_{vv}(k,\mu) + \frac{\sigma^2}{n})- P_{vg}(k,\mu)^2}
+ \begin{bmatrix}
+   1 &
+   0\\
+   0 &
+  0
+   \end{bmatrix}\\
+   & - 
+ \frac{ (P_{gg}(k,\mu) + \frac{1}{n}) }
+ {\left((P_{gg}(k,\mu) + \frac{1}{n})(P_{vv}(k,\mu) + \frac{\sigma^2}{n})- P_{vg}(k,\mu)^2\right)^2} \\
+& \times 
+     \begin{bmatrix}
+     P_{vv}(k,\mu) + \frac{\sigma^2}{n} &
+   -P_{vg}(k,\mu)  \\
+  - P_{vg}(k,\mu)  &
+  P_{gg}(k,\mu) + \frac{1}{n} 
+   \end{bmatrix})\\
+ \end{align}
 %

### 2019-01-24 doc/src/partials.py

@@ -87,18 +87,20 @@ def finvp(f00,f11,f01,f00p,f11p,f01p):
 def Cmatrices(z,mu,ng,duration,sigm):
     a = 1/(1.+z)
     OmegaM_a = OmegaM(a)
-    n = duration*restrate
+    n = duration*restrate/(1+z)
     sigv_factor = numpy.log(10)/5*z/(1+z)
     sigv =sigm * sigv_factor *3e5
     lnfgfactor = dlnfs8dg(a)
 
     sigv2 = sigv**2
-    ninv = sigv2/n
+    nvinv = sigv2/n
+    ninv = 1./n
 
-    nginv = 1./ng
+    # nginv = 1./ng
 
     f = OmegaM(a)**.55
     D_ = D(a)
+
     pggs = Pgg(mu,f,D_)
     pgvs = Pgv(mu,f,D_)
     pvvs = Pvv(mu,f,D_)
@@ -116,18 +118,21 @@ def Cmatrices(z,mu,ng,duration,sigm):
     dCdb = []
     # Cinvs = []
     Cinvn = []
+    Cinvsigmam = []
     for pgg, pgv, pvv, pgg_l, pgv_l, pvv_l, pgg_b, pgv_b  in zip(pggs,pgvs,pvvs,pggs_l,pgvs_l,pvvs_l,pggs_b,pgvs_b):
-        C.append(numpy.array([[pgg+ninv,pgv],[pgv,pvv+ninv]]))
+        C.append(numpy.array([[pgg+ninv,pgv],[pgv,pvv+nvinv]]))
         Cinv.append(numpy.linalg.inv(C[-1]))
         dCdl.append(numpy.array([[pgg_l,pgv_l],[pgv_l,pvv_l]])*OmegaM_a**0.55*lnfgfactor) # second term converts from fsigma8 to gamma
         dCdb.append(numpy.array([[pgg_b,pgv_b],[pgv_b,0]]))
-        den = (pgg+ninv)*(pvv+ninv)-pgv**2
-        # Cinvs.append(1./den * numpy.array([[1,0],[0,0]]) - (pgg+nginv)/den**2 * numpy.array([[pvv+ninv,-pgv],[-pgv,pgg+nginv]]))
+        den = (pgg+ninv)*(pvv+nvinv)-pgv**2
[…]

### 2019-01-24 doc/src/partials.py
fix C_sigma_M

@@ -131,7 +131,7 @@ def Cmatrices(z,mu,ng,duration,sigm):
         Cinvn.append(-1./den * numpy.array([[sigv2,0],[0,1]]) + (sigv2*(pgg+ninv)+(pvv+nvinv))/den**2 * numpy.array([[pvv+nvinv,-pgv],[-pgv,pgg+ninv]]))
         Cinvn[-1]=Cinvn[-1]/n #in terms of lnt  n**-2 * n = 1/n
         Cinvsigmam.append(1./den * numpy.array([[1,0],[0,0]]) + (pgg+ninv)/den**2 * numpy.array([[pvv+nvinv,-pgv],[-pgv,pgg+ninv]]))
-        Cinvsigmam[-1] = Cinvsigmam[-1] * sigv_factor 
+        Cinvsigmam[-1] = Cinvsigmam[-1] * sigv_factor *3e5 * 2 * sigv / n
     return C, Cinv, dCdl,dCdb,Cinvn, Cinvsigmam
 
 def traces(z,mu,ng,duration,sigm):
@@ -436,13 +436,13 @@ def set1():
     plt.savefig('dvardlnt.png')
     plt.clf()
 
-    plt.plot(zmaxs,dvardsigM[0]/2/numpy.sqrt(var[0]),label='Two Years',color='red')
-    plt.plot(zmaxs,dvardsigM[1]/2/numpy.sqrt(var[1]),label='Ten Years',color='black')
+    plt.plot(zmaxs,-dvardsigM[0]/2/numpy.sqrt(var[0]),label='Two Years',color='red')
+    plt.plot(zmaxs,-dvardsigM[1]/2/numpy.sqrt(var[1]),label='Ten Years',color='black')
     plt.legend()
     # plt.yscale("log", nonposy='clip')
     plt.ylim((0,plt.ylim()[1]))
     plt.xlabel(r'$z_{max}$')
-    plt.ylabel(r'$\frac{d\sigma_{\gamma}}{d\sigma_M$')
+    plt.ylabel(r'$|\frac{d\sigma_{\gamma}}{d\sigma_M}|$')
     plt.tight_layout()
     plt.savefig('dvardsigM.png')
     plt.clf()

### 2019-01-24 doc/article.tex
forging on

@@
 To be precise, the peculiar
-velocity power spectrum has dependence $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
+velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
 $\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
 To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
-at low redshift.   The $H$-dependence of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
+at low redshift.   The $H$-dependences of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
 }
@@
 \centering
-\includegraphics[width=0.5\textwidth]{src/new2.png}
+\includegraphics[width=0.47\textwidth]{src/new2.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
@@
 hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
- for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$
- (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)
+ for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$.
   This sample has comparable
@@
 
-Our scientific interest is in probing gravity, so while most of the literature cast peculiar velocity results in terms of 
-$fD$, we now focus on $\gamma$.  Survey design should be tailored toward minimizing the uncertainty in $\gamma$
-$\sigma_\gamma$ rather than $\sigma_{\ln{fD}}$.  To illustrate the distinction,
- $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega} d\ln{a}$, which for $\Lambda$CDM
- $\approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively; given two surveys with the same percentage precision  in $fD$, the one at lower redshift
-better constrains $\gamma$. 
-The uncertainties in $\gamma$ are estimated using a
-Fisher matrix analysis
-similar  to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
-The analysis is performed in Fourier space, 
-where the Fisher information matrix of parameters $\lambda \in \{\gamma,bD\}$ is
+Our scientific interest is in probing gravity so we here focus  on $\gamma$ rather than on $fD$.
+To illustrate the distinction,
+ $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega}\, d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively
+ in $\Lambda$CDM;  two surveys with the same fractional precision  in $fD$ will have different precision in $\gamma$,
+ with the one at lower redshift providing the tighter constraint. 
+We project uncertainties in $\gamma$, $\sigma_\gamma$ for a suite of idealized surveys
+using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
+The Fisher information matrix is
 \begin{align}
@@
   P_{vv}(k,\mu) + \frac{\sigma^2}{n}
-   \end{bmatrix}.
+   \end{bmatrix}
 \label{cov:eq}
 \end{equation}
-The parameter dependence enters through $fD$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia overdensity
+and the two parameters considered are $\lambda \in \{\gamma,bD\}$.
+The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy overdensity
 power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
 \propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
-and the line of sight.  Note that while the $bD$ term does contain information on $\gamma$, we do not use it in this analysis.
-The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma,\gamma}}$.
+and the line of sight.  While the $bD$ term does contain information on $\gamma$, it is not used for making constraints.
+The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
 Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
@@
 
-The survey parameters that enter the calculation are the solid angle $\Omega$, depth of the survey $r_{max}$ (or $z_{max}$), the intrinsic SN~Ia magnitude dispersion 
-through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$, and the survey duration $t$ through $n=\epsilon rt$ where $r$ is the observer SN~Ia rate
-and $\epsilon$ is the sample-selection efficiency.
-An analysis in which the RSD and velocity surveys are treated independently is considered by
+The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth $r_{max}$ (or $z_{max}$), duration
+$t$ through $n=\epsilon rt$ where $r$ is the observer-frame SN~Ia rate
+and $\epsilon$ is the sample-selection efficiency, and
+the intrinsic SN~Ia magnitude dispersion 
+through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$.
+An analysis in which the RSD and velocity surveys are treated independently is achieved by
 setting the off-diagonal elements of $C$ to zero.
 
-We calculate  $\sigma_\gamma$ of  peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations of $t=2$ and 10~years, covering $3\pi$  of sky, and with 
-analyses accounting for or neglecting
-cross-correlations.
+We consider peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
 The other survey parameters
-$\Omega=3$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
-The bias of SN~Ia hosts is taken to be $b=1.2$.
-The uncertainties in $\gamma$ are shown in Figure~\ref{var:fig}.
+$\Omega=3\pi$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
+The $k$-limits are  taken to be $k_{min} = (2 r_{max})^{-1}$  and 
+$k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
+The input bias of SN~Ia host galaxies is set as $b=1.2$.
+
+All the surveys considered can provide meaning tests of gravity.
+The projected uncertainty in $\gamma$ achieved by the suite of surveys are shown in Figure~\ref{var:fig}.
+For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities, one that uses
+both overdensities (RSD) and peculiar velocities independently, and one  that uses
+both RSD, peculiar velocities, and their cross-correlation.
+The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
+which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $2\sigma$ level.
+The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
+The figure also reveals an interesting interplay between RSD and velocity measurements.  Peculiar velocities alone
+can account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
+However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  factors of $\sim 0.8$.
+There are important
+$k$-modes that are  sample variance limited either in  overdensity and/or peculiar velocity who benefit from the
+noise suppression engendered by cross-correlations.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{src/var.png}
+\includegraphics[width=0.47\textwidth]{src/var.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Projected $\sigma_\gamma$ of  peculiar velocity surveys for a range of redshift depths $z_{max}$, with survey durations of $t=2$ and 10~years, and with 
-analyses accounting for or neglecting
-cross-correlations.
+\caption{The projected uncertainty in $\gamma$ achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
+For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed) , one that uses
+both overdensities (RSD) and peculiar velocities independently (dotted), and one  that uses
+both RSD, peculiar velocities, and their cross-correlation (solid).
 \label{var:fig}}
@@
 
-Shallow, short-duration peculiar-velocity surveys already achieve meaningful constraints on gravity.   A 2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
-which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $2\sigma$ level.
-Using cross-correlation information decreases $\sigma_\gamma$ by  factors of $\sim 0.8$.   
- This modest improvement 
-comes from the suppression of  sample variance.
-It is the density measurement that is sample-variance
-limited,  the peculiar velocity measurement is shot-noise limited.
-
-Having established that peculiar velocity surveys can produce a precise test of gravity, we now consider survey optimization through the 
-survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also show the relative importance of different
-redshifts with a given survey.  An important survey parameter that is not considered is cost.
+Survey performance is examined in more detail by considering how $\sigma_\gamma$ changes with respect to the 
+survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
+within a given survey.  
 \begin{description}
@@
 \item[Differential Redshift Bin $z$] Certain differential redshift bins  constrain $\gamma$ more strongly than others.  
-The left panel
-of Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
+Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
 In both cases the shape of the curve is the same, with the amplitude monotonically decreasing from $z=0.01$ out to $z=0.2$.  If we had
@@
 \centering
-\includegraphics[width=0.47\textwidth]{src/dvardz.png}
-\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
+\includegraphics[width=0.5\textwidth]{src/dvardz.png}
+%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Left: $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.  Right: $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
- $z_{max}$ for two- and ten-year surveys.
+\caption{$\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$. 
 \label{dvar:fig}}
@@
  \item[Survey duration $t$] There is a significant decrease on $\sigma_\gamma$ when going from a 2- to ten-year survey, as shown
- in  Figure~\ref{var:fig}.   The right panel of Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
+ in  Figure~\ref{var:fig}.   The left panel of Figure~\ref{dvar2:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
  $z_{max}$ for two- and ten-year surveys.  The non-zero amplitudes show that shot-noise affects both two- and ten-year surveys, but
  much less so for the latter more so at low redshifts.  After 10-years, the effect of sample-variance important for shallower surveys.
+ \begin{figure}
+\centering
+\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
+\includegraphics[width=0.47\textwidth]{src/dvardsigM.png}
+%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\caption{Left: $\partial \sigma_\gamma / \partial \ln{t}$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
+ $z_{max}$ for two- and ten-year surveys.
+\label{dvar2:fig}}
+\end{figure}
 \item[Intrinsic Magnitude Dispersion $\sigma_M$]
@@
 is important  Figure~\ref{power:fig}.
+ (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)
 
@@
 %
-\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
-In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
-The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
-The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
-The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.
-
-It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
-In linear theory, the overdensity field can be expressed as
-$
-\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
-where $D(a)$ is  the linear growth factor.
-The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
-where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
-Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
-from the CMB, provide a measure of $fD$
-(or  equivalently $f\sigma_8$\footnote{
-The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
-commonly used in place of $D$ to normalize the
-overall amplitude of  overdensities.}).
-
-The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
-Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
-though this detail is not considered here.
-The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
-at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
-maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
-$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
-scale where  theory predictions are reliable.
-Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
-
-The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
-The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
-2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag.
-
-The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
-Nevertheless, their low intrinsic-magnitude uncertainties 
-can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
-of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
-have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
-\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
-However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
+%\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
+%In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
+%The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
+%The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
+%The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.
+%
+%It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
+%In linear theory, the overdensity field can be expressed as
+%$
+%\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
+%where $D(a)$ is  the linear growth factor.
+%The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
+%where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
+%Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
+%from the CMB, provide a measure of $fD$
+%(or  equivalently $f\sigma_8$\footnote{
+%The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
+%commonly used in place of $D$ to normalize the
+%overall amplitude of  overdensities.}).
+%
+%
+%The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
+%Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
+%though this detail is not considered here.
+%The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
+%at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
+%maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
+%$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
+%scale where  theory predictions are reliable.
+%Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
+%
+%The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
+%The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
+%2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
+%
+%
+%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
+%Nevertheless, their low intrinsic-magnitude uncertainties 
+%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
+%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
+%have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
+%\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
+%However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
+%
 
@@
 %\end{table}
-
-Baryonic structures are sensitive to $fD$ through redshift space distortions (RSD), 
-which on large scales gives the combination $(b + f \mu^2)D$ where $b$ is the bias between the tracer and dark matter
-and $\mu$ gives the angular separation of Fourier modes to the line-of-sight. 
-The precision in  $fD$ derived from the baseline LSST WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
-from RSD of supernovae and peculiar velocities.
+%
+%Baryonic structures are sensitive to $fD$ through redshift space distortions (RSD), 
+%which on large scales gives the combination $(b + f \mu^2)D$ where $b$ is the bias between the tracer and dark matter
+%and $\mu$ gives the angular separation of Fourier modes to the line-of-sight. 
+%The precision in  $fD$ derived from the baseline LSST WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
+%from RSD of supernovae and peculiar velocities.
+%
 
@@
 
-\section{Probing Gravity}
+%\section{Probing Gravity}
 %The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
@@
 
-\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
-\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
-
-\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
- obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
-Their analysis is performed in Fourier space, 
-where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
-\begin{align}
-F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-\frac{\partial C}{\partial \lambda_j} \right]\\
-& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
-\label{fisher:eqn}
-\end{align}
- and covariance
-\begin{equation}
-C(k,\mu)  =
-  \begin{bmatrix}
-   P_{gg}(k,\mu) + \frac{1}{n_g} &
-   P_{vg}(k,\mu)  \\
-   P_{vg}(k,\mu)  &
-  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
-   \end{bmatrix}.
-\label{cov:eq}
-\end{equation}
-
-The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
-\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
-The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
-is $\sigma$. 
-The integrand in the Fisher-element calculation
-\begin{equation}
-I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
-\frac{\partial C}{\partial  (f\sigma_8)} \right]
-\end{equation}
-represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
-differential Fourier
-volume element.
-
 %
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\textwidth]{src/new.png}
-%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
-%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
-%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
-%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
-%redshift depths $z_{max}$.
-%\label{shot:fig}}
-%\end{figure}
-
-\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
-First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets.
-
-The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
-Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
-Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
-is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
+%\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
+%\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
+%
+%
+%
+%\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
+% obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
+%Their analysis is performed in Fourier space, 
+%where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
+%\begin{align}
+%F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+%\frac{\partial C}{\partial \lambda_j} \right]\\
+%& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+%\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
+%\label{fisher:eqn}
+%\end{align}
+% and covariance
+%\begin{equation}
+%C(k,\mu)  =
+%  \begin{bmatrix}
+%   P_{gg}(k,\mu) + \frac{1}{n_g} &
+%   P_{vg}(k,\mu)  \\
+%   P_{vg}(k,\mu)  &
+%  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+%   \end{bmatrix}.
+%\label{cov:eq}
+%\end{equation}
+%
+%
+%The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
+%\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
+%The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
+%is $\sigma$. 
+%The integrand in the Fisher-element calculation
+%\begin{equation}
+%I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
+%\frac{\partial C}{\partial  (f\sigma_8)} \right]
+%\end{equation}
+%represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
+%differential Fourier
+%volume element.  
+%
+%%
+%%\begin{figure}
+%%\centering
+%%\includegraphics[width=0.5\textwidth]{src/new.png}
+%%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+%%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
+%%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
+%%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
+%%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
+%%redshift depths $z_{max}$.
+%%\label{shot:fig}}
+%%\end{figure}
+%
+%
+%\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
+%First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
+%
+%The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
+%Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
+%Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
+%is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
 Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
@@
 $f\sigma_8$ at $z\sim 0.1$.
+
+An important survey parameter that is not considered is cost.
 

### 2019-01-24 doc/src/partials.py

@@ -152,7 +152,9 @@ def traces(z,mu,ng,duration,sigm):
 
     llsigM=[]
     lbsigM=[]
-    bbsigM=[]   
+    bbsigM=[]
+
+    ll_vonly=[]
 
     for C, Cinv, C_l,C_b,Cinvn,CinvsigM in zip(cmatrices[0],cmatrices[1],cmatrices[2],cmatrices[3],cmatrices[4],cmatrices[5]):
         ll.append(numpy.trace(Cinv@C_l@Cinv@C_l))
@@ -171,8 +173,11 @@ def traces(z,mu,ng,duration,sigm):
         bb_ind.append(C[0][0]**(-2)*C_b[0][0]**2+C[1][1]**(-2)*C_b[1][1]**2)
         lb_ind.append(C[0][0]**(-2)*C_l[0][0]*C_b[0][0]+C[1][1]**(-2)*C_l[1][1]*C_b[1][1])
 
+        ll_vonly.append(C[1][1]**(-2)*C_l[1][1]**2)
+
     return numpy.array(ll),numpy.array(bb),numpy.array(lb),numpy.array(lls),numpy.array(bbs),numpy.array(lbs), \
-        numpy.array(ll_ind),numpy.array(bb_ind),numpy.array(lb_ind),numpy.array(llsigM),numpy.array(bbsigM),numpy.array(lbsigM)
+        numpy.array(ll_ind),numpy.array(bb_ind),numpy.array(lb_ind),numpy.array(llsigM),numpy.array(bbsigM),numpy.array(lbsigM), \
+        numpy.array(ll_vonly)
 
 def muintegral(z,ng,duration,sigm):
     mus=numpy.arange(-1,1.001,0.1)
@@ -189,10 +194,11 @@ def muintegral(z,ng,duration,sigm):
     llsigM=numpy.zeros((len(mus),len(matter[:,0])))
     bbsigM=numpy.zeros((len(mus),len(matter[:,0])))
     blsigM=numpy.zeros((len(mus),len(matter[:,0])))
+    ll_vonly=numpy.zeros((len(mus),len(matter[:,0])))  
 
     for i in range(len(mus)):
         dum = traces(z,mus[i],ng,duration,sigm)
[…]

### 2019-01-25 doc/article.tex
going on

@@
 $k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
+The sample-selection efficiency $\epsilon$ is redshift-independent, i.e.\ the native redshift distribution is not sculpted.
 The input bias of SN~Ia host galaxies is set as $b=1.2$.
@@
 \caption{The projected uncertainty in $\gamma$ achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
-For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed) , one that uses
+For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
 both overdensities (RSD) and peculiar velocities independently (dotted), and one  that uses
@@
 
-Survey performance is examined in more detail by considering how $\sigma_\gamma$ changes with respect to the 
+Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
+changes with respect to the 
 survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
@@
 \item[Solid Angle $\Omega$]
-The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$.  The baseline
-LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
-the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-coverage of the southern equatorial pole could significantly increase
-the sky coverage with a corresponding decrease in the variance of $\gamma$.
-\item[Differential Redshift Bin $z$] Certain differential redshift bins  constrain $\gamma$ more strongly than others.  
-Figure~\ref{dvar:fig} shows $\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$.
-In both cases the shape of the curve is the same, with the amplitude monotonically decreasing from $z=0.01$ out to $z=0.2$.  If we had
-to sculpt the distribution (say due to limited follow-up resources), the preference would be to sculpt out the highest redshift bins
-resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the SN-discovery distribution and the sample is unambiguously
-parameterized by $z_{max}$.  Note that this result is based on constraining a constant redshift-independent $\gamma$.
+The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \sqrt{\Omega}$.
+\item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
+a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
+This is demonstrated to be the case at the end of both 2- and 10-year surveys and $z_{max}=0.2$.
+Figure~\ref{dvar:fig} shows
+ $|\partial \sigma_\gamma / \partial z|$, which
+for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
+to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
+resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
 \begin{figure}
 \centering
-\includegraphics[width=0.5\textwidth]{src/dvardz.png}
+\includegraphics[width=0.47\textwidth]{src/dvardz.png}
 %\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{$\partial \sigma_\gamma / \partial z$  for two- and ten-year surveys with limiting $z_{max}=0.2$. 
+\caption{$|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
 \label{dvar:fig}}
@@
 \item[Redshift Depth $z_{max}$]
-Increasing the survey redshift depth increases the $\gamma$ precision though that precision asymptotes
+Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
 $z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
- \item[Survey duration $t$] There is a significant decrease on $\sigma_\gamma$ when going from a 2- to ten-year survey, as shown
- in  Figure~\ref{var:fig}.   The left panel of Figure~\ref{dvar2:fig} shows $\partial \sigma_\gamma / \partial \ln{t}$ as a function of
- $z_{max}$ for two- and ten-year surveys.  The non-zero amplitudes show that shot-noise affects both two- and ten-year surveys, but
- much less so for the latter more so at low redshifts.  After 10-years, the effect of sample-variance important for shallower surveys.
+ \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] Increasing the survey duration to increase the number density improves constraints.  The benefit is stronger for
+shallow compared to deep surveys, and the strength of the  benefit drops significantly over the course of 10 years.
+These trends are seen in the left panel of Figure~\ref{dvar2:fig}, which plots $|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
+ $z_{max}$ for two- and ten-year surveys, and
+reflect those $k$-modes that do not benefit more data because they are sample-variance limited, either because they are at low $k$ for shallow surveys
+or have already been saturated over the course of the survey.
  \begin{figure}
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Left: $\partial \sigma_\gamma / \partial \ln{t}$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
+\caption{Left: $|\partial \sigma_\gamma / \partial \ln{t}|$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
  $z_{max}$ for two- and ten-year surveys.
@@
 \end{figure}
-\item[Intrinsic Magnitude Dispersion $\sigma_M$]
-The relationship between intrinsic magnitude dispersion and precision in $\gamma$ can be seen through the 
-previous result, as $\partial \sigma_\gamma/\partial \sigma_M = \partial \sigma_\gamma / \partial \ln{t}$.
-Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
-Thus, a decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
-%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
-%limited so the $\sigma^2_M/n$-term
-%in the Fisher matrix is important.
-Survey follow-up strategy,
-which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
-This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
+Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
+has similar relationships with $\sigma_\gamma$.
+This is evident in the plot of $\partial \sigma_\gamma/\partial \sigma_M$ shown in the right panel of Figure~\ref{dvar2:fig}.
+The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
  in number density, or equivalently in survey duration. 
@@
 
-Shot-noise limit touched upon several times in the preceding discussion.  Go back to Figure.
-
-There are a number of interesting features to note:
-\begin{itemize}
-\item Ten-year survey has lower  $\sigma_\gamma$  relative to  two-year surveys with factors that are redshift-dependent.  For $z_{max}=0.3$
-the uncertainties are reduced by a factor $\sim 0.5$ whereas at  $z_{max}=0.7$ the factor is $\sim 0.7$
-\end{itemize}
-
- Shot noise
-is important  Figure~\ref{power:fig}.
- (the total rest-frame volumetric SN~Ia rate being $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.)
+From the above calculations, we find empirically that shot noise is important since higher number densities and lower intrinsic magnitude dispersion
+significantly reduce $\gamma$ uncertainty, but that sample variance is also relevant  by the end of a ten-year survey.
+Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
+First note that the  rest-frame volumetric SN~Ia rate (full selection efficiency) is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
+after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
+$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
+sare shown in  Figure~\ref{power:fig}.  For all of these survey scenarios, the transition between sample variance and shot-noise equality 
+occurs at a $k$ squarely within the $k$-range that carries the most signal.
 
@@
 %is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
-Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
+%Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
 %The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
@@
 reach a targeted $\sigma_M$.
+
+The baseline
+LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
+the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+coverage of the southern equatorial pole could significantly increase
+the sky coverage with a corresponding decrease in the variance of $\gamma$.
+
 \item Follow-up resources to determine per-SN distances:
@@
 An important survey parameter that is not considered is cost.
+
+Effect of 10-years.
 

### 2019-01-25 doc/src/partials.py

@@ -130,7 +130,7 @@ def Cmatrices(z,mu,ng,duration,sigm):
 
         Cinvn.append(-1./den * numpy.array([[sigv2,0],[0,1]]) + (sigv2*(pgg+ninv)+(pvv+nvinv))/den**2 * numpy.array([[pvv+nvinv,-pgv],[-pgv,pgg+ninv]]))
         Cinvn[-1]=Cinvn[-1]/n #in terms of lnt  n**-2 * n = 1/n
-        Cinvsigmam.append(1./den * numpy.array([[1,0],[0,0]]) + (pgg+ninv)/den**2 * numpy.array([[pvv+nvinv,-pgv],[-pgv,pgg+ninv]]))
+        Cinvsigmam.append(1./den * numpy.array([[1,0],[0,0]]) - (pgg+ninv)/den**2 * numpy.array([[pvv+nvinv,-pgv],[-pgv,pgg+ninv]]))
         Cinvsigmam[-1] = Cinvsigmam[-1] * sigv_factor *3e5 * 2 * sigv / n
     return C, Cinv, dCdl,dCdb,Cinvn, Cinvsigmam
 
@@ -359,7 +359,7 @@ def zintegral(zmax,ng,duration,sigm):
 
 
 def set2():
-
+    fig,(ax) = plt.subplots(1, 1, figsize=(8*.85,6*.85))
     durations = [2,10]
     labels = ['Two Years','Ten Years']
     colors = ['red','black']
@@ -376,12 +376,12 @@ def set2():
 
         for duration,label,color in zip(durations,labels,colors):
             dvardz=[]
-            f00,f11,f10, _,_,_,_,_,_,_,_,_ = zintegral(zmax,ng,duration,sigm)
+            f00,f11,f10, _,_,_,_,_,_,_, _,_,_ = zintegral(zmax,ng,duration,sigm)
             var= numpy.linalg.inv(numpy.array([[f00,f10],[f10,f11]]))[0,0]*2*3.14/.75
             for z,r in zip(zs,rs):
                 a=1./(1+z)
                 drdz = 1/numpy.sqrt(OmegaM0/a**3 + (1-OmegaM0)) # 1/H
-                _,_,_, f00s,f11s,f10s,_,_,_ = kintegral(z,zmax,ng,duration,sigm)
+                _,_,_, f00s,f11s,f10s,_,_,_,_,_,_,_ = kintegral(z,zmax,ng,duration,sigm)
[…]

### 2019-01-25 doc/article.tex
forging on

@@
 \begin{abstract}
-Cadence wide-field imaging surveys 
+Cadenced wide-field imaging surveys 
 in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
 hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
-their absolute magnitudes, now make them competitive probes of the growth of structure.  The peculiar velocity power spectrum
-is sensitive to $f\sigma_8$,
-the product of the linear growth and amplitude of density perturbations.  Cross-correlation with synergistic galaxy surveys further constrains $f \sigma_8$ and the galaxy bias. 
-Thus, in the next decade the peculiar velocities of
-SNe~Ia will provide the growth of structure  in the local $z<0.3$ Universe as a powerful test of General Relativity and other models of gravity.
+their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
+is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
+In the next decade the peculiar velocities of
+SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ 
+down to $0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
 \end{abstract}
@@
 Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
-First note that the  rest-frame volumetric SN~Ia rate (full selection efficiency) is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
+The  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
 after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
 $\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
-sare shown in  Figure~\ref{power:fig}.  For all of these survey scenarios, the transition between sample variance and shot-noise equality 
-occurs at a $k$ squarely within the $k$-range that carries the most signal.
+can be compared using  Figure~\ref{power:fig}.  For all of these survey scenarios, sample variance and shot-noise equality 
+occurs at a $k$  within the range that carries the most signal.  However, the ten-year surveys do reach the sample variance
+limit if   $k_{max} < 0.2$~$h$\,Mpc$^{-1}$.
 
@@
 calibration may be sufficient.  Progenitor population  is not expected to evolve significantly in the $0.01<z<0.2$ Universe.
+
+Modeling $k_{max}$ is a high as possible.
 

### 2019-01-25 doc/src/partials.py

@@ -270,7 +270,11 @@ def kintegral(z,zmax,ng,duration,sigm):
     blsigM = numpy.trapz(matter[:,0][w]**2*blsigM[w],matter[:,0][w])
     ll_vonly = numpy.trapz(matter[:,0][w]**2*ll_vonly[w],matter[:,0][w])
 
-    return ll,bb,bl, lls,bbs,bls,ll_ind,bb_ind,bl_ind, llsigM,bbsigM,blsigM,ll_vonly
+    llkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*ll)
+    bbkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*bb)
+    blkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*bl)
+
+    return ll,bb,bl, lls,bbs,bls,ll_ind,bb_ind,bl_ind, llsigM,bbsigM,blsigM,ll_vonly, llkmax,bbkmax,blkmax
 
 # utility numbers
 zmax_zint = 0.3
@@ -301,9 +305,12 @@ def zintegral(zmax,ng,duration,sigm):
     bbsigM=numpy.zeros(len(rs))
     blsigM=numpy.zeros(len(rs))
     ll_vonly=numpy.zeros(len(rs))
+    llkmax=numpy.zeros(len(rs))
+    bbkmax=numpy.zeros(len(rs))
+    blkmax=numpy.zeros(len(rs))
 
     for i in range(len(rs)):
-        ll[i],bb[i],bl[i],lls[i],bbs[i],bls[i],ll_ind[i],bb_ind[i],bl_ind[i],llsigM[i],bbsigM[i],blsigM[i],ll_vonly[i] = kintegral(zs[i],zmax,ng,duration,sigm)
+        ll[i],bb[i],bl[i],lls[i],bbs[i],bls[i],ll_ind[i],bb_ind[i],bl_ind[i],llsigM[i],bbsigM[i],blsigM[i],ll_vonly[i],llkmax[i],bbkmax[i],blkmax[i] = kintegral(zs[i],zmax,ng,duration,sigm)
 
     ll = numpy.trapz(rs**2*ll,rs)
     bb = numpy.trapz(rs**2*bb,rs)
@@ -318,6 +325,9 @@ def zintegral(zmax,ng,duration,sigm):
     bbsigM = numpy.trapz(rs**2*bbsigM,rs)
[…]

### 2019-01-26 doc/article.tex
added k_max derivative and continued on with text changes

@@
 survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
-within a given survey.  
+within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.
 \begin{description}
@@
 This is demonstrated to be the case at the end of both 2- and 10-year surveys and $z_{max}=0.2$.
-Figure~\ref{dvar:fig} shows
+The top panel of Figure~\ref{dvar:fig} shows
  $|\partial \sigma_\gamma / \partial z|$, which
@@
 \centering
-\includegraphics[width=0.47\textwidth]{src/dvardz.png}
+\includegraphics[width=0.4\textwidth]{src/dvardz.png}\\
+\includegraphics[width=0.4\textwidth]{src/dvardlnt.png}
+\includegraphics[width=0.4\textwidth]{src/dvardsigM.png}
+%\includegraphics[width=0.4\textwidth]{src/dvardkmax.png}
 %\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{$|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
+\caption{Top: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
+Bottom Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$; Bottom Right: $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$each as a function of
+ $z_{max}$ for two- and ten-year surveys.
 \label{dvar:fig}}
@@
 $z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
- \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] Increasing the survey duration to increase the number density improves constraints.  The benefit is stronger for
-shallow compared to deep surveys, and the strength of the  benefit drops significantly over the course of 10 years.
-These trends are seen in the left panel of Figure~\ref{dvar2:fig}, which plots $|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
- $z_{max}$ for two- and ten-year surveys, and
-reflect those $k$-modes that do not benefit more data because they are sample-variance limited, either because they are at low $k$ for shallow surveys
-or have already been saturated over the course of the survey.
- \begin{figure}
-\centering
-\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
-\includegraphics[width=0.47\textwidth]{src/dvardsigM.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Left: $|\partial \sigma_\gamma / \partial \ln{t}|$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
+ \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] 
+Increasing the survey duration  increases the number density of discovered supernovae. As
+shot-noise is relevant for all the surveys considered, the increased survey duration improves the
+precision in $\gamma$. The surveys that benefit most are those more susceptible to shot noise:
+higher redshifts  due to their higher velocity dispersion and shorter surveys  due to their lower number 
+densities.
+These trends are shown in the bottom-left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
  $z_{max}$ for two- and ten-year surveys.
-\label{dvar2:fig}}
-\end{figure}
 Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
-has similar relationships with $\sigma_\gamma$.
-This is evident in the plot of $\partial \sigma_\gamma/\partial \sigma_M$ shown in the right panel of Figure~\ref{dvar2:fig}.
-The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
- in number density, or equivalently in survey duration. 
+has a similar relationship with $\sigma_\gamma$.
+This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ shown in the bottom-right panel of Figure~\ref{dvar:fig}.
+ \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory and
+ are Gaussian-distributed.  Changes in this scale engender changes in the $\gamma$ precision, as    $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$
+and is survey-independent.
  \end{description}
 
-From the above calculations, we find empirically that shot noise is important since higher number densities and lower intrinsic magnitude dispersion
-significantly reduce $\gamma$ uncertainty, but that sample variance is also relevant  by the end of a ten-year survey.
-Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
-The  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
-after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
+From the above calculations, we find empirically that shot noise is important
+but that sample variance is non-negligible at the end of longer surveys or when the intrinsic magnitude dispersion is small.
+Fine tuning
+survey design is aided by understanding when surveys become sample variance limited.
+After a few years the RSD measurement is sample variance limited,
+the  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$  
+For peculiar velocities,
+the  contributions of $P_{vv}$ and
 $\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
-can be compared using  Figure~\ref{power:fig}.  For all of these survey scenarios, sample variance and shot-noise equality 
-occurs at a $k$  within the range that carries the most signal.  However, the ten-year surveys do reach the sample variance
-limit if   $k_{max} < 0.2$~$h$\,Mpc$^{-1}$.
+can be compared using  Figure~\ref{power:fig}.  For these survey scenarios, sample variance and shot-noise equality 
+occurs at a $k$  within the range that carries the most signal; both sample variance and shot-noise limited modes contribute to the constraints.
+If $k_{max}$ were to decrease moderately some of the surveys could become sample variance limited, whereas increases in $k_{max}$ introduce additional
+ shot-limited modes.
 
-%
-%
-%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
-%Nevertheless, their low intrinsic-magnitude uncertainties 
-%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
-%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
-%
-%\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
-%In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
-%The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
-%The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
-%The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.
-%
-%It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
-%In linear theory, the overdensity field can be expressed as
-%$
-%\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
-%where $D(a)$ is  the linear growth factor.
-%The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
-%where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
-%Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
-%from the CMB, provide a measure of $fD$
-%(or  equivalently $f\sigma_8$\footnote{
-%The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
-%commonly used in place of $D$ to normalize the
-%overall amplitude of  overdensities.}).
-%
-%
-%The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
-%Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
-%though this detail is not considered here.
-%The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
-%at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
-%maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
-%$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
-%scale where  theory predictions are reliable.
-%Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
-%
-%The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
-%The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
-%2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
-%
-%
-%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
-%Nevertheless, their low intrinsic-magnitude uncertainties 
-%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
-%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
-%have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
-%\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
-%However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
-%
-
-%The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
-%Their Figure~4b) shows how well these measurements can distinguish between $\gamma=0.42, 0.55, 0.68$
-%\citep[$f(R)$, GR,  and DGP gravities,][]{2007APh....28..481L} gravities respectively.
-%SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
-%kinematic S-Z effect)
-%RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
-%SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
-%
-%%{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
-%% Requires the booktabs if the memoir class is not being used
-%\begin{table}
-%   \centering
-%   %\topcaption{Table captions are better up top} % requires the topcapt package
-%   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
-%	\hline
-%	& RSD & RSD + PV & RSD+PV\\
-%	Redshift 	 & ($z$-bin) & ($z$-bin)& (cumulative)\\  \hline
-%      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
-%     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
-%     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
-%     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
-%      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
-%     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
-%      \hline
-%   \end{tabular}
-%   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
-%   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
-%   The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.}
-%   \label{tab:howlett}
-%\end{table}
-%
-%Baryonic structures are sensitive to $fD$ through redshift space distortions (RSD), 
-%which on large scales gives the combination $(b + f \mu^2)D$ where $b$ is the bias between the tracer and dark matter
-%and $\mu$ gives the angular separation of Fourier modes to the line-of-sight. 
-%The precision in  $fD$ derived from the baseline LSST WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
-%from RSD of supernovae and peculiar velocities.
-%
-
-%Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
-%insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
-%than their naive sum.
-
-%\section{Probing Gravity}
-%The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
-%It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the light-emittting structures
-%in the Universe and from the peculiar velocities of test masses therein.
-%Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
-%and repulsion of density inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
-%can be determined from its observer-frame magnitude and redshift. For a given background cosmology, the observed magnitude provides
-%an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and CMB-frame redshifts.
-
-%Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
-%Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and so are bias-free.
-%As
-%mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
-%leads to the dependence on $H_0 d_L$. (Note that the combination $H_0 d_L$ is independent of $H_0$.)
-
-%
-%The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
-%Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
-%there is little benefit in going beyond
-%$z=0.3$.
-%At $z<0.2$ the constraining power on  $f\sigma_8$ 
-%comes primarily from peculiar velocities
-%but by $z=0.3$ RSD is more important.
-
-%
-%\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
-%\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
-%
-%
-%
-%\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
-% obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
-%Their analysis is performed in Fourier space, 
-%where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
-%\begin{align}
-%F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-%\frac{\partial C}{\partial \lambda_j} \right]\\
-%& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
-%\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
-%\label{fisher:eqn}
-%\end{align}
-% and covariance
-%\begin{equation}
-%C(k,\mu)  =
-%  \begin{bmatrix}
-%   P_{gg}(k,\mu) + \frac{1}{n_g} &
-%   P_{vg}(k,\mu)  \\
-%   P_{vg}(k,\mu)  &
-%  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
-%   \end{bmatrix}.
-%\label{cov:eq}
-%\end{equation}
-%
-%
-%The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
-%\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
-%The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
-%is $\sigma$. 
-%The integrand in the Fisher-element calculation
-%\begin{equation}
-%I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
-%\frac{\partial C}{\partial  (f\sigma_8)} \right]
-%\end{equation}
-%represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
-%differential Fourier
-%volume element.  
-%
-%%
-%%\begin{figure}
-%%\centering
-%%\includegraphics[width=0.5\textwidth]{src/new.png}
-%%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-%%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
-%%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
-%%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
-%%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
-%%redshift depths $z_{max}$.
-%%\label{shot:fig}}
-%%\end{figure}
-%
-%
-%\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
-%First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
-%
-%The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
-%Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
-%Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
-%is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
-%Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
-%The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
-%\subsection{Survey Parameters that Determine Scientific Performance}
-
-%The survey parameters that  enter the Fisher matrix are the volume $V$, which
-%in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
-%magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
-
-%
-%
-%\begin{itemize}
-%\item Solid Angle $\Omega$:
-%The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
-%LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
-%the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-%coverage of the southern equatorial pole, could significantly increase
-%the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
-%\item Redshift Depth $z_{max}$:
-%Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is it worth doing so at the expense of neglecting lower redshift
-%SNe? This is answered in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
-%stronger probative power and benefit most from peculiar velocities.  Although
-%a differential increase in redshift depth increases the volume in Eq,~\ref{fisher:eqn}, the degradation in  intrinsic peculiar velocity $\propto z/(1+z)$
-%and the fact that shot noise is still important leads to the conclusion that a low-redshift tracer is more valuable than one at higher redshift.
-%
-%%For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
-%%limited.  We have run simulations of LSST SN surveys
-%%and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
-%%of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
-%%It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
-%%
-%%\begin{figure}
-%%\centering
-%%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
-%%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
-%%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
-%%realizations made in simulating the surveys.
-%%For several limiting redshift depths, the STON is plotted as a
-%%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
-%%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
-%%\label{notshot:fig}}
-%%\end{figure}
-%%
-%%The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
-%%$z_{max}$.  This is shown in the
-%%right plot of   Figure~\ref{notshot:fig}.
-%%
-%
-%
-%\item Intrinsic Magnitude Dispersion $\sigma_M$:
-%Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
-%A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
-%%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
-%%limited so the $\sigma^2_M/n$-term
-%%in the Fisher matrix is important.
-%Survey follow-up strategy,
-%which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
-%This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
-% in number density, or equivalently in survey duration. 
-%\end{itemize}
-
-\subsection{Designing an SN~Ia Peculiar Velocity Survey}
 There are different options to consider in the design of a SN~Ia peculiar velocity survey, though at this point it is difficult to assess which is optimal
@@
 available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
-\item Role of LSST:
+\item Role of the rolling search (e.g.\ LSST):
 LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.
@@
 For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
- \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.
+ \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.  The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
+ in number density, or equivalently in survey duration.
 
@@
 
-Modeling $k_{max}$ is a high as possible.
-
-\section{Conclusions}
 SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.

### 2019-01-26 doc/src/partials.py

@@ -270,11 +270,11 @@ def kintegral(z,zmax,ng,duration,sigm):
     blsigM = numpy.trapz(matter[:,0][w]**2*blsigM[w],matter[:,0][w])
     ll_vonly = numpy.trapz(matter[:,0][w]**2*ll_vonly[w],matter[:,0][w])
 
-    llkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*ll)
-    bbkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*bb)
-    blkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*bl)
+    # llkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*ll)
+    # bbkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*bb)
+    # blkmax = numpy.interp(kmax,matter[:,0], matter[:,0]**2*bl)
 
-    return ll,bb,bl, lls,bbs,bls,ll_ind,bb_ind,bl_ind, llsigM,bbsigM,blsigM,ll_vonly, llkmax,bbkmax,blkmax
+    return ll,bb,bl, lls,bbs,bls,ll_ind,bb_ind,bl_ind, llsigM,bbsigM,blsigM,ll_vonly
 
 # utility numbers
 zmax_zint = 0.3
@@ -305,12 +305,12 @@ def zintegral(zmax,ng,duration,sigm):
     bbsigM=numpy.zeros(len(rs))
     blsigM=numpy.zeros(len(rs))
     ll_vonly=numpy.zeros(len(rs))
-    llkmax=numpy.zeros(len(rs))
-    bbkmax=numpy.zeros(len(rs))
-    blkmax=numpy.zeros(len(rs))
+    # llkmax=numpy.zeros(len(rs))
+    # bbkmax=numpy.zeros(len(rs))
+    # blkmax=numpy.zeros(len(rs))
 
     for i in range(len(rs)):
-        ll[i],bb[i],bl[i],lls[i],bbs[i],bls[i],ll_ind[i],bb_ind[i],bl_ind[i],llsigM[i],bbsigM[i],blsigM[i],ll_vonly[i],llkmax[i],bbkmax[i],blkmax[i] = kintegral(zs[i],zmax,ng,duration,sigm)
[…]

### 2019-01-27 doc/article.tex
forging on almost done

@@
 
-\section{}
+\section{Introduction}
 
@@
 $0<z<0.2$ most of the signal comes from peculiar velocities and at higher redshifts $RSD$ is more constraining.
+
+\section{Testing Gravity with Peculiar Velocity Surveys}
 
@@
  \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory and
- are Gaussian-distributed.  Changes in this scale engender changes in the $\gamma$ precision, as    $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$
-and is survey-independent.
+ are Gaussian-distributed.  Changes in this scale engender changes in the $\gamma$ precision as the survey-independent
+  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$.
  \end{description}
@@
 
-There are different options to consider in the design of a SN~Ia peculiar velocity survey, though at this point it is difficult to assess which is optimal
-or most cost effective.  While there are a variety
-of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
-there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
-Here we  make several general conclusions
-based upon the above findings, while noting that 
-constructing such a model is a parallel
-endeavor that will benefit peculiar-velocity  and other SN~Ia science.
-
-%Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
-%Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.
-
+\section{Components and Considerations of a Peculiar Velocity Survey}
+In the above we have established how survey performance depends on the parameters $\Omega$, $z_{max}$, $t$, and $\sigma_M$.
+We now touch on how survey choices affect $\Omega$, $z_{max}$, and $\sigma_M$. 
 \begin{itemize}
-\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
-The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
-a pure SN~Ia sample obtained through
-spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
-The feasibility of using photometric classification is an important subject of research
-that can have implications for survey planning, particularly in extending the redshift range.
+\item Spectroscopic transient classification:  SNe~Ia are defined based on spectroscopic features,
+through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
+Maintaining a low sample intrinsic magnitude dispersion $\sigma_M$ relies on a the survey having spectroscopic classification to provide a pure sample
+(e.g.\ SNIFS, SED Machine; the surface density of active transients is low making MOS unnecessary).
+Photometric classification is an alternative survey choice, and the amount of sample dispersion that can be achieved photometrically is
+an active subject of  research.  Note that photometric classification benefits from having spectroscopic redshifts.
 \item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
-to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
-available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
-\item Role of the rolling search (e.g.\ LSST):
-LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.
+to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  These redshifts could come as a
+natural byproduct of the spectroscopic transient classification if the spectrograph has sufficient resolution.  Alternatively, 
+bright-galaxy moderate-resolution redshift surveys (e.g.\ DESI, 4MOST) can also provide multiplexed observations of 
+a significant fraction of nearby SN hosts before or well after SN discovery.
+\item Rolling imaging surveys: SNe~Ia must be discovered and their distances determined.  Wide-field cadenced imaging surveys (e.g.\ ZTF, LSST) can deliver
+both, though there is a question whether the survey light curves are sufficient to derive precision distances.  For example, 
 LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
 The nominal cadence  
- produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
-The wide-field of LSST is
- well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
-$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
-The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
-reach a targeted $\sigma_M$.
-
-The baseline
-LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
-the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-coverage of the southern equatorial pole could significantly increase
-the sky coverage with a corresponding decrease in the variance of $\gamma$.
-
+ produces sparse per-band light curves with limited ability to constrain distances.
+The imaging survey has a trade between  the solid angle ($\Omega$) it monitors and  the light-light quality ($\sigma_M$) it generates, whether it should serve
+as a SN~Ia discovery and distance machine, or a discovery machine only and relying on a complementary suite of instruments to get distances.
+% , which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
+%The wide-field of LSST is
+% well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
+%$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
+%The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
+%reach a targeted $\sigma_M$.
+%
+%The baseline
+%LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
+%the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
+%coverage of the southern equatorial pole could significantly increase
+%the sky coverage with a corresponding decrease in the variance of $\gamma$.
 \item Follow-up resources to determine per-SN distances:
@@
  in number density, or equivalently in survey duration. 
-
-Scientific leverage dictates that low-redshift sources are most valuable.
-This aligns with observational considerations, for which closer and hence brighter objects require more modest follow-up resources.
-Therefore, $z_{max}$ depends on the follow-up program as the redshift where complete follow-up saturates resources.
-
-A follow-up survey with $z_{max}=0.2$ would have $\sim 150,000$ targets over 10 years.  These sources will be observed when
-brighter than $r \sim 20.5$~mag and so will be accessible to 2--4m-class telescopes.  
+ \item Telescope aperture--Exposure Time:  Light-collection capabilities depend on the fluxes of the targets.
+ Closer and hence brighter objects require more modest telescope apertures and/or exposure times.  For our science case,
+ low-redshift sources are most valuable.  $z_{max}$ occurs when complete target follow-up saturates resources.
  \end{itemize}
 
-A low-redshift supernova survey is insensitive to two important uncertainties
+We note that a low-redshift supernova survey is insensitive to two important uncertainties
 associated with the SN~Ia Hubble diagram. The measurement is relatively insensitive to absolute color calibration, due to the limited
@@
 
+To give an idea of scale, a follow-up survey with $z_{max}=0.2$ would have $\sim 150,000$ targets over 10 years.  These sources will be observed when
+brighter than $r \sim 20.5$~mag.
+
+%While there are a variety
+%of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
+%there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
+%Constructing such a model is a parallel
+%endeavor that will benefit peculiar-velocity  and other SN~Ia science.
+
+\section{Conclusions}
+
 SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
 high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
-of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ significantly better
-than galaxy surveys, at lower redshifts that provide better leverage to test gravity models.
+of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $\gamma$ significantly better
+than galaxy surveys.
 While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
-their probative power.  The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
+their probative power.
+
+The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
 access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
-are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
-$f\sigma_8$ at $z\sim 0.1$.
+are of the highest interest,
 
-An important survey parameter that is not considered is cost.
+%so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
+%$f\sigma_8$ at $z\sim 0.1$.
 
-Effect of 10-years.
-
+\begin{comment}
 \newpage
@@
  \end{align}
+ \end{comment}
 %

### 2019-01-27 doc/article.tex
version for release

@@
 \begin{description}
-\item[Title]Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Survey
+\item[Title]Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys
 \item[Thematic Science Area] Cosmology and Fundamental Physics
@@
 \begin{abstract}
-Cadenced wide-field imaging surveys 
-in the upcoming decade will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands.  The increase in the number density of SNe~Ia, in parallel with improvements in the standardization of
+In the upcoming decade cadenced wide-field imaging surveys 
+will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+hundreds of thousands.  The increase in the number density  and solid-angle coverage 
+of SNe~Ia, in parallel with improvements in the standardization of
 their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
@@
 In the next decade the peculiar velocities of
-SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ 
-down to $0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
+SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ with as low as
+$0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
 \end{abstract}
@@
 is dominated by a ``dark energy'' that is gravitationally repulsive, or General Relativity is inadequate and needs to be replaced by a modified theory of
-gravity.
-It is only appropriate that in the upcoming decade, with their sheer numbers and improved distance precisions, SNe~Ia will provide measurements of the {\it inhomogeneous} motions of structures in the Universe
+gravity.  It is only appropriate that in the upcoming decade, with their sheer numbers, solid-angle coverage,
+and improved distance precisions, SNe~Ia will provide measurements of the {\it inhomogeneous} motions of structures in the Universe
 that will provide an unmatched test of whether dark energy or modified gravity is responsible for the accelerating expansion of the Universe.
@@
 }
-The $\Lambda$CDM prediction for the $z=0$ velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
+The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
 \citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
@@
 $f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization to model gravity, peculiar velocity
-surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose $\Lambda$CDM prediction is plotted 
+surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
 in Figure~2 of  \citet{1475-7516-2013-04-031}.
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Volume-weighted velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
-Volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$, expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions.
+\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
+Overplotted are volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$ expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions.
 The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
@@
  \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane method to
-get a 15\% uncertainty in $fD$ at $z\approx 0$ using density and velocity cross-correlations.
+get a 15\% uncertainty in $fD$ at $z\approx 0$ using combined density and velocity information.
 The upcoming 
@@
 and the WALLABY+WNSHS surveys will obtain Tully-Fisher galaxies at densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
-$z=0-0.1$.
-Covering 75\% of the sky to a maximum depth of $z=0.1$, these surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
-(For reference, DESI projects 10\% precision of $fD$ at $z>0.3$ redshifts  using RSD alone.)
+$z=0-0.1$ covering 75\% of the sky.
+These surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
+(For reference, DESI projects 10\% precision of $fD$ at $z>0.3$ redshifts  using Redshift Space Distortions (RSD) alone.)
 Existing SN~Ia samples
@@
 dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
-Recent studies indicate that with the right data, SNe absolute
+Recent studies indicate that with the right data, SN absolute
 magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
@@
   will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands, over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
+hundreds of thousands; over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
  for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$.
@@
 
-Given these two advances, over the course of a decade an LSST SN survey can produce interesting $fD$ measurements in several
-redshift bins.   
-\citet{2017ApJ...847..128H} find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, where at
-$0<z<0.2$ most of the signal comes from peculiar velocities and at higher redshifts $RSD$ is more constraining.
+Given these two advances, supernovae discovered by wide-field searches in the next decade will measure the growth of structure.
+For example, 
+over the course of a decade an LSST SN survey can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, where at
+$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from $RSD$  \citep{2017ApJ...847..128H}.
 
@@
 \propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
-and the line of sight.  While the $bD$ term does contain information on $\gamma$, it is not used for making constraints.
+and the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
 The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
 Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
-but are not considered here.
+but are not considered.
 
@@
 through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$.
-An analysis in which the RSD and velocity surveys are treated independently is achieved by
-setting the off-diagonal elements of $C$ to zero.
+An analysis in which the RSD and velocity surveys are treated independently 
+sets the off-diagonal elements of $C$ to zero.
 
@@
 
-All the surveys considered can provide meaning tests of gravity.
+All the surveys considered  provide meaningful tests of gravity.
 The projected uncertainty in $\gamma$ achieved by the suite of surveys are shown in Figure~\ref{var:fig}.
-For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities, one that uses
-both overdensities (RSD) and peculiar velocities independently, and one  that uses
-both RSD, peculiar velocities, and their cross-correlation.
+The primary analysis of interest uses overdensities (RSD), peculiar velocities, and their cross-correlations.
 The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
@@
 The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
-The figure also reveals an interesting interplay between RSD and velocity measurements.  Peculiar velocities alone
+Figure~\ref{var:fig} also shows uncertainties based on two other analyses, one that only uses peculiar velocities, and
+one that uses both RSD and peculiar velocities independently.
+Peculiar velocities alone
 can account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
-However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  factors of $\sim 0.8$.
-There are important
+However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  $\sim 20$\%.
+The implication is that there are important
 $k$-modes that are  sample variance limited either in  overdensity and/or peculiar velocity who benefit from the
-noise suppression engendered by cross-correlations.
+sample-noise suppression engendered by cross-correlations.
 
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{The projected uncertainty in $\gamma$ achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
+\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
 For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
-both overdensities (RSD) and peculiar velocities independently (dotted), and one  that uses
+both RSD and peculiar velocities independently (dotted), and one  that uses
 both RSD, peculiar velocities, and their cross-correlation (solid).
@@
 a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
-This is demonstrated to be the case at the end of both 2- and 10-year surveys and $z_{max}=0.2$.
+This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{max}=0.2$.
 The top panel of Figure~\ref{dvar:fig} shows
@@
 \caption{Top: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
-Bottom Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$; Bottom Right: $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$each as a function of
+Bottom Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$; Bottom Right: $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ each as a function of
  $z_{max}$ for two- and ten-year surveys.
@@
  \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] 
-Increasing the survey duration  increases the number density of discovered supernovae. As
-shot-noise is relevant for all the surveys considered, the increased survey duration improves the
-precision in $\gamma$. The surveys that benefit most are those more susceptible to shot noise:
-higher redshifts  due to their higher velocity dispersion and shorter surveys  due to their lower number 
+Increasing the survey duration  increases the number density of discovered supernovae. 
+The increased survey duration improves the
+precision in $\gamma$ as
+shot-noise is relevant for all the surveys considered. The surveys with the highest fractional benefit are those that are more susceptible to shot noise:
+those that monitor higher redshifts that have higher velocity dispersion, and shorter surveys  due to their lower number 
 densities.
@@
 This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ shown in the bottom-right panel of Figure~\ref{dvar:fig}.
- \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory and
- are Gaussian-distributed.  Changes in this scale engender changes in the $\gamma$ precision as the survey-independent
-  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$.
+ \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
+Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$,
+ which  is survey-independent.
  \end{description}
@@
 From the above calculations, we find empirically that shot noise is important
-but that sample variance is non-negligible at the end of longer surveys or when the intrinsic magnitude dispersion is small.
+but that sample variance is non-negligible toward the end of longer surveys or when the intrinsic magnitude dispersion is small.
 Fine tuning
@@
 After a few years the RSD measurement is sample variance limited,
-the  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$  
+the  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.
 For peculiar velocities,
@@
 In the above we have established how survey performance depends on the parameters $\Omega$, $z_{max}$, $t$, and $\sigma_M$.
-We now touch on how survey choices affect $\Omega$, $z_{max}$, and $\sigma_M$. 
+We now touch on how survey elements affect $\Omega$, $z_{max}$, and $\sigma_M$. 
 \begin{itemize}
@@
 Maintaining a low sample intrinsic magnitude dispersion $\sigma_M$ relies on a the survey having spectroscopic classification to provide a pure sample
-(e.g.\ SNIFS, SED Machine; the surface density of active transients is low making MOS unnecessary).
+(e.g.\ SNIFS, SED Machine; the low surface density of active low-$z$ transients makes MOS superfluous).
 Photometric classification is an alternative survey choice, and the amount of sample dispersion that can be achieved photometrically is
-an active subject of  research.  Note that photometric classification benefits from having spectroscopic redshifts.
+an active subject of  research.  Photometric classification benefits from having spectroscopic redshifts.
 \item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
@@
 natural byproduct of the spectroscopic transient classification if the spectrograph has sufficient resolution.  Alternatively, 
-bright-galaxy moderate-resolution redshift surveys (e.g.\ DESI, 4MOST) can also provide multiplexed observations of 
-a significant fraction of nearby SN hosts before or well after SN discovery.
+moderate-resolution (bright-)galaxy  redshift surveys (e.g.\ DESI, 4MOST) can  provide multiplexed observations of 
+a significant fraction of nearby SN hosts before  discovery or after transient light has faded away.
 \item Rolling imaging surveys: SNe~Ia must be discovered and their distances determined.  Wide-field cadenced imaging surveys (e.g.\ ZTF, LSST) can deliver
-both, though there is a question whether the survey light curves are sufficient to derive precision distances.  For example, 
+both, though there is a question as to distances will be derived solely from the light curves they generate.
+For example, 
 LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
@@
  produces sparse per-band light curves with limited ability to constrain distances.
-The imaging survey has a trade between  the solid angle ($\Omega$) it monitors and  the light-light quality ($\sigma_M$) it generates, whether it should serve
-as a SN~Ia discovery and distance machine, or a discovery machine only and relying on a complementary suite of instruments to get distances.
+The observing strategy of the imaging survey thus has a trade between  the solid angle ($\Omega$) monitored and  the light-light quality ($\sigma_M$) generated.
+The wide-field imager may better serve as a discovery machine leaving the distance determinations to other single-object follow-up instruments.
 % , which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
@@
  in number density, or equivalently in survey duration. 
- \item Telescope aperture--Exposure Time:  Light-collection capabilities depend on the fluxes of the targets.
+ \item Telescope Aperture $\times$ Exposure Time:  Light-collection capabilities depend on the fluxes of the targets.
  Closer and hence brighter objects require more modest telescope apertures and/or exposure times.  For our science case,
- low-redshift sources are most valuable.  $z_{max}$ occurs when complete target follow-up saturates resources.
+ low-redshift sources are most valuable and  a finite amount of follow-up should be spent on saturating redshifts lower than $z_{max}$.
  \end{itemize}
 
-We note that a low-redshift supernova survey is insensitive to two important uncertainties
-associated with the SN~Ia Hubble diagram. The measurement is relatively insensitive to absolute color calibration, due to the limited
+Focusing on low-redshift supernovae  mitigates  two important uncertainties
+associated with the high-$z$ SN~Ia cosmology. The measurement is relatively insensitive to absolute color calibration, due to the limited
 range of {\it observer} wavelengths used to standardize restframe distance moduli.
-Depending on which cameras are used to derive distances, instrumental (rather than absolute)
-calibration may be sufficient.  Progenitor population  is not expected to evolve significantly in the $0.01<z<0.2$ Universe.
-
-To give an idea of scale, a follow-up survey with $z_{max}=0.2$ would have $\sim 150,000$ targets over 10 years.  These sources will be observed when
-brighter than $r \sim 20.5$~mag.
+Depending on which cameras are used to derive distances, instrumental, as opposed to absolute,
+calibration may be sufficient.  The progenitor population  is not expected to evolve significantly in the $0.01<z<0.2$ Universe.
 
@@
 their probative power.  
-
 The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
 access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
-are of the highest interest,
+are of the highest interest.
+As a point of reference, a practically complete (most fields have 
+disruptive temporal gaps) follow-up survey with $z_{max}=0.2$ would have $\sim 150,000$, $r \lesssim 20.5$~mag targets over 10 years.
 

### 2019-01-28 doc/article.tex
add anze comments

@@
 Lyon, France} 
+\author{A.~Slo\u{z}ar}
+\affiliation{Brookhaven National Laboratory, Physics Department, Upton, NY
+11973, USA} 
 \author{others}
@@
 surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
-in Figure~2 of  \citet{1475-7516-2013-04-031}.
+in Figure~2 of  \citet{1475-7516-2013-04-031}.\footnote{The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
+commonly used in place of $D$ to normalize the
+overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.}
 
@@
 have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
-\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
+\citep{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
 %Measurement of the velocity field using LSST-discovered

### 2019-01-28 doc/article.tex
Eric's comments

@@
 \affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+\author{H.~Courtois}
+\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
+1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+Lyon, France} 
 \author{S.~Gontcho A Gontcho}
@@
 These surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
-(For reference, DESI projects 10\% precision of $fD$ at $z>0.3$ redshifts  using Redshift Space Distortions (RSD) alone.)
+(For reference, DESI projects 10\% precision of $fD$ at $z \approx 0.3$  using Redshift Space Distortions (RSD) alone.)
 Existing SN~Ia samples
@@
 For example, 
-over the course of a decade an LSST SN survey can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, where at
+over the course of a decade an LSST SN survey can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
+2.2\% uncertainty assuming redshift-independent $fD$, where at
 $0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from $RSD$  \citep{2017ApJ...847..128H}.
@@
 \section{Testing Gravity with Peculiar Velocity Surveys}
-
-Our scientific interest is in probing gravity so we here focus  on $\gamma$ rather than on $fD$.
+While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest here is in probing gravity
+so we here focus  on $\gamma$.
 To illustrate the distinction,
- $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega + \int \Omega^{\gamma} \ln{\Omega}\, d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively
+ $\frac{d(\ln{fD})}{d\gamma} = \ln\Omega_M + \int \Omega_M^{\gamma} \ln{\Omega_M}\, d\ln{a} \approx -1.68, -0.75,-0.37$ at $z=0,0.5,1.0$ respectively
  in $\Lambda$CDM;  two surveys with the same fractional precision  in $fD$ will have different precision in $\gamma$,
@@
 
-The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth $r_{max}$ (or $z_{max}$), duration
+The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth $r_{max}= \chi(z_{max})$, duration
 $t$ through $n=\epsilon rt$ where $r$ is the observer-frame SN~Ia rate
@@
 The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
-which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $2\sigma$ level.
+which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $>3\sigma$ level.
 The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
@@
 \item[Solid Angle $\Omega$]
-The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \sqrt{\Omega}$.
+The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
 \item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
@@
  produces sparse per-band light curves with limited ability to constrain distances.
-The observing strategy of the imaging survey thus has a trade between  the solid angle ($\Omega$) monitored and  the light-light quality ($\sigma_M$) generated.
+The observing strategy of the imaging survey thus has a trade between  the solid angle ($\Omega$) monitored and  the light-curve quality ($\sigma_M$) generated.
 The wide-field imager may better serve as a discovery machine leaving the distance determinations to other single-object follow-up instruments.
@@
  Closer and hence brighter objects require more modest telescope apertures and/or exposure times.  For our science case,
- low-redshift sources are most valuable and  a finite amount of follow-up should be spent on saturating redshifts lower than $z_{max}$.
+ low-redshift sources are most valuable and the finite amount of follow-up should be spent on saturating redshifts lower than $z_{max}$.
  \end{itemize}

### 2019-01-28 doc/article.tex
update author list
Fix k_{min}

@@
 Lyon, France} 
+\author{T.~Davis}
+\affiliation{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
 \author{S.~Gontcho A Gontcho}
@@
 over the course of a decade an LSST SN survey can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
-2.2\% uncertainty assuming redshift-independent $fD$, where at
+2.2\% uncertainty on fD  within this interval , where at
 $0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from $RSD$  \citep{2017ApJ...847..128H}.
@@
 The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth $r_{max}= \chi(z_{max})$, duration
-$t$ through $n=\epsilon rt$ where $r$ is the observer-frame SN~Ia rate
+$t$ through $n=\epsilon \phi t$ where $\phi$ is the observer-frame SN~Ia rate
 and $\epsilon$ is the sample-selection efficiency, and
@@
 $\Omega=3\pi$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
-The $k$-limits are  taken to be $k_{min} = (2 r_{max})^{-1}$  and 
+The $k$-limits are  taken to be $k_{min} = \pi/r_{max}$  and 
 $k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
@@
 After a few years the RSD measurement is sample variance limited,
-the  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.
+the  rest-frame volumetric SN~Ia rate is $\phi=7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.
 For peculiar velocities,
@@
 
-\begin{comment}
 \newpage
@@
  \end{align}
- \end{comment}
+
 %

### 2019-01-28 doc/src/cov.py

@@ -144,52 +144,52 @@ def zintegral(zmax,sigm, n,ng,independent=False):
     return numpy.trapz(zs**2*arr0,zs),numpy.trapz(zs**2*arr1,zs),numpy.trapz(zs**2*arr2,zs),numpy.trapz(zs**2*arr3,zs)
 
 
-zmaxs = numpy.arange(0.05,0.301,0.025)
-# zmaxs = numpy.log(numpy.arange(1,10.01,2))*0.25/numpy.log(10)+0.05
-noiseless = []
-twoyear = []
-tenyear = []
-twoyearind = []
-tenyearind = []
-independent = []
-for zmax in zmaxs:
-    zi = zintegral(zmax,0, tenyeardensity,ng)
-    F = numpy.array([[zi[0],zi[1]],[zi[2],zi[3]]])
-    noiseless.append(numpy.linalg.inv(F)[0,0])
-    # zi = zintegral(zmax,0, tenyeardensity, pgg,pvg,pvv,pgg_l,pvg_l,pvv_l,pgg_b,pvg_b,ng,independent=True)
-    # F = numpy.array([[zi[0],zi[1]],[zi[2],zi[3]]])
-    # independent.append(numpy.linalg.inv(F)[0,0])        
-    zi = zintegral(zmax,sigm, twoyeardensity,ng)
-    F = numpy.array([[zi[0],zi[1]],[zi[2],zi[3]]])
-    twoyear.append(numpy.linalg.inv(F)[0,0])
-    zi = zintegral(zmax,sigm, tenyeardensity,ng)
-    F = numpy.array([[zi[0],zi[1]],[zi[2],zi[3]]])
-    tenyear.append(numpy.linalg.inv(F)[0,0])
-    zi = zintegral(zmax,sigm, twoyeardensity,ng,independent=True)
-    F = numpy.array([[zi[0],zi[1]],[zi[2],zi[3]]])
-    twoyearind.append(numpy.linalg.inv(F)[0,0])
-    zi = zintegral(zmax,sigm, tenyeardensity,ng,independent=True)
-    F = numpy.array([[zi[0],zi[1]],[zi[2],zi[3]]])
-    tenyearind.append(numpy.linalg.inv(F)[0,0])
+# zmaxs = numpy.arange(0.05,0.301,0.025)
[…]

### 2019-01-28 doc/src/partials.py

@@ -251,7 +251,7 @@ def muintegral(z,ng,duration,sigm):
 
 
 def kintegral(z,zmax,ng,duration,sigm):
-    kmin = 1./(zmax*3e3*2)
+    kmin = numpy.pi/(zmax*3e3)
     kmax = 0.1
     w = numpy.logical_and(matter[:,0] >= kmin, matter[:,0]< kmax)
 
@@ -391,7 +391,7 @@ def set2():
             for z,r in zip(zs,rs):
                 a=1./(1+z)
                 drdz = 1/numpy.sqrt(OmegaM0/a**3 + (1-OmegaM0)) # 1/H
-                _,_,_, f00s,f11s,f10s,_,_,_,_,_,_,_,_,_,_ = kintegral(z,zmax,ng,duration,sigm)
+                _,_,_, f00s,f11s,f10s,_,_,_,_,_,_,_ = kintegral(z,zmax,ng,duration,sigm)
                 dvardz.append(finvp(f00,f11,f10,f00s,f11s,f10s))
                 dvardz[-1] = dvardz[-1]/r**2
                 dvardz[-1] = dvardz[-1]*drdz  # z now has units of 100 km/s
@@ -406,7 +406,7 @@ def set2():
     plt.savefig('dvardz.png')
     plt.clf()
 
-# set2()
+set2()
 
 
 def set1():

### 2019-01-30 doc/article.tex
changes due to Dragan's comments

@@
 \usepackage{natbib,verbatim}
-\usepackage{grffile}
-\usepackage{textcase}
+%\usepackage{grffile}
+%\usepackage{textcase}
 \usepackage{hyperref}
-\usepackage[overload]{textcase}
+%\usepackage[overload]{textcase}
 
@@
 \begin{document}
+%\raggedright
+\begin{raggedright}
+\huge
+Astro2020 Science White Paper \linebreak
 
-\thispagestyle{empty}
-\pagenumbering{gobble} 
-\begin{description}
-\item[Title]Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys
-\item[Thematic Science Area] Cosmology and Fundamental Physics
-\item[Authors] and a list of authors
-\item[Lead Author Contact Information]Alex Kim; Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720;
-1-510-486-4621;
-\href{mailto:agkim@lbl.gov}{agkim@lbl.gov}
-\end{description}
+Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys \linebreak
+\normalsize
 
-\clearpage
+\noindent \textbf{Thematic Areas:} \hspace*{60pt} $\square$ Planetary Systems \hspace*{10pt} $\square$ Star and Planet Formation \hspace*{20pt}\linebreak
+$\square$ Formation and Evolution of Compact Objects \hspace*{31pt} $\blacksquare$ Cosmology and Fundamental Physics \linebreak
+  $\square$  Stars and Stellar Evolution \hspace*{1pt} $\square$ Resolved Stellar Populations and their Environments \hspace*{40pt} \linebreak
+  $\square$    Galaxy Evolution   \hspace*{45pt} $\square$             Multi-Messenger Astronomy and Astrophysics \hspace*{65pt} \linebreak
+
+\textbf{Principal Author:}
+
+Name:	
+ Alex Kim\linebreak						
+Institution:  
+Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720 \linebreak
+Email:  \url{agkim@lbl.gov}
+ \linebreak
+Phone:  1-510-486-4621
+ \linebreak
+
+\textbf{Co-authors:} (names and institutions)
+  \linebreak
+
+\textbf{Abstract:}
+In the upcoming decade cadenced wide-field imaging surveys 
+will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+hundreds of thousands.  The increase in the number density  and solid-angle coverage 
+of SNe~Ia, in parallel with improvements in the standardization of
+their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
+is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
+In the next decade the peculiar velocities of
+SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ with as low as
+$0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
+\end{raggedright}
+\pagebreak
 \newpage
@@
 The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
-\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravities 
-follow the relation
-$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively.  Adopting this parameterization to model gravity, peculiar velocity
+\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravity follow the relation
+$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively \citep[see][for a review of
+these gravity models within a cosmological context]{HUTERER201523}.  Adopting this parameterization to model gravity, peculiar velocity
 surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
@@
 For example, 
-over the course of a decade an LSST SN survey can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
+over the course of a decade a SN survey relying on LSST discoveries 
+plus spectroscopic redshifts can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
 2.2\% uncertainty on fD  within this interval , where at
@@
  with the one at lower redshift providing the tighter constraint. 
+ In this section, we focus on how well peculiar velocity surveys in the upcoming decade can measure $\gamma$ and the sensitivity of the measurement
+ on survey parameter choices.
+
 We project uncertainties in $\gamma$, $\sigma_\gamma$ for a suite of idealized surveys
@@
 Figure~\ref{var:fig} also shows uncertainties based on two other analyses, one that only uses peculiar velocities, and
-one that uses both RSD and peculiar velocities independently.
+one that combines independent RSD and peculiar velocity results.
 Peculiar velocities alone
@@
 \centering
-\includegraphics[width=0.47\textwidth]{src/var.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
+\includegraphics[width=0.49\textwidth]{src/var.png}
 \caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
@@
 within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.
-\begin{description}
-\item[Solid Angle $\Omega$]
+
+{\bf Solid Angle $\Omega$:}
 The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
-\item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
+
+{\bf Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
 a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
 This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{max}=0.2$.
-The top panel of Figure~\ref{dvar:fig} shows
+The leftpanel of Figure~\ref{dvar:fig} shows
  $|\partial \sigma_\gamma / \partial z|$, which
@@
 \centering
-\includegraphics[width=0.4\textwidth]{src/dvardz.png}\\
-\includegraphics[width=0.4\textwidth]{src/dvardlnt.png}
-\includegraphics[width=0.4\textwidth]{src/dvardsigM.png}
-%\includegraphics[width=0.4\textwidth]{src/dvardkmax.png}
-%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
-%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Top: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
-Bottom Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$; Bottom Right: $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ each as a function of
- $z_{max}$ for two- and ten-year surveys.
+\includegraphics[width=0.49\textwidth]{src/dvardz.png}
+\includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
+\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
+Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
+ $z_{max}$ for two- (dashed) and ten-year (solid) surveys.
 \label{dvar:fig}}
 \end{figure}
-\item[Redshift Depth $z_{max}$]
+
+{\bf Redshift Depth $z_{max}$:}
 Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
 $z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
- \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] 
-Increasing the survey duration  increases the number density of discovered supernovae. 
-The increased survey duration improves the
-precision in $\gamma$ as
-shot-noise is relevant for all the surveys considered. The surveys with the highest fractional benefit are those that are more susceptible to shot noise:
-those that monitor higher redshifts that have higher velocity dispersion, and shorter surveys  due to their lower number 
-densities.
-These trends are shown in the bottom-left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
+
+{\bf Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
+An increased survey duration accumulates more supernovae, decreasing shot noise and  increasing the 
+precision in $\gamma$ for all the surveys considered.
+The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution,
+shorter surveys and those with higher $z_{max}$ (because of the higher $\sigma$), benefit more from extending survey duration.
+These trends are shown in the left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
  $z_{max}$ for two- and ten-year surveys.
@@
 has a similar relationship with $\sigma_\gamma$.
-This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ shown in the bottom-right panel of Figure~\ref{dvar:fig}.
- \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
+This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ also shown in the left panel of Figure~\ref{dvar:fig}.
+
+{\bf Minimum length scale, maximum wavenumber $k_{max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
 Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$,
  which  is survey-independent.
- \end{description}
 
-From the above calculations, we find empirically that shot noise is important
-but that sample variance is non-negligible toward the end of longer surveys or when the intrinsic magnitude dispersion is small.
-Fine tuning
-survey design is aided by understanding when surveys become sample variance limited.
-After a few years the RSD measurement is sample variance limited,
-the  rest-frame volumetric SN~Ia rate is $\phi=7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$.
+The condition of sample variance--shot noise equality of a $k$-mode is now examined.
+This is important in tuning survey design because, as mentioned above, the range of surveys of interest have different relative contributions of sample variance and shot noise.
+The RSD measurement is sample variance limited, as within a few years supernova discoveries from
+the  rest-frame volumetric rate of $\phi=7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$ make shot noise unimportant.
 For peculiar velocities,
@@
 
-\section{Components and Considerations of a Peculiar Velocity Survey}
-In the above we have established how survey performance depends on the parameters $\Omega$, $z_{max}$, $t$, and $\sigma_M$.
-We now touch on how survey elements affect $\Omega$, $z_{max}$, and $\sigma_M$. 
-\begin{itemize}
-\item Spectroscopic transient classification:  SNe~Ia are defined based on spectroscopic features,
+\section{Elements of a Peculiar Velocity Survey}
+A successful peculiar velocity survey must accomplish several things, not just measure distances and host-galaxy redshifts.  These elements are as follows:
+
+{\bf SNe Must Be Discovered:} Cadenced wide-field imaging surveys have proven to be efficient SN~Ia discovery machines.  ZTF and LSST
+are examples, but not the sole representatives, of current and future surveys that will generate high surface-densities of SNe~Ia $n$ over a large solid angle $\Omega$. 
+For example, 
+LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
+
+{\bf SNe~Ia Must Be Classified:}  Cadenced wide-field imaging surveys discover SNe~Ia together with a background of other transient events.  The peculiar velocity
+analysis is performed using a subset of all of these transients, with the intrinsic magnitude dispersion dependent on  sample selection criteria.  Tailored SN~Ia subsets
+and pure SN~Ia samples will have lower $\sigma_M$ compared to samples that suffer from non-Ia contamination.
+Spectroscopic classification can ensure the purity of the analysis sample.  SNe~Ia are defined based on spectroscopic features,
 through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
-Maintaining a low sample intrinsic magnitude dispersion $\sigma_M$ relies on a the survey having spectroscopic classification to provide a pure sample
-(e.g.\ SNIFS, SED Machine; the low surface density of active low-$z$ transients makes MOS superfluous).
-Photometric classification is an alternative survey choice, and the amount of sample dispersion that can be achieved photometrically is
-an active subject of  research.  Photometric classification benefits from having spectroscopic redshifts.
-\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
-to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  These redshifts could come as a
-natural byproduct of the spectroscopic transient classification if the spectrograph has sufficient resolution.  Alternatively, 
+SNIFS and SED Machine are examples of single-object spectrographs built with the purpose of classifying transients discovered in imaging surveys
+(the low surface density of active low-$z$ transients makes MOS superfluous).
+Photometric classification is an alternative approach,  which benefits from the light-curve data already in hand from the search.  The level of
+$\sigma_M$ that can be achieved, SN~Ia completeness, and the effect of classification-distance covariance  are subjects of
+active  research.  Photometric classification benefits from having spectroscopic redshifts.
+
+{\bf Host-galaxies Need Precise Redshifts:} Peculiar velocities come from the difference between host-galaxy and cosmological redshifts.  Host redshift uncertainties need to be
+ $<0.5$\% in order not to contribute significantly
+to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  These redshifts could come naturally
+from galaxy signal in  the spectroscopic transient classification if the spectrograph had sufficient resolution.  Alternatively, 
 moderate-resolution (bright-)galaxy  redshift surveys (e.g.\ DESI, 4MOST) can  provide multiplexed observations of 
 a significant fraction of nearby SN hosts before  discovery or after transient light has faded away.
-\item Rolling imaging surveys: SNe~Ia must be discovered and their distances determined.  Wide-field cadenced imaging surveys (e.g.\ ZTF, LSST) can deliver
-both, though there is a question as to distances will be derived solely from the light curves they generate.
-For example, 
-LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
-The nominal cadence  
- produces sparse per-band light curves with limited ability to constrain distances.
-The observing strategy of the imaging survey thus has a trade between  the solid angle ($\Omega$) monitored and  the light-curve quality ($\sigma_M$) generated.
-The wide-field imager may better serve as a discovery machine leaving the distance determinations to other single-object follow-up instruments.
-% , which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
-%The wide-field of LSST is
-% well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
-%$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
-%The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
-%reach a targeted $\sigma_M$.
-%
-%The baseline
-%LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage beyond
-%the LSST baseline benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
-%coverage of the southern equatorial pole could significantly increase
-%the sky coverage with a corresponding decrease in the variance of $\gamma$.
-\item Follow-up resources to determine per-SN distances:
-Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
-expanded (UV, NIR) wavelength coverage, and spectral features
-can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
-For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
+
+{\bf SNe~Ia Need Precise Distances:} Cosmological redshifts are inferred from SN~Ia distances. Distances are calculated from multi-band light curves, redshifts (when available),
+SN spectroscopic features, and
+host-galaxy properties.
+Distance uncertainties here have been characterized by $\sigma_M$,
+which includes both measurement uncertainty and intrinsic magnitude dispersion.  The latter depends on the data-quality and methodology used to calculate distances.
+For the same supernovae, the intrinsic distance dispersion derived from a dataset with sparse light curves, a small number of bands, limited wavelength coverage, and
+lacking spectral-feature measurements is larger compared to the dispersion derived from a broader dataset  with dense light curves, a large number of bands, 
+broad wavelength coverage, and with spectral-feature measurements.  Historically surveys with 2 or 3 well-sampled optical light curves obtained $\sigma_M \approx 0.15$~mag,
+with dense, $griz$-band optical light curves DES \citep{2018arXiv181102377B} gets $\sigma_M\approx 0.10$~mag, 
+standardization studies show that supplemental
+infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
  \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag. The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
  in number density, or equivalently in survey duration.
- \item Telescope Aperture $\times$ Exposure Time:  Light-collection capabilities depend on the fluxes of the targets.
- Closer and hence brighter objects require more modest telescope apertures and/or exposure times.  For our science case,
- low-redshift sources are most valuable and the finite amount of follow-up should be spent on saturating redshifts lower than $z_{max}$.
- \end{itemize}
 
+As a point of reference, a practically complete (accounting for temporal observing gaps) discovery up to $z_{max}=0.2$ would have $\sim 150,000$, $r \lesssim 20.5$~mag targets over 10 years.  
 Focusing on low-redshift supernovae  mitigates  two important uncertainties
@@
 
-%While there are a variety
-%of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
-%there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
-%Constructing such a model is a parallel
-%endeavor that will benefit peculiar-velocity  and other SN~Ia science.
-
 \section{Conclusions}
 
-SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
-high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
-of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $\gamma$ significantly better
-than galaxy surveys.
-While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
-their probative power.  
-The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
-access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
-are of the highest interest.
-As a point of reference, a practically complete (most fields have 
-disruptive temporal gaps) follow-up survey with $z_{max}=0.2$ would have $\sim 150,000$, $r \lesssim 20.5$~mag targets over 10 years.
+%SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  
+In the next decade,
+the high number of SN discoveries together with improved precision in their distance precisions will make $z<0.2$ SNe~Ia, more so than
+galaxies,  powerful probes of gravity through their effect  on the growth of structure. 
+A peculiar velocity survey must discover, classify, and get distances of SNe~Ia, as well as obtain redshifts of their host galaxies.
+Choices in survey design impact the precision obtained on $\gamma$.  SN~Ia peculiar velocity surveys are of the highest scientific
+interest and we encourage
+the community to develop aggressive surveys in the pursuit of testing  General Relativity and probing gravity.
 
-%so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
-%$f\sigma_8$ at $z\sim 0.1$.
-
+\begin{comment}
 \newpage
@@
  \end{align}
-
+\end{comment}
 %

### 2019-01-30 doc/src/cov.py

@@ -3,8 +3,8 @@
 import numpy
 import matplotlib.pyplot as plt
 import matplotlib
-
-
+matplotlib.rcParams['font.size'] = 14
+matplotlib.rcParams['lines.linewidth'] = 2.0
 matter = numpy.loadtxt("../../pv/dragan/matterpower.dat")
 
 # vv
@@ -281,7 +281,7 @@ plt.subplots_adjust(hspace=0.05)
 plt.savefig('pvv.png')
 
 plt.clf()
-fig, axs = plt.subplots(1,1,sharex=True,figsize=(8,5))
+fig, axs = plt.subplots(1,1,sharex=True)
 z=0.1
 restrate = 0.65*  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
 # sigm = 0.08
@@ -293,29 +293,29 @@ colors=['blue','green','red']
 zs = numpy.array([0.05,0.1,.2])
 axs.plot(matter[:,0],matter[:,0]**3*Pvv(1),label=r'$k^3P_{vv}(\mu=1)$',color='black')
 axs.plot(matter[:,0],matter[:,0]**3*Pvv(0.5),label=r'$k^3P_{vv}(\mu=0.5)$',color='cyan')
-axs.plot(matter[:,0],matter[:,0]**3*(0.08 * sigv_factor *3e5)**2 / (2*restrate/(1+z)), label=r'$k^3\sigma^2/n$ $\sigma_M=0.08$ 2 years $z=0.1$',ls='--',color='red')
-axs.plot(matter[:,0],matter[:,0]**3*(0.15 * sigv_factor *3e5)**2 / (2*restrate/(1+z)), label=r'$k^3\sigma^2/n$ $\sigma_M=0.15$ 2 years $z=0.1$',ls=':',color='red')
-axs.plot(matter[:,0],matter[:,0]**3*(0.08 * sigv_factor *3e5)**2 / (10*restrate/(1+z)), label=r'$k^3\sigma^2/n$ $\sigma_M=0.08$ 10 years $z=0.1$',ls='--',color='black')
-axs.plot(matter[:,0],matter[:,0]**3*(0.15 * sigv_factor *3e5)**2 / (10*restrate/(1+z)), label=r'$k^3\sigma^2/n$ $\sigma_M=0.15$ 10 years $z=0.1$',ls=':',color='black')
[…]

### 2019-01-30 doc/src/partials.py

@@ -5,6 +5,8 @@ import matplotlib.pyplot as plt
 import matplotlib
 import scipy.integrate as integrate
 from astropy.cosmology import FlatLambdaCDM
+matplotlib.rcParams['font.size'] = 14
+matplotlib.rcParams['lines.linewidth'] = 2.0
 
 # cosmology
 OmegaM0 = 0.28
@@ -180,7 +182,7 @@ def traces(z,mu,ng,duration,sigm):
         numpy.array(ll_vonly)
 
 def muintegral(z,ng,duration,sigm):
-    mus=numpy.arange(-1,1.001,0.1)
+    mus=numpy.arange(0,1.001,0.05)
 
     ll=numpy.zeros((len(mus),len(matter[:,0])))
     bb=numpy.zeros((len(mus),len(matter[:,0])))
@@ -247,7 +249,7 @@ def muintegral(z,ng,duration,sigm):
     # bb_ind=bb_ind+h/2
     # bl_ind=bl_ind+i/2
     
-    return ll,bb,bl, lls,bbs,bls,ll_ind,bb_ind,bl_ind, llsigM,bbsigM,blsigM,ll_vonly
+    return 2*ll,2*bb,2*bl, 2*lls,2*bbs,2*bls,2*ll_ind,2*bb_ind,2*bl_ind, 2*llsigM,2*bbsigM,2*blsigM,2*ll_vonly
 
 
 def kintegral(z,zmax,ng,duration,sigm):
@@ -369,7 +371,7 @@ def zintegral(zmax,ng,duration,sigm):
 
 
 def set2():
-    fig,(ax) = plt.subplots(1, 1, figsize=(8*.8,6*.8))
+    fig,(ax) = plt.subplots(1, 1)
     durations = [2,10]
     labels = ['Two Years','Ten Years']
     colors = ['red','black']
@@ -406,11 +408,11 @@ def set2():
     plt.savefig('dvardz.png')
     plt.clf()
 
-set2()
+# set2()
 
 
 def set1():
-    fig,(ax) = plt.subplots(1, 1, figsize=(8*.8,6*.8))
+    fig,(ax) = plt.subplots(1, 1)
     zmaxs = numpy.exp(numpy.arange(numpy.log(0.05),numpy.log(.300001),numpy.log(.3/.05)/8))
     durations = [2,10]
     labels = ['Two Years','Ten Years']
@@ -459,27 +461,49 @@ def set1():
[…]

### 2019-02-01 doc/article.tex
Helene, Dragan-2 comments in

@@
 Astro2020 Science White Paper \linebreak
-
+proj
 Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys \linebreak
@@
 Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
-First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields $\sigma_M \gtrsim 0.12$~mag absolute magnitude
-dispersion.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
+First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields  absolute magnitude
+dispersion $\sigma_M \gtrsim 0.12$~mag.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
 Recent studies indicate that with the right data, SN absolute
@@
 
-Given these two advances, supernovae discovered by wide-field searches in the next decade will measure the growth of structure.
+Given these two advances, supernovae discovered by wide-field searches in the next decade will 
+ be able to stringently constrain the growth of structure in the low-redshift Universe.
 For example, 
@@
 While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest here is in probing gravity
-so we here focus  on $\gamma$.
+so we here focus  on the growth index $\gamma$.
 To illustrate the distinction,
@@
 
-We project uncertainties in $\gamma$, $\sigma_\gamma$ for a suite of idealized surveys
-using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}.
+We project uncertainties  on the growth index, $\sigma_\gamma$ for a suite of idealized surveys
+using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}
+\citep[there is an alternative approach using an estimator for the mean pairwise velocity][]{2011PhRvD..83d3004B}.
 The Fisher information matrix is
@@
 
-The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth $r_{max}= \chi(z_{max})$, duration
+The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth
+given by the comoving distance out to the maximum redshift $r_{max}=r(z_{max})$, duration
 $t$ through $n=\epsilon \phi t$ where $\phi$ is the observer-frame SN~Ia rate
@@
 within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.
+
+The dependence on the uncertainty in the growth index on the survey parameters is as follows:
 
@@
 
-{\bf SNe~Ia Need Precise Distances:} Cosmological redshifts are inferred from SN~Ia distances. Distances are calculated from multi-band light curves, redshifts (when available),
+{\bf SNe~Ia Need Precise Distances:} SN~Ia distances, along with their measured redshifts, are required to infer the peculiar velocities of their hosts.
+Distances are calculated from multi-band light curves, redshifts (when available),
 SN spectroscopic features, and
 host-galaxy properties.
-Distance uncertainties here have been characterized by $\sigma_M$,
+Distance uncertainties here have been characterized by magnitude dispersion $\sigma_M$,
 which includes both measurement uncertainty and intrinsic magnitude dispersion.  The latter depends on the data-quality and methodology used to calculate distances.
@@
 lacking spectral-feature measurements is larger compared to the dispersion derived from a broader dataset  with dense light curves, a large number of bands, 
-broad wavelength coverage, and with spectral-feature measurements.  Historically surveys with 2 or 3 well-sampled optical light curves obtained $\sigma_M \approx 0.15$~mag,
+broad wavelength coverage, and with spectral-feature measurements.  Historically surveys with two or three well-sampled optical light curves obtained $\sigma_M \approx 0.15$~mag,
 with dense, $griz$-band optical light curves DES \citep{2018arXiv181102377B} gets $\sigma_M\approx 0.10$~mag, 
@@
 As a point of reference, a practically complete (accounting for temporal observing gaps) discovery up to $z_{max}=0.2$ would have $\sim 150,000$, $r \lesssim 20.5$~mag targets over 10 years.  
-Focusing on low-redshift supernovae  mitigates  two important uncertainties
+Focusing on low-redshift supernovae  mitigates  several important uncertainties
 associated with the high-$z$ SN~Ia cosmology. The measurement is relatively insensitive to absolute color calibration, due to the limited
@@
 A peculiar velocity survey must discover, classify, and get distances of SNe~Ia, as well as obtain redshifts of their host galaxies.
-Choices in survey design impact the precision obtained on $\gamma$.  SN~Ia peculiar velocity surveys are of the highest scientific
+Choices in survey design impact the precision obtained on the growth index $\gamma$.  
+No other probe of growth of structure or tracer of peculiar velocity can provide comparable precision on  $\gamma$ in the next decade.
+SN~Ia peculiar velocity surveys are of the highest scientific
 interest and we encourage
@@
 Each supernova is assigned a magnitude based on the distance modulus of its cosmological
-redshift plus a random term drawn from a Normal distribution $\sigma_M$, which for simplicity is the same for all supernovae and captures both
+redshift plus a random term drawn from a Normal distribution with standard deviation $\sigma_M$, which for simplicity is the same for all supernovae and captures both
 intrinsic magnitude dispersion and measurement uncertainty.  No other corrections are applied to the observed magnitude.  Dipole effects of the heliocentric

### 2019-02-01 doc/article.tex
fix stupid typo

@@
 Astro2020 Science White Paper \linebreak
-proj
 Testing Gravity Using Type Ia Supernovae Discovered by Next-Generation Wide-Field Imaging Surveys \linebreak

### 2019-02-07 doc/src/partials.py
for debugging relative to master

@@ -554,14 +554,14 @@ def set1():
             f00,f11,f10, f00s,f11s,f10s,f00_ind,f11_ind,f10_ind, f00sigM,f11sigM,f10sigM,f00_vonly,f02,f12,f22, \
                 f02s,f12s,f22s,f02_ind,f12_ind,f22_ind, f02sigM,f12sigM,f22sigM, f01_vonly, f11_vonly = zintegral(zmax,ng,duration,sigm_Ia,restrate_Ia)
             # dv_.append(finvp(f00,f11,f10,f00s,f11s,f10s))
-            print(finvp(f00,f11,f10,f00s,f11s,f10s))
             dv_.append(finvp3d(f00,f10,f02,f11,f12,f22+sigOM0sqinv, f00s,f10s,f02s,f11s,f12s,f22s))
-            print (finvp3d(f00,f10,f02,f11,f12,f22+sigOM0sqinv, f00s,f10s,f02s,f11s,f12s,f22s))
-            wefwe
             v_.append(numpy.linalg.inv(numpy.array([[f00,f10,f02],[f10,f11,f12],[f02,f12,f22+sigOM0sqinv]]))[0,0])
             vind_.append(numpy.linalg.inv(numpy.array([[f00_ind,f10_ind,f02_ind],[f10_ind,f11_ind,f12_ind],[f02_ind,f12_ind,f22_ind+sigOM0sqinv]]))[0,0])
             vvonly_.append(numpy.linalg.inv(numpy.array([[f00_vonly,f01_vonly],[f01_vonly,f11_vonly+sigOM0sqinv]]))[0,0])
             dvsigM_.append(finvp3d(f00,f10,f02,f11,f12,f22+sigOM0sqinv,f00sigM,f10sigM,f02sigM,f11sigM,f12sigM,f22sigM))
+            print(v_[-1],dv_[-1],dvsigM_[-1])
+
+            wefe
             # dvdkmax_.append(finvp(f00,f11,f10,f00kmax,f11kmax,f10kmax))
 
 

### 2019-02-07 doc/src/partials.py
debug

@@ -436,6 +436,10 @@ def set1():
             vind_.append(numpy.linalg.inv(numpy.array([[f00_ind,f10_ind],[f10_ind,f11_ind]]))[0,0])
             vvonly_.append(1./f00_vonly)
             dvsigM_.append(finvp(f00,f11,f10,f00sigM,f11sigM,f10sigM))
+            print(f00,f11,f10,f00s,f11s,f10s)
+            print(v_[-1],dv_[-1],dvsigM_[-1])
+
+            wefew
             # dvdkmax_.append(finvp(f00,f11,f10,f00kmax,f11kmax,f10kmax))
 
 

### 2019-02-07 doc/src/partials.py
debug

@@ -377,6 +377,8 @@ def kintegral(z,zmax,ng,duration,sigm,restrate):
     lO = numpy.trapz(matter[:,0][w]**2*lO[w],matter[:,0][w])
     bO = numpy.trapz(matter[:,0][w]**2*bO[w],matter[:,0][w])
     OO = numpy.trapz(matter[:,0][w]**2*OO[w],matter[:,0][w])
+    print(lls[w])
+    wefwe
     lls = numpy.trapz(matter[:,0][w]**2*lls[w],matter[:,0][w])
     bbs = numpy.trapz(matter[:,0][w]**2*bbs[w],matter[:,0][w])
     bls = numpy.trapz(matter[:,0][w]**2*bls[w],matter[:,0][w])
@@ -559,6 +561,7 @@ def set1():
             vind_.append(numpy.linalg.inv(numpy.array([[f00_ind,f10_ind,f02_ind],[f10_ind,f11_ind,f12_ind],[f02_ind,f12_ind,f22_ind+sigOM0sqinv]]))[0,0])
             vvonly_.append(numpy.linalg.inv(numpy.array([[f00_vonly,f01_vonly],[f01_vonly,f11_vonly+sigOM0sqinv]]))[0,0])
             dvsigM_.append(finvp3d(f00,f10,f02,f11,f12,f22+sigOM0sqinv,f00sigM,f10sigM,f02sigM,f11sigM,f12sigM,f22sigM))
+
             print(v_[-1],dv_[-1],dvsigM_[-1])
 
             wefe

### 2019-02-07 doc/src/partials.py
fix wrong off diag in nversematrix

@@ -30,12 +30,12 @@ sigm_Ia = 0.08
 ng = 1e-3
 
 def Cinverse(m00,m11,m01):
-    return numpy.array([[m11,m01**2],[m01**2,m00]])/(m00*m11 - m01**2)
+    return numpy.array([[m11,-m01],[-m01,m00]])/(m00*m11 - m01**2)
 
 def Cinverse_partial(m00,m11,m01,dm00,dm11,dm01):
     den = (m00*m11 - m01**2)
     return numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den \
-        - numpy.array([[m11,m01**2],[m01**2,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
+        - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
 
 def OmegaM(a,OmegaM0=OmegaM0):
     return OmegaM0/(OmegaM0 + (1-OmegaM0)*a**3)

### 2019-02-07 doc/src/partials.py
state for debugging

@@ -34,6 +34,10 @@ def Cinverse(m00,m11,m01):
 
 def Cinverse_partial(m00,m11,m01,dm00,dm11,dm01):
     den = (m00*m11 - m01**2)
+    print (numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den,  \
+        - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01) )
+
+    fwfe
     return numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den \
         - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
 
@@ -150,6 +154,9 @@ def finvp3d(f00,f01,f02,f11,f12,f22,f00p,f01p,f02p,f11p,f12p,f22p):
 
 
 def Cmatrices(z,mu,ng,duration,sigm,restrate):
+    duration = 10
+    mu=0.5
+
     a = 1/(1.+z)
     OmegaM_a = OmegaM(a)
     n = duration*restrate/(1+z)
@@ -377,8 +384,6 @@ def kintegral(z,zmax,ng,duration,sigm,restrate):
     lO = numpy.trapz(matter[:,0][w]**2*lO[w],matter[:,0][w])
     bO = numpy.trapz(matter[:,0][w]**2*bO[w],matter[:,0][w])
     OO = numpy.trapz(matter[:,0][w]**2*OO[w],matter[:,0][w])
-    print(lls[w])
-    wefwe
     lls = numpy.trapz(matter[:,0][w]**2*lls[w],matter[:,0][w])
     bbs = numpy.trapz(matter[:,0][w]**2*bbs[w],matter[:,0][w])
     bls = numpy.trapz(matter[:,0][w]**2*bls[w],matter[:,0][w])
@@ -495,7 +500,7 @@ def zintegral(zmax,ng,duration,sigm,restrate):
 
 def set2():
     fig,(ax) = plt.subplots(1, 1)
-    durations = [2,10]
+    durations = [2.,10.]
     labels = ['Two Years','Ten Years']
     colors = ['red','black']
 
@@ -561,10 +566,6 @@ def set1():
[…]

### 2019-02-07 doc/src/partials.py
fix bug in invers

@@ -34,10 +34,6 @@ def Cinverse(m00,m11,m01):
 
 def Cinverse_partial(m00,m11,m01,dm00,dm11,dm01):
     den = (m00*m11 - m01**2)
-    print (numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den,  \
-        - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01) )
-
-    fwfe
     return numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den \
         - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
 
@@ -122,7 +118,7 @@ def finvp3d(f00,f01,f02,f11,f12,f22,f00p,f01p,f02p,f11p,f12p,f22p):
         - f00p*f12**2 - 2*f00*f12*f12p \
         - f11p*f02**2 - 2*f11*f02*f02p \
         - f22p*f01**2 - 2* f22*f01*f01p \
-        + 2*(f01p*f02*f12 + f01*f02p*f12 + f01*f-2*f12p)
+        + 2*(f01p*f02*f12 + f01*f02p*f12 + f01*f02*f12p)
 
     # full matrix
     # mat = numpy.array([ \
@@ -154,9 +150,6 @@ def finvp3d(f00,f01,f02,f11,f12,f22,f00p,f01p,f02p,f11p,f12p,f22p):
 
 
 def Cmatrices(z,mu,ng,duration,sigm,restrate):
-    duration = 10
-    mu=0.5
-
     a = 1/(1.+z)
     OmegaM_a = OmegaM(a)
     n = duration*restrate/(1+z)
@@ -516,13 +509,15 @@ def set2():
 
         for duration,label,color in zip(durations,labels,colors):
             dvardz=[]
-            f00,f11,f10, _,_,_,_,_,_,_, _,_,_,f02,f12,f22 = zintegral(zmax,ng,duration,sigm_Ia,restrate_Ia)
+            f00,f11,f10, f00s,f11s,f10s,f00_ind,f11_ind,f10_ind, f00sigM,f11sigM,f10sigM,f00_vonly,f02,f12,f22, \
[…]

### 2019-02-07 doc/src/partials.py
fixed now!

@@ -515,7 +515,7 @@ def set2():
             for z,r in zip(zs,rs):
                 a=1./(1+z)
                 drdz = 1/numpy.sqrt(OmegaM0/a**3 + (1-OmegaM0)) # 1/H
-                f00,f11,f10, f00s,f11s,f10s,f00_ind,f11_ind,f10_ind, f00sigM,f11sigM,f10sigM,f00_vonly,f02,f12,f22, \
+                _,_,_, f00s,f11s,f10s,f00_ind,f11_ind,f10_ind, f00sigM,f11sigM,f10sigM,f00_vonly,f02,f12,f22, \
                     f02s,f12s,f22s,f02_ind,f12_ind,f22_ind, f02sigM,f12sigM,f22sigM, f01_vonly, f11_vonly= kintegral(z,zmax,ng,duration,sigm_Ia,restrate_Ia)
                 dvardz.append(finvp3d(f00,f10,f02,f11,f12,f22+sigOM0sqinv, f00s,f10s,f02s,f11s,f12s,f22s))
                 dvardz[-1] = dvardz[-1]/r**2
@@ -528,7 +528,7 @@ def set2():
     plt.ylabel(r'$\sigma_{\gamma}^{-1}|\frac{d\sigma_{\gamma}}{dz}|$')
     plt.legend()
     plt.tight_layout()
-    plt.savefig('/tmp/dvardz.png')
+    plt.savefig('dvardz.png')
     plt.clf()
 
 set2()
@@ -583,7 +583,7 @@ def set1():
     plt.ylabel(r'$\sigma_{\gamma}$')
     plt.legend()
     plt.tight_layout()
-    plt.savefig('/tmp/var.png')
+    plt.savefig('var.png')
 
     plt.clf()
     fig, ax1 = plt.subplots()
@@ -605,7 +605,7 @@ def set1():
     ax2.set_ylabel(r'$\sigma_{\gamma}^{-1}\frac{d\sigma_{\gamma}}{d\sigma_M}$')
     fig.tight_layout()
     plt.legend()
-    plt.savefig('/tmp/dvardxxx.png')
+    plt.savefig('dvardxxx.png')
     plt.clf()
 
 set1()

### 2019-02-07 doc/src/partials.py
merge from omegam0

@@ -12,6 +12,10 @@ matplotlib.rcParams['lines.linewidth'] = 2.0
 OmegaM0 = 0.28
 cosmo = FlatLambdaCDM(H0=100, Om0=OmegaM0)
 
+gamma=0.55
+
+sigOM0sqinv=1/0.005**2
+
 # power spectrum from CAMB
 matter = numpy.loadtxt("../../pv/dragan/matterpower.dat")
 
@@ -19,47 +23,61 @@ matter = numpy.loadtxt("../../pv/dragan/matterpower.dat")
 f=OmegaM0**0.55
 
 # SN properties
-restrate = 0.65*  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
-sigm = 0.08
+restrate_Ia = 0.65*  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
+sigm_Ia = 0.08
 
 # galaxy density (TAIPAN Howlett et al.)
 ng = 1e-3
 
+def Cinverse(m00,m11,m01):
+    return numpy.array([[m11,-m01],[-m01,m00]])/(m00*m11 - m01**2)
 
+def Cinverse_partial(m00,m11,m01,dm00,dm11,dm01):
+    den = (m00*m11 - m01**2)
+    return numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den \
+        - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
 
-def OmegaM(a):
-    return OmegaM0/(OmegaM0 + 0.72*a**3)
+def OmegaM(a,OmegaM0=OmegaM0):
+    return OmegaM0/(OmegaM0 + (1-OmegaM0)*a**3)
 
-def integrand_dlnfs(a):
-    return numpy.log(OmegaM(a))* OmegaM(a)**0.55 / a
+def integrand_dlnfs(a,*args):
+    return numpy.log(OmegaM(a,*args))* OmegaM(a,*args)**0.55 / a
 
-def dlnfs8dg(a):
-    return numpy.log(OmegaM(a)) + integrate.quad(integrand_dlnfs, 1e-8, a)[0]
+def dlnfs8dg(a,*args):
[…]

### 2019-02-07 pv/view.py
remove commented lines

@@ -95,11 +95,7 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
     chain=  pickle.load(open('/Users/akim/project/PeculiarVelocity/outcosmo/'+name,'rb'))
     cutchain = chain[:,500:,:]
     # plotter(cutchain)
-    # c = ChainConsumer()
-    # dum = numpy.reshape(cutchain,(cutchain.shape[0]*cutchain.shape[1],cutchain.shape[2]))
-    # dum[:,3] = dum[:,3]*3e5
-    # c.add_chain(dum, parameters=["$A$", "$\mathcal{M}$","$\sigma_M$","$\sigma_{v}$"])
-    # c.plotter.plot(filename="/Users/akim/project/PeculiarVelocity/outcosmo/"+name+".png", figsize="column",truth=[None,0,0.08,None])
+
     fsigma8 = numpy.sqrt(cutchain[:,:,0])
     effston.append(fsigma8.mean()/fsigma8.std()*numpy.sqrt(18000./760))
     # print("{:4.2f} & {:4.2f} & {:4.2f} & {} & {:6.2f} & {:6.2f} & {:6.3f} & {:6.2f} \\\\".format( \
@@ -111,6 +107,7 @@ for name,nsn,zma,fra,snsi in zip(names,nsne,zmax,frac,snsig):
         effston[-1]/numpy.sqrt(18000./760*nsn)\
         ))
 
+plt.clf()
 effston = numpy.array(effston)
 mkrs=['o','s','*','D']
 for z_,m in zip(zmax_,mkrs):
@@ -131,7 +128,6 @@ plt.legend(fontsize=14)
 plt.tight_layout()
 plt.savefig('notshot.png')
 plt.clf()
-
 
 for z_,m in zip(zmax_,mkrs):
     fs = []

### 2019-02-10 doc/article.tex
add r_min
putting in some of greg's comments

@@
 Peculiar velocity surveys have already been  used to measure $fD$, though not to a level where  gravity models can be precisely distinguished.
- \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane method to
+ \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane distances of spiral galaxies, using absolute magnitudes
+ with
+ $\sim 0.43$~mag  precision calibrated by emission line velocity dispersions, to
 get a 15\% uncertainty in $fD$ at $z\approx 0$ using combined density and velocity information.
@@
 TAIPAN survey will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
-and the WALLABY+WNSHS surveys will obtain Tully-Fisher galaxies at densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
+and the WALLABY+WNSHS surveys will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
+of galaxies with densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
 $z=0-0.1$ covering 75\% of the sky.
@@
 Recent studies indicate that with the right data, SN absolute
-magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}.  Though not yet
+magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}. 
+Though not yet
 established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
 that is observed using current calibrations.  
-One such SN is worth $\gtrsim 25$ galaxies with 0.4~mag absolute magnitude uncertainty.
+At this precision the intrinsic velocity dispersion  at $z=0.028$ is  of $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
+provide $S/N \sim 1$.
 Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
@@
   This sample has comparable
- number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.
+ number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.  With similar densities,
+ the 10-year SN~Ia survey will have
+ a 29$\times$ reduction in shot-noise ($\sigma^2_M/n$) relative to the Fundamental Plane survey of TAIPAN.
 
@@
 \citep[there is an alternative approach using an estimator for the mean pairwise velocity][]{2011PhRvD..83d3004B}.
+***Super-sample vairaince***
 The Fisher information matrix is
@@
 %\frac{\partial C}{\partial \lambda_j} \right]\\
-& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+& = \frac{\Omega}{8\pi^2} \int_{r_{min}}^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
 \frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
@@
 \end{equation}
-and the two parameters considered are $\lambda \in \{\gamma,bD\}$.
+and the parameters considered are $\lambda \in \{\gamma,bD, \Omega_{M0}\}$.
 The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy overdensity
@@
 and the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
+Both $f$ and $D$ depend on $\Omega_M=\frac{\Omega_{M0}}{\Omega_{M0} + (1-\Omega_{M0})a^3}$.
 The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
@@
 $k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
+A minimum distance $r_{min}=r(z=0.01)$ is imposed as the physics derivation is taken at the limit
+where peculiar velocities are significantly  smaller than
+the cosmological redshift.
 The sample-selection efficiency $\epsilon$ is redshift-independent, i.e.\ the native redshift distribution is not sculpted.
-The input bias of SN~Ia host galaxies is set as $b=1.2$.
+The input bias of SN~Ia host galaxies is set as $b=1.2$.  An independent measurement of $\Omega_{M0}=0.3\pm0.005$
+is included and is important contributor to the $\gamma$ constraint.  ***Uniform sampling ***
 
@@
 The primary analysis of interest uses overdensities (RSD), peculiar velocities, and their cross-correlations.
-The short and shallow,  2-year, $z_{max}=0.1$ survey has $\sigma_\gamma \sim 0.038$,
+The short and shallow,  2-year, $z_{max}=0.11$ survey has $\sigma_\gamma \sim 0.038$,
 which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $>3\sigma$ level.
@@
 This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{max}=0.2$.
-The leftpanel of Figure~\ref{dvar:fig} shows
+The left panel of Figure~\ref{dvar:fig} shows
  $|\partial \sigma_\gamma / \partial z|$, which
@@
 shorter surveys and those with higher $z_{max}$ (because of the higher $\sigma$), benefit more from extending survey duration.
-These trends are shown in the left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
+These trends are shown in the right panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
  $z_{max}$ for two- and ten-year surveys.
@@
 to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  These redshifts could come naturally
-from galaxy signal in  the spectroscopic transient classification if the spectrograph had sufficient resolution.  Alternatively, 
+from galaxy signal in  the spectroscopic transient classification if the spectrograph had sufficient resolution.  ***Check number and explain. ***  Alternatively, 
 moderate-resolution (bright-)galaxy  redshift surveys (e.g.\ DESI, 4MOST) can  provide multiplexed observations of 
@@
 Choices in survey design impact the precision obtained on the growth index $\gamma$.  
-No other probe of growth of structure or tracer of peculiar velocity can provide comparable precision on  $\gamma$ in the next decade.
+No other probe of growth of structure or tracer of peculiar velocity can alone provide comparable precision on  $\gamma$ in the next decade.
+Combined low-redshift peculiar velocity and high-redshift RSD $fD$ measurements are highly complimentary as together they probe the
+$\gamma$-dependent shape of $fD(z)$ (not just its normalization) and potential scale-dependent influence of gravitational models, since low-
+and high-redshift surveys are weighted by lower and higher $k$-modes respectively.
 SN~Ia peculiar velocity surveys are of the highest scientific
@@
 
-\begin{comment}
+***
+DESI and 4MOST will exist in the 2020 decade, (the former having been endorsed by Astro 2010), but at low redshift their constraints on fs8 will be limited by cosmic variance, Since these two redshifts surveys will cover most of the usable sky, they will saturate the RSD technique, making PVs the only way to improve on fs8 at low z.
+TAIPAN and DESI BGS will be able to measure FP distances for nearly all usable galaxies nearby, so at low z the FP PV technique will also saturate in the coming decade.
+Thus, SN PV will be the most powerful way to further constrain $\sigma_\gamma$.
+Moreover, SN PV directly benefits from DESI and 4MOST through the cross-correlation of the galaxy density and velocity fields, and the in the redshifts provided for many of the SN host galaxies.
+
+t TAIPAN+(+4MOST)+DESI are *capable* of saturating the FP technique nearby.
+If 4MOST or DESI chose not to do FP or fail in the attempt, that is just more reason to do SN PVs!
+But I think it's a good idea for readers to know that SN PVs are needed even if 4MOST and DESI do the best job they can on FP, simply because one runs out of suitable nearby FP galaxies.
+***
+%\begin{comment}
 \newpage
@@
   P_{vv}(k,\mu) + \frac{\sigma^2}{n}
+   \end{bmatrix}.
+\end{equation}
+\begin{equation}
+C^{-1} =
+ \frac{1}{(P_{gg}(k,\mu) + \frac{1}{n})(P_{vv}(k,\mu) + \frac{\sigma^2}{n})- P_{vg}(k,\mu)^2}
+  \begin{bmatrix}
+    P_{vv}(k,\mu) + \frac{\sigma^2}{n}&
+   P_{vg}(k,\mu)^2  \\
+   P_{vg}(k,\mu)^2 &
+   P_{gg}(k,\mu) + \frac{1}{n} 
    \end{bmatrix}.
@@
  \end{align}
-\end{comment}
+
+ \begin{align}
+ P_{vv,\Omega_M} & = 2(f_{,\Omega_M}D + fD_{,\Omega_M})(fD)^{-1}P_{vv}\\
+ P_{\delta \delta,\Omega_M} & = 2(bD_{,\Omega_M} +f_{,\Omega_M}D\mu^2 + fD_{,\Omega_M}\mu^2)(bD+fD\mu^2)^{-1}P_{\delta \delta}\\
+ P_{\delta v,\Omega_M} & =\left( (f_{,\Omega_M}D + fD_{,\Omega_M})(fD)^{-1} + 
+( bD_{,\Omega_M} +f_{,\Omega_M}D\mu^2 + fD_{,\Omega_M}\mu^2)(bD+fD\mu^2)^{-1}\right)
+ P_{\delta v}
+ \end{align}
+
+  \begin{align}
+D_{,\Omega_M} & =\gamma D \int_a^1 \Omega_M^{\gamma-1} d\ln{a}\\
+ f_{,\Omega_M} & = \gamma \Omega_M^{\gamma -1} \\
+ \Omega_{M,\Omega_{M_0}} & =\frac{\Omega_{M_0}(1-a^3) + 2a^3-1}{(\Omega_{M_0}+(1-\Omega_{M_0}a^3))^2}
+ \end{align}
+%\end{comment}
 %

### 2019-02-11 doc/article.tex
Greg's comments in

@@
 \affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+\author{A.~Bahmanyar}
+\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
+50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
 \author{S.~BenZvi}
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
+\author{R.~Hlo\u{z}ek}
+\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
+50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
 \author{C.~Howlett}
@@
 \caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
-Overplotted are volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$ expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions.
+Overplotted are volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$ expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions, and TAIPAN (green, dash-dotted).
 The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
@@
  number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.  With similar densities,
- the 10-year SN~Ia survey will have
- a 29$\times$ reduction in shot-noise ($\sigma^2_M/n$) relative to the Fundamental Plane survey of TAIPAN.
+ the (two) 10-year SN~Ia survey will have
+ a (6) 29$\times$ reduction in shot-noise, $\sigma^2_M/n$, relative to the Fundamental Plane survey of TAIPAN.
 
@@
 \citep[there is an alternative approach using an estimator for the mean pairwise velocity][]{2011PhRvD..83d3004B}.
-***Super-sample vairaince***
 The Fisher information matrix is
@@
 The input bias of SN~Ia host galaxies is set as $b=1.2$.  An independent measurement of $\Omega_{M0}=0.3\pm0.005$
-is included and is important contributor to the $\gamma$ constraint.  ***Uniform sampling ***
+is included and is important contributor to the $\gamma$ constraint.  Number densities are taken to be
+direction-independent, neglecting the slight declination-dependence of SN-survey time windows.
 
@@
 
-{\bf Host-galaxies Need Precise Redshifts:} Peculiar velocities come from the difference between host-galaxy and cosmological redshifts.  Host redshift uncertainties need to be
- $<0.5$\% in order not to contribute significantly
-to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  These redshifts could come naturally
-from galaxy signal in  the spectroscopic transient classification if the spectrograph had sufficient resolution.  ***Check number and explain. ***  Alternatively, 
+{\bf Host-galaxies Need Precise Redshifts:} Peculiar velocities come from the difference between host-galaxy and cosmological redshifts.  Fractional
+host-redshift uncertainties
+need to satisfy $\sigma_z/z \ll \sigma_M$ in order not to contribute significantly
+to the error budget.   These redshifts could come naturally
+from galaxy signal in  the spectroscopic transient classification provided the spectrograph had sufficient resolution $R \gg \sigma_M^{-1}$.  Alternatively, 
 moderate-resolution (bright-)galaxy  redshift surveys (e.g.\ DESI, 4MOST) can  provide multiplexed observations of 
@@
 
+We have assumed that the galaxy and peculiar-velocity surveys as being the one and the same, since
+the same galaxies are used for both overdensity and velocity measurements.  Though not considered in our calculations,
+there would be benefit in increasing the overdensity sample to cover the larger volume that
+affect the velocities within $z_{max}$, based on the expectations of linear theory and super-sample covariance.  Attaching these higher-redshift supernova hosts to
+a peculiar-velocity survey would
+not significantly strain the requirements, as they would not need precise distances.
+
 \section{Conclusions}
@@
 Choices in survey design impact the precision obtained on the growth index $\gamma$.
+
 No other probe of growth of structure or tracer of peculiar velocity can alone provide comparable precision on  $\gamma$ in the next decade.
+At low redshift, the RSD measurement is quickly sample variance limited (as are the planned DESI BGS and 4MOST surveys) making peculiar velocities the only 
+precision probe of $fD$.
+TAIPAN and a TAIPAN-like DESI BGS will be able to measure FP distances for nearly all usable nearby galaxies, so at low-$z$ the Fundamental Plane peculiar-velocity
+technique will  saturate at a level that is not competitive with a modes 2-year SN survey.
+
 Combined low-redshift peculiar velocity and high-redshift RSD $fD$ measurements are highly complimentary as together they probe the
@@
 
-***
-DESI and 4MOST will exist in the 2020 decade, (the former having been endorsed by Astro 2010), but at low redshift their constraints on fs8 will be limited by cosmic variance, Since these two redshifts surveys will cover most of the usable sky, they will saturate the RSD technique, making PVs the only way to improve on fs8 at low z.
-TAIPAN and DESI BGS will be able to measure FP distances for nearly all usable galaxies nearby, so at low z the FP PV technique will also saturate in the coming decade.
-Thus, SN PV will be the most powerful way to further constrain $\sigma_\gamma$.
-Moreover, SN PV directly benefits from DESI and 4MOST through the cross-correlation of the galaxy density and velocity fields, and the in the redshifts provided for many of the SN host galaxies.
-
-t TAIPAN+(+4MOST)+DESI are *capable* of saturating the FP technique nearby.
-If 4MOST or DESI chose not to do FP or fail in the attempt, that is just more reason to do SN PVs!
-But I think it's a good idea for readers to know that SN PVs are needed even if 4MOST and DESI do the best job they can on FP, simply because one runs out of suitable nearby FP galaxies.
-***
 %\begin{comment}

### 2019-02-11 doc/src/cov.py

@@ -297,6 +297,8 @@ axs.plot(matter[:,0],matter[:,0]**3*(0.08 * sigv_factor *3e5)**2 / (2*restrate/(
 axs.plot(matter[:,0],matter[:,0]**3*(0.15 * sigv_factor *3e5)**2 / (2*restrate/(1+z)), label=r'$k^3\sigma^2/n$; $\sigma_M=0.15$, 2 years, $z=0.1$',ls=':',color='red')
 axs.plot(matter[:,0],matter[:,0]**3*(0.08 * sigv_factor *3e5)**2 / (10*restrate/(1+z)), label=r'$k^3\sigma^2/n$; $\sigma_M=0.08$, 10 years, $z=0.1$',ls='--',color='black')
 axs.plot(matter[:,0],matter[:,0]**3*(0.15 * sigv_factor *3e5)**2 / (10*restrate/(1+z)), label=r'$k^3\sigma^2/n$; $\sigma_M=0.15$, 10 years, $z=0.1$',ls=':',color='black')
+axs.plot(matter[:,0],matter[:,0]**3*(0.2*0.1*3e5)**2 / 1e-3, label=r'$k^3\sigma^2/n$; TAIPAN, $z=0.1$',ls='-.',color='green')
+
 axs.set_ylabel(r'[km$^2$s$^{-2}$]')
 # axs.ticklabel_format(style='sci', axis='y',scilimits=(0,0))
 dum=1.07
@@ -317,7 +319,25 @@ axs.legend(loc=2, prop={'size': 'x-small'})
 plt.subplots_adjust(hspace=0.05)
 plt.tight_layout()
 plt.savefig('new2.png')
-wefwe
+plt.clf()
+fig, axs = plt.subplots(1,1,sharex=True)
+colors=['blue','green','red']
+zs = numpy.array([0.05,0.1,.2])
+axs.plot(matter[:,0],matter[:,0]**3*Pgg(1),label=r'$k^3P_{vv}(\mu=1)$',color='black')
+axs.plot(matter[:,0],matter[:,0]**3*Pgg(0.5),label=r'$k^3P_{vv}(\mu=0.5)$',color='cyan')
+axs.plot(matter[:,0],matter[:,0]**3/1e-3, label=r'$k^3\sigma^2/n$; TAIPAN, $z=0.1$',ls='-.',color='green')
+
+# axs.ticklabel_format(style='sci', axis='y',scilimits=(0,0))
[…]

### 2019-02-15 doc/article.tex
remove appendix
antilogus comments

@@
 \affiliation{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
-\author{S.~Perlmutter}
+\author{P.~Nugent}
+\affiliation{Computational Research Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}\author{S.~Perlmutter}
 \affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
@@
 
+A search tailored to discover SN~Ia to a certain depth will also discover intrinsically fainter SNe~IIP to a shallower
+redshifts.  Advances are being made to calibrate these SN distances, with a 0.2~mag distance modulus uncertainty seemingly feasable.
+With a  $\sim 4.5$ higher number density than SNe~Ia   we project uncertainties in the growth index of
+$\sigma_\gamma = 0.067, 0.046$ for  $z_{max}=0.05$, 2- and 10-year surveys respectively.  Including follow-up of
+SNe~IIP witin a SN~Ia program could be beneficial, particularly for checks of systematics.
+
 \section{Conclusions}
@@
 
-%\begin{comment}
+\begin{comment}
 \newpage
@@
  \end{align}
-%\end{comment}
+\end{comment}
 %

### 2019-03-05 doc/article.tex
after Tam and Or comments

@@
 
+\textbf{Endorsers:} (names and institutions)
+  \linebreak
+ Behzad Ansarinejad $^{1,2}$, 
+ Robert Armstrong $^{}$, 
+ Carlo Baccigalupi $^{3,4,5}$, 
+ Maciej Bilicki $^{6}$, 
+ Julian Borrill $^{7}$, 
+ Simone Ferraro $^{7}$, 
+ Llu\'is Galbany $^{8}$, 
+ Juan Garc\'ia-Bellido $^{9}$, 
+ Juan Garc\'ia-Bellido $^{10}$, 
+ Martina Gerbino $^{}$, 
+ Mandeep S.S. Gill $^{11,12,13}$, 
+ Larry Gladney $^{14}$, 
+ Marc Kamionkowski $^{1,19}$, 
+ Kazuya Koyama $^{20}$, 
+ Michele Liguori $^{21}$, 
+ Joel Meyers $^{22}$, 
+ "Surhud More" $^{23}$, 
+ Jeffrey A. Newman $^{8}$, 
+ Gustavo Niz $^{24}$, 
+ "Antonella Palmese" $^{25}$, 
+ Francesco Piacentini $^{26}$, 
+ Francesco Piacentni  $^{26,27}$, 
+ Steven Rodney $^{28}$, 
+ Benjamin Rose $^{29}$, 
+ Lado Samushia $^{30}$, 
+ Neelima Sehgal $^{31}$, 
+ Leonardo Senatore $^{11}$, 
+ Huanyuan Shan $^{1,33}$, 
+ Scott Watson $^{35}$, 
+ Martin White $^{36,7}$, 
+ Weishuang Xu $^{37}$, 
+ Gong-Bo Zhao $^{38,39,20}$,
+
+%%% 
+%\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
+% $^{1}$ \ED \\
+%$^{2}$ \Durham \\
+%$^{3}$ \SISSA \\
+%$^{4}$ \IFPU \\
+%$^{5}$ \INFN \\
+%$^{6}$ \CFT \\
+%$^{7}$ \LBL \\
+%$^{8}$ \Pitt \\
+%$^{9}$ \IFT \\
+%$^{10}$ \UAM \\
+%$^{11}$ \KIPAC \\
+%$^{12}$ \Stanford \\
+%$^{13}$ \SLAC \\
+%$^{14}$ \Yale \\
+%$^{15}$ \UR \\
+%$^{16}$ \dunlap \\
+%$^{17}$ \daa \\
+%$^{18}$ \Umich \\
+%$^{19}$ \JHU \\
+%$^{20}$ \Port \\
+%$^{21}$ \UNIPD \\
+%$^{22}$ \SMU \\
+%$^{23}$ \IUCAA \\
+%$^{24}$ \UGTO \\
+%$^{25}$ \FNAL \\
+%$^{26}$ \RomaS \\
+%$^{27}$ \INFNRM \\
+%$^{28}$ \USC \\
+%$^{29}$ \STSCI \\
+%$^{30}$ \KSU \\
+%$^{31}$ \StonyBrook \\
+%$^{32}$ \KASSI \\
+%$^{33}$ \SHAO\\
+%$^{34}$ \BNL \\
+%$^{35}$ \Syracuse \\
+%$^{36}$ \UCB \\
+%$^{37}$ \HarvardPhys \\
+%$^{38}$ \NAOC \\
+%$^{39}$ \UCAS \\
+%
 \textbf{Abstract:}
@@
 \affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+\author{O.~Graur}
+\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138, USA}
+\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
+\affiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
 \author{R.~Graziani}
@@
 50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
+\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON, M5S 3H4}
 \author{C.~Howlett}
@@
 Lyon, France} 
+\author{A.~Shafieloo}
+\affiliation{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
 \author{A.~Slo\u{z}ar}
@@
 In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
-The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the linear growth factor and
+The SN~Ia Hubble diagram has  scatter along the redshift axis since peculiar velocities induce differing observed and cosmological redshifts.  The contribution
+of this scatter is pronounced at low redshifts and when the scatter along the magnitude axis (i.e.\ intrinsic magnitude dispersion) is small.
+The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the spatially-independent
+growth factor for the linear evolution of density perturbations and
 $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
@@
 The upcoming 
-TAIPAN survey will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
-and the WALLABY+WNSHS surveys will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
+TAIPAN survey \citep{2017PASA...34...47D} will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
+and the WALLABY+WNSHS surveys \citep{2008ExA....22..151J} will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
 of galaxies with densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
@@
 These surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
-(For reference, DESI projects 10\% precision of $fD$ at $z \approx 0.3$  using Redshift Space Distortions (RSD) alone.)
+For reference, DESI projects 10\% precision of $fD$ at $z \approx 0.3$  by statistically looking 
+for expected signatures of galaxies infalling toward mass overdensities, Redshift Space Distortions (RSD).
+Relative to galaxies with  Fundamental Plane or Tully-Fisher distances, 
+SN~Ia host galaxies currently have significantly lower number density but have better per-object peculiar velocity precision.
 Existing SN~Ia samples
@@
 \citep{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
+SNe~Ia discovered by ZTF over the next several years will provide first probative measures of $fD$ at $z<0.1$.
 %Measurement of the velocity field using LSST-discovered
@@
 2.2\% uncertainty on fD  within this interval , where at
-$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from $RSD$  \citep{2017ApJ...847..128H}.
+$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from Redshift Space Distortions  \citep{2017ApJ...847..128H}.
 
@@
 and the parameters considered are $\lambda \in \{\gamma,bD, \Omega_{M0}\}$.
-The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy overdensity
+The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy count overdensity
 power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
@@
 
-We consider peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
+We consider SN peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
 The other survey parameters
@@
 \includegraphics[width=0.49\textwidth]{src/var.png}
-\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) surveys of varying depth $z_{max}$.
+\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) SN~Ia surveys of varying depth $z_{max}$.
 For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses

### 2019-03-05 doc/density.tex
after Tam and Or comments

\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath}
\usepackage{natbib,verbatim}
%\usepackage{grffile}
%\usepackage{textcase}
\usepackage{hyperref}
%\usepackage[overload]{textcase}

%SetFonts

%SetFonts

\begin{document}
Consider a galaxy survey that is complete in a volume.  Suppose there are $N_{gal}$ galaxies in the survey.  The volume is divided into $N_{cell} \gg N_{gal}$
equal-volume cells, such that the number of galaxies that occupy cell $i$, $N_i=0$ or $1$.  Suppose that the expected number density is constant over the volume.
The number of expected galaxies per cell and its variance (from Poisson statistics) is $\bar{N} = \frac{N_{\text{gal}}}{N_{\text{cell}}}$.  
Denote the dark matter overdensity correlation function as $\xi^{DM}_{\delta \delta}$. The galaxy overdensity correlation function is
$\xi_{\delta \delta} =  b^2 \xi^{DM}_{\delta \delta}$.  The number-per-cell correlation function 
is $\xi_{NN}  =\bar{N}^2 \xi_{\delta \delta}$.

We are interested in the amplitude of the correlation functions, so we parameterize
 this is $\xi_{NN} =A\bar{N}^2S$.
The expected covariance in galaxy numbers is
\begin{align*}
C &= \bar{N} I + A\bar{N}^2S.
\end{align*}

The likelihood is written as $-2\ln{L} = \ln{\det{C}} + (N-\bar{N})^T C^{-1}(N-\bar{N}) + const$.   Lets Taylor expand about $A=0$ and collect terms that
depend on $A$.  The second term gives
\begin{align*}
A (N-\bar{N})^T S(N-\bar{N})
\end{align*}
The first term gives (check this)
\begin{align*}
A\text{Tr}(\bar{N}S),
\end{align*}
which is goes to zero as we go to infinitely small cells so $\bar{N} \rightarrow 0$.  So we only have to worry about the first term, which when the
cells are very small
\begin{align*}
A (N-\bar{N})^T S(N-\bar{N}) &= A \left( \sum_{i,j \in {galaxy}} S_{ij} -2 \sum_{i \in {galaxy}, j\in {cell}} \bar{N}N_i S_{ij}  +  \sum_{ i,j\in {cell}}  \bar{N}^2 S_{ij}\right).
\end{align*}
The first term is the ``standard'' piece.  The second two pieces are new, which involve integrals over the model correlations.

Lets consider the velocities at the same time.  The covariance matrix looks like
\begin{align*}
C &=
\begin{pmatrix}
\bar{N} I &  0 \\
0&\vec{\sigma}_v^2 I \\
\end{pmatrix}
+A\begin{pmatrix}
\bar{N}^2\xi_{\delta\delta} &  \bar{N}\xi_{\delta v}\\
 \bar{N}\xi_{\delta v}& \xi_{vv} \\
  \end{pmatrix}\\
  & = \left(\sqrt{\bar{N}}I, \vec{\sigma}_v I \right) \left(I + 
  A\begin{pmatrix}
\bar{N}\xi_{\delta\delta} &  \sqrt)\bar{N}) \vec{\sigma}_v \xi_{\delta v}\\
 \bar{N}\xi_{\delta v}& \xi_{vv} \\
  \end{pmatrix}
  \right)\left(\sqrt{\bar{N}}I, \vec{\sigma}_v I \right).
\end{align*}
Again the likelihood is written as $-2\ln{L} = \ln{\det{C}} + (N-\bar{N})^T C^{-1}(N-\bar{N}) + const$.   Lets Taylor expand about $A=0$ and collect terms that
depend on $A$.

The second term gives
\begin{align*}
A (N-\bar{N})^T S(N-\bar{N})
\end{align*}
The first term gives (check this)
\begin{align*}
A\text{Tr}(\bar{N}S),
\end{align*}
which is goes to zero as we go to infinitely small cells so $\bar{N} \rightarrow 0$.  So we only have to worry about the first term, which when the
cells are very small
\begin{align*}
A (N-\bar{N})^T S(N-\bar{N}) &= A \left( \sum_{i,j \in {galaxy}} S_{ij} -2 \sum_{i \in {galaxy}, j\in {cell}} \bar{N}N_i S_{ij}  +  \sum_{ i,j\in {cell}}  \bar{N}^2 S_{ij}\right).
\end{align*}
The first term is the ``standard'' piece.  The second two pieces are new, which involve integrals over the model correlations.

Taylor expand this around $A=0$.  Then reduce based on cells that have $\sigma_v \rightarrow \infty$.
\begin{comment}

%The overdensity is $\delta_{gg}=-1$ for cells with no galaxies and 
%$\delta_{gg}=\frac{N_{\text{cell}}}{N_{\text{gal}}}-1$ with one galaxy.  The variance in overdensity is $\frac{N_{\text{cell}}}{N_{\text{gal}}}$.
%We approximate the Poisson distribution as being Gaussian (the fact that this is a poor approximation we have to check).
The probability for the data, $\mathbf{x}$ (a vector composed of 0's and 1's) 
is modeled to come from a Gaussian distribution (despite the fact that Poisson statistics better describes the shot noise) with covariance matrix
\begin{align*}
C &= \left(\frac{N_{\text{gal}}}{N_{\text{cell}}}\right)^2\xi_{\delta \delta} + \frac{N_{\text{gal}}}{N_{\text{cell}}}I\\
 &= \left(\frac{N_{\text{gal}}}{N_{\text{cell}}}\right)
 \left(I + \frac{N_{\text{gal}}}{N_{\text{cell}}} \xi_{\delta \delta}\right),
\end{align*}
where $\xi_{\delta \delta}$ is the overdensity correlation.

The challenge is to approximate the likelihood not using the full matrix but only the small fraction of cells with a galaxy because the cells with zero galaxies drop out,
in the limit of $N_{\text{cell}} \rightarrow \infty$.

For galaxies counts, the likelihood is the probability that there are $N$ galaxies in a certain configuration described by the distances between all of the pairs $\mathbf{r}_{ij}$,
given the model expected galaxy density $n$ and the galaxy two-point correlation function $\xi(r)$, as
$p(\mathbf{r}_{ij};n, \xi(r))$.

Recall the relationship between the joint probability of two galaxies and  the two-point correlation function
$$p(r;n,\xi) = n^2[1+\xi(r)].$$

The relationship between the joint probability of three and four galaxies and three- and four-point correlation functions are introduced in \citet{1975ApJ...196....1P} and \citet{1978ApJ...221...19F}.  If the underlying distribution is Normal, the joint probabilities depend only on the two-point function, such that
$$p(r_{12},r_{23},r_{13};n,\xi) = n^3[1+\xi_{12}+\xi_{23}+\xi_{13}]$$
$$p(r_{12},r_{23},r_{13},r_{14},r_{23},r_{24},r_{34};n,\xi) = n^4[1+\xi_{12}+\xi_{23}+\xi_{13}+\xi_{14}+\xi_{23}+\xi_{24}+\xi_{34}+
\xi_{12}\xi_{34}+
\xi_{13}\xi_{24}+
\xi_{14}\xi_{23}].$$
The joint probability of $N$ galaxies is constructed in an analogous manner and is composed 
of the $\xi_{ij}$ of all pairs.

Now consider the case of a simultaneous analysis including both galaxy positions and velocities.
Divide the Universe into $N_{\text{cell}}$ cells whose sizes are small enough that there will only will be one galaxy in each.  The cell centers
are configured angularly (e.g.\ with HEALPIX) such that any given cell is separated from the others by the same set of quantized azimuthal distances.  With
the volume binned based on radius, the same set of distances separate any cell from the others.
Convert the survey containing $N_{\text{gal}}$ galaxies into a galaxy overdensity map, where $\delta_{gg}=-1$ for cells with no galaxies and 
$\delta_{gg}=\frac{N_{\text{cell}}}{N_{\text{gal}}}-1$.  Then we can write a big covariance matrix $C$ for the overdensity and velocity fields. 
The covariance matrix is composed of 4 quadrants.  The density piece is $C_{\delta \delta}[i,j] = A\xi_{\delta \delta}(i,j) + \frac{N_{\text{gal}}}{N_{\text{cell}}}\delta_{ij}$ where $i$ and $j$ are cell indices,
the cross-term piece is $C_{\delta v}[i,j] = A\xi_{\delta v}(i,j)$, and the velocity piece is  $C_{vv}[i,j] =A\xi_{vv}(i,j) +  \sigma^2_v\delta_{ij}$.  For cells
that do not contain a galaxy $\sigma^2_v=\infty$.  Recall the $\xi$ is a function of distance between two points, thus each of these matrices can be ordered
to be ``banded'' matrix.  The eigenvalues of banded matrices are known\footnote{\url{http://lampx.tugraz.at/~hadley/ss2/appendix/bandedmatrices/bandedmatrix.php}}.
The likelihood depends on
$$-2\ln{L} = 
\begin{pmatrix}
\delta_{gg}, v
\end{pmatrix}
\begin{pmatrix}
C_{\delta \delta} & C_{\delta v}\\
C_{\delta v} & C_{vv}
\end{pmatrix}^{-1}
\begin{pmatrix}
\delta_{gg}\\
 v
\end{pmatrix}
+
\ln{\left(\det{\begin{pmatrix}
C_{\delta \delta} & C_{\delta v}\\
C_{\delta v} & C_{vv}
\end{pmatrix}}\right)} + const.
$$
We are not interested in the absolute normalization of the likelihood, so let us consider perturbations around the $A=1$, 
$$-2\ln{L(A)} +  2\ln{L(A=1)}.$$
So we are interested in a Taylor expansion of $L$ about $A=1$ taken to the limit of $N_{\text{cell}} \rightarrow \infty$.

 \begin{thebibliography}{1}
\bibitem[Peebles \& Groth(1975)]{1975ApJ...196....1P} Peebles, P.~J.~E., \& Groth, E.~J.\ 1975, \apj, 196, 1 
\bibitem[Fry \& Peebles(1978)]{1978ApJ...221...19F} Fry, J.~N., \& Peebles, P.~J.~E.\ 1978, \apj, 221, 19 
 \end{thebibliography}
 \end{comment}
\end{document}

### 2019-03-06 doc/article.tex
get more into white paper format

@@
 \usepackage{natbib,verbatim}
+\usepackage[multiple]{footmisc}
 %\usepackage{grffile}
@@
 \usepackage{hyperref}
+
 %\usepackage[overload]{textcase}
@@
 
-\textbf{Co-authors:} (names and institutions)
+\textbf{Co-authors:}
+%\author[0000-0001-6315-8743]{A.~G.~Kim}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+G.~Aldering\footnotemark[1],
+\footnotetext[1]{Physics Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+P.~Antilogus\footnotemark[2],
+\footnotetext[2]{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+A.~Bahmanyar\footnotemark[3],
+\footnotetext[3]{Department of Astronomy and Astrophysics, University of Toronto,
+50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
+S.~BenZvi\footnotemark[4],
+\footnotetext[4]{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+H.~Courtois\footnotemark[5],
+\footnotetext[5]{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
+1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+Lyon, France} 
+T.~Davis\footnotemark[6],
+\footnotetext[6]{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
+S.~Gontcho A Gontcho\footnotemark[4],
+%\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+O.~Graur\footnotemark[7]\textsuperscript{,}\footnotemark[8]\textsuperscript{,}\footnotemark[9],
+\footnotetext[7]{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138, USA}
+\footnotetext[8]{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
+\footnotetext[9]{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+R.~Graziani\footnotemark[10],
+\footnotetext[10]{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
+Physique de Clermont, F-63000 Clermont-Ferrand, France}    
+C.~Harper\footnotemark[1],
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+R.~Hlo\u{z}ek\footnotemark[3]\textsuperscript{,}\footnotemark[11],
+%\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
+%50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
+\footnotetext[11]{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON, M5S 3H4}
+C.~Howlett\footnotemark[12],
+\footnotetext[12]{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
+D.~Huterer\footnotemark[13],
+\footnotetext[13]{Department of Physics, University of Michigan, 450 Church Street, Ann
+Arbor, MI 48109, USA }
+C.~Ju\footnotemark[12],
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+P.-F.~Leget\footnotemark[14],
+\footnotetext[14]{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+E.~V.~Linder\footnotemark[1],
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+P.~McDonald\footnotemark[1],
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+J.~Nordin\footnotemark[15],
+\footnotetext[15]{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
+P.~Nugent\footnotemark[16],
+\footnotetext[16]{Computational Research Division, Lawrence Berkeley National Laboratory, 
+    1 Cyclotron Road, Berkeley, CA, 94720}
+S.~Perlmutter\footnotemark[1],
+N.~Regnault\footnotemark[14],
+%\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+M.~Rigault\footnotemark[5],
+%\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
+%1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+%Lyon, France} 
+A.~Shafieloo\footnotemark[17],
+\footnotetext[17]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
+A.~Slo\u{z}ar\footnotemark[18],
+\footnotetext[18]{Brookhaven National Laboratory, Physics Department, Upton, NY
+11973, USA} 
+M.~Wood-Vasey\footnotemark[19]
+\footnotetext[19]{PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA} 
 \linebreak
 
-\textbf{Endorsers:} (names and institutions)
-  \linebreak
+\textbf{Endorsers:}
  Behzad Ansarinejad $^{1,2}$, 
@@
  Weishuang Xu $^{37}$, 
- Gong-Bo Zhao $^{38,39,20}$,
+ Gong-Bo Zhao $^{38,39,20}$
+  \linebreak
 
@@
 \setcounter{page}{1}
-\title{Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys}
-\author[0000-0001-6315-8743]{A.~G.~Kim}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\author{G.~Aldering}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\author{P.~Antilogus}
-\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
-\author{A.~Bahmanyar}
-\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
-50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
-\author{S.~BenZvi}
-\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
-\author{H.~Courtois}
-\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
-1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
-Lyon, France} 
-\author{T.~Davis}
-\affiliation{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
-\author{S.~Gontcho A Gontcho}
-\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
-\author{O.~Graur}
-\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138, USA}
-\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
-\affiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
-\author{R.~Graziani}
-\affiliation{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
-Physique de Clermont, F-63000 Clermont-Ferrand, France}    
-\author{C.~Harper}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\author{R.~Hlo\u{z}ek}
-\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
-50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
-\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON, M5S 3H4}
-\author{C.~Howlett}
-\affiliation{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
-\author{D.~Huterer}
-\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
-Arbor, MI 48109, USA }
-\author{C.~Ju}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\author{P.-F.~Leget}
-\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
-\author{E.~V.~Linder}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\author{P.~McDonald}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\author{J.~Nordin}
-\affiliation{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
-\author{P.~Nugent}
-\affiliation{Computational Research Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}\author{S.~Perlmutter}
-\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
-    1 Cyclotron Road, Berkeley, CA, 94720}
-\affiliation{Department of Physics, University of California Berkeley, 366 LeConte Hall MC 7300, Berkeley, CA, 94720-7300}
-\author{N.~Regnault}
-\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
-\author{M.~Rigault}
-\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
-1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
-Lyon, France} 
-\author{A.~Shafieloo}
-\affiliation{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
-\author{A.~Slo\u{z}ar}
-\affiliation{Brookhaven National Laboratory, Physics Department, Upton, NY
-11973, USA} 
-\author{others}
+%\title{Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys}
+%\author[0000-0001-6315-8743]{A.~G.~Kim}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\author{G.~Aldering}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\author{P.~Antilogus}
+%\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+%\author{A.~Bahmanyar}
+%\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
+%50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
+%\author{S.~BenZvi}
+%\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+%\author{H.~Courtois}
+%\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
+%1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+%Lyon, France} 
+%\author{T.~Davis}
+%\affiliation{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
+%\author{S.~Gontcho A Gontcho}
+%\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
+%\author{O.~Graur}
+%\affiliation{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138, USA}
+%\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
+%\affiliation{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+%\author{R.~Graziani}
+%\affiliation{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
+%Physique de Clermont, F-63000 Clermont-Ferrand, France}    
+%\author{C.~Harper}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\author{R.~Hlo\u{z}ek}
+%\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
+%50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
+%\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON, M5S 3H4}
+%\author{C.~Howlett}
+%\affiliation{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
+%\author{D.~Huterer}
+%\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
+%Arbor, MI 48109, USA }
+%\author{C.~Ju}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\author{P.-F.~Leget}
+%\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+%\author{E.~V.~Linder}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\author{P.~McDonald}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\author{J.~Nordin}
+%\affiliation{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
+%\author{P.~Nugent}
+%\affiliation{Computational Research Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}\author{S.~Perlmutter}
+%\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
+%    1 Cyclotron Road, Berkeley, CA, 94720}
+%\affiliation{Department of Physics, University of California Berkeley, 366 LeConte Hall MC 7300, Berkeley, CA, 94720-7300}
+%\author{N.~Regnault}
+%\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+%\author{M.~Rigault}
+%\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
+%1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+%Lyon, France} 
+%\author{A.~Shafieloo}
+%\affiliation{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
+%\author{A.~Slo\u{z}ar}
+%\affiliation{Brookhaven National Laboratory, Physics Department, Upton, NY
+%11973, USA} 
+%\author{M.~Wood-Vasey}
+%\affiliation{PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA} 
+%\author{others}
 
@@
 
-\begin{abstract}
-In the upcoming decade cadenced wide-field imaging surveys 
-will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
-hundreds of thousands.  The increase in the number density  and solid-angle coverage 
-of SNe~Ia, in parallel with improvements in the standardization of
-their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
-is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
-In the next decade the peculiar velocities of
-SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ with as low as
-$0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
-\end{abstract}
+%\begin{abstract}
+%In the upcoming decade cadenced wide-field imaging surveys 
+%will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+%hundreds of thousands.  The increase in the number density  and solid-angle coverage 
+%of SNe~Ia, in parallel with improvements in the standardization of
+%their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
+%is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
+%In the next decade the peculiar velocities of
+%SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ with as low as
+%$0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
+%\end{abstract}
 

### 2019-03-08 doc/article.tex
really add Tam's comments and Dave's

@@
 \footnotetext[6]{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
+Simone Ferraro\footnotemark[1],
 S.~Gontcho A Gontcho\footnotemark[4],
@@
 Physique de Clermont, F-63000 Clermont-Ferrand, France}
+J.~Guy\footnotemark[1],
 C.~Harper\footnotemark[1],
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
-S.~Perlmutter\footnotemark[1],
+S.~Perlmutter\footnotemark[1]\textsuperscript{,}\footnotemark[20],
+\footnotetext[20]{Department of Physics, University of California Berkeley, Berkeley, CA 94720}
 N.~Regnault\footnotemark[14],
@@
 11973, USA} 
+M.~White\footnotemark[1],\textsuperscript{,}\footnotemark[20],
 M.~Wood-Vasey\footnotemark[19]
@@
 their absolute magnitudes, now make them competitive probes of the growth of structure and hence of gravity.  The peculiar velocity power spectrum
-is sensitive to $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
+is sensitive to the growth index $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
 In the next decade the peculiar velocities of
-SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ with as low as
-$0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
+SNe~Ia in the local $z<0.3$ Universe will provide a measure of $\gamma$ to $\pm 0.01$ precision that can definitively distinguish  between General Relativity and leading models of alternative gravity.
 \end{raggedright}
@@
 growth factor for the linear evolution of density perturbations and
-$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate with $a$ the scale factor  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
 To be precise, the peculiar
 velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
-$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ is the observed distance and $\bar{d}_L(z)$ is the expected distance at the observed redshift $z$.
+$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ and  $\bar{d}_L(z)$ are the measured and expected distances at the observed redshift $z$.
 To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
@@
 surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
-in Figure~2 of  \citet{1475-7516-2013-04-031}.\footnote{The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
-commonly used in place of $D$ to normalize the
-overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.}
+in Figure~2 of  \citet{1475-7516-2013-04-031}.\footnote{Note that we prefer the notation $fD$ in place of $f\sigma_8$, which is commonly used in the literature.}
+%
+%The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
+%commonly used in place of $D$ to normalize the
+%overall amplitude of  overdensities, so the standard parameterization used by the community is $f\sigma_8$.}
 
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=1, 0.5$ (solid black, cyan)  as predicted for General Relativity in the linear regime.
-Overplotted are volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$ expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions, and TAIPAN (green, dash-dotted).
-The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
-redshift depths $z_{max}$.
+\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=\cos{(\hat{k} \cdot \hat{r})}=1, 0.5$ (magenta, cyan)  where 
+$\hat{r}$ is the line of sight, as predicted for General Relativity in the linear regime.
+Overplotted are volume-weighted peculiar-velocity power-spectrum shot noise predictions (diagonal lines) for various observing parameters.  Red shows the shot noise expected from a 2-year LSST survey
+while black shows a 10-year LSST survey.  The dotted and dashed lines indicate the assumed intrinsic magnitude dispersion, using 0.08 (dashed) or 0.15 mag (dotted).  The expected shot
+noise from TAIPAN is shown in green (dash-dotted). 
+%
+%Overplotted are volume-weighted peculiar velocity shot-noise  $k^3\sigma^2/n$ at $z=0.1$ expected from 2- and 10-year (red, black) supernova densities, 0.08 and 0.15~mag (dashed, dotted) intrinsic magnitude dispersions, and TAIPAN (green, dash-dotted).
+The bottom solid grey horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
+redshift depths $z_{\rm max}$.
 \label{power:fig}}
@@
 Peculiar velocity surveys have already been  used to measure $fD$, though not to a level where  gravity models can be precisely distinguished.
- \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane distances of spiral galaxies, using absolute magnitudes
+ \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane distances of elliptical galaxies, to estimate absolute magnitudes
  with
- $\sim 0.43$~mag  precision calibrated by emission line velocity dispersions, to
+ $\sim 0.43$~mag  precision to
 get a 15\% uncertainty in $fD$ at $z\approx 0$ using combined density and velocity information.
@@
 Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
-First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter SED model yields  absolute magnitude
+First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter spectral model yields  absolute magnitude
 dispersion $\sigma_M \gtrsim 0.12$~mag.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
@@
 established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
-that is observed using current calibrations.  
+that is observed using current calibrations.  If corrections of all SNe are not possible, the use of SN~Ia subclasses is an option though at the expense of reducing the sample.
 At this precision the intrinsic velocity dispersion  at $z=0.028$ is  of $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
-provide $S/N \sim 1$.
+provide a peculiar velocity with $S/N \sim 1$.
 Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
@@
 over the course of a decade a SN survey relying on LSST discoveries 
-plus spectroscopic redshifts can produce find 4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
-2.2\% uncertainty on fD  within this interval , where at
+plus spectroscopic redshifts can produce  4--14\% uncertainties in $fD$ in $0.05$ redshift bins from $z=0$ to 0.3, cumulatively giving
+2.2\% uncertainty on $fD$  within this interval, where at
 $0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from Redshift Space Distortions  \citep{2017ApJ...847..128H}.
@@
 \section{Testing Gravity with Peculiar Velocity Surveys}
-While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest here is in probing gravity
-so we here focus  on the growth index $\gamma$.
+While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest is in probing gravity,
+so here we focus  on the growth index $\gamma$.
 To illustrate the distinction,
@@
 %\frac{\partial C}{\partial \lambda_j} \right]\\
-& = \frac{\Omega}{8\pi^2} \int_{r_{min}}^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+& = \frac{\Omega}{8\pi^2} \int_{r_{\rm min}}^{r_{\rm max}}  \int_{k_{\rm min}}^{k_{\rm max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
 \frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
@@
 \end{equation}
-and the parameters considered are $\lambda \in \{\gamma,bD, \Omega_{M0}\}$.
+and the parameters considered are $\lambda \in \{\gamma,bD, \Omega_{M_0}\}$.
 The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy count overdensity
@@
 and the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
-Both $f$ and $D$ depend on $\Omega_M=\frac{\Omega_{M0}}{\Omega_{M0} + (1-\Omega_{M0})a^3}$.
+Both $f$ and $D$ depend on $\Omega_M=\frac{\Omega_{M_0}}{\Omega_{M_0} + (1-\Omega_{M_0})a^3}$.
 The uncertainty in $\gamma$ is $\sigma_\gamma = \sqrt{\left(F^{-1}\right)_{\gamma\gamma}}$.
 Non-GR models may  also predict a change in the scale-dependence of the growth or non-constant $\gamma$, such observations provide additional leverage in probing gravity
-but are not considered.
+but are not considered here.
 
 The uncertainty  $\sigma_\gamma$ of a survey depends on its solid angle $\Omega$, depth
-given by the comoving distance out to the maximum redshift $r_{max}=r(z_{max})$, duration
+given by the comoving distance out to the maximum redshift $r_{\rm max}=r(z_{\rm max})$, duration
 $t$ through $n=\epsilon \phi t$ where $\phi$ is the observer-frame SN~Ia rate
@@
 the intrinsic SN~Ia magnitude dispersion 
-through $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$.
+through the resulting peculiar velocity intrinsic dispersion $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$.
 An analysis in which the RSD and velocity surveys are treated independently 
@@
 
-We consider SN peculiar velocity surveys for a range of redshift depths $z_{max}$ for survey durations of $t=2$ and 10~years.
+We consider SN peculiar velocity surveys for a range of redshift depths $z_{\rm max}$ for survey durations of $t=2$ and 10~years.
 The other survey parameters
 $\Omega=3\pi$, $\epsilon=0.65$, $\sigma_M=0.08$~mag are fixed.
-The $k$-limits are  taken to be $k_{min} = \pi/r_{max}$  and 
-$k_{max} = 0.1$~$h$\,Mpc$^{-1}$.
-A minimum distance $r_{min}=r(z=0.01)$ is imposed as the physics derivation is taken at the limit
+The $k$-limits are  taken to be $k_{\rm min} = \pi/r_{\rm max}$  and 
+$k_{\rm max} = 0.1$~$h$\,Mpc$^{-1}$.
+A minimum distance $r_{\rm min}=r(z=0.01)$ is imposed as the physics derivation is taken at the limit
 where peculiar velocities are significantly  smaller than
@@
 is included and is important contributor to the $\gamma$ constraint.  Number densities are taken to be
-direction-independent, neglecting the slight declination-dependence of SN-survey time windows.
+direction-independent, neglecting the slight declination-dependence of SN-survey time windows.\footnote{Code used for the analysis in this article available
+at \url{https://github.com/LSSTDESC/SNPeculiarVelocity/tree/master/doc/src}.}
 
@@
 The primary analysis of interest uses overdensities (RSD), peculiar velocities, and their cross-correlations.
-The short and shallow,  2-year, $z_{max}=0.11$ survey has $\sigma_\gamma \sim 0.038$,
+The short and shallow,  2-year, $z_{\rm max}=0.11$ survey has $\sigma_\gamma \sim 0.038$,
 which can distinguish  between  General Relativity, $f(R)$,  and DGP gravities at the $>3\sigma$ level.
-The 10-year survey performance asymptotes at $z_{max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
+The 10-year survey performance asymptotes at $z_{\rm max} \sim0.2$ at a precision of  $\sigma_\gamma \sim 0.01$.
 Figure~\ref{var:fig} also shows uncertainties based on two other analyses, one that only uses peculiar velocities, and
@@
 \includegraphics[width=0.49\textwidth]{src/var.png}
-\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two- (red) and ten-year (black) SN~Ia surveys of varying depth $z_{max}$.
+\caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two-year (red) and ten-year (black) SN~Ia surveys of varying depth $z_{\rm max}$.
 For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
@@
 changes with respect to the 
-survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
-within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.
+survey parameters  $\Omega$, $z_{\rm max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
+within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{\rm max}$.
 
@@
 a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
-This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{max}=0.2$.
+This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{\rm max}=0.2$.
 The left panel of Figure~\ref{dvar:fig} shows
@@
 to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
-resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
+resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$.
 \begin{figure}
@@
 \includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
-\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
+\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{\rm max}=0.2$. 
 Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
- $z_{max}$ for two- (dashed) and ten-year (solid) surveys.
+ $z_{\rm max}$ for two- (dashed) and ten-year (solid) surveys.
 \label{dvar:fig}}
@@
 
-{\bf Redshift Depth $z_{max}$:}
+{\bf Redshift Depth $z_{\rm max}$:}
 Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
-$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
+$z_{\rm max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 
@@
 The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution,
-shorter surveys and those with higher $z_{max}$ (because of the higher $\sigma$), benefit more from extending survey duration.
+shorter surveys and those with higher $z_{\rm max}$ (because of the higher $\sigma$), benefit more from extending survey duration.
 These trends are shown in the right panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
- $z_{max}$ for two- and ten-year surveys.
+ $z_{\rm max}$ for two- and ten-year surveys.
 Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
 has a similar relationship with $\sigma_\gamma$.
-This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ also shown in the left panel of Figure~\ref{dvar:fig}.
+This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$, which is shown in the left panel of Figure~\ref{dvar:fig}.
 
-{\bf Minimum length scale, maximum wavenumber $k_{max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
-Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$,
+{\bf Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
+Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{\rm max} = 0.0050$ at $k_{\rm max}=0.1h$\,Mpc$^{-1}$,
  which  is survey-independent.
@@
 occurs at a $k$  within the range that carries the most signal; both sample variance and shot-noise limited modes contribute to the constraints.
-If $k_{max}$ were to decrease moderately some of the surveys could become sample variance limited, whereas increases in $k_{max}$ introduce additional
+If $k_{\rm max}$ were to decrease moderately some of the surveys could become sample variance limited, whereas increases in $k_{\rm max}$ introduce additional
  shot-limited modes.
@@
 SNIFS and SED Machine are examples of single-object spectrographs built with the purpose of classifying transients discovered in imaging surveys
-(the low surface density of active low-$z$ transients makes MOS superfluous).
+(the low surface density of active low-$z$ transients makes multi-object spectroscopy superfluous).
 Photometric classification is an alternative approach,  which benefits from the light-curve data already in hand from the search.  The level of
@@
 from galaxy signal in  the spectroscopic transient classification provided the spectrograph had sufficient resolution $R \gg \sigma_M^{-1}$.  Alternatively, 
-moderate-resolution (bright-)galaxy  redshift surveys (e.g.\ DESI, 4MOST) can  provide multiplexed observations of 
+moderate-resolution bright-galaxy  redshift surveys such as already planned by DESI can  provide multiplexed observations of 
 a significant fraction of nearby SN hosts before discovery or after transient light has faded away.
@@
 infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
- \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag. The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
- in number density, or equivalently in survey duration.
+ \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag. 
+ A survey with $\sigma_M=0.08$~mag probes provides the same constraints in $\gamma$ in 3.52$\times$ less time than if 0.15~mag dispersions were used.
 
-As a point of reference, a practically complete (accounting for temporal observing gaps) discovery up to $z_{max}=0.2$ would have $\sim 150,000$, $r \lesssim 20.5$~mag targets over 10 years.  
+As a point of reference, a practically complete (accounting for temporal observing gaps) discovery up to $z_{\rm max}=0.2$ would have $\sim 150,000$, $r \lesssim 20.5$~mag targets over 10 years.  
 Focusing on low-redshift supernovae  mitigates  several important uncertainties
@@
 there would be benefit in increasing the overdensity sample to cover the larger volume that
-affect the velocities within $z_{max}$, based on the expectations of linear theory and super-sample covariance.  Attaching these higher-redshift supernova hosts to
+affect the velocities within $z_{\rm max}$, based on the expectations of linear theory and super-sample covariance.  Attaching these higher-redshift supernova hosts to
 a peculiar-velocity survey would
@@
 With a  $\sim 4.5$ higher number density than SNe~Ia   we project uncertainties in the growth index of
-$\sigma_\gamma = 0.067, 0.046$ for  $z_{max}=0.05$, 2- and 10-year surveys respectively.  Including follow-up of
+$\sigma_\gamma = 0.067, 0.046$ for  $z_{\rm max}=0.05$, 2- and 10-year surveys respectively.  Including follow-up of
 SNe~IIP witin a SN~Ia program could be beneficial, particularly for checks of systematics.
@@
 \begin{align}
-F_{ij} &  = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+F_{ij} &  = \frac{\Omega}{8\pi^2} \int_0^{r_{\rm max}}  \int_{k_{\rm min}}^{k_{\rm max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
 \frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
@@
 
-$$F_{ij,z} =  \frac{\Omega}{8\pi^2} \frac{dr}{dz} r^2  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1}  k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
+$$F_{ij,z} =  \frac{\Omega}{8\pi^2} \frac{dr}{dz} r^2  \int_{k_{\rm min}}^{k_{\rm max}}  \int_{-1}^{1}  k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
 \frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk$$
 
-$$F_{ij,s} =  \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[
+$$F_{ij,s} =  \frac{\Omega}{8\pi^2} \int_0^{r_{\rm max}}  \int_{k_{\rm min}}^{k_{\rm max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[
 C^{-1}_{,s} \frac{\partial C}{\partial \lambda_i} C^{-1}\frac{\partial C}{\partial \lambda_j} +
@@
 
-For all analyses we remove supernovae below observed $z_{min}=0.01$ in order to
+For all analyses we remove supernovae below observed $z_{\rm min}=0.01$ in order to
 reduce errors made in the first-order transformation between velocity and
@@
 Different parameters describing the input simulated data define the subsamples, 
-which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{max}$, intrinsic
+which are used to explore the range of possible survey strategies.   The varied parameters are the redshift range, $z_{\rm max}$, intrinsic
 magnitude dispersion, $\sigma_M$, and SN number density, $n = \phi n_{10}$, where $ n_{10}$ is the  10-year total number density.  In general a given point
 of sky is not monitored over the full year.  
-The fit posterior for one representative subsample, with input  $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
+The fit posterior for one representative subsample, with input  $z_{\rm max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
 For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
@@
 absolute magnitude, $\sigma_M$ is the magnitude dispersion, and $\sigma_v$ is the non-linear contribution
-to peculiar velocity.  This result is for the subset $z_{max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
+to peculiar velocity.  This result is for the subset $z_{\rm max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$.  The dotted lines represent the inputs of the supernova
 part of the simulation.
@@
    \hline
-$z_{max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
+$z_{\rm max}$ & $\phi$ & $\sigma_{M}$  &  STON $f\sigma_8$ & STON per SN \\
 \hline
@@
    \end{tabular}
-   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{max}$,
+   \caption{ston in $A$ or equivalently $f\sigma_8$ for surveys parameterized by the maximum redshift $z_{\rm max}$,
    the fraction of total supernovae in 10 years ``fraction'', intrinsic magnitude dispersion $\sigma_{SN}$.  The number of galaxies
@@
 In Fourier space, the data covariance is a combination of sample and shot noise (Eq.~\ref{cov:eq}).
-In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{max}$.
+In Figure~\ref{scaling:fig} the effective LSST ston is plotted as a function of input $n \sigma^{-2}_M$ for different $z_{\rm max}$.
 For all cases, the ston improves with larger $n \sigma^{-2}_M$ showing that the sample-variance limit has not been reached,
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{max}$.
+\caption{ Effective LSST ston as a function of input  $n \sigma^{-2}_M$.  The points are color-coded for different $z_{\rm max}$.
 \label{scaling:fig}}
@@
 
-A SN peculiar survey benefits from greater redshift depth $z_{max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
+A SN peculiar survey benefits from greater redshift depth $z_{\rm max}$ due to the increase in volume and numbers of supernovae.  However,  for fixed magnitude uncertainty
 peculiar-velocity uncertainties increase with redshift and the follow-up of fainter distant SNe requires more resources. 
@@
 shows that the effective worth of each supernova ($\sigma_A N_{gal}^{-0.5}$) diminishes with increasing survey depth.
-The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
-a sample with $z_{max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
+The table also shows that for $\sigma_M=0.08$~mag, the subsample with $z_{\rm max}=0.1$ with 1294 supernovae yields a ston of 15.96, whereas
+a sample with $z_{\rm max}=0.2$ with 2021 supernovae has a lower ston of 13.61.  Within the range of surveys considered, a lower-redshift supernova
 is more valuable than one at high redshift.  For both scientific and resource reasons, it is best to devote resources to the lowest possible redshifts.

### 2019-03-08 doc/src/cov.py

@@ -289,9 +289,9 @@ sigv_factor = numpy.log(10)/5*z/(1+z)
 # sigv =sigm * sigv_factor *3e5
 
 
-colors=['blue','green','red']
+colors=['grey','grey','grey']
 zs = numpy.array([0.05,0.1,.2])
-axs.plot(matter[:,0],matter[:,0]**3*Pvv(1),label=r'$k^3P_{vv}(\mu=1)$',color='black')
+axs.plot(matter[:,0],matter[:,0]**3*Pvv(1),label=r'$k^3P_{vv}(\mu=1)$',color='magenta')
 axs.plot(matter[:,0],matter[:,0]**3*Pvv(0.5),label=r'$k^3P_{vv}(\mu=0.5)$',color='cyan')
 axs.plot(matter[:,0],matter[:,0]**3*(0.08 * sigv_factor *3e5)**2 / (2*restrate/(1+z)), label=r'$k^3\sigma^2/n$; $\sigma_M=0.08$, 2 years, $z=0.1$',ls='--',color='red')
 axs.plot(matter[:,0],matter[:,0]**3*(0.15 * sigv_factor *3e5)**2 / (2*restrate/(1+z)), label=r'$k^3\sigma^2/n$; $\sigma_M=0.15$, 2 years, $z=0.1$',ls=':',color='red')
@@ -319,6 +319,8 @@ axs.legend(loc=2, prop={'size': 'x-small'})
 plt.subplots_adjust(hspace=0.05)
 plt.tight_layout()
 plt.savefig('new2.png')
+
+wef
 plt.clf()
 fig, axs = plt.subplots(1,1,sharex=True)
 colors=['blue','green','red']

### 2019-03-08 doc/src/partials.py

@@ -531,8 +531,8 @@ def set2():
     plt.savefig('dvardz.png')
     plt.clf()
 
-set2()
-wefwe
+# set2()
+# wefwe
 
 def set1():
     fig,(ax) = plt.subplots(1, 1)
@@ -571,12 +571,12 @@ def set1():
         var_vonly.append(numpy.array(vvonly_)*2*3.14/.75) #3/4   
         # dvardkmax.append(numpy.array(dvdkmax_)*2*3.14/.75)
  
-    plt.plot(zmaxs,numpy.sqrt(var[0]),label='Two Years',color='red')
-    plt.plot(zmaxs,numpy.sqrt(var_ind[0]),label='Two Years, Independent Surveys',color='red',ls=':')  
     plt.plot(zmaxs,numpy.sqrt(var_vonly[0]),label='Two Years, Velocity Only',color='red',ls='--')  
-    plt.plot(zmaxs,numpy.sqrt(var[1]),label='Ten Years',color='black')
-    plt.plot(zmaxs,numpy.sqrt(var_ind[1]),label='Ten Years, Independent Surveys',color='black',ls=':')   
+    plt.plot(zmaxs,numpy.sqrt(var_ind[0]),label='Two Years, Independent Surveys',color='red',ls=':')  
+    plt.plot(zmaxs,numpy.sqrt(var[0]),label='Two Years, Cross-Correlated',color='red')
     plt.plot(zmaxs,numpy.sqrt(var_vonly[1]),label='Ten Years, Velocity Only',color='black',ls='--') 
+    plt.plot(zmaxs,numpy.sqrt(var_ind[1]),label='Ten Years, Independent Surveys',color='black',ls=':')   
+    plt.plot(zmaxs,numpy.sqrt(var[1]),label='Ten Years, Cross-Correlated',color='black')
 
     plt.xlabel(r'$z_{max}$')
     plt.ylim((0,0.07))
@@ -610,7 +610,7 @@ def set1():
 
 set1()
 
-
+wefwe
 def snIIp():
     fig,(ax) = plt.subplots(1, 1)
     zmax = 0.05
@@ -632,27 +632,22 @@ def snIIp():
         dvsigM_=[]
         dvdkmax_=[]
         for sigm in sigs:
[…]

### 2019-03-08 doc/article.tex
significantly shrunk to conform to White paper page limits

@@
-\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
+%\documentclass{article} 
+\documentclass[12pt]{article}
+\usepackage{times}
+\usepackage{geometry}
+\geometry{letterpaper, portrait, margin=1in}
+\usepackage[utf8]{inputenc}
+\usepackage{enumitem,amssymb}
+\usepackage{ragged2e}
+\newlist{thematic}{itemize}{8}
+\setlist[thematic]{label=$\square$}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+\hspace{-2.5pt}}
+\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
+
+%\documentclass{aastex62}   	% use "amsart" instead of "article" for AMSLaTeX format
 \usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
@@
 \usepackage[multiple]{footmisc}
+\usepackage{aas_macros}
 %\usepackage{grffile}
@@
 \usepackage{hyperref}
+%\usepackage{geometry}
 
@@
 \footnotetext[5]{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
-1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucléaire de
+1, Villeurbanne; CNRS/IN2P3, Institut de Physique Nucl\'eaire de
 Lyon, France} 
@@
 %\end{abstract}
-
-\section{Introduction}
+%\section{Introduction}
 
@@
 growth factor for the linear evolution of density perturbations and
-$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate with $a$ the scale factor  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.\footnote{
-To be precise, the peculiar
-velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
-$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ and  $\bar{d}_L(z)$ are the measured and expected distances at the observed redshift $z$.
-To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
-at low redshift.   The $H$-dependences of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
-}
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate with $a$ the scale factor  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
+%\footnote{
+%To be precise, the peculiar
+%velocity power spectrum also depends on the Hubble parameter as $P_{vv}\propto (HfD)^2$.  A supernova  survey  measures luminosity distance fluctuations
+%$\delta_{d_L} = (d_L- \bar{d}_L(z))/\bar{d}_L(z)$, where $d_L$ and  $\bar{d}_L(z)$ are the measured and expected distances at the observed redshift $z$.
+%To first order in peculiar velocity along the line of sight $v$, $\delta_{d_L} =  v \left(1-\frac{1}{H\bar{d}_L(z)}\right) \approx -\frac{v}{H\bar{d}_L(z)}$
+%at low redshift.   The $H$-dependences of  $P_{vv}$ and the conversion from distances to velocities cancel, making  peculiar velocity surveys sensitive to $(fD)^2$.
+%}
 The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
@@
 surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
-in Figure~2 of  \citet{1475-7516-2013-04-031}.\footnote{Note that we prefer the notation $fD$ in place of $f\sigma_8$, which is commonly used in the literature.}
+in Figure~2 of  \citet{1475-7516-2013-04-031}. (Note that we prefer the notation $fD$ in place of $f\sigma_8$, which is commonly used in the literature.)
 %
@@
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=\cos{(\hat{k} \cdot \hat{r})}=1, 0.5$ (magenta, cyan)  where 
-$\hat{r}$ is the line of sight, as predicted for General Relativity in the linear regime.
-Overplotted are volume-weighted peculiar-velocity power-spectrum shot noise predictions (diagonal lines) for various observing parameters.  Red shows the shot noise expected from a 2-year LSST survey
+\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=\cos{(\hat{k} \cdot \hat{r})}=1, 0.5$ (magenta, cyan) 
+($\hat{r}$ is the line of sight), as predicted for General Relativity in the linear regime.
+Overplotted are volume-weighted peculiar-velocity power-spectrum shot noise  (diagonal lines) for various observing parameters.  Red shows the shot noise expected from a 2-year LSST survey
 while black shows a 10-year LSST survey.  The dotted and dashed lines indicate the assumed intrinsic magnitude dispersion, using 0.08 (dashed) or 0.15 mag (dotted).  The expected shot
@@
 
-\section{Testing Gravity with Peculiar Velocity Surveys}
+%\section{Testing Gravity with Peculiar Velocity Surveys}
 While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest is in probing gravity,
@@
 is included and is important contributor to the $\gamma$ constraint.  Number densities are taken to be
-direction-independent, neglecting the slight declination-dependence of SN-survey time windows.\footnote{Code used for the analysis in this article available
-at \url{https://github.com/LSSTDESC/SNPeculiarVelocity/tree/master/doc/src}.}
+direction-independent, neglecting the slight declination-dependence of SN-survey time windows. Code used for the analysis in this article available
+at \url{http:tiny.cc/PVScience}.
 
@@
 \caption{The projected uncertainty in $\gamma$, $\sigma_\gamma$, achieved by two-year (red) and ten-year (black) SN~Ia surveys of varying depth $z_{\rm max}$.
-For each survey uncertainties are based on three types of analyses, one that only uses peculiar velocities (dashed), one that uses
-both RSD and peculiar velocities independently (dotted), and one  that uses
+For each survey uncertainties are based on three types of analyses:  using only peculiar velocities (dashed); using
+both RSD and peculiar velocities independently (dotted);  using
 both RSD, peculiar velocities, and their cross-correlation (solid).
@@
 survey parameters  $\Omega$, $z_{\rm max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
-within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{\rm max}$.
-
-The dependence on the uncertainty in the growth index on the survey parameters is as follows:
-
+within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{\rm max}$. \newline
+%
+%The dependence on the uncertainty in the growth index on the survey parameters is as follows:
+%
 {\bf Solid Angle $\Omega$:}
 The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
-
+%
 {\bf Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
@@
 to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
-resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$.
+resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$. \newline
 \begin{figure}
@@
 \end{figure}
-
 {\bf Redshift Depth $z_{\rm max}$:}
 Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
-$z_{\rm max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
-
+$z_{\rm max} \sim 0.2$ as seen in Figure~\ref{var:fig}. \newline
 {\bf Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
@@
 has a similar relationship with $\sigma_\gamma$.
-This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$, which is shown in the left panel of Figure~\ref{dvar:fig}.
-
+This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$, which is shown in the left panel of Figure~\ref{dvar:fig}.\newline
 {\bf Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
@@
  which  is survey-independent.
+
+\begin{comment}
 
@@
 
-\section{Elements of a Peculiar Velocity Survey}
+% \section{Elements of a Peculiar Velocity Survey}
 A successful peculiar velocity survey must accomplish several things, not just measure distances and host-galaxy redshifts.  These elements are as follows:
@@
 
-\section{Conclusions}
+\end{comment}
+% \section{Conclusions}
 
@@
 
-\bibliographystyle{aasjournal}
+\bibliographystyle{plain}
 \bibliography{/Users/akim/Documents/alex}

### 2019-03-08 doc/article.tex
clean up sections, bibiliography, better conclusions

@@
 
-In the late 1990's, Type~Ia supernovae (SNe~Ia) were used as distance probes to measure the homogeneous expansion history of the Universe.  The remarkable discovery
+{\bf Introduction:} In the late 1990's, Type~Ia supernovae (SNe~Ia) were used as distance probes to measure the homogeneous expansion history of the Universe.  The remarkable discovery
 that the expansion is accelerating  has called into question our basic understanding of the gravitational forces within the Universe.  Either it
@@
 growth factor for the linear evolution of density perturbations and
-$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate with $a$ the scale factor  \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate with $a$ the scale factor  \cite{2006PhRvD..73l3526H,2011ApJ...741...67D}.
 %\footnote{
@@
 The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
-\citet{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravity follow the relation
-$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively \citep[see][for a review of
+\cite{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravity follow the relation
+$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively \cite[see][for a review of
 these gravity models within a cosmological context]{HUTERER201523}.  Adopting this parameterization to model gravity, peculiar velocity
 surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
-in Figure~2 of  \citet{1475-7516-2013-04-031}. (Note that we prefer the notation $fD$ in place of $f\sigma_8$, which is commonly used in the literature.)
+in Figure~2 of  \cite{1475-7516-2013-04-031}. (Note that we prefer the notation $fD$ in place of $f\sigma_8$, which is commonly used in the literature.)
 %
@@
 Peculiar velocity surveys have already been  used to measure $fD$, though not to a level where  gravity models can be precisely distinguished.
- \citet{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane distances of elliptical galaxies, to estimate absolute magnitudes
+ \cite{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane distances of elliptical galaxies, to estimate absolute magnitudes
  with
@@
 The upcoming 
-TAIPAN survey \citep{2017PASA...34...47D} will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
-and the WALLABY+WNSHS surveys \citep{2008ExA....22..151J} will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
+TAIPAN survey \cite{2017PASA...34...47D} will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
+and the WALLABY+WNSHS surveys \cite{2008ExA....22..151J} will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
 of galaxies with densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
 $z=0-0.1$ covering 75\% of the sky.
-These surveys combined are projected to have 3\% $fD$ uncertainties \citep{2017MNRAS.464.2517H}.
+These surveys combined are projected to have 3\% $fD$ uncertainties \cite{2017MNRAS.464.2517H}.
 For reference, DESI projects 10\% precision of $fD$ at $z \approx 0.3$  by statistically looking 
@@
 have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
-\citep{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
+\cite{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
 SNe~Ia discovered by ZTF over the next several years will provide first probative measures of $fD$ at $z<0.1$.
 %Measurement of the velocity field using LSST-discovered
-%SNe~Ia has been quantified by \citet{2011PhRvD..83d3004B,2017JCAP...01..060O}
+%SNe~Ia has been quantified by \cite{2011PhRvD..83d3004B,2017JCAP...01..060O}
 
@@
 Recent studies indicate that with the right data, SN absolute
-magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \citep[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}. 
+magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \cite[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}. 
 Though not yet
@@
 2.2\% uncertainty on $fD$  within this interval, where at
-$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from Redshift Space Distortions  \citep{2017ApJ...847..128H}.
+$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from Redshift Space Distortions  \cite{2017ApJ...847..128H}.
 
 %\section{Testing Gravity with Peculiar Velocity Surveys}
-While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest is in probing gravity,
+{\bf Testing Gravity with Peculiar Velocity Surveys:} While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest is in probing gravity,
 so here we focus  on the growth index $\gamma$.
@@
 We project uncertainties  on the growth index, $\sigma_\gamma$ for a suite of idealized surveys
-using a Fisher matrix analysis similar to that of \citet{2017ApJ...847..128H, 2017MNRAS.464.2517H}
-\citep[there is an alternative approach using an estimator for the mean pairwise velocity][]{2011PhRvD..83d3004B}.
+using a Fisher matrix analysis similar to that of \cite{2017ApJ...847..128H, 2017MNRAS.464.2517H}
+\cite[there is an alternative approach using an estimator for the mean pairwise velocity][]{2011PhRvD..83d3004B}.
 The Fisher information matrix is
@@
 %
-{\bf Solid Angle $\Omega$:}
+{\it Solid Angle $\Omega$:}
 The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
 %
-{\bf Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
+{\it Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
 a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
@@
 to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
-resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$. \newline
+resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$. 
 \begin{figure}
@@
 \label{dvar:fig}}
-\end{figure}
-{\bf Redshift Depth $z_{\rm max}$:}
+\end{figure}\newline
+{\it Redshift Depth $z_{\rm max}$:}
 Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
 $z_{\rm max} \sim 0.2$ as seen in Figure~\ref{var:fig}. \newline
-{\bf Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
+{\it Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
 An increased survey duration accumulates more supernovae, decreasing shot noise and  increasing the 
@@
 This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$, which is shown in the left panel of Figure~\ref{dvar:fig}.\newline
-{\bf Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
+{\it Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
 Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{\rm max} = 0.0050$ at $k_{\rm max}=0.1h$\,Mpc$^{-1}$,
@@
 broad wavelength coverage, and with spectral-feature measurements.  Historically surveys with two or three well-sampled optical light curves obtained $\sigma_M \approx 0.15$~mag,
-with dense, $griz$-band optical light curves DES \citep{2018arXiv181102377B} gets $\sigma_M\approx 0.10$~mag, 
+with dense, $griz$-band optical light curves DES \cite{2018arXiv181102377B} gets $\sigma_M\approx 0.10$~mag, 
 standardization studies show that supplemental
-infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
- \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag. 
+infrared data \cite{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
+ \cite{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag. 
  A survey with $\sigma_M=0.08$~mag probes provides the same constraints in $\gamma$ in 3.52$\times$ less time than if 0.15~mag dispersions were used.
@@
 %SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  
-In the next decade,
+{\bf Conclusions:} In the next decade,
 the high number of SN discoveries together with improved precision in their distance precisions will make $z<0.2$ SNe~Ia, more so than
-galaxies,  powerful probes of gravity through their effect  on the growth of structure. 
-A peculiar velocity survey must discover, classify, and get distances of SNe~Ia, as well as obtain redshifts of their host galaxies.
-Choices in survey design impact the precision obtained on the growth index $\gamma$.
+galaxies,  powerful probes of gravity through their effect  on the growth of structure.  Different survey strategies can be adopted to take advantage of these
+supernovae, we provide a formalism and code by which their scientific merits can be assessed and here sketch results for a range of options.
+%A peculiar velocity survey must discover, classify, and get distances of SNe~Ia, as well as obtain redshifts of their host galaxies.
+%Choices in survey design impact the precision obtained on the growth index $\gamma$.
 
@@
 quantify the suppression of sample variance achieved when considering matter-densities and velocities 
-within the same volume \citep{2007PhRvL..99h1301G}.
+within the same volume \cite{2007PhRvL..99h1301G}.
 
@@
 supernova rate of each host is determined using 
-\citet{2012ApJ...755...61S}.   This host-based rate
+\cite{2012ApJ...755...61S}.   This host-based rate
 underestimates the expected discovery based on the volumetric
-rate of \citet{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
+rate of \cite{2010ApJ...713.1026D}.  The volumetric rate is more robust, so  we boost the SN production
 by a factor 2.4 to get the 10-year volumetric rate expectation. The total number of simulated SNe 
@@
 motion with respect to the CMB and of the galaxies with respect to the CMB  are ignored: referring to
-\citet{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
+\cite{2011ApJ...741...67D}, the effects in Eq.~18 are not included.  Including these effects are straightforward but is computationally expensive
 and does not affect our conclusions.
@@
 \section{Analysis}
-The analysis is  almost identical to that of \citet{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
+The analysis is  almost identical to that of \cite{2015JCAP...12..033H, 2017JCAP...05..015H}, except as noted later.  
 The peculiar magnitude is given by 
@@
 The peculiar magnitude correlation
-function $\xi_{\delta m \delta m}$ \citep{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
-matter density power spectrum is calculated using CAMB \citep{Lewis:2002ah}  assuming the same
+function $\xi_{\delta m \delta m}$ \cite{2011ApJ...741...67D,2015JCAP...12..033H} expected from General Relativity acting on the CMB
+matter density power spectrum is calculated using CAMB \cite{Lewis:2002ah}  assuming the same
 cosmological parameters used for the cosmoDC2 catalog.  Our model for the data covariance
@@
 where $\bar{z}$ is the cosmological redshift of the host galaxy.
-We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \citep{2015MNRAS.454.3920H}. 
-Such subtleties are unimportant as \citet{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
+We note that the CAMB calculation of $\xi$ includes non-linear corrections, which actually may be a bad thing for peculiar velocities \cite{2015MNRAS.454.3920H}. 
+Such subtleties are unimportant as \cite{2015JCAP...12..033H} find that their results are insensitive to whether non-linear corrections are or are not included.
 
@@
 The fit posterior for one representative subsample, with input  $z_{\rm max}=0.2$, $\sigma_M=0.08$, $\phi=0.65$, is shown in Figure~\ref{zmax:fig}.
-For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \citet{2015JCAP...12..033H, 2017JCAP...05..015H},
+For all subsamples, there is no evidence for non-convergence of the MCMC chains.  The parameter $\sigma_v$, not included in  \cite{2015JCAP...12..033H, 2017JCAP...05..015H},
 is found to be consistent with zero, poorly constrained, and very weakly correlated with $A$.  Table~\ref{tab:subsets} gives the results for the different subsamples.
@@
 %
-%\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
+%\cite{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
 %They perform their analysis in Fourier space, 
@@
 %is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
-%\citep{2017MNRAS.464.2517H}.  
+%\cite{2017MNRAS.464.2517H}.  
 %The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 

### 2019-03-08 doc/article.tex
clean up to fit page limit

@@
 %\usepackage{geometry}
+\usepackage{titlesec}
 
+\titleformat*{\section}{\large\bfseries}
 %\usepackage[overload]{textcase}
@@
 \footnotetext[6]{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
-Simone Ferraro\footnotemark[1],
+S.~Ferraro\footnotemark[1],
 S.~Gontcho A Gontcho\footnotemark[4],
@@
     1 Cyclotron Road, Berkeley, CA, 94720}
-S.~Perlmutter\footnotemark[1]\textsuperscript{,}\footnotemark[20],
-\footnotetext[20]{Department of Physics, University of California Berkeley, Berkeley, CA 94720}
+S.~Perlmutter\footnotemark[1]\textsuperscript{,}\footnotemark[17],
+\footnotetext[17]{Department of Physics, University of California Berkeley, Berkeley, CA 94720}
 N.~Regnault\footnotemark[14],
@@
 %Lyon, France} 
-A.~Shafieloo\footnotemark[17],
-\footnotetext[17]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
-A.~Slo\u{z}ar\footnotemark[18],
-\footnotetext[18]{Brookhaven National Laboratory, Physics Department, Upton, NY
+A.~Shafieloo\footnotemark[18],
+\footnotetext[18]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
+A.~Slo\u{z}ar\footnotemark[19],
+\footnotetext[19]{Brookhaven National Laboratory, Physics Department, Upton, NY
 11973, USA} 
+R.~B.~Tully\footnotemark[20],
+\footnotetext[20]{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}
 M.~White\footnotemark[1],\textsuperscript{,}\footnotemark[20],
-M.~Wood-Vasey\footnotemark[19]
-\footnotetext[19]{PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA} 
+M.~Wood-Vasey\footnotemark[21]
+\footnotetext[21]{PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA} 
 \linebreak
@@
 \textbf{Endorsers:}
- Behzad Ansarinejad $^{1,2}$, 
- Robert Armstrong $^{}$, 
- Carlo Baccigalupi $^{3,4,5}$, 
- Maciej Bilicki $^{6}$, 
- Julian Borrill $^{7}$, 
- Simone Ferraro $^{7}$, 
- Llu\'is Galbany $^{8}$, 
- Juan Garc\'ia-Bellido $^{9}$, 
- Juan Garc\'ia-Bellido $^{10}$, 
- Martina Gerbino $^{}$, 
- Mandeep S.S. Gill $^{11,12,13}$, 
- Larry Gladney $^{14}$, 
- Marc Kamionkowski $^{1,19}$, 
- Kazuya Koyama $^{20}$, 
- Michele Liguori $^{21}$, 
- Joel Meyers $^{22}$, 
- "Surhud More" $^{23}$, 
- Jeffrey A. Newman $^{8}$, 
- Gustavo Niz $^{24}$, 
- "Antonella Palmese" $^{25}$, 
- Francesco Piacentini $^{26}$, 
- Francesco Piacentni  $^{26,27}$, 
- Steven Rodney $^{28}$, 
- Benjamin Rose $^{29}$, 
- Lado Samushia $^{30}$, 
- Neelima Sehgal $^{31}$, 
- Leonardo Senatore $^{11}$, 
- Huanyuan Shan $^{1,33}$, 
- Scott Watson $^{35}$, 
- Martin White $^{36,7}$, 
- Weishuang Xu $^{37}$, 
- Gong-Bo Zhao $^{38,39,20}$
  \linebreak
@@
 is sensitive to the growth index $\gamma$, which captures the effect of gravity on the linear growth of structure through the relation $f=\Omega_M^\gamma$.
+We present the first  projections for the precision in $\gamma$ for a range of realistic SN peculiar-velocity survey scenarios.
 In the next decade the peculiar velocities of
@@
 %\end{abstract}
-%\section{Introduction}
-
-{\bf Introduction:} In the late 1990's, Type~Ia supernovae (SNe~Ia) were used as distance probes to measure the homogeneous expansion history of the Universe.  The remarkable discovery
+\section{Introduction}
+In the late 1990's, Type~Ia supernovae (SNe~Ia) were used as distance probes to measure the homogeneous expansion history of the Universe.  The remarkable discovery
 that the expansion is accelerating  has called into question our basic understanding of the gravitational forces within the Universe.  Either it
@@
 In the next decade, SNe~Ia will be used as peculiar-velocity probes to measure  the influence of gravity on structure formation within the Universe.
-The SN~Ia Hubble diagram has  scatter along the redshift axis since peculiar velocities induce differing observed and cosmological redshifts.  The contribution
-of this scatter is pronounced at low redshifts and when the scatter along the magnitude axis (i.e.\ intrinsic magnitude dispersion) is small.
+Peculiar velocities induce scatter along the redshift axis of the SN Hubble diagram, which is
+pronounced at low redshifts and when the magnitude scatter (e.g.\ due to intrinsic magnitude dispersion) is small.
 The peculiar velocity power spectrum is sensitive to the growth of structure as $P_{vv}\propto (fD)^2$, where $D$ is  the spatially-independent
-growth factor for the linear evolution of density perturbations and
-$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate with $a$ the scale factor  \cite{2006PhRvD..73l3526H,2011ApJ...741...67D}.
+``growth factor'' in the linear evolution of density perturbations and
+$f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate where $a$ is the scale factor  \cite{2006PhRvD..73l3526H,2011ApJ...741...67D}.
 %\footnote{
@@
 %}
-The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity, indeed 
+The $\Lambda$CDM prediction for the $z=0$ peculiar velocity power spectrum is shown in Figure~\ref{power:fig}. The growth of structure depends on gravity;
 \cite{2007APh....28..481L} find that General Relativity, $f(R)$,  and DGP gravity follow the relation
-$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively \cite[see][for a review of
-these gravity models within a cosmological context]{HUTERER201523}.  Adopting this parameterization to model gravity, peculiar velocity
-surveys are  sensitive to $fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$, whose prediction for $\Lambda$CDM  is plotted 
-in Figure~2 of  \cite{1475-7516-2013-04-031}. (Note that we prefer the notation $fD$ in place of $f\sigma_8$, which is commonly used in the literature.)
+$f \approx \Omega_M^\gamma$ with $\gamma=0.55, 0.42, 0.68$ respectively (see \cite{HUTERER201523} for a review
+or these  models).  Using this parameterization to model gravity, peculiar velocity
+surveys probe $\gamma$ through $fD$, whose $\gamma$-dependence is plotted 
+in Figure~2 of  \cite{1475-7516-2013-04-031}.
+%$fD=\Omega_M^{\gamma} \exp{\left(\int_a^1 \Omega_M^{\gamma} d\ln{a} \right)}$
 %
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.47\textwidth]{src/new2.png}
+\includegraphics[width=0.49\textwidth]{src/new2.png}
 %\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
-\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu=\cos{(\hat{k} \cdot \hat{r})}=1, 0.5$ (magenta, cyan) 
-($\hat{r}$ is the line of sight), as predicted for General Relativity in the linear regime.
-Overplotted are volume-weighted peculiar-velocity power-spectrum shot noise  (diagonal lines) for various observing parameters.  Red shows the shot noise expected from a 2-year LSST survey
+\caption{Volume-weighted peculiar velocity power spectrum $k^3P_{vv}(z=0)$ for $\mu \equiv \cos{(\hat{k} \cdot \hat{r})}=1, 0.5$ (magenta, cyan) 
+where $\hat{r}$ is the line of sight, as predicted for General Relativity in the linear regime.
+Overplotted are peculiar-velocity power-spectrum shot noise  (diagonal lines) for various observing parameters.  Red shows the shot noise expected from a 2-year LSST survey
 while black shows a 10-year LSST survey.  The dotted and dashed lines indicate the assumed intrinsic magnitude dispersion, using 0.08 (dashed) or 0.15 mag (dotted).  The expected shot
@@
 
-Peculiar velocity surveys have already been  used to measure $fD$, though not to a level where  gravity models can be precisely distinguished.
- \cite{2017MNRAS.471..839A} use 6dFGS peculiar velocities using the Fundamental Plane distances of elliptical galaxies, to estimate absolute magnitudes
+Peculiar velocity surveys have already been  used to measure $fD$ (also referred to as $f\sigma_8$), though not to a level where  gravity models can be precisely distinguished.
+ \cite{2017MNRAS.471..839A} use 6dFGS peculiar velocities using  Fundamental Plane distances of elliptical galaxies to estimate absolute magnitudes
  with
- $\sim 0.43$~mag  precision to
-get a 15\% uncertainty in $fD$ at $z\approx 0$ using combined density and velocity information.
+ $\sim 0.43$~mag  precision, yielding a 15\% uncertainty in $fD$ at $z\approx 0$.
 The upcoming 
@@
 $z=0-0.1$ covering 75\% of the sky.
-These surveys combined are projected to have 3\% $fD$ uncertainties \cite{2017MNRAS.464.2517H}.
-For reference, DESI projects 10\% precision of $fD$ at $z \approx 0.3$  by statistically looking 
-for expected signatures of galaxies infalling toward mass overdensities, Redshift Space Distortions (RSD).
+These surveys combined are projected to have 3\% uncertainties in $fD$ \cite{2017MNRAS.464.2517H}.
+For reference, DESI projects a 10\% precision of $fD$ at $z \approx 0.3$  by looking 
+for signatures (Redshift Space Distortions; RSD) expected from galaxies infalling toward mass overdensities.
 Relative to galaxies with  Fundamental Plane or Tully-Fisher distances, 
@@
 \cite{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
-SNe~Ia discovered by ZTF over the next several years will provide first probative measures of $fD$ at $z<0.1$.
+SNe~Ia discovered by ZTF \cite{2019PASP..131a8002B} over the next several years will provide first probative measures of $fD$ at $z<0.1$.
 %Measurement of the velocity field using LSST-discovered
@@
 established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
-that is observed using current calibrations.  If corrections of all SNe are not possible, the use of SN~Ia subclasses is an option though at the expense of reducing the sample.
-At this precision the intrinsic velocity dispersion  at $z=0.028$ is  of $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
-provide a peculiar velocity with $S/N \sim 1$.
+that is observed using current calibrations.  At this precision the intrinsic velocity dispersion  at $z=0.028$ is  of $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
+measure a peculiar velocity with $S/N \sim 1$.
+ If corrections of all SNe~Ia are not possible, the use of SN~Ia subclasses is an option though at the expense of reducing the
+numbers of velocity probes.
 Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
-  will increase the number of identified  $z<0.3$ Type~Ia supernovae (SNe~Ia)  from the hundreds to the
+  will increase the number of identified  $z<0.3$ Type~Ia supernovae from the hundreds to the
 hundreds of thousands; over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
@@
  number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.  With similar densities,
- the (two) 10-year SN~Ia survey will have
+ the (two) ten-year SN~Ia survey will have
  a (6) 29$\times$ reduction in shot-noise, $\sigma^2_M/n$, relative to the Fundamental Plane survey of TAIPAN.
 
-Given these two advances, supernovae discovered by wide-field searches in the next decade will 
- be able to stringently constrain the growth of structure in the low-redshift Universe.
+Given these  advances, supernovae discovered by wide-field searches in the next decade will 
+ be able to tightly constrain the growth of structure in the low-redshift Universe.
 For example, 
@@
 2.2\% uncertainty on $fD$  within this interval, where at
-$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from Redshift Space Distortions  \cite{2017ApJ...847..128H}.
+$0<z<0.2$ most of the probative power comes from peculiar velocities and at higher redshifts from RSD  \cite{2017ApJ...847..128H}.
 
-%\section{Testing Gravity with Peculiar Velocity Surveys}
-{\bf Testing Gravity with Peculiar Velocity Surveys:} While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest is in probing gravity,
+\section{Testing Gravity with Peculiar Velocity Surveys}
+While the growth rate $fD$ can be used to test several aspects of physics beyond the standard cosmological model (e.g.\ dark matter clustering, dark energy evolution), our scientific interest is in probing gravity,
 so here we focus  on the growth index $\gamma$.
@@
  with the one at lower redshift providing the tighter constraint. 
- In this section, we focus on how well peculiar velocity surveys in the upcoming decade can measure $\gamma$ and the sensitivity of the measurement
- on survey parameter choices.
+ In this section, we demonstrate that peculiar velocity surveys in the upcoming decade can measure $\gamma$ precisely
+for a range of survey-parameter choices.
 
@@
 using a Fisher matrix analysis similar to that of \cite{2017ApJ...847..128H, 2017MNRAS.464.2517H}
-\cite[there is an alternative approach using an estimator for the mean pairwise velocity][]{2011PhRvD..83d3004B}.
+(there is an alternative approach using an estimator for the mean pairwise velocity
+\cite{2011PhRvD..83d3004B}).
+The ``cross-correlation'' analysis incorporates both galaxy overdensities and peculiar velocities.
 The Fisher information matrix is
@@
 and the parameters considered are $\lambda \in \{\gamma,bD, \Omega_{M_0}\}$.
-The parameter dependence enters through $(fD)(\gamma)$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy count overdensity
+The parameter dependence enters through $fD$ in the relations $P_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy count overdensity
 power spectrum $P_{\delta \delta }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
-\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu$ is the cosine of the angle between the $k$-mode
-and the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
+\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu\equiv \cos{(\hat{k} \cdot \hat{r})}$ where $\hat{r}$ is the direction of
+the line of sight.  While the $bD$ term does contain information on $\gamma$, its constraining power is not used here.
 Both $f$ and $D$ depend on $\Omega_M=\frac{\Omega_{M_0}}{\Omega_{M_0} + (1-\Omega_{M_0})a^3}$.
@@
 through the resulting peculiar velocity intrinsic dispersion $\sigma \approx (\frac{5}{\ln{10}} \frac{1+z}{z})^{-1} \sigma_M$.
-An analysis in which the RSD and velocity surveys are treated independently 
-sets the off-diagonal elements of $C$ to zero.
+%An analysis in which the RSD and velocity surveys are treated independently 
+%sets the off-diagonal elements of $C$ to zero.
 
-We consider SN peculiar velocity surveys for a range of redshift depths $z_{\rm max}$ for survey durations of $t=2$ and 10~years.
+We consider SN peculiar velocity surveys for a range of redshift depths $z_{\rm max}$ for durations of $t=2$ and 10~years.
 The other survey parameters
@@
 $k_{\rm max} = 0.1$~$h$\,Mpc$^{-1}$.
-A minimum distance $r_{\rm min}=r(z=0.01)$ is imposed as the physics derivation is taken at the limit
-where peculiar velocities are significantly  smaller than
+A minimum distance $r_{\rm min}=r(z=0.01)$ is imposed as our analysis assumes that peculiar velocities are significantly  smaller than
 the cosmological redshift.
 The sample-selection efficiency $\epsilon$ is redshift-independent, i.e.\ the native redshift distribution is not sculpted.
-The input bias of SN~Ia host galaxies is set as $b=1.2$.  An independent measurement of $\Omega_{M0}=0.3\pm0.005$
-is included and is important contributor to the $\gamma$ constraint.  Number densities are taken to be
-direction-independent, neglecting the slight declination-dependence of SN-survey time windows. Code used for the analysis in this article available
-at \url{http:tiny.cc/PVScience}.
+The input bias of SN~Ia host galaxies is set as $b=1.2$.  An independent measurement of $\Omega_{M_0}=0.3\pm0.005$
+is included and is a non-trivial contributor to the $\gamma$ constraint.  Number densities are taken to be
+direction-independent, neglecting the slight declination-dependence of SN-survey time windows
 
@@
 The projected uncertainty in $\gamma$ achieved by the suite of surveys are shown in Figure~\ref{var:fig}.
-The primary analysis of interest uses overdensities (RSD), peculiar velocities, and their cross-correlations.
+The primary result is for the cross-correlation analysis that uses overdensities (RSD), peculiar velocities, and their cross-correlations.
 The short and shallow,  2-year, $z_{\rm max}=0.11$ survey has $\sigma_\gamma \sim 0.038$,
@@
 Peculiar velocities alone
-can account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
+account for much of the probative power of the surveys. RSD alone do not provide significant constraints.
 However, considering RSD and velocity cross-correlations decreases $\sigma_\gamma$ by  $\sim 20$\%.
@@
 
-Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
+Survey performance is examined in more detail by considering how $\sigma_\gamma$ in the cross-correlation analysis
 changes with respect to the 
 survey parameters  $\Omega$, $z_{\rm max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
-within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{\rm max}$. \newline
+within a given survey.  Though not directly a survey parameter, we also examine changes with respect to our fiducial choice of  $k_{\rm max}$. \newline
 %
@@
 for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
-to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
+to sculpt the distribution, the preference would be to cut out the highest redshift bins
 resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$. 
@@
 precision in $\gamma$ for all the surveys considered.
-The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution,
-shorter surveys and those with higher $z_{\rm max}$ (because of the higher $\sigma$), benefit more from extending survey duration.
-These trends are shown in the right panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
+The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution (i.e.\ 
+shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration, as shown
+n the right-panel plot of Figure~\ref{dvar:fig} that shows $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
  $z_{\rm max}$ for two- and ten-year surveys.
 Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
-has a similar relationship with $\sigma_\gamma$.
-This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$, which is shown in the left panel of Figure~\ref{dvar:fig}.\newline
+has a similar relationship with $\sigma_\gamma$ as shown
+in the left-panel plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$  of Figure~\ref{dvar:fig}.\newline
 {\it Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.
@@
 \end{comment}
-% \section{Conclusions}
+\section{Conclusions}
 
 %SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  
-{\bf Conclusions:} In the next decade,
-the high number of SN discoveries together with improved precision in their distance precisions will make $z<0.2$ SNe~Ia, more so than
+In the next decade,
+the high number of SN discoveries together with improved precision in their distance precisions will make $z<0.3$ SNe~Ia, more so than
 galaxies,  powerful probes of gravity through their effect  on the growth of structure.  Different survey strategies can be adopted to take advantage of these
-supernovae, we provide a formalism and code by which their scientific merits can be assessed and here sketch results for a range of options.
+supernovae, and in this White Paper we present a formalism and code (available
+at \url{http:tiny.cc/PVScience})
+by which their scientific merits can be assessed and present results for a range of options.
 %A peculiar velocity survey must discover, classify, and get distances of SNe~Ia, as well as obtain redshifts of their host galaxies.
@@
 TAIPAN and a TAIPAN-like DESI BGS will be able to measure FP distances for nearly all usable nearby galaxies, so at low-$z$ the Fundamental Plane peculiar-velocity
-technique will  saturate at a level that is not competitive with a modes 2-year SN survey.
+technique will  saturate at a level that is not competitive with a  2-year SN survey.
 

### 2019-03-08 doc/article.tex
spelling errors

@@
 \footnotetext[18]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
-A.~Slo\u{z}ar\footnotemark[19],
+A.~Slo\v{z}ar\footnotemark[19],
 \footnotetext[19]{Brookhaven National Laboratory, Physics Department, Upton, NY
@@
 \textbf{Endorsers:}
+Behzad Ansarinejad$^{1}$, 
+Robert Armstrong$^{2}$, 
+Jacobo Asorey$^{3}$, 
+Carlo Baccigalupi$^{4,5,6}$, 
+Maciej Bilicki$^{7}$, 
+Julian Borrill$^{8}$, 
+Elizabeth Buckley-Geer$^{9}$, 
+Tamara M Davis$^{10}$, 
+Kelly A. Douglass$^{11}$, 
+Cora Dvorkin$^{12}$, 
+Simone Ferraro$^{8}$, 
+Simon Foreman$^{13}$, 
+Llu\'is Galbany$^{14}$, 
+Juan Garc\'ia-Bellido$^{15}$, 
+Juan Garc\'ia-Bellido$^{16}$, 
+Martina Gerbino$^{17}$, 
+Mandeep S.S. Gill$^{18,19,20}$, 
+Larry Gladney$^{21}$, 
+Satya {Gontcho A Gontcho}$^{11}$, 
+Julien Guy$^{8}$, 
+Ren\'ee Hlo\v{z}ek$^{22,23}$, 
+Dragan Huterer$^{24}$, 
+Saurabh W.~Jha$^{25}$, 
+Johann Cohen-Tanugi$^{26}$, 
+D.\ O.\ Jones$^{27}$, 
+Marc Kamionkowski$^{28}$, 
+Ryan E. Keeley$^{3}$, 
+Robert Kehoe$^{29}$, 
+Savvas M. Koushiappas$^{30}$, 
+Ely D.~Kovetz$^{31}$, 
+Kazuya Koyama$^{32}$, 
+Benjamin L'Huillier$^{3}$, 
+Ofer Lahav$^{33}$, 
+Chien-Hsiu Lee$^{34}$, 
+Michele Liguori$^{35}$, 
+Axel de la Macorra$^{36}$, 
+Patrick McDonald$^{8}$, 
+Joel Meyers$^{29}$, 
+Mickael Rigault$^{37}$, 
+Surhud More$^{38}$, 
+Jeffrey A. Newman$^{14}$, 
+Gustavo Niz$^{39}$, 
+Antonella Palmese$^{9}$, 
+Francesco Piacentini$^{40}$, 
+Francesco Piacentni$^{40,41}$, 
+Steven Rodney$^{42}$, 
+Benjamin Rose$^{43}$, 
+Masao Sako$^{44}$, 
+Lado Samushia$^{45}$, 
+Neelima Sehgal$^{46}$, 
+Leonardo Senatore$^{18}$, 
+Arman Shafieloo$^{3}$, 
+Sara Simon$^{24}$, 
+An\v{z}e Slosar$^{47}$, 
+Stephanie Escoffier$^{48}$, 
+Scott Watson$^{49}$, 
+Martin White$^{50,8}$, 
+Weishuang Xu$^{12}$, 
+Gong-Bo Zhao$^{51,32}$, 
+Yi Zheng$^{52}$, 
 \linebreak
@@
 %%% 
-%\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
-% $^{1}$ \ED \\
-%$^{2}$ \Durham \\
-%$^{3}$ \SISSA \\
-%$^{4}$ \IFPU \\
-%$^{5}$ \INFN \\
-%$^{6}$ \CFT \\
-%$^{7}$ \LBL \\
-%$^{8}$ \Pitt \\
-%$^{9}$ \IFT \\
-%$^{10}$ \UAM \\
-%$^{11}$ \KIPAC \\
-%$^{12}$ \Stanford \\
-%$^{13}$ \SLAC \\
-%$^{14}$ \Yale \\
-%$^{15}$ \UR \\
-%$^{16}$ \dunlap \\
-%$^{17}$ \daa \\
-%$^{18}$ \Umich \\
-%$^{19}$ \JHU \\
-%$^{20}$ \Port \\
-%$^{21}$ \UNIPD \\
-%$^{22}$ \SMU \\
-%$^{23}$ \IUCAA \\
-%$^{24}$ \UGTO \\
-%$^{25}$ \FNAL \\
-%$^{26}$ \RomaS \\
-%$^{27}$ \INFNRM \\
-%$^{28}$ \USC \\
-%$^{29}$ \STSCI \\
-%$^{30}$ \KSU \\
-%$^{31}$ \StonyBrook \\
-%$^{32}$ \KASSI \\
-%$^{33}$ \SHAO\\
-%$^{34}$ \BNL \\
-%$^{35}$ \Syracuse \\
-%$^{36}$ \UCB \\
-%$^{37}$ \HarvardPhys \\
-%$^{38}$ \NAOC \\
-%$^{39}$ \UCAS \\
-%
+\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
+ $^{1}$ \Durham \\
+$^{2}$ \LLNL \\
+$^{3}$ \KASSI \\
+$^{4}$ \SISSA \\
+$^{5}$ \IFPU \\
+$^{6}$ \INFN \\
+$^{7}$ \CFT \\
+$^{8}$ \LBL \\
+$^{9}$ \FNAL \\
+$^{10}$ \Queensland \\
+$^{11}$ \UR \\
+$^{12}$ \HarvardPhys \\
+$^{13}$ \CITA \\
+$^{14}$ \Pitt \\
+$^{15}$ \IFT \\
+$^{16}$ \UAM \\
+$^{17}$ \ANLHEP \\
+$^{18}$ \KIPAC \\
+$^{19}$ \Stanford \\
+$^{20}$ \SLAC \\
+$^{21}$ \Yale \\
+$^{22}$ \dunlap \\
+$^{23}$ \daa \\
+$^{24}$ \Umich \\
+$^{25}$ \Rutgers \\
+$^{26}$ \LUPM \\
+$^{27}$ \UCSC \\
+$^{28}$ \JHU \\
+$^{29}$ \SMU \\
+$^{30}$ \Brown \\
+$^{31}$ \BenGurion \\
+$^{32}$ \Port \\
+$^{33}$ \UCL \\
+$^{34}$ \NOAO \\
+$^{35}$ \UNIPD \\
+$^{36}$ \IFUNAM \\
+$^{37}$ \LPC \\
+$^{38}$ \IUCAA \\
+$^{39}$ \UGTO \\
+$^{40}$ \RomaS \\
+$^{41}$ \INFNRM \\
+$^{42}$ \USC \\
+$^{43}$ \STSCI \\
+$^{44}$ \UPenn \\
+$^{45}$ \KSU \\
+$^{46}$ \StonyBrook \\
+$^{47}$ \BNL \\
+$^{48}$ \CPPM \\
+$^{49}$ \Syracuse \\
+$^{50}$ \UCB \\
+$^{51}$ \NAOC \\
+$^{52}$ \KIAS \\
+
 \textbf{Abstract:}
@@
 \includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
-\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{\rm max}=0.2$. 
+\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survey with limiting  depth $z_{\rm max}=0.2$. 
 Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
@@
 
-Combined low-redshift peculiar velocity and high-redshift RSD $fD$ measurements are highly complimentary as together they probe the
+Combined low-redshift peculiar velocity and high-redshift RSD $fD$ measurements are highly complementary as together they probe the
 $\gamma$-dependent shape of $fD(z)$ (not just its normalization) and potential scale-dependent influence of gravitational models, since low-

### 2019-03-10 doc/article.tex
anze fix
last comments from Greg and affiliations

--- 2019-03-08 doc/article.tex
@@
 Arbor, MI 48109, USA }
-C.~Ju\footnotemark[12],
+C.~Ju\footnotemark[1],
 %\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
@@
 %\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
-M.~Rigault\footnotemark[5],
+M.~Rigault\footnotemark[10],
 %\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
@@
 \footnotetext[18]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
-A.~Slo\u{z}ar\footnotemark[19],
+A.~Slosar\footnotemark[19],
 \footnotetext[19]{Brookhaven National Laboratory, Physics Department, Upton, NY
@@
 
+\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
 \textbf{Endorsers:}
- \linebreak
+Behzad Ansarinejad$^{1}$, 
+Robert Armstrong$^{2}$, 
+Jacobo Asorey$^{3}$, 
+Carlo Baccigalupi$^{4,5,6}$, 
+Maciej Bilicki$^{7}$, 
+Julian Borrill$^{8}$, 
+Elizabeth Buckley-Geer$^{9}$, 
+Kelly A. Douglass$^{11}$, 
+Cora Dvorkin$^{12}$, 
+Simon Foreman$^{13}$, 
+Llu\'is Galbany$^{14}$, 
+Juan Garc\'ia-Bellido$^{15,16}$, 
+Martina Gerbino$^{17}$, 
+Mandeep S.S. Gill$^{18,19,20}$, 
+Larry Gladney$^{21}$, 
+Saurabh W.~Jha$^{25}$, 
+Johann Cohen-Tanugi$^{26}$, 
+D.\ O.\ Jones$^{27}$, 
+Marc Kamionkowski$^{28}$, 
+Ryan E. Keeley$^{3}$, 
+Robert Kehoe$^{29}$, 
+Savvas M. Koushiappas$^{30}$, 
+Ely D.~Kovetz$^{31}$, 
+Kazuya Koyama$^{32}$, 
+Benjamin L'Huillier$^{3}$, 
+Ofer Lahav$^{33}$, 
+Chien-Hsiu Lee$^{34}$, 
+Michele Liguori$^{35}$, 
+Axel de la Macorra$^{36}$, 
+Joel Meyers$^{29}$, 
+Surhud More$^{38}$, 
+Jeffrey A. Newman$^{14}$, 
+Gustavo Niz$^{39}$, 
+Antonella Palmese$^{9}$, 
+Francesco Piacentini$^{40,41}$, 
+Steven Rodney$^{42}$, 
+Benjamin Rose$^{43}$, 
+Masao Sako$^{44}$, 
+Lado Samushia$^{45}$, 
+Neelima Sehgal$^{46}$, 
+Leonardo Senatore$^{18}$, 
+Sara Simon$^{24}$, 
+Stephanie Escoffier$^{48}$, 
+Scott Watson$^{49}$, 
+Weishuang Xu$^{12}$, 
+Gong-Bo Zhao$^{51,32}$, 
+Yi Zheng$^{52}$,
 
 %%%
-%\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
-% $^{1}$ \ED \\
-%$^{2}$ \Durham \\
-%$^{3}$ \SISSA \\
-%$^{4}$ \IFPU \\
-%$^{5}$ \INFN \\
-%$^{6}$ \CFT \\
-%$^{7}$ \LBL \\
-%$^{8}$ \Pitt \\
-%$^{9}$ \IFT \\
-%$^{10}$ \UAM \\
-%$^{11}$ \KIPAC \\
-%$^{12}$ \Stanford \\
-%$^{13}$ \SLAC \\
-%$^{14}$ \Yale \\
-%$^{15}$ \UR \\
-%$^{16}$ \dunlap \\
-%$^{17}$ \daa \\
-%$^{18}$ \Umich \\
-%$^{19}$ \JHU \\
-%$^{20}$ \Port \\
-%$^{21}$ \UNIPD \\
-%$^{22}$ \SMU \\
-%$^{23}$ \IUCAA \\
-%$^{24}$ \UGTO \\
-%$^{25}$ \FNAL \\
-%$^{26}$ \RomaS \\
-%$^{27}$ \INFNRM \\
-%$^{28}$ \USC \\
-%$^{29}$ \STSCI \\
-%$^{30}$ \KSU \\
-%$^{31}$ \StonyBrook \\
-%$^{32}$ \KASSI \\
-%$^{33}$ \SHAO\\
-%$^{34}$ \BNL \\
-%$^{35}$ \Syracuse \\
-%$^{36}$ \UCB \\
-%$^{37}$ \HarvardPhys \\
-%$^{38}$ \NAOC \\
-%$^{39}$ \UCAS \\
-%
+
+ $^{1}$ \Durham \\
+$^{2}$ \LLNL \\
+$^{3}$ \KASSI \\
+$^{4}$ \SISSA \\
+$^{5}$ \IFPU \\
+$^{6}$ \INFN \\
+$^{7}$ \CFT \\
+$^{8}$ \LBL \\
+$^{9}$ \FNAL \\
+$^{10}$ \Queensland \\
+$^{11}$ \UR \\
+$^{12}$ \HarvardPhys \\
+$^{13}$ \CITA \\
+$^{14}$ \Pitt \\
+$^{15}$ \IFT \\
+$^{16}$ \UAM \\
+$^{17}$ \ANLHEP \\
+$^{18}$ \KIPAC \\
+$^{19}$ \Stanford \\
+$^{20}$ \SLAC \\
+$^{21}$ \Yale \\
+$^{22}$ \dunlap \\
+$^{23}$ \daa \\
+$^{24}$ \UMich \\
+$^{25}$ \Rutgers \\
+$^{26}$ \LUPM \\
+$^{27}$ \UCSC \\
+$^{28}$ \JHU \\
+$^{29}$ \SMU \\
+$^{30}$ \Brown \\
+$^{31}$ \BenGurion \\
+$^{32}$ \Port \\
+$^{33}$ \UCL \\
+$^{34}$ \NOAO \\
+$^{35}$ \UNIPD \\
+$^{36}$ \IFUNAM \\
+$^{37}$ \LPC \\
+$^{38}$ \IUCAA \\
+$^{39}$ \UGTO \\
+$^{40}$ \RomaS \\
+$^{41}$ \INFNRM \\
+$^{42}$ \USC \\
+$^{43}$ \STSCI \\
+$^{44}$ \UPenn \\
+$^{45}$ \KSU \\
+$^{46}$ \StonyBrook \\
+$^{47}$ \BNL \\
+$^{48}$ \CPPM \\
+$^{49}$ \Syracuse \\
+$^{50}$ \UCB \\
+$^{51}$ \NAOC \\
+$^{52}$ \KIAS \\
+
 \textbf{Abstract:}
@@
 established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
-that is observed using current calibrations.  At this precision the intrinsic velocity dispersion  at $z=0.028$ is  of $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
+that is observed using current calibrations.  At this precision the intrinsic velocity dispersion  at $z=0.028$ is  $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
 measure a peculiar velocity with $S/N \sim 1$.
@@
 The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution (i.e.\ 
-shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration, as shown
-n the right-panel plot of Figure~\ref{dvar:fig} that shows $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
- $z_{\rm max}$ for two- and ten-year surveys.
+shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration,  shown as  $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
+ $z_{\rm max}$ for two- and ten-year surveys
+in the right-panel plot of Figure~\ref{dvar:fig}.
 Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus

### 2019-03-11 doc/article.tex
affiliates done

--- 2019-03-08 doc/article.tex
@@
 \footnotetext[20]{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}
-M.~White\footnotemark[1],\textsuperscript{,}\footnotemark[20],
+M.~White\footnotemark[1]\textsuperscript{,}\footnotemark[17],
 M.~Wood-Vasey\footnotemark[21]
@@
 
+\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
 \textbf{Endorsers:}
- \linebreak
+Behzad Ansarinejad
+\footnotemark[22],
+\footnotetext[22]{Department of Physics, Lower Mountjoy, South Rd, Durham DH1 3LE, United Kingdom} 
+Robert Armstrong\footnotemark[23],
+\footnotetext[23]{Lawrence Livermore National Laboratory, Livermore, CA, 94550} 
+Jacobo Asorey\footnotemark[18],
+Carlo Baccigalupi\footnotemark[24]\textsuperscript{,}\footnotemark[25]\textsuperscript{,}\footnotemark[26],
+\footnotetext[24]{SISSA - International School for Advanced Studies, Via Bonomea 265, 34136 Trieste, Italy} 
+\footnotetext[25]{IFPU - Institute for Fundamental Physics of the Universe, Via Beirut 2, 34014 Trieste, Italy} 
+\footnotetext[26]{INFN - National Institute for Nuclear Physics, Via Valerio 2, I-34127 Trieste, Italy} 
+Maciej Bilicki\footnotemark[27],
+\footnotetext[27]{Center for Theoretical Physics, Polish Academy of Sciences, al. Lotnik\'{o}w 32/46, 02-668, Warsaw, Poland} 
+Julian Borrill\footnotemark[16], 
+Elizabeth Buckley-Geer\footnotemark[28],
+\footnotetext[28]{Fermi National Accelerator Laboratory, Batavia, IL 60510} 
+Kelly A. Douglass\footnotemark[4],
+Cora Dvorkin\footnotemark[29],
+\footnotetext[29]{Department of Physics, Harvard University, Cambridge, MA 02138, USA} 
+Simon Foreman\footnotemark[30],
+\footnotetext[30]{Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5S 3H8, Canada} 
+Llu\'is Galbany\footnotemark[21],
+Juan Garc\'ia-Bellido\footnotemark[31]\textsuperscript{,}\footnotemark[32],
+\footnotetext[31]{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain} 
+\footnotetext[32]{Universidad Aut\'{o}noma de Madrid, 28049, Madrid, Spain} 
+Martina Gerbino\footnotemark[33],
+\footnotetext[33]{HEP Division, Argonne National Laboratory, Lemont, IL 60439, USA} 
+Mandeep S.S. Gill\footnotemark[34]\textsuperscript{,}\footnotemark[35]\textsuperscript{,}\footnotemark[36],
+\footnotetext[34]{Kavli Institute for Particle Astrophysics and Cosmology, Stanford 94305} 
+\footnotetext[35]{Stanford University, Stanford, CA 94305} 
+\footnotetext[36]{SLAC National Accelerator Laboratory, Menlo Park, CA 94025} 
+Larry Gladney\footnotemark[37],
+\footnotetext[37]{Department of Physics, Yale University, New Haven, CT 06520} 
+Saurabh W.~Jha\footnotemark[38],
+\footnotetext[38]{Department of Physics and Astronomy, Rutgers, the State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854, USA} 
+Johann Cohen-Tanugi\footnotemark[39],
+\footnotetext[39]{Laboratoire Univers et Particules de Montpellier, Univ. Montpellier and CNRS, 34090 Montpellier, France} 
+D.\ O.\ Jones\footnotemark[40],
+\footnotetext[40]{University of California at Santa Cruz, Santa Cruz, CA 95064} 
+Marc Kamionkowski\footnotemark[41],
+\footnotetext[41]{Johns Hopkins University, Baltimore, MD 21218} 
+Ryan E. Keeley\footnotemark[18],
+Robert Kehoe\footnotemark[42],
+\footnotetext[42]{Southern Methodist University, Dallas, TX 75275} 
+Savvas M. Koushiappas\footnotemark[43],
+\footnotetext[43]{Brown University, Providence, RI 02912} 
+Ely D.~Kovetz\footnotemark[44],
+\footnotetext[44]{Department of Physics, Ben-Gurion University, Be'er Sheva 84105, Israel} 
+Kazuya Koyama\footnotemark[45],
+\footnotetext[45]{Institute of Cosmology \& Gravitation, University of Portsmouth, Dennis Sciama Building, Burnaby Road, Portsmouth PO1 3FX, UK} 
+Benjamin L'Huillier\footnotemark[18],
+Ofer Lahav\footnotemark[46],
+\footnotetext[46]{University College London, WC1E 6BT London, United Kingdom} 
+Chien-Hsiu Lee\footnotemark[47],
+\footnotetext[47]{National Optical Astronomy Observatory, 950 N. Cherry Ave., Tucson, AZ 85719 USA} 
+Michele Liguori\footnotemark[48],
+\footnotetext[48]{Dipartimento di Fisica e Astronomia ``G. Galilei'',Universit\`a degli Studi di Padova, via Marzolo 8, I-35131, Padova, Italy} 
+Axel de la Macorra\footnotemark[49],
+\footnotetext[49]{IFUNAM - Instituto de F\'{i}sica, Universidad Nacional Aut\'onoma de M\'exico, 04510 CDMX, M\'exico} 
+Joel Meyers\footnotemark[42],
+Surhud More\footnotemark[50],
+\footnotetext[50]{The Inter-University Centre for Astronomy and Astrophysics, Pune, 411007, India} 
+Jeffrey A. Newman\footnotemark[21],
+Gustavo Niz\footnotemark[51],
+\footnotetext[51]{Divisi\'{o}n de Ciencias e Ingenier\'{i}as, Universidad de Guanajuato, Le\'{o}n 37150, M\'{e}xico} 
+Antonella Palmese\footnotemark[28],
+Francesco Piacentini\footnotemark[52]\textsuperscript{,}\footnotemark[53],
+\footnotetext[52]{Dipartimento di Fisica, Universit\`{a} La Sapienza, P. le A. Moro 2, Roma, Italy} 
+\footnotetext[53]{Istituto Nazionale di Fisica Nucleare, Sezione di Roma, 00185 Roma, Italy} 
+Steven Rodney\footnotemark[54],
+\footnotetext[54]{The University of South Carolina, Columbia, SC 29208} 
+Benjamin Rose\footnotemark[55],
+\footnotetext[55]{Space Telescope Science Institute, Baltimore, MD 21218} 
+Masao Sako\footnotemark[56],
+\footnotetext[56]{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, Pennsylvania 19104, USA} 
+Lado Samushia\footnotemark[57],
+\footnotetext[57]{Kansas State University, Manhattan, KS 66506} 
+Neelima Sehgal\footnotemark[58],
+\footnotetext[58]{Stony Brook University, Stony Brook, NY 11794} 
+Leonardo Senatore\footnotemark[34],
+Sara Simon\footnotemark[13],
+Stephanie Escoffier\footnotemark[59],
+\footnotetext[59]{Aix Marseille Univ, CNRS/IN2P3, CPPM, Marseille, France} 
+Scott Watson\footnotemark[60],
+\footnotetext[60]{Syracuse University, Syracuse, NY 13244} 
+Weishuang Xu\footnotemark[29],
+Gong-Bo Zhao\footnotemark[61]\textsuperscript{,}\footnotemark[45],
+\footnotetext[61]{National Astronomical Observatories, Chinese Academy of Sciences, PR China} 
+Yi Zheng\footnotemark[62]
+\footnotetext[62]{School of Physics, Korea Institute for Advanced Study, 85 Hoegiro, Dongdaemun-gu, Seoul 130-722, Korea}
 
 %%%
-%\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
-% $^{1}$ \ED \\
-%$^{2}$ \Durham \\
-%$^{3}$ \SISSA \\
-%$^{4}$ \IFPU \\
-%$^{5}$ \INFN \\
-%$^{6}$ \CFT \\
-%$^{7}$ \LBL \\
-%$^{8}$ \Pitt \\
-%$^{9}$ \IFT \\
-%$^{10}$ \UAM \\
-%$^{11}$ \KIPAC \\
-%$^{12}$ \Stanford \\
-%$^{13}$ \SLAC \\
-%$^{14}$ \Yale \\
-%$^{15}$ \UR \\
-%$^{16}$ \dunlap \\
-%$^{17}$ \daa \\
-%$^{18}$ \Umich \\
-%$^{19}$ \JHU \\
-%$^{20}$ \Port \\
-%$^{21}$ \UNIPD \\
-%$^{22}$ \SMU \\
-%$^{23}$ \IUCAA \\
-%$^{24}$ \UGTO \\
-%$^{25}$ \FNAL \\
-%$^{26}$ \RomaS \\
-%$^{27}$ \INFNRM \\
-%$^{28}$ \USC \\
-%$^{29}$ \STSCI \\
-%$^{30}$ \KSU \\
-%$^{31}$ \StonyBrook \\
-%$^{32}$ \KASSI \\
-%$^{33}$ \SHAO\\
-%$^{34}$ \BNL \\
-%$^{35}$ \Syracuse \\
-%$^{36}$ \UCB \\
-%$^{37}$ \HarvardPhys \\
-%$^{38}$ \NAOC \\
-%$^{39}$ \UCAS \\
-%
+
+% $^{1}$ \Durham \\
+%$^{2}$ \LLNL \\
+%$^{3}$ \KASSI \\
+%$^{4}$ \SISSA \\
+%$^{5}$ \IFPU \\
+%$^{6}$ \INFN \\
+%$^{7}$ \CFT \\
+%$^{8}$ \LBL \\
+%$^{9}$ \FNAL \\
+%$^{10}$ \Queensland \\
+%$^{11}$ \UR \\
+%$^{12}$ \HarvardPhys \\
+%$^{13}$ \CITA \\
+%$^{14}$ \Pitt \\
+%$^{15}$ \IFT \\
+%$^{16}$ \UAM \\
+%$^{17}$ \ANLHEP \\
+%$^{18}$ \KIPAC \\
+%$^{19}$ \Stanford \\
+%$^{20}$ \SLAC \\
+%$^{21}$ \Yale \\
+%$^{22}$ \dunlap \\
+%$^{23}$ \daa \\
+%$^{24}$ \UMich \\
+%$^{25}$ \Rutgers \\
+%$^{26}$ \LUPM \\
+%$^{27}$ \UCSC \\
+%$^{28}$ \JHU \\
+%$^{29}$ \SMU \\
+%$^{30}$ \Brown \\
+%$^{31}$ \BenGurion \\
+%$^{32}$ \Port \\
+%$^{33}$ \UCL \\
+%$^{34}$ \NOAO \\
+%$^{35}$ \UNIPD \\
+%$^{36}$ \IFUNAM \\
+%$^{37}$ \LPC \\
+%$^{38}$ \IUCAA \\
+%$^{39}$ \UGTO \\
+%$^{40}$ \RomaS \\
+%$^{41}$ \INFNRM \\
+%$^{42}$ \USC \\
+%$^{43}$ \STSCI \\
+%$^{44}$ \UPenn \\
+%$^{45}$ \KSU \\
+%$^{46}$ \StonyBrook \\
+%$^{47}$ \BNL \\
+%$^{48}$ \CPPM \\
+%$^{49}$ \Syracuse \\
+%$^{50}$ \UCB \\
+%$^{51}$ \NAOC \\
+%$^{52}$ \KIAS \\
+
 \textbf{Abstract:}
@@
 \cite{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
-SNe~Ia discovered by ZTF \cite{2019PASP..131a8002B} over the next several years will provide first probative measures of $fD$ at $z<0.1$.
+SNe~Ia discovered by ASAS-SN, ATLAS, and ZTF \cite{2014ApJ...788...48S,2018PASP..130f4505T,2019PASP..131a8002B} over the next several years will provide first probative measures of $fD$ at $z<0.1$.
 %Measurement of the velocity field using LSST-discovered
@@
 \includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
-\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{\rm max}=0.2$. 
+\caption{Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survey with limiting  depth $z_{\rm max}=0.2$. 
 Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
@@
 The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution (i.e.\ 
-shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration, as shown
-n the right-panel plot of Figure~\ref{dvar:fig} that shows $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
- $z_{\rm max}$ for two- and ten-year surveys.
+shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration.
+% in the right-panel plot of Figure~\ref{dvar:fig}.
 Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
-has a similar relationship with $\sigma_\gamma$ as shown
-in the left-panel plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$  of Figure~\ref{dvar:fig}.\newline
+has a similar relationship with $\sigma_\gamma$; the effect of duration and magnitude dispersion are shown in the right-panel plot of Figure~\ref{dvar:fig} as  $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$
+and $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ 
+as a function of
+ $z_{\rm max}$ for two- and ten-year surveys.\newline
 {\it Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.

### 2019-03-18 doc/article.tex
changes to author list requested
new author list

@@
 \footnotetext[6]{School of Mathematics and Physics, University of Queensland, Brisbane, QLD 4072, Australia}
+H.~Feldman\footnotemark[7],
+\footnotetext[7]{Department of Physics \& Astronomy, University of Kansas, Lawrence, KS 66045-7550 USA}
 S.~Ferraro\footnotemark[1],
@@
 %\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
-O.~Graur\footnotemark[7]\textsuperscript{,}\footnotemark[8]\textsuperscript{,}\footnotemark[9],
-\footnotetext[7]{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138, USA}
-\footnotetext[8]{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
-\footnotetext[9]{NSF Astronomy and Astrophysics Postdoctoral Fellow}
-R.~Graziani\footnotemark[10],
-\footnotetext[10]{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
+O.~Graur\footnotemark[8]\textsuperscript{,}\footnotemark[9]\textsuperscript{,}\footnotemark[10],
+\footnotetext[8]{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138, USA}
+\footnotetext[9]{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
+\footnotetext[10]{NSF Astronomy and Astrophysics Postdoctoral Fellow}
+R.~Graziani\footnotemark[11],
+\footnotetext[11]{Universit\'e Clermont Auvergne, CNRS/IN2P3, Laboratoire de
 Physique de Clermont, F-63000 Clermont-Ferrand, France}
@@
 %    1 Cyclotron Road, Berkeley, CA, 94720}
-R.~Hlo\u{z}ek\footnotemark[3]\textsuperscript{,}\footnotemark[11],
+R.~Hlo\u{z}ek\footnotemark[3]\textsuperscript{,}\footnotemark[12],
 %\affiliation{Department of Astronomy and Astrophysics, University of Toronto,
 %50 St. George Street, Toronto, Ontario, Canada M5S 3H4}
-\footnotetext[11]{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON, M5S 3H4}
-C.~Howlett\footnotemark[12],
-\footnotetext[12]{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
-D.~Huterer\footnotemark[13],
-\footnotetext[13]{Department of Physics, University of Michigan, 450 Church Street, Ann
+\footnotetext[12]{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, ON, M5S 3H4}
+C.~Howlett\footnotemark[13],
+\footnotetext[13]{International Centre for Radio Astronomy Research, The University of Western Australia, Crawley, WA 6009, Australia}
+D.~Huterer\footnotemark[14],
+\footnotetext[14]{Department of Physics, University of Michigan, 450 Church Street, Ann
 Arbor, MI 48109, USA }
@@
 %    1 Cyclotron Road, Berkeley, CA, 94720}
-P.-F.~Leget\footnotemark[14],
-\footnotetext[14]{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
+P.-F.~Leget\footnotemark[15],
+\footnotetext[15]{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
 E.~V.~Linder\footnotemark[1],
@@
 %    1 Cyclotron Road, Berkeley, CA, 94720}
-J.~Nordin\footnotemark[15],
-\footnotetext[15]{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
-P.~Nugent\footnotemark[16],
-\footnotetext[16]{Computational Research Division, Lawrence Berkeley National Laboratory, 
+J.~Nordin\footnotemark[16],
+\footnotetext[16]{Institut fur Physik, Humboldt-Universitat zu Berlin, Newtonstr. 15, 12489 Berlin, Germany}
+P.~Nugent\footnotemark[17],
+\footnotetext[17]{Computational Research Division, Lawrence Berkeley National Laboratory, 
     1 Cyclotron Road, Berkeley, CA, 94720}
-S.~Perlmutter\footnotemark[1]\textsuperscript{,}\footnotemark[17],
-\footnotetext[17]{Department of Physics, University of California Berkeley, Berkeley, CA 94720}
-N.~Regnault\footnotemark[14],
+S.~Perlmutter\footnotemark[1]\textsuperscript{,}\footnotemark[18],
+\footnotetext[18]{Department of Physics, University of California Berkeley, Berkeley, CA 94720}
+N.~Regnault\footnotemark[15],
 %\affiliation{Laboratoire de Physique Nucl\'eaire et de Hautes Energies, Sorbonne Universit\'e, CNRS-IN2P3, 4 Place Jussieu, 75005 Paris, France}    
-M.~Rigault\footnotemark[10],
+M.~Rigault\footnotemark[11],
 %\affiliation{Universit\'e de Lyon, F-69622, Lyon, France; Universit\'e de Lyon
@@
 %Lyon, France} 
-A.~Shafieloo\footnotemark[18],
-\footnotetext[18]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
-A.~Slosar\footnotemark[19],
-\footnotetext[19]{Brookhaven National Laboratory, Physics Department, Upton, NY
+A.~Shafieloo\footnotemark[19],
+\footnotetext[19]{Korea Astronomy and Space Science Institute, Yuseong-gu, Daedeok-daero 776, Daejeon 34055, Korea}
+A.~Slosar\footnotemark[20],
+\footnotetext[20]{Brookhaven National Laboratory, Physics Department, Upton, NY
 11973, USA} 
-R.~B.~Tully\footnotemark[20],
-\footnotetext[20]{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}
-M.~White\footnotemark[1]\textsuperscript{,}\footnotemark[17],
-M.~Wood-Vasey\footnotemark[21]
-\footnotetext[21]{PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA} 
+R.~B.~Tully\footnotemark[21],
+\footnotetext[21]{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}
+L.~Wang\footnotemark[22],
+\footnotetext[22]{Texas A\&M University, College Station, TX 77843}
+M.~White\footnotemark[1]\textsuperscript{,}\footnotemark[18],
+M.~Wood-Vasey\footnotemark[23]
+\footnotetext[23]{PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA} 
 \linebreak
@@
 Behzad Ansarinejad
-\footnotemark[22],
-\footnotetext[22]{Department of Physics, Lower Mountjoy, South Rd, Durham DH1 3LE, United Kingdom} 
-Robert Armstrong\footnotemark[23],
-\footnotetext[23]{Lawrence Livermore National Laboratory, Livermore, CA, 94550} 
-Jacobo Asorey\footnotemark[18],
-Carlo Baccigalupi\footnotemark[24]\textsuperscript{,}\footnotemark[25]\textsuperscript{,}\footnotemark[26],
-\footnotetext[24]{SISSA - International School for Advanced Studies, Via Bonomea 265, 34136 Trieste, Italy} 
-\footnotetext[25]{IFPU - Institute for Fundamental Physics of the Universe, Via Beirut 2, 34014 Trieste, Italy} 
-\footnotetext[26]{INFN - National Institute for Nuclear Physics, Via Valerio 2, I-34127 Trieste, Italy} 
-Maciej Bilicki\footnotemark[27],
-\footnotetext[27]{Center for Theoretical Physics, Polish Academy of Sciences, al. Lotnik\'{o}w 32/46, 02-668, Warsaw, Poland} 
-Julian Borrill\footnotemark[16], 
-Elizabeth Buckley-Geer\footnotemark[28],
-\footnotetext[28]{Fermi National Accelerator Laboratory, Batavia, IL 60510} 
+\footnotemark[24],
+\footnotetext[24]{Department of Physics, Lower Mountjoy, South Rd, Durham DH1 3LE, United Kingdom} 
+Robert Armstrong\footnotemark[25],
+\footnotetext[25]{Lawrence Livermore National Laboratory, Livermore, CA, 94550} 
+Jacobo Asorey\footnotemark[19],
+Carlo Baccigalupi\footnotemark[26]\textsuperscript{,}\footnotemark[27]\textsuperscript{,}\footnotemark[28],
+\footnotetext[26]{SISSA - International School for Advanced Studies, Via Bonomea 265, 34136 Trieste, Italy} 
+\footnotetext[27]{IFPU - Institute for Fundamental Physics of the Universe, Via Beirut 2, 34014 Trieste, Italy} 
+\footnotetext[28]{INFN - National Institute for Nuclear Physics, Via Valerio 2, I-34127 Trieste, Italy} 
+Maciej Bilicki\footnotemark[29],
+\footnotetext[29]{Center for Theoretical Physics, Polish Academy of Sciences, al. Lotnik\'{o}w 32/46, 02-668, Warsaw, Poland} 
+Julian Borrill\footnotemark[17], 
+Elizabeth Buckley-Geer\footnotemark[30],
+\footnotetext[30]{Fermi National Accelerator Laboratory, Batavia, IL 60510} 
 Kelly A. Douglass\footnotemark[4],
-Cora Dvorkin\footnotemark[29],
-\footnotetext[29]{Department of Physics, Harvard University, Cambridge, MA 02138, USA} 
-Simon Foreman\footnotemark[30],
-\footnotetext[30]{Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5S 3H8, Canada} 
-Llu\'is Galbany\footnotemark[21],
-Juan Garc\'ia-Bellido\footnotemark[31]\textsuperscript{,}\footnotemark[32],
-\footnotetext[31]{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain} 
-\footnotetext[32]{Universidad Aut\'{o}noma de Madrid, 28049, Madrid, Spain} 
-Martina Gerbino\footnotemark[33],
-\footnotetext[33]{HEP Division, Argonne National Laboratory, Lemont, IL 60439, USA} 
-Mandeep S.S. Gill\footnotemark[34]\textsuperscript{,}\footnotemark[35]\textsuperscript{,}\footnotemark[36],
-\footnotetext[34]{Kavli Institute for Particle Astrophysics and Cosmology, Stanford 94305} 
-\footnotetext[35]{Stanford University, Stanford, CA 94305} 
-\footnotetext[36]{SLAC National Accelerator Laboratory, Menlo Park, CA 94025} 
-Larry Gladney\footnotemark[37],
-\footnotetext[37]{Department of Physics, Yale University, New Haven, CT 06520} 
-Saurabh W.~Jha\footnotemark[38],
-\footnotetext[38]{Department of Physics and Astronomy, Rutgers, the State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854, USA} 
-Johann Cohen-Tanugi\footnotemark[39],
-\footnotetext[39]{Laboratoire Univers et Particules de Montpellier, Univ. Montpellier and CNRS, 34090 Montpellier, France} 
-D.\ O.\ Jones\footnotemark[40],
-\footnotetext[40]{University of California at Santa Cruz, Santa Cruz, CA 95064} 
-Marc Kamionkowski\footnotemark[41],
-\footnotetext[41]{Johns Hopkins University, Baltimore, MD 21218} 
-Ryan E. Keeley\footnotemark[18],
-Robert Kehoe\footnotemark[42],
-\footnotetext[42]{Southern Methodist University, Dallas, TX 75275} 
-Savvas M. Koushiappas\footnotemark[43],
-\footnotetext[43]{Brown University, Providence, RI 02912} 
-Ely D.~Kovetz\footnotemark[44],
-\footnotetext[44]{Department of Physics, Ben-Gurion University, Be'er Sheva 84105, Israel} 
-Kazuya Koyama\footnotemark[45],
-\footnotetext[45]{Institute of Cosmology \& Gravitation, University of Portsmouth, Dennis Sciama Building, Burnaby Road, Portsmouth PO1 3FX, UK} 
-Benjamin L'Huillier\footnotemark[18],
-Ofer Lahav\footnotemark[46],
-\footnotetext[46]{University College London, WC1E 6BT London, United Kingdom} 
-Chien-Hsiu Lee\footnotemark[47],
-\footnotetext[47]{National Optical Astronomy Observatory, 950 N. Cherry Ave., Tucson, AZ 85719 USA} 
-Michele Liguori\footnotemark[48],
-\footnotetext[48]{Dipartimento di Fisica e Astronomia ``G. Galilei'',Universit\`a degli Studi di Padova, via Marzolo 8, I-35131, Padova, Italy} 
-Axel de la Macorra\footnotemark[49],
-\footnotetext[49]{IFUNAM - Instituto de F\'{i}sica, Universidad Nacional Aut\'onoma de M\'exico, 04510 CDMX, M\'exico} 
-Joel Meyers\footnotemark[42],
-Surhud More\footnotemark[50],
-\footnotetext[50]{The Inter-University Centre for Astronomy and Astrophysics, Pune, 411007, India} 
-Jeffrey A. Newman\footnotemark[21],
-Gustavo Niz\footnotemark[51],
-\footnotetext[51]{Divisi\'{o}n de Ciencias e Ingenier\'{i}as, Universidad de Guanajuato, Le\'{o}n 37150, M\'{e}xico} 
-Antonella Palmese\footnotemark[28],
-Francesco Piacentini\footnotemark[52]\textsuperscript{,}\footnotemark[53],
-\footnotetext[52]{Dipartimento di Fisica, Universit\`{a} La Sapienza, P. le A. Moro 2, Roma, Italy} 
-\footnotetext[53]{Istituto Nazionale di Fisica Nucleare, Sezione di Roma, 00185 Roma, Italy} 
-Steven Rodney\footnotemark[54],
-\footnotetext[54]{The University of South Carolina, Columbia, SC 29208} 
-Benjamin Rose\footnotemark[55],
-\footnotetext[55]{Space Telescope Science Institute, Baltimore, MD 21218} 
-Masao Sako\footnotemark[56],
-\footnotetext[56]{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, Pennsylvania 19104, USA} 
-Lado Samushia\footnotemark[57],
-\footnotetext[57]{Kansas State University, Manhattan, KS 66506} 
-Neelima Sehgal\footnotemark[58],
-\footnotetext[58]{Stony Brook University, Stony Brook, NY 11794} 
-Leonardo Senatore\footnotemark[34],
-Sara Simon\footnotemark[13],
-Stephanie Escoffier\footnotemark[59],
-\footnotetext[59]{Aix Marseille Univ, CNRS/IN2P3, CPPM, Marseille, France} 
-Scott Watson\footnotemark[60],
-\footnotetext[60]{Syracuse University, Syracuse, NY 13244} 
-Weishuang Xu\footnotemark[29],
-Gong-Bo Zhao\footnotemark[61]\textsuperscript{,}\footnotemark[45],
-\footnotetext[61]{National Astronomical Observatories, Chinese Academy of Sciences, PR China} 
-Yi Zheng\footnotemark[62]
-\footnotetext[62]{School of Physics, Korea Institute for Advanced Study, 85 Hoegiro, Dongdaemun-gu, Seoul 130-722, Korea}
+Cora Dvorkin\footnotemark[31],
+\footnotetext[31]{Department of Physics, Harvard University, Cambridge, MA 02138, USA} 
+Simon Foreman\footnotemark[32],
+\footnotetext[32]{Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5S 3H8, Canada} 
+Llu\'is Galbany\footnotemark[23],
+Juan Garc\'ia-Bellido\footnotemark[33]\textsuperscript{,}\footnotemark[34],
+\footnotetext[33]{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain} 
+\footnotetext[34]{Universidad Aut\'{o}noma de Madrid, 28049, Madrid, Spain} 
+Martina Gerbino\footnotemark[35],
+\footnotetext[35]{HEP Division, Argonne National Laboratory, Lemont, IL 60439, USA} 
+Mandeep S.S. Gill\footnotemark[36]\textsuperscript{,}\footnotemark[37]\textsuperscript{,}\footnotemark[38],
+\footnotetext[36]{Kavli Institute for Particle Astrophysics and Cosmology, Stanford 94305} 
+\footnotetext[37]{Stanford University, Stanford, CA 94305} 
+\footnotetext[38]{SLAC National Accelerator Laboratory, Menlo Park, CA 94025} 
+Larry Gladney\footnotemark[39],
+\footnotetext[39]{Department of Physics, Yale University, New Haven, CT 06520} 
+Saurabh W.~Jha\footnotemark[40],
+\footnotetext[40]{Department of Physics and Astronomy, Rutgers, the State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854, USA} 
+Johann Cohen-Tanugi\footnotemark[41],
+\footnotetext[41]{Laboratoire Univers et Particules de Montpellier, Univ. Montpellier and CNRS, 34090 Montpellier, France} 
+D.\ O.\ Jones\footnotemark[42],
+\footnotetext[42]{University of California at Santa Cruz, Santa Cruz, CA 95064} 
+Marc Kamionkowski\footnotemark[43],
+\footnotetext[43]{Johns Hopkins University, Baltimore, MD 21218} 
+Ryan E. Keeley\footnotemark[19],
+Robert Kehoe\footnotemark[44],
+\footnotetext[44]{Southern Methodist University, Dallas, TX 75275} 
+Savvas M. Koushiappas\footnotemark[45],
+\footnotetext[45]{Brown University, Providence, RI 02912} 
+Ely D.~Kovetz\footnotemark[46],
+\footnotetext[46]{Department of Physics, Ben-Gurion University, Be'er Sheva 84105, Israel} 
+Kazuya Koyama\footnotemark[47],
+\footnotetext[47]{Institute of Cosmology \& Gravitation, University of Portsmouth, Dennis Sciama Building, Burnaby Road, Portsmouth PO1 3FX, UK} 
+Benjamin L'Huillier\footnotemark[19],
+Ofer Lahav\footnotemark[48],
+\footnotetext[48]{University College London, WC1E 6BT London, United Kingdom} 
+Chien-Hsiu Lee\footnotemark[49],
+\footnotetext[49]{National Optical Astronomy Observatory, 950 N. Cherry Ave., Tucson, AZ 85719 USA} 
+Michele Liguori\footnotemark[50],
+\footnotetext[50]{Dipartimento di Fisica e Astronomia ``G. Galilei'',Universit\`a degli Studi di Padova, via Marzolo 8, I-35131, Padova, Italy} 
+Axel de la Macorra\footnotemark[51],
+\footnotetext[51]{IFUNAM - Instituto de F\'{i}sica, Universidad Nacional Aut\'onoma de M\'exico, 04510 CDMX, M\'exico} 
+Joel Meyers\footnotemark[44],
+Surhud More\footnotemark[52],
+\footnotetext[52]{The Inter-University Centre for Astronomy and Astrophysics, Pune, 411007, India} 
+Jeffrey A. Newman\footnotemark[23],
+Gustavo Niz\footnotemark[53],
+\footnotetext[53]{Divisi\'{o}n de Ciencias e Ingenier\'{i}as, Universidad de Guanajuato, Le\'{o}n 37150, M\'{e}xico} 
+Antonella Palmese\footnotemark[30],
+Francesco Piacentini\footnotemark[54]\textsuperscript{,}\footnotemark[55],
+\footnotetext[54]{Dipartimento di Fisica, Universit\`{a} La Sapienza, P. le A. Moro 2, Roma, Italy} 
+\footnotetext[55]{Istituto Nazionale di Fisica Nucleare, Sezione di Roma, 00185 Roma, Italy} 
+Steven Rodney\footnotemark[56],
+\footnotetext[56]{The University of South Carolina, Columbia, SC 29208} 
+Benjamin Rose\footnotemark[57],
+\footnotetext[57]{Space Telescope Science Institute, Baltimore, MD 21218} 
+Masao Sako\footnotemark[58],
+\footnotetext[58]{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, Pennsylvania 19104, USA} 
+Lado Samushia\footnotemark[59],
+\footnotetext[59]{Kansas State University, Manhattan, KS 66506} 
+Neelima Sehgal\footnotemark[60],
+\footnotetext[60]{Stony Brook University, Stony Brook, NY 11794} 
+Sara Simon\footnotemark[14],
+Stephanie Escoffier\footnotemark[61],
+\footnotetext[61]{Aix Marseille Univ, CNRS/IN2P3, CPPM, Marseille, France} 
+Scott Watson\footnotemark[62],
+\footnotetext[62]{Syracuse University, Syracuse, NY 13244} 
+Weishuang Xu\footnotemark[31],
+Gong-Bo Zhao\footnotemark[63]\textsuperscript{,}\footnotemark[47],
+\footnotetext[63]{National Astronomical Observatories, Chinese Academy of Sciences, PR China} 
+Yi Zheng\footnotemark[64]
+\footnotetext[64]{School of Physics, Korea Institute for Advanced Study, 85 Hoegiro, Dongdaemun-gu, Seoul 130-722, Korea}
 

### 2019-05-07 doc/point.tex
add Chao's draft

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{physics}
\title{Complete draft}
%\author{Chao Ju, Alex Kim}
\begin{document}
\maketitle

There is sufficient information in the distribution of peculiar velocities and overdensities of galaxies to test general relativity on the largest scale. Suppose we have a region that encloses $N_{gal}$ number of galaxies in the universe. We can divide the region into $N_{cell}\gg 1$ number of equal volume cells whose size can be made so small that either one or no galaxy is contained in one cell. This is encoded in a vector $\textbf{N}$ of size $N_{cell}$ whose entries are either 0 or 1. For any 2 cells, we can compute their correlation, a function of the distance $r$ between the cells only. In our application, we consider $\xi_{\delta\delta}$ (the overdensity-overdensity correlation), $\xi_{\delta v}$ (overdensity-velocity correlation), $\xi_{v\delta}$ (velocity-overdensity correlation), and $\xi_{vv}$ (velocity-velocity correlation).\par
We postulate that deviations from expected variance $\bar{N}$ and expected peculiar velocity $\bar{v}$ are Gaussian. Then the above correlation matrices can be used to construct a block matrix $\textbf{C}$, parameterized by $A$ and $\bm{\sigma}_v$, a vector of shot noise for peculiar velocity measurement:
\begin{align}
\textbf{C} &\equiv \textbf{K} + A\textbf{S} \\ 
&=
\left(
\begin{array}{cc}
\bar{N}\textbf{I} & 0 \nonumber \\
0 & \bm{\sigma}_v^2\textbf{I}
\end{array}
\right)
+
A
\left(
\begin{array}{cc}
\bar{N}^2\xi_{\delta\delta} & \bar{N}\xi_{\delta v} \\
\bar{N}\xi_{\delta v} & \xi_{vv}
\end{array}
\right).
\end{align}\par
The reason we have $\bar{N}$ as the variance for overdensity is that our small galaxy cells are best modeled by Poisson statistics, in which the mean equals the variance. In addition, note that the dimension of $C$ is $N_{cell}+N_{gal}$ by $N_{cell}+N_{gal}$. For the peculiar velocity covariance, we assign $\sigma^2=\infty$ to cells that do not contain any galaxy, and one can show that removing rows and columns that contain those cells has no effect but a constant offset on what we will compute later. \par
Given the block matrix $C$ and our data $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected, we can form the Gaussian likelihood as a function of $A$
\[
\mathcal{L}(A) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
\]\par
To make the expression cleaner, we take the log of the likelihood and focus on the $A$-dependent part:
\begin{equation}
-2\ln\mathcal{L}(A) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
\end{equation}\par
Given the enormous size of $\textbf{C}$, computing the log likelihood directly is not possible. In the following, we seek an approximation method accurate to first order in $A$. \par
First, we focus on the second term, $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{K}$ into an identity matrix. To do so, we introduce matrix $\textbf{M}$ and $\textbf{X}$. We let $\textbf{M}=\sqrt{\textbf{K}}$, and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$. Then $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$. Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
\begin{equation}
\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M},
\end{equation}
and its inverse, to first order, is
\[
\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} + \text{const}.
\] \par
Having this expression, we can now compute the second term in equation (2). Denoting the value of this term by $s$, we have
\begin{align*}
s &=\textbf{x}^T \textbf{C}^{-1}\textbf{x} \\
&= -A\left(\textbf{x}^T \textbf{K}^{-1}\right)\textbf{S}\left(\textbf{K}^{-1}\textbf{x}\right).
\end{align*} \par
Using the equations for covariance calculation from \cite{adam}, we can break the above expression into a sum of 6 terms and write out the explicit solution for each (apart from a zeroth-order constant). Let 
\[
-\left(\textbf{x}^T \textbf{K}^{-1}\right)\textbf{S}\left(\textbf{K}^{-1}\textbf{x}\right) = s_1 +s_2+\hdots +s_6
\]
\begin{align*}
s_1&= -\sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij},  \\
s_2 &= 2 \sum_{i\in N_{gal},j\in N_{cell}} \bar{N} \xi_{\delta\delta,ij}, \\
&= 2\bar{N} \frac{b^2}{2\pi} \sum_{i\in N_{gal}} \int r^2drd\Omega  \int dk P(k) k^2 j_0 (k |\textbf{r}-\textbf{r}_i|), \\
s_3 &= -\sum_{i,j\in N_{cell}} \bar{N}^2  \xi_{\delta\delta,ij}, \\ 
&=-\bar{N}^2 \frac{b^2}{2\pi} \int r_1^2dr_1d\Omega_1 \int r_2^2dr_2 d\Omega_2 \int dk P(k) k^2 j_0 (k |\textbf{r}_1-\textbf{r}_2|), \\
s_4 &=- \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}, \\
s_5 &= -2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}, \\
s_6 &= 2\sum_{i\in N_{cell}, j\in N_{gal}} \bar{N}\frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij},  \\
&= 2\bar{N}\sum_{j\in N_{gal}} \frac{aHfb\Delta v_j}{2\pi^2 \sigma^2_{v,j}}\int r^2 dr d\Omega \int dk P(k) k ((\hat{\textbf{r}}-\hat{\textbf{r}_j})\cdot\hat{\textbf{r}_j}) j_1(k|\textbf{r}-\textbf{r}_j|). 
%s_7 &= A  \sum_{j\in N_{cell}, i\in N_{gal}} \bar{N}\frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v\delta,ij}  \\
%&= A\bar{N} \sum_{i\in N_{gal}} \frac{+aHfb\Delta v_i}{2\pi^2 \sigma^2_{v,i}}\int r^2 dr d\Omega \int dk P(k) k (\textbf{r}\cdot\textbf{r}_i) j_1(k|\textbf{r}-\textbf{r}_i|)
\end{align*} \par
We note that in the above expressions, $P(k)$ is the power spectrum, $a$ the expansion factor, $H$ the Hubble constant, $b$ the galaxy bias, $f$ the growth rate of structure, and $j_l$ the spherical Bessel function. \par
Having computed the second term in equation (2), we now turn to the first term, $\ln\det\textbf{C}$. Using equation (3) to factor $\textbf{C}$, we have
\begin{align*}
\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\textbf{X})\textbf{M})) \\
&= \ln\det(\textbf{I}+A\textbf{X}) +\text{const} \\
&\approx A \tr \textbf{X} \\
&= A\tr(\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}) \\
&= A\tr 
\left(
\begin{array}{cc}
\bar{N}\xi_{\delta\delta} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\delta v} \\
\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\delta} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
\end{array}
\right),
\end{align*}
The trace can be broken up into 2 sums, $d_1+d_2$:
\begin{align*}
d_1 = \frac{b^2\bar{N}}{2\pi} \int r^2drd\Omega  \int dk P(k) k^2 
\end{align*}
since $j_0(0)=1$.
\[
d_2 = \sum_{i}\xi_{vv,ii}\sigma_i^{-2}
\] \par
In the limit as $N_{cell}$ goes to infinity, $\bar{N}\propto 1/N_{cell}$ goes to zero. Therefore, terms that contain $k$ factors of $N_{cell}$ will vanish in the $N_{cell}\to\infty$ limit unless they also contain $k$ summations over $N_{cell}$: $\sum_{N_{cell}}$. The reason is that $\sum_{N_{cell}}\bar{N}$ stays finite when $N_{cell}\to \infty$. Looking at the above $s$ and $d$ terms, we see that all of them are finite in the large $\bar{N}$ limit, and the the full likelihood to first order in $\bar{N}$ is
\[
\boxed{A(s_1+s_2+s_4+s_5+s_6+d_1+d_2)}
\]\par
$s_3$ is not selected because it is quadratic in $\bar{N}$.
\section{Inversion to second order}
The $O(A^2)$ prefactor in $\textbf{C}^{-1}$ is 
\begin{align*}
\textbf{M}^{-1}\textbf{X}^2\textbf{M}^{-1} &= \textbf{M}^{-1} \textbf{M}^{-1} \textbf{S} \textbf{M}^{-1}\textbf{M}^{-1} \textbf{S} \textbf{M}^{-1}\textbf{M}^{-1} \\
&=\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}
\end{align*}\par
I first sight I wanted to move $\textbf{S}$ across the $\textbf{K}^{-1}$ by using a commutator, thinking that it might simplify things a lot. But it turned out that the commutator did not lead to the expected cancelation, so let's do it the hard way. We first do $\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}$, henceforth denoting the inverse of $\bm{\sigma}^2_v\textbf{I}$ as $\bm{\sigma}^{-1}$:
\begin{align*}
\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S} &=
\left(
\begin{array}{cc}
\bar{N}\xi_{\delta\delta} & \xi_{\delta v} \\
\bar{N}\bm{\sigma}^{-1} \xi_{v\delta} & \bm{\sigma}^{-1} \xi_{vv}
\end{array}
\right)
\left(
\begin{array}{cc}
\bar{N}\xi_{\delta\delta} & \xi_{\delta v} \\
\bar{N}\bm{\sigma}^{-1} \xi_{v\delta} & \bm{\sigma}^{-1} \xi_{vv}
\end{array}
\right) \\
&=
\left(
\begin{array}{cc}
\bar{N}^2 \xi_{\delta\delta}^2 + \bar{N}\xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\xi_{\delta\delta}\xi_{\delta v}+\xi_{\delta v}\bm{\sigma}^{-1}\xi_{vv} \\
\bar{N}^2\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta\delta}+\bar{N}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta v} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}
\end{array}
\right)
\end{align*}\par
Next, we multiply $\textbf{K}^{-1}$ from the right and get
\begin{align*}
\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} &=
\left(
\begin{array}{cc}
\bar{N}^2 \xi_{\delta\delta}^2 + \bar{N}\xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\xi_{\delta\delta}\xi_{\delta v}+\xi_{\delta v}\bm{\sigma}^{-1}\xi_{vv} \\
\bar{N}^2\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta\delta}+\bar{N}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta v} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}
\end{array}
\right)
\left(
\begin{array}{cc}
\bar{N}^{-1}\textbf{I} & 0 \\
0 & \bm{\sigma}^{-1}
\end{array}
\right) \\
&=
\left(
\begin{array}{cc}
\bar{N} \xi_{\delta\delta}^2 + \xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\xi_{\delta\delta}\xi_{\delta v}\bm{\sigma}^{-1}+\xi_{\delta v}\bm{\sigma}^{-1}\xi_{vv}\bm{\sigma}^{-1} \\
\bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta\delta}+\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta v}\bm{\sigma}^{-1} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1}
\end{array}
\right)
\end{align*}\par
Let's denote the 4 blocks in the above matrix by
\[
\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}=
\left(
\begin{array}{cc}
\textbf{A} & \textbf{B} \\
\textbf{C} & \textbf{D}
\end{array}
\right)
\]\par
Let's now compute each block (the summations below are taken from $0$ to $N_{gal}$ unless otherwise specified). 
\begin{align*}
\textbf{A}_{ij} &=(\bar{N} \xi_{\delta\delta}^2 + \xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta})_{ij} \\
&= \frac{\bar{N}b^4}{4\pi^4} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|) +\sum_{m}\sigma_{m}^{-2}\xi_{\delta v,im}\xi_{v\delta,mj} \\
&=\frac{\bar{N}b^4}{4\pi^4} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|)\\
&+\frac{a^2H^2f^2b^2}{4\pi^2} \sum_m \sigma_{m}^{-2} \iint dk_1dk_2 P(k_1)P(k_2) k_1k_2j_1(k_1|\textbf{r}_i-\textbf{r}_m|)j_1(k_2|\textbf{r}_m-\textbf{r}_j|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m(\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m\\
\textbf{B}_{ij} &= \frac{\bar{N}aHfb^3\sigma_j^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|)\\
&+ \sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{\delta v,im}\xi_{vv,mj} \\
&=\frac{\bar{N}aHfb^3\sigma_j^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) \\
&+\frac{\sigma_j^{-2}aHfb}{2\pi^2}\sum_m \sigma_m^{-2} \xi_{vv,mj}\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m \\
\textbf{C}_{ij} &=\frac{\bar{N}aHfb^3\sigma_i^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_ij_0(k_2|\textbf{r}-\textbf{r}_j|)j_1(k_1|\textbf{r}-\textbf{r}_i|)\\
&+ \sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{vv,im}\xi_{v\delta,mj} \\
&=\frac{\bar{N}aHfb^3\sigma_i^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_ij_0(k_2|\textbf{r}-\textbf{r}_j|)j_1(k_1|\textbf{r}-\textbf{r}_i|)\\
&+\frac{\sigma_j^{-2}aHfb}{2\pi^2}\sum_m \sigma_m^{-2} \xi_{vv,im}\int dk P(k)k j_1(k|\textbf{r}_m-\textbf{r}_j|)(\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m\\
\textbf{D}_{ij} &=\frac{\bar{N}a^2H^2f^2b^2\sigma_i^{-2}\sigma_j^{-2}}{4\pi^4} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_i(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_j\times \\ 
&\times j_1(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) +\sigma_i^{-2}\sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{vv,im}\xi_{vv,mj}
\end{align*}\par
Having obtained the 4 block matrices, we now proceed with the calculation of $\textbf{x}^T\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{x}$. Similar to what we did for the first order, we break the sum into the following terms (again, summation is within the range of $N_{gal}$ by default unless otherwise stated):
\begin{align*}
S_1 &= \sum_{i, j} \textbf{A}_{ij} \\
&=\frac{\bar{N}b^4}{4\pi^4}\sum_{i, j} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|)  \\
&+ \frac{a^2H^2f^2b^2}{4\pi^2} \sum_{m,i,j} \sigma_{m}^{-2} \iint dk_1dk_2 P(k_1)P(k_2) k_1k_2j_1(k_1|\textbf{r}_i-\textbf{r}_m|)j_1(k_2|\textbf{r}_m-\textbf{r}_j|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m(\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m\\
S_2 &=-2\bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \textbf{A}_{ij}\\
&=\frac{-\bar{N}^2 b^4}{2\pi^4}\sum_{i}\int r^2r_j^2drd\Omega dr_j d\Omega_j \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|)-\frac{a^2H^2f^2b^2\bar{N}}{2\pi^2}\times \\
&\times  \sum_{m,i} \sigma_{m}^{-2}\int r_j^2dr_j d\Omega_j \iint dk_1dk_2 P(k_1)P(k_2) k_1k_2j_1(k_1|\textbf{r}_i-\textbf{r}_m|)j_1(k_2|\textbf{r}_m-\textbf{r}_j|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m \times \\
&\times (\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m \\
S_3 &= \bar{N}^2 \sum_{i, j\in N_{cell}} \textbf{A}_{ij} \\
&=\frac{\bar{N}^3b^4}{4\pi^4}\int r^2r_i^2r_j^2 drd\Omega d\Omega_i d\Omega_j \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|)  \\
&+ \frac{a^2H^2f^2b^2 \bar{N}^2}{4\pi^2} \sum_{m} \sigma_{m}^{-2}\int r_i^2r_j^2 dr_idr_j d\Omega_i d\Omega_j  \iint dk_1dk_2 P(k_1)P(k_2) k_1k_2j_1(k_1|\textbf{r}_i-\textbf{r}_m|)j_1(k_2|\textbf{r}_m-\textbf{r}_j|)\times\\
&\times (\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m(\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m \\
S_4 &= \sum_{i,j} \textbf{D}_{ij}\Delta v_i \Delta v_j \\
&= \frac{\bar{N}a^2H^2f^2b^2}{4\pi^4}\sum_{i,j}\sigma_i^{-2}\sigma_j^{-2}\Delta v_i \Delta v_j \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_i(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_j\times \\ 
&\times j_1(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) +\sum_{m,i,j} \sigma_i^{-2}\sigma_j^{-2}\sigma_m^{-2}\Delta v_i \Delta v_j\xi_{vv,im}\xi_{vv,mj}\\
S_5 &= 2\sum_{i,j}\textbf{B}_{ij}\Delta v_j \\
&=\frac{\bar{N}aHfb^3}{2\pi^4}\sum_{i,j}\sigma_j^{-2}\Delta v_j\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) \\
&+\frac{\sigma_j^{-2}aHfb}{\pi^2}\sum_{m,i,j}\Delta v_j \sigma_m^{-2} \xi_{vv,mj}\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m\\
S_6 &=-2\bar{N} \sum_{i\in N_{cell},j\in N_{gal}}\textbf{B}_{ij} \Delta v_j\\
&=\frac{-\bar{N}^2aHfb^3}{2\pi^4}\sum_{j}\sigma_j^{-2}\Delta v_j\int r^2r_i^2drdr_id\Omega d\Omega_i \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)\times \\
&\times j_1(k_2|\textbf{r}-\textbf{r}_j|)-\frac{\bar{N}aHfb}{\pi^2}\sum_{m,j}\sigma_j^{-2} \Delta v_j \sigma_m^{-2} \xi_{vv,mj}\int r_i^2 dr_id\Omega_i\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m
\end{align*}\par
The seconder-order contribution from $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$ to the likelihood is thus
\[
A^2 (S_1 +S_2 +S_3 +S_4 +S_5 +S_6)
\]\par
\section{Log det to second order}
We now calculate the second-order contribution from the log det term. The $O(A^2)$ prefactor in $\ln \det(\textbf{I}+A\textbf{X})$ is
\[
-\frac{1}{2}\tr (\textbf{X}^2) = -\frac{1}{2}\tr(\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1})
\]\par
Since the trace is cyclic, we can permute the matrix multiplication and make it into something we've calculated on page 3:
\begin{align*}
-\frac{1}{2}\tr (\textbf{X}^2)&=-\frac{1}{2}\tr(\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1})\\
&=-\frac{1}{2}\tr(\textbf{M}^{-1}\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}\textbf{M}^{-1}\textbf{S}) \\
&= -\frac{1}{2}\tr(\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}) \\
&= -\frac{1}{2}\tr\left\{
\left(
\begin{array}{cc}
\bar{N}^2 \xi_{\delta\delta}^2 + \bar{N}\xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\xi_{\delta\delta}\xi_{\delta v}+\xi_{\delta v}\bm{\sigma}^{-1}\xi_{vv} \\
\bar{N}^2\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta\delta}+\bar{N}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta v} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}
\end{array}
\right)
\right\}
\end{align*}\par
Expanding the trace, we get (using $D_i$ to label each term)
\begin{align*}
D_1 + D_2 + D_3 + D_4 &\equiv -\frac{\bar{N}^2}{2} \sum_{i,j\in N_{cell}} \xi_{\delta\delta,ij}\xi_{\delta\delta,ji} -\frac{\bar{N}}{2}\sum_{i \in N_{cell},j\in N_{gal}} \sigma_j^{-2}\xi_{\delta v,ij} \xi_{v\delta,ji} \\
&-\frac{\bar{N}}{2} \sum_{i\in N_{gal},j\in N_{cell}} \sigma_i^{-2} \xi_{v\delta,ij}\xi_{\delta v,ji} - \frac{1}{2} \sum_{i,j} \sigma_i^{-2}\sigma_j^{-2} \xi_{vv,ij}\xi_{vv,ji}
\end{align*}\par
Let's calculate each $D$:
\begin{align*}
D_1 &= -\frac{\bar{N}^2 b^4}{8\pi^4} \int r^2r_i^2drdr_id\Omega d\Omega_i d \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_i|) \\
D_2 &= -\frac{a^2H^2f^2b^2\bar{N}}{8\pi^2} \sum_j \sigma_{j}^{-2}\int r_i^2 dr_i d\Omega_i \iint dk_1dk_2 P(k_1)P(k_2) k_1k_2j_1(k_1|\textbf{r}_i-\textbf{r}_j|)j_1(k_2|\textbf{r}_j-\textbf{r}_i|)\times \\
&\times (\hat{\textbf{r}}_i-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_j(\hat{\textbf{r}}_j-\hat{\textbf{r}}_i)\cdot\hat{\textbf{r}}_j\\
D_3 &= -\frac{\bar{N}a^2H^2f^2b^2}{8\pi^4} \sum_{i}\sigma_i^{-2} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_i(\hat{\textbf{r}}-\hat{\textbf{r}}_i)\cdot\hat{\textbf{r}}_i\times \\ 
&\times j_1(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_i|)\\
D_4 &=- \frac{1}{2} \sum_{i,j} \sigma_i^{-2}\sigma_j^{-2} \xi_{vv,ij}\xi_{vv,ji}
\end{align*}\par
By similar argument as we did at the end of the previous section, all of the above stay finite in the limit $N_{cell}\to\infty$. In summary, the full likelihood at second order (in $A$, not in $\bar{N}$) is
\[
\boxed{A^2 (S_1 +S_2' +S_4 +S_5 +S_6'+D_2+D_3+D_4)}
\]
where
\begin{align*}
S_2' &=-\frac{a^2H^2f^2b^2\bar{N}}{2\pi^2}\times \\
&\times  \sum_{m,i} \sigma_{m}^{-2}\int r_j^2dr_j d\Omega_j \iint dk_1dk_2 P(k_1)P(k_2) k_1k_2j_1(k_1|\textbf{r}_i-\textbf{r}_m|)j_1(k_2|\textbf{r}_m-\textbf{r}_j|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m \times \\
&\times (\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m \\
S_6' &=-\frac{\bar{N}aHfb}{\pi^2}\sum_{m,j}\sigma_j^{-2} \Delta v_j \sigma_m^{-2} \xi_{vv,mj}\int r_i^2 dr_id\Omega_i\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m
\end{align*}
after removing the $O(\bar{N}^2)$ terms. Note that $S_3$ and $D_1$ dropped out completely.
\section{Full likelihood to second order}
Adding up both the first and the second order contribution, we have
\[
\mathcal{L} =\boxed{ A^2 (S_1 +S_2'  +S_4 +S_5 +S_6'+D_2+D_3+D_4) +A(s_1+s_2+s_4+s_5+s_6+d_1+d_2)}
\]
with extremum at
\[
\boxed{A^*= -\frac{s_1+s_2+s_4+s_5+s_6+d_1+d_2}{2(S_1 +S_2'  +S_4 +S_5 +S_6'+D_2+D_3+D_4)}}
\]

\begin{thebibliography}{9}
\bibitem{adam}
C. Adams, C. Blake. 
Improving constraints on the growth rate of structure by modelling the density-velocity cross-correlation in the 6dF Galaxy Survey. arXiv:1706.05205 [astro-ph.CO].
\end{thebibliography}

\end{document}

### 2019-05-07 doc/src/fom_aksgg.py
SNPV FoM for OSTF v1

@@ -1,0 +1,91 @@
+#!/usr/bin/env python
+
+import numpy as np 
+import matplotlib.pyplot as plt
+#import healpy as hp
+import os
+import re
+import partials
+
+name = 'nsn_tot.npy'
+skyfrac=[]
+rt=[]
+for root, dirs, files in os.walk("Footprints/"): # for 10 years footprints
+    if name in files:
+    	# to read the map 
+        m = np.load(os.path.join(root, name))
+
+    	# to display it 
+    	# hp.mollview(m, nest=1)
+
+    	# to get the total number of SNe
+        nsn_tot = np.sum(m)
+
+    	# to get the total footprint (in pixels)
+        npixnon0=(m>0).sum()
+
+        # get the total number of pixels
+        npixtot = npixnon0 + (m==0).sum()
+
+        if root[root.rfind('/')+1:] == 'baseline2018a' :
+            skyfrac_ref =  npixnon0/npixtot
+
+        else:
+
+            skyfrac.append(npixnon0/npixtot)
+
+            dum=[m.start() for m in re.finditer('_', root)]
+            rt.append(root[root.rfind('/')+1:])
+
+rt = np.array(rt)
+#print('rt',rt)
+
+skyfrac=np.array(skyfrac)
+#print("skyfrac",skyfrac)
+
+#w=np.where(rt == 'baseline2018a')[0]
+#skyfrac_ref= skyfrac[w]
+#print(skyfrac_ref)
+
+fom_ref=partials.surveyFOM(skyfrac_ref,10)
+#print("fom_ref",fom_ref)
+fom_10 = partials.surveyFOM(skyfrac,10) / fom_ref
+#print(fom_10)
+
+rt=rt
+
+so =np.argsort(rt)
+dtp = []
+list = []
+
+for index in range(len(rt)):
+    list= (so[index],fom_10[so[index]])
+    dtp.append((list))
+
+#print("liste=",dtp)
+dtype = [('cadence_argsort', float), ('fom_10', float)]
+to_sort = np.array(dtp, dtype=dtype)
+to_plot = np.sort(to_sort,order="fom_10") 
+
+tp_cadence = []
+tp_fom_10 = [] 
+
+for index in range(len(to_plot)):
+    i = len(to_plot) - index
+    tp_cadence.append(rt[int(to_plot[i][0])])
[…]

### 2019-05-07 doc/src/fom_ostf.py

@@ -1,0 +1,28 @@
+ #!/usr/bin/env python
+
+import numpy as np 
+import matplotlib.pyplot as plt
+
+tp_cadence=['rolling_mix_10yrs_opsim', 'pontus_2489', 'rolling_10yrs_opsim', 'kraken_2042', 'alt_sched_rolling', 'kraken_2044', 'kraken_2026', 'kraken_2035', 'pontus_2002', 'colossus_2667', 'mothra_2049', 'nexus_2097', 'pontus_2573', 'alt_sched'] 
+
+tp_fom_10=[0.82523797,0.86024946,0.99581343,0.97263434,0.90679374,0.71793979,0.99543942,1.01288421,0.72109293,0.82840686,0.71807872,0.71547569,0.73722483,0.91633328] 
+
+idx=[11,5,10,8,12,0,9,1,4,13,3,6,2,7] 
+p1=[]
+p2=[]
+
+for i in idx: 
+	p1.append(tp_cadence[i])
+	p2.append(tp_fom_10[i])
+
+fig = plt.figure(figsize=(10,15))
+ax = fig.add_subplot(1, 1, 1)
+#plt.grid()
+plt.barh(p1,p2, alpha=0.65, color="black")
+plt.xticks(rotation=0,fontsize=20)
+plt.yticks(fontsize=20)
+plt.tick_params(length=10,width=1.2)
+plt.xlabel("$\sigma_\gamma$(cadence) / $\sigma_\gamma$(baseline2018a)",fontsize=20)
+plt.ylim(0.5,13.5)
+plt.show()
+plt.savefig('FOM_Gamma_OSTF_new.pdf')

### 2019-05-08 doc/point.tex
version for Pat

@@
 \usepackage{physics}
+\usepackage{verbatim}
 \title{Complete draft}
@@
 -2\ln\mathcal{L}(A) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
+\label{log:eqn}
 \end{equation}\par
+
+%  ---- Change
+%NEW STUFF
+
+We new expand the log-likelihood as a Taylor series for small $A$ around $A=0$.
+
+First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{K}$ term
+by   introducing the matrices $\textbf{M}$ and $\textbf{X}$,
+where $\textbf{M}=\sqrt{\textbf{K}}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
+Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
+\begin{equation}
+\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M}.
+\end{equation}
+Expanded as a Taylor series
+\begin{align}
+\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\textbf{X}+A^2 \textbf{X}^2 + \ldots) \textbf{M}^{-1} \\
+& = \textbf{K}^{-1}   -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} +\frac{A^2}{2} \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} +O(A^3) + \ldots 
+\end{align}
+%= -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} + \text{const}.
+%\] \par
+
+We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
+ $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
+and $\xi_{vv}$.  In this manner
+\begin{align}
+\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
+\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+\end{align}
+Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
+for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
+\begin{align*}
+\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
+& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
+& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
+& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
+& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
+\end{align*}
+
+Now consider the next term $\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}$.  The middle section is
+\begin{align}
+\textbf{S}\textbf{K}^{-1}\textbf{S} & =
+\begin{pmatrix}
+ \bar{N}^2  \xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}  & \bar{N} \xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
+ \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
+\end{pmatrix}
+\end{align}
+to leading order of $\bar{N}$ anticipating that we will take the limit where it goes to zero.
+Then
+\begin{align}
+\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}
+& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
+ + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV\\
+& +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2.
+\end{align}
+
+We have assembled the contributions of $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ to second order in $A$.
+
+Returning to the other contribution to the log-likelihood in Eq.~\ref{log:eqn} we have
+\begin{align*}
+\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\textbf{X})\textbf{M})) \\
+&= \ln\det(\textbf{I}+A\textbf{X}) +\text{const} \\
+&= A \tr \textbf{X} -\frac{A^2}{2}\tr \textbf{X}^2+ O(A^3) + \ldots\\
+\end{align*}
+Now
+\begin{equation}
+\textbf{X} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
+\begin{pmatrix}
+\bar{N}\xi_{\delta\delta} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\delta v} \\
+\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\delta} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
+\end{pmatrix}.
+\end{equation}
+The contribution to the trace of the upper diagonal in the limit of small cell volumes is
+\begin{equation}
+\bar{N} \tr \xi_{\delta\delta} = n \int \xi_{\delta\delta,VV} dV
+\end{equation}
+and the lower diagonal is
+\begin{equation}
+\tr  \frac{1}{\sigma_{v}^2} \xi_{vv} =  \sum_{i \in N_{gal}}   \frac{1}{\sigma_{v,i}^2} \xi_{vv,ii}.
+\end{equation}
+
+The trace of $\textbf{X}^2$ has terms, to leading order in $\bar{N}$, has terms
+\begin{equation}
+\bar{N}  \tr  \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} = n \int \left( \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} \right)_{VV} dV
+\end{equation}
+and 
+\begin{equation}
+\tr   \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} =  \sum_{i \in N_{gal}} \left(  \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} \right)_{ii}
+\end{equation}
+
+Putting everything together, the expansion of the log-likelihood in $A$ has the first order coefficient
+\begin{align}
+c_1 &=- \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+& +\sum_{i \in N_{gal}} \left( \frac{1}{\sigma_{v,i}^2} \xi_{vv,ii}   + 2n \int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV  \right)\\
+& +n \int \xi_{\delta\delta,VV} dV  -n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2
+\end{align}
+and the second order coefficient
+\begin{align}
+c_2 & = \frac{1}{2} \Biggl( 
+ \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
+ + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+&- \sum_{i \in N_{gal}}\left( \left(  \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} \right)_{ii}  +2n \int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV  \right)\\
+&-n  \int \left( \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} \right)_{VV} dV  +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2 \Biggr).
+\end{align}
+
+In this second-order approximation of the log-likelihood, the minimum occurs at $A_{\text{min}}=-c_1/2c_2$ and the second derivative is $2c_2$.
+
+\begin{comment}
+END NEW STUFF
+
 Given the enormous size of $\textbf{C}$, computing the log likelihood directly is not possible. In the following, we seek an approximation method accurate to first order in $A$. \par
@@
 \]
-
+\end{comment}
 \begin{thebibliography}{9}

### 2019-05-09 doc/point.tex
update but need to check Fisher matrix

@@
 In our application, we consider $\xi_{\delta\delta}$ (the overdensity-overdensity correlation), $\xi_{\delta v}$ (overdensity-velocity correlation), $\xi_{v\delta}$ (velocity-overdensity correlation), and $\xi_{vv}$ (velocity-velocity correlation).\par
+
+The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,
+
 We postulate that deviations from expected variance $\bar{N}$ and expected peculiar velocity $\bar{v}$ are Gaussian.
@@
 \begin{align}
-\textbf{C} &\equiv \textbf{K} + A\textbf{S} \\ 
+\textbf{C} &\equiv \textbf{C}_0 + \textbf{S}\\
+%\textbf{C} &\equiv \textbf{C}_0 + A\textbf{S} \\ 
 &=
@@
 +
-A
-\left(
-\begin{array}{cc}
-\bar{N}^2\xi_{\delta\delta} & \bar{N}\xi_{\delta v} \\
-\bar{N}\xi_{\delta v} & \xi_{vv}
-\end{array}
-\right).
-\end{align}\par
+\begin{pmatrix}
+\bar{N}^2\xi_{gg} & \bar{N}\xi_{g v} \\
+\bar{N}\xi_{v g} & \xi_{vv}
+\end{pmatrix},
+\end{align}
+where $\xi_{gg}$, $\xi_{vv}$, and $\bar{N}\xi_{g v}$ are the galaxy overdensity and peculiar velocity correlation functions and their cross-correlation
+respectively.
+
 The reason we have $\bar{N}$ as the variance for overdensity is that our small galaxy cells are best modeled by Poisson statistics, in which the mean equals the variance.
@@
 For the peculiar velocity covariance, we assign $\sigma^2=\infty$ to cells that do not contain any galaxy, and one can show that removing rows and columns that contain those cells has no effect but a constant offset on what we will compute later.
-\par
-Given the block matrix $C$ and our data $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected, we can form the Gaussian likelihood as a function of $A$
+
+The science parameters $\theta$ enter through these correlation functions, and may include $\theta \in \{fD, bD, \gamma, \Omega_{M_0},\ldots\}$ and others
+that describe the matter power spectrum.
+For example, the parameter $fD$ enters in through the relations $\xi_{vv}\propto (fD\mu)^2$, the SN~Ia host-galaxy count overdensity
+power spectrum $\xi_{g g }\propto (bD + fD\mu^2)^2$, and the galaxy-velocity cross-correlation $\xi_{vg}
+\propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu\equiv \cos{(\hat{k} \cdot \hat{r})}$ where $\hat{r}$ is the direction of
+the line of sight.  
+Assuming linear dependence of the parameters,
+\begin{align}
+\textbf{C} 
+&=
+\left(
+\begin{array}{cc}
+\bar{N}\textbf{I} & 0 \nonumber \\
+0 & \bm{\sigma}_v^2\textbf{I}
+\end{array}
+\right)
++ \sum_\theta \theta
+\begin{pmatrix}
+\bar{N}^2\xi_{gg,\theta} & \bar{N}\xi_{g v,\theta} \\
+\bar{N}\xi_{v g,\theta} & \xi_{vv,\theta}
+\end{pmatrix}\\
+&=
+\left(
+\begin{array}{cc}
+\bar{N}\textbf{I} & 0 \nonumber \\
+0 & \bm{\sigma}_v^2\textbf{I}
+\end{array}
+\right)
++ \sum_\alpha \theta_\alpha \textbf{S}_\alpha .
+\end{align}
+
+%\left(
+%\begin{array}{cc}
+%\bar{N}\textbf{I} & 0 \nonumber \\
+%0 & \bm{\sigma}_v^2\textbf{I}
+%\end{array}
+%\right)
+%+
+%A
+%\left(
+%\begin{array}{cc}
+%\bar{N}^2\xi_{\delta\delta} & \bar{N}\xi_{\delta v} \\
+%\bar{N}\xi_{\delta v} & \xi_{vv}
+%\end{array}
+%\right).
+%\end{align}\par
+
+The Gaussian likelihood as a function of $\theta$
 \[
-\mathcal{L}(A) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
+\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
 \]\par
@@
 \begin{equation}
--2\ln\mathcal{L}(A) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
+-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
 \label{log:eqn}
-\end{equation}\par
+\end{equation}
+
+The estimator and Fisher Matrix for this Gaussian likelihood are known
+\begin{equation}
+\hat{\theta}_\alpha = \frac{1}{2} F^{-1}_{\alpha \beta} [(\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]
+\end{equation}
+and
+\begin{equation}
+F_{\alpha\beta} = \frac{1}{2} \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0 \textbf{S}_\beta).
+\end{equation}
+
+The term
+\begin{align}
+\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \textbf{x}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \textbf{x}\\
+&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
+\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+\end{align}
+Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
+for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = \int n dV$, where $n$ is the galaxy number density,
+\begin{align*}
+\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
+& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
+& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
+& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
+&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
+\end{align*}
+
+The trace term is
+\begin{align}
+ \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\beta) \\
+ & = \bar{N} \tr(\xi_{\delta\delta,\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
+ & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i)
+\end{align}
+
+The remaining term to the leading term in $\bar{N}$ is
+\begin{align}
+ \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1} \textbf{S}_\beta) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\alpha \textbf{C}_0^{-1} \textbf{S}_\beta)  \\
+ & = \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
+ & = \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
+ & +  \sum_{i,j\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\alpha}(x_i,x_j)  \frac{1}{\sigma^2_{v,j}}  \xi_{vv,\beta}(x_j,x_i).
+\end{align}
 
@@
 %NEW STUFF
+\begin{comment}
+We now expand the log-likelihood as a Taylor series for small $\theta$ around $\theta=0$.
 
-We new expand the log-likelihood as a Taylor series for small $A$ around $A=0$.
-
-First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{K}$ term
+First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
 by   introducing the matrices $\textbf{M}$ and $\textbf{X}$,
-where $\textbf{M}=\sqrt{\textbf{K}}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
+where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
 Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
@@
 \textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\textbf{X}+A^2 \textbf{X}^2 + \ldots) \textbf{M}^{-1} \\
-& = \textbf{K}^{-1}   -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} +\frac{A^2}{2} \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} +O(A^3) + \ldots 
+& = \textbf{C}_0^{-1}   -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +\frac{A^2}{2} \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +O(A^3) + \ldots 
 \end{align}
-%= -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} + \text{const}.
+%= -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
 %\] \par
 
-We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
- $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
-and $\xi_{vv}$.  In this manner
+A term of interest is 
 \begin{align}
-\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
+\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
 \frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
@@
 \begin{align*}
-\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
+\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
 & +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
@@
 
-Now consider the next term $\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}$.  The middle section is
+We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
+ $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
+and $\xi_{vv}$.  In this manner
 \begin{align}
-\textbf{S}\textbf{K}^{-1}\textbf{S} & =
+\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
+\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+\end{align}
+Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
+for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
+\begin{align*}
+\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
+& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
+& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
+& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
+& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
+\end{align*}
+
+Now consider the next term $\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}$.  The middle section is
+\begin{align}
+\textbf{S}\textbf{C}_0^{-1}\textbf{S} & =
 \begin{pmatrix}
@@
 \begin{align}
-\textbf{x}^T \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} \textbf{x}
+\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}
 & = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
@@
 
-\begin{comment}
 END NEW STUFF
@@
 First, we focus on the second term, $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$.
-Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{K}$ into an identity matrix.
+Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{C}_0$ into an identity matrix.
 To do so, we introduce matrix $\textbf{M}$ and $\textbf{X}$.
-We let $\textbf{M}=\sqrt{\textbf{K}}$, and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$.
+We let $\textbf{M}=\sqrt{\textbf{C}_0}$, and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$.
 Then $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
@@
 \[
-\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} + \text{const}.
+\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
 \] \par
@@
 s &=\textbf{x}^T \textbf{C}^{-1}\textbf{x} \\
-&= -A\left(\textbf{x}^T \textbf{K}^{-1}\right)\textbf{S}\left(\textbf{K}^{-1}\textbf{x}\right).
+&= -A\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right).
 \end{align*} \par
@@
 \[
--\left(\textbf{x}^T \textbf{K}^{-1}\right)\textbf{S}\left(\textbf{K}^{-1}\textbf{x}\right) = s_1 +s_2+\hdots +s_6
+-\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right) = s_1 +s_2+\hdots +s_6
 \]
@@
 \textbf{M}^{-1}\textbf{X}^2\textbf{M}^{-1} &= \textbf{M}^{-1} \textbf{M}^{-1} \textbf{S} \textbf{M}^{-1}\textbf{M}^{-1} \textbf{S} \textbf{M}^{-1}\textbf{M}^{-1} \\
-&=\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}
+&=\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}
 \end{align*}\par
-I first sight I wanted to move $\textbf{S}$ across the $\textbf{K}^{-1}$ by using a commutator, thinking that it might simplify things a lot.
+I first sight I wanted to move $\textbf{S}$ across the $\textbf{C}_0^{-1}$ by using a commutator, thinking that it might simplify things a lot.
 But it turned out that the commutator did not lead to the expected cancelation, so let's do it the hard way.
-We first do $\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}$, henceforth denoting the inverse of $\bm{\sigma}^2_v\textbf{I}$ as $\bm{\sigma}^{-1}$:
+We first do $\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}$, henceforth denoting the inverse of $\bm{\sigma}^2_v\textbf{I}$ as $\bm{\sigma}^{-1}$:
 \begin{align*}
-\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S} &=
+\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S} &=
 \left(
@@
 \end{align*}\par
-Next, we multiply $\textbf{K}^{-1}$ from the right and get
+Next, we multiply $\textbf{C}_0^{-1}$ from the right and get
 \begin{align*}
-\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1} &=
+\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} &=
 \left(
@@
 \[
-\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}=
+\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}=
 \left(
@@
 \end{align*}\par
-Having obtained the 4 block matrices, we now proceed with the calculation of $\textbf{x}^T\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{x}$.
+Having obtained the 4 block matrices, we now proceed with the calculation of $\textbf{x}^T\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{x}$.
 Similar to what we did for the first order, we break the sum into the following terms (again, summation is within the range of $N_{gal}$ by default unless otherwise stated):
@@
 &=-\frac{1}{2}\tr(\textbf{M}^{-1}\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}\textbf{M}^{-1}\textbf{S}) \\
-&= -\frac{1}{2}\tr(\textbf{K}^{-1}\textbf{S}\textbf{K}^{-1}\textbf{S}) \\
+&= -\frac{1}{2}\tr(\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}) \\
 &= -\frac{1}{2}\tr\left\{

### 2019-05-10 doc/src/fom_aksgg.py
SNPV FoM for OSTF v2

@@ -38,54 +38,50 @@ for root, dirs, files in os.walk("Footprints/"): # for 10 years footprints
             rt.append(root[root.rfind('/')+1:])
 
 rt = np.array(rt)
+#print('len(rt)',len(rt))
 #print('rt',rt)
 
 skyfrac=np.array(skyfrac)
 #print("skyfrac",skyfrac)
 
-#w=np.where(rt == 'baseline2018a')[0]
-#skyfrac_ref= skyfrac[w]
-#print(skyfrac_ref)
-
-fom_ref=partials.surveyFOM(skyfrac_ref,10)
+fom_ref=1./partials.surveyFOM(skyfrac_ref,10)
 #print("fom_ref",fom_ref)
-fom_10 = partials.surveyFOM(skyfrac,10) / fom_ref
+fom_10 = (1./partials.surveyFOM(skyfrac,10)) / fom_ref
 #print(fom_10)
-
-rt=rt
 
 so =np.argsort(rt)
 dtp = []
 list = []
 
-for index in range(len(rt)):
+for index in range(len(rt)-1):
     list= (so[index],fom_10[so[index]])
     dtp.append((list))
 
 #print("liste=",dtp)
 dtype = [('cadence_argsort', float), ('fom_10', float)]
 to_sort = np.array(dtp, dtype=dtype)
-to_plot = np.sort(to_sort,order="fom_10") 
+to_plot = np.sort(to_sort,order="fom_10")
 
 tp_cadence = []
 tp_fom_10 = [] 
 
-for index in range(len(to_plot)):
-    i = len(to_plot) - index
-    tp_cadence.append(rt[int(to_plot[i][0])])
-    tp_fom_10.append(to_plot[i][1])
+for index in range(len(dtp)):
+    i = len(dtp) - index -1
+    a=to_plot[i]
+    tp_cadence.append(rt[int(a[0])])
+    tp_fom_10.append(a[1])
 
-print("Footprints : ", tp_cadence)
-print("FoM over 10 years : ", tp_fom_10)
+#print("Footprints : ", tp_cadence)
+#print("FoM over 10 years : ", tp_fom_10)
 
-#fig = plt.figure(figsize=(10,15))
-#ax = fig.add_subplot(1, 1, 1)
-#plt.barh(tp_cadence,tp_fom_10, alpha=0.7, color="grey")
[…]

### 2019-05-10 doc/src/fom_ostf.py
Delete fom_ostf.py

@@ -1,28 +1,0 @@
- #!/usr/bin/env python
-
-import numpy as np 
-import matplotlib.pyplot as plt
-
-tp_cadence=['rolling_mix_10yrs_opsim', 'pontus_2489', 'rolling_10yrs_opsim', 'kraken_2042', 'alt_sched_rolling', 'kraken_2044', 'kraken_2026', 'kraken_2035', 'pontus_2002', 'colossus_2667', 'mothra_2049', 'nexus_2097', 'pontus_2573', 'alt_sched'] 
-
-tp_fom_10=[0.82523797,0.86024946,0.99581343,0.97263434,0.90679374,0.71793979,0.99543942,1.01288421,0.72109293,0.82840686,0.71807872,0.71547569,0.73722483,0.91633328] 
-
-idx=[11,5,10,8,12,0,9,1,4,13,3,6,2,7] 
-p1=[]
-p2=[]
-
-for i in idx: 
-	p1.append(tp_cadence[i])
-	p2.append(tp_fom_10[i])
-
-fig = plt.figure(figsize=(10,15))
-ax = fig.add_subplot(1, 1, 1)
-#plt.grid()
-plt.barh(p1,p2, alpha=0.65, color="black")
-plt.xticks(rotation=0,fontsize=20)
-plt.yticks(fontsize=20)
-plt.tick_params(length=10,width=1.2)
-plt.xlabel("$\sigma_\gamma$(cadence) / $\sigma_\gamma$(baseline2018a)",fontsize=20)
-plt.ylim(0.5,13.5)
-plt.show()
-plt.savefig('FOM_Gamma_OSTF_new.pdf')

### 2019-05-10 doc/src/fom_aksgg.py
SNPV_FoM_Gamma Figure for OSTF

@@ -2,7 +2,6 @@
 
 import numpy as np 
 import matplotlib.pyplot as plt
-#import healpy as hp
 import os
 import re
 import partials
@@ -53,7 +52,7 @@ so =np.argsort(rt)
 dtp = []
 list = []
 
-for index in range(len(rt)-1):
+for index in range(len(rt)):
     list= (so[index],fom_10[so[index]])
     dtp.append((list))
 
@@ -82,6 +81,8 @@ plt.yticks(fontsize=20)
 plt.tick_params(length=10,width=1.2)
 plt.xlabel("$\sigma_\gamma^{-1}$(cadence) / $\sigma_\gamma^{-1}$(baseline2018a)",fontsize=20)
 plt.ylim(0.5,13.5)
+plt.vlines(1.,0.5,13.5,colors='k',linestyles='dashed',label="baseline2018a") 
+plt.legend()
 plt.show()
-plt.savefig('FOM_Gamma_OSTF_new.pdf')
+plt.savefig('sn_pv_FOM_Gamma.pdf')
 

### 2019-05-10 doc/src/fom_aksgg.py
SNPV_FoM_Gamma Figure for OSTF

@@ -84,5 +84,5 @@ plt.ylim(0.5,13.5)
 plt.vlines(1.,0.5,13.5,colors='k',linestyles='dashed',label="baseline2018a") 
 plt.legend()
 plt.show()
-plt.savefig('sn_pv_FOM_Gamma.pdf')
+plt.savefig('sn_pv_FoM_Gamma.pdf')
 

### 2019-05-10 doc/src/fom.py
rectified missing ^2 exponent

@@ -79,10 +79,10 @@ plt.barh(tp_cadence,tp_fom_10, alpha=0.65, color="black")
 plt.xticks(rotation=0,fontsize=20)
 plt.yticks(fontsize=20)
 plt.tick_params(length=10,width=1.2)
-plt.xlabel("$\sigma_\gamma^{-1}$(cadence) / $\sigma_\gamma^{-1}$(baseline2018a)",fontsize=20)
+plt.xlabel("$\sigma_\gamma^{-2}$(cadence) / $\sigma_\gamma^{-2}$(baseline2018a)",fontsize=20)
 plt.ylim(0.5,13.5)
 plt.vlines(1.,0.5,13.5,colors='k',linestyles='dashed',label="baseline2018a") 
 plt.legend()
 plt.show()
-plt.savefig('sn_pv_FoM_Gamma.pdf')
+#plt.savefig('sn_pv_FoM_Gamma.pdf')
 

### 2019-05-10 doc/src/fom.py
rm ylim that removes 'worst' cadence

@@ -70,8 +70,8 @@ for index in range(len(dtp)):
     tp_cadence.append(rt[int(a[0])])
     tp_fom_10.append(a[1])
 
-#print("Footprints : ", tp_cadence)
-#print("FoM over 10 years : ", tp_fom_10)
+print("Footprints : ", tp_cadence)
+print("FoM over 10 years : ", tp_fom_10)
 
 fig = plt.figure(figsize=(10,15))
 ax = fig.add_subplot(1, 1, 1)
@@ -80,8 +80,8 @@ plt.xticks(rotation=0,fontsize=20)
 plt.yticks(fontsize=20)
 plt.tick_params(length=10,width=1.2)
 plt.xlabel("$\sigma_\gamma^{-2}$(cadence) / $\sigma_\gamma^{-2}$(baseline2018a)",fontsize=20)
-plt.ylim(0.5,13.5)
-plt.vlines(1.,0.5,13.5,colors='k',linestyles='dashed',label="baseline2018a") 
+#plt.ylim(0.5,13.5)
+plt.vlines(1.,0.,14.,colors='k',linestyles='dashed',label="baseline2018a") 
 plt.legend()
 plt.show()
 #plt.savefig('sn_pv_FoM_Gamma.pdf')

### 2019-05-10 doc/src/fom.py
rm ylim that removes 'worst' cadence

@@ -70,18 +70,18 @@ for index in range(len(dtp)):
     tp_cadence.append(rt[int(a[0])])
     tp_fom_10.append(a[1])
 
-print("Footprints : ", tp_cadence)
-print("FoM over 10 years : ", tp_fom_10)
+#print("Footprints : ", tp_cadence)
+#print("FoM over 10 years : ", tp_fom_10)
 
 fig = plt.figure(figsize=(10,15))
 ax = fig.add_subplot(1, 1, 1)
-plt.barh(tp_cadence,tp_fom_10, alpha=0.65, color="black")
+plt.barh(tp_cadence,tp_fom_10, alpha=0.5, color="black")
 plt.xticks(rotation=0,fontsize=20)
 plt.yticks(fontsize=20)
 plt.tick_params(length=10,width=1.2)
 plt.xlabel("$\sigma_\gamma^{-2}$(cadence) / $\sigma_\gamma^{-2}$(baseline2018a)",fontsize=20)
-#plt.ylim(0.5,13.5)
-plt.vlines(1.,0.,14.,colors='k',linestyles='dashed',label="baseline2018a") 
+plt.ylim(-0.5,13.5)
+plt.vlines(1.,-0.5,13.5,colors='k',linestyles='dashed',label="baseline2018a") 
 plt.legend()
 plt.show()
 #plt.savefig('sn_pv_FoM_Gamma.pdf')

### 2019-05-10 doc/point.tex
cleaned up version

@@
 
-The trace term is
+The trace term depends on only the zero-lag correlation
 \begin{align}
@@
  & = \bar{N} \tr(\xi_{\delta\delta,\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
- & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i)
+ & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i).
 \end{align}
 
-The remaining term to the leading term in $\bar{N}$ is
+The term in the Fisher matrix, to the leading term in $\bar{N}$ is
 \begin{align}
  \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1} \textbf{S}_\beta) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\alpha \textbf{C}_0^{-1} \textbf{S}_\beta)  \\
- & = \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
- & = \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
+ & = \bar{N}^2 \tr(\xi_{\delta \delta,\alpha}\xi_{\delta \delta,\beta})+
+ \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
+ & = \int n(x) n(y) \xi_{\delta \delta,\alpha}(x,y)\xi_{\delta \delta,\beta}(y,x) d^3x d^3y \\
+ & + \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
  & +  \sum_{i,j\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\alpha}(x_i,x_j)  \frac{1}{\sigma^2_{v,j}}  \xi_{vv,\beta}(x_j,x_i).
@@
 
+\end{comment}
+
 We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
@@
 \end{align*}
+\begin{comment}
 

### 2019-05-27 doc/point.tex
added appendix deriving estimators.  Need to check minus sign

@@
 
-The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,
+%The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,
+The data themselves are denoted as $\textbf{y} = [N_1, N_2, \hdots, N_{N_{cell}}, v_1, \hdots, v_{N_{gal}}]$
+and are modeled as $\textbf{y}_0 =[\bar{N}, \bar{N}, \hdots, \bar{N}, v^0_{1}, \hdots, v^0_{N_{gal}}]$.  For now we assume that the model is not fixed but depends
+on linear parameters $p_\alpha$ such that
+ $\textbf{y}_0 = p_\alpha \textbf{x}_\alpha$.
+Their difference is  denoted as $\Delta \textbf{y}= \textbf{y} - \textbf{y}_0 $ and $\Delta v= v-v^0$.  (Although the note is written in terms of velocities
+in practice we expect to work with magnitudes as the are observed.)
 
@@
 \[
-\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\textbf{x}^T\textbf{C}^{-1}\textbf{x}\right\}.
+\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}\right\}.
 \]\par
@@
 \begin{equation}
--2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \textbf{x}^T\textbf{C}^{-1}\textbf{x} + \text{const}.
+-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y} + \text{const}.
 \label{log:eqn}
@@
 
-The estimator and Fisher Matrix for this Gaussian likelihood are known
-\begin{equation}
-\hat{\theta}_\alpha = \frac{1}{2} F^{-1}_{\alpha \beta} [(\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]
-\end{equation}
-and
-\begin{equation}
-F_{\alpha\beta} = \frac{1}{2} \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0 \textbf{S}_\beta).
-\end{equation}
+The estimator and Fisher matrix for this Gaussian likelihood are
+\begin{align}
+\hat{\theta}_\alpha & = \frac{1}{2} F^{(\theta \theta)\,-1}_{\alpha \beta} [(\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]\\
+\hat{p}_\alpha  &=  F^{(pp)\, -1}_{\alpha \beta}  (\textbf{y}^T  \textbf{C}^{-1} x_\beta)
+\end{align}
+where  
+\begin{align}
+F^{(\theta\theta)}_{\alpha\beta}& = \frac{1}{2} \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1}  \textbf{S}_\beta)\\
+F^{(pp)}_{\alpha\beta} & =x^T_\alpha \textbf{C}^{-1} x_\beta\\
+F^{(\theta p)}_{\alpha\beta} & = 0.
+\end{align}
 
+We focus on the parameters $\theta$ related to the covariance matrix .
 The term
 \begin{align}
-\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \textbf{x}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \textbf{x}\\
+\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \Delta \textbf{y}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \Delta \textbf{y}\\
 &=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
@@
 \begin{align*}
-\textbf{x}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{x} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
+\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
 & +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
@@
 
-First, we focus on the  term $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
-by   introducing the matrices $\textbf{M}$ and $\textbf{X}$,
-where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$ such that $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
-Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
+First, we focus on the  term $\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
+by   introducing the matrices $\textbf{M}$ and $\Delta \textbf{y}$,
+where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \Delta \textbf{y} \textbf{M}$ such that $\Delta \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
+Now, our matrix $\textbf{C}$ can be re-written in terms of $\Delta \textbf{y}$ and $\textbf{M}$:
 \begin{equation}
-\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M}.
+\textbf{C} = \textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M}.
 \end{equation}
@@
 \begin{align}
-\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\textbf{X}+A^2 \textbf{X}^2 + \ldots) \textbf{M}^{-1} \\
+\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\Delta \textbf{y}+A^2 \Delta \textbf{y}^2 + \ldots) \textbf{M}^{-1} \\
 & = \textbf{C}_0^{-1}   -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +\frac{A^2}{2} \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +O(A^3) + \ldots 
@@
 \begin{align}
-\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
+\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
 \frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
@@
 \begin{align*}
-\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
+\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
 & +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
@@
 
-\end{comment}
-
-We evaluate $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ by splitting up the counts and velocities.  The data vector $\textbf{x}$ is composed of the counts
+We evaluate $\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y}$ by splitting up the counts and velocities.  The data vector $\Delta \textbf{y}$ is composed of the counts
  $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
@@
 \end{align*}
-\begin{comment}
 
@@
 \begin{align}
-\textbf{x}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \textbf{x}
+\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}
 & = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
@@
 
-We have assembled the contributions of $\textbf{x}^T \textbf{C}^{-1}\textbf{x}$ to second order in $A$.
+We have assembled the contributions of $\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y}$ to second order in $A$.
 
@@
 \begin{align*}
-\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\textbf{X})\textbf{M})) \\
-&= \ln\det(\textbf{I}+A\textbf{X}) +\text{const} \\
-&= A \tr \textbf{X} -\frac{A^2}{2}\tr \textbf{X}^2+ O(A^3) + \ldots\\
+\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M})) \\
+&= \ln\det(\textbf{I}+A\Delta \textbf{y}) +\text{const} \\
+&= A \tr \Delta \textbf{y} -\frac{A^2}{2}\tr \Delta \textbf{y}^2+ O(A^3) + \ldots\\
 \end{align*}
@@
 \begin{equation}
-\textbf{X} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
+\Delta \textbf{y} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
 \begin{pmatrix}
@@
 
-The trace of $\textbf{X}^2$ has terms, to leading order in $\bar{N}$, has terms
+The trace of $\Delta \textbf{y}^2$ has terms, to leading order in $\bar{N}$, has terms
 \begin{equation}
@@
 Given the enormous size of $\textbf{C}$, computing the log likelihood directly is not possible. In the following, we seek an approximation method accurate to first order in $A$. \par
-First, we focus on the second term, $\textbf{x}^T\textbf{C}^{-1}\textbf{x}$.
+First, we focus on the second term, $\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}$.
 Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{C}_0$ into an identity matrix.
-To do so, we introduce matrix $\textbf{M}$ and $\textbf{X}$.
-We let $\textbf{M}=\sqrt{\textbf{C}_0}$, and $\textbf{S} = \textbf{M} \textbf{X} \textbf{M}$.
-Then $\textbf{X}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
-Now, our matrix $\textbf{C}$ can be re-written in terms of $\textbf{X}$ and $\textbf{M}$:
+To do so, we introduce matrix $\textbf{M}$ and $\Delta \textbf{y}$.
+We let $\textbf{M}=\sqrt{\textbf{C}_0}$, and $\textbf{S} = \textbf{M} \Delta \textbf{y} \textbf{M}$.
+Then $\Delta \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
+Now, our matrix $\textbf{C}$ can be re-written in terms of $\Delta \textbf{y}$ and $\textbf{M}$:
 \begin{equation}
-\textbf{C} = \textbf{M}(\textbf{I}+A\textbf{X})\textbf{M},
+\textbf{C} = \textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M},
 \end{equation}
@@
 \[
-\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\textbf{X}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
+\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\Delta \textbf{y}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
 \] \par
@@
 \begin{align*}
-s &=\textbf{x}^T \textbf{C}^{-1}\textbf{x} \\
-&= -A\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right).
+s &=\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y} \\
+&= -A\left(\Delta \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\Delta \textbf{y}\right).
 \end{align*} \par
@@
 \[
--\left(\textbf{x}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\textbf{x}\right) = s_1 +s_2+\hdots +s_6
+-\left(\Delta \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\Delta \textbf{y}\right) = s_1 +s_2+\hdots +s_6
 \]
@@
 \end{comment}
+
+\appendix
+\section{Estimator and Fisher Matrix}
+Suppose we have a likelihood that has parameters $\theta$ and $p$ such that
+$$f\equiv-\ln{\mathcal{L}(\theta,p)} = \frac{1}{2}\left(\ln{(\det(C(\theta)))} + (y-y_0 )^T C^{-1} (y-y_0)\right),$$
+where $C$ is invertible (the covariance matrix is positive definite),
+$$y_0 = p_\alpha x_\alpha,$$ and
+ $$C=C_0 + \theta_\alpha S_\alpha.$$
+Then
+\begin{align}
+\frac{df}{d\theta_\alpha} &= \frac{1}{2}\left(\tr(C^{-1}S_\alpha)   + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right)\\ 
+\frac{df}{dp_{\alpha}} &= -(y-y_0)^T C^{-1} x_\alpha\\ 
+\frac{df^2}{d\theta_{\alpha} d\theta_\beta} &= \frac{1}{2}  \tr (C^{-1} S_\beta C^{-1}  S_\alpha) +  (y-y_0)^T (
+C^{-1} S_{\beta } C^{-1} S_\alpha C^{-1}
+)
+(y-y_0) \\ 
+\frac{df^2}{dp_{\alpha} dp_\beta} &= x^T_\beta C^{-1} x_\alpha\\ 
+\frac{df^2}{dp_\alpha \,d\theta_\beta} &=-  (y-y_0)^T C^{-1}  S _\beta C^{-1} x_\alpha\\
+F_{\alpha \beta}^{(\theta \theta)} & \equiv \left \langle  \frac{df^2}{d\theta_{\alpha} d\theta_\beta} \right \rangle = \frac{1}{2} \tr (C^{-1} S_\beta C^{-1}  S_\alpha) \\ 
+F_{\alpha \beta}^{(p p)}  & \equiv \left \langle  \frac{df^2}{dp_{\alpha} dp_\beta} \right \rangle= x^T_\beta C^{-1} x_\alpha\\ 
+F_{\alpha \beta}^{(\theta p)}  & \equiv \left \langle  \frac{df^2}{dp_\alpha \,d\theta_\beta} \right \rangle =0.
+\end{align}
+
+We are interested in the properties of the function at its minimum, i.e. at $(\hat{\theta},\hat{p})$ occurs where $df/d\theta_\alpha=0 $ and $df/dp{_\alpha}=0$.
+\begin{align}
+0 & = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C)  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+& = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} (C_0 + \hat{\theta}_\beta S_\beta) )  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+& =\hat{ \theta}_\beta F^{(\theta \theta)}_{\alpha \beta} +  \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C_0)  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+\hat{ \theta}_\beta  & = - \frac{1}{2} F^{(\theta \theta)\, -1}_{\beta \alpha} \left(\tr(C^{-1}S_\alpha C^{-1} C_0)  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+0 &=  (y-\hat{p}_\beta x_\beta )^T C^{-1} x_\alpha \\
+& = y^TC^{-1} x_\alpha - \hat{p}_\beta F^{(pp)}_{\alpha \beta} \\
+ \hat{p}_\beta  & =   F^{(pp)\,-1}_{\beta \alpha} (y^TC^{-1} x_\alpha ).
+\end{align}
+
 \begin{thebibliography}{9}

### 2019-06-10 doc/point.tex
more edits before sending to Chao

@@
 
-There is sufficient information in the distribution of peculiar velocities and overdensities of galaxies to test general relativity on the largest scale.
+%There is sufficient information in the distribution of peculiar velocities and overdensities of galaxies to test general relativity on the largest scale.
 Suppose we have a region that encloses $N_{gal}$ number of galaxies in the universe.
@@
 
+Consider a galaxy count and peculiar velocity survey that covers a volume, which is divided into $N_{cell}$ equal-volume cells.  Each cell will contain some
+number of survey galaxies; when the cell volume is large then each cell is likely to contain many galaxies whereas when the cell volume is small each cell is likely
+to have either one or no galaxies.  Similarly, when the cell volume is large  its  peculiar velocity is determined combining all the velocity measurements therein,
+whereas when the volume is small the cell velocity is likely to be that of the one galaxy therein or there is no measurement, alternatively viewed as an infinite
+uncertainty measurement.  We denote $N_{gal}$ as the number of cells that contains at least one galaxy, at the large cell volume limit
+$N_{gal}=N_{cell}$ whereas in the limit of infinitely small cell volumes it is
+the number of galaxies in the survey.
+
 %The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,
-The data themselves are denoted as $\textbf{y} = [N_1, N_2, \hdots, N_{N_{cell}}, v_1, \hdots, v_{N_{gal}}]$
-and are modeled as $\textbf{y}_0 =[\bar{N}, \bar{N}, \hdots, \bar{N}, v^0_{1}, \hdots, v^0_{N_{gal}}]$.  For now we assume that the model is not fixed but depends
+The data in the cells are denoted as $\textbf{y} = [N_1, N_2, \hdots, N_{N_{cell}}, v_1, \hdots, v_{N_{gal}}]$
+and are modeled as $\textbf{y}_0 =[\bar{N}, \bar{N}, \hdots, \bar{N}, v_{0,1}, \hdots, v_{0,N_{gal}}]$.  For now we assume that the model is not fixed but depends
 on linear parameters $p_\alpha$ such that
  $\textbf{y}_0 = p_\alpha \textbf{x}_\alpha$.
-Their difference is  denoted as $\Delta \textbf{y}= \textbf{y} - \textbf{y}_0 $ and $\Delta v= v-v^0$.  (Although the note is written in terms of velocities
-in practice we expect to work with magnitudes as the are observed.)
+Their difference is  denoted as $\Delta \textbf{y}= \textbf{y} - \textbf{y}_0 $ and $\Delta v= v-v_0$.  (Although the note is written in terms of velocities
+in practice we expect to work with magnitudes.)
 
-We postulate that deviations from expected variance $\bar{N}$ and expected peculiar velocity $\bar{v}$ are Gaussian.
-Then the above correlation matrices can be used to construct a block matrix $\textbf{C}$, parameterized by $A$ and $\bm{\sigma}_v$, a vector of shot noise for peculiar velocity measurement:
+We model the observations as being drawn from a Gaussian distribution.
+The variance in number counts in a cell is $\bar{N}$ and the vector of peculiar velocity measurement in the cells is denoted $\bm{\sigma}_v$,
+  Then the above correlation matrices can be used to construct a block matrix $\textbf{C}$:
 \begin{align}
@@
 \end{align}
-where $\xi_{gg}$, $\xi_{vv}$, and $\bar{N}\xi_{g v}$ are the galaxy overdensity and peculiar velocity correlation functions and their cross-correlation
+where $\xi_{gg}$, $\xi_{vv}$, and $\xi_{g v}$ are the galaxy overdensity and peculiar velocity correlation functions and their cross-correlation
 respectively.  
-
-The reason we have $\bar{N}$ as the variance for overdensity is that our small galaxy cells are best modeled by Poisson statistics, in which the mean equals the variance.
-In addition, note that the dimension of $C$ is $N_{cell}+N_{gal}$ by $N_{cell}+N_{gal}$.
+The dimension of $C$ is $N_{cell}+N_{gal}$ by $N_{cell}+N_{gal}$.
 For the peculiar velocity covariance, we assign $\sigma^2=\infty$ to cells that do not contain any galaxy, and one can show that removing rows and columns that contain those cells has no effect but a constant offset on what we will compute later.
 
-The science parameters $\theta$ enter through these correlation functions, and may include $\theta \in \{fD, bD, \gamma, \Omega_{M_0},\ldots\}$ and others
+The science parameters $\theta$ enter through the correlation functions, and may include $\theta \in \{fD, bD, \gamma, \Omega_{M_0},\ldots\}$ and others
 that describe the matter power spectrum.
@@
 \propto  (bD + fD\mu^2)fD$, where $b$ is the galaxy bias and $\mu\equiv \cos{(\hat{k} \cdot \hat{r})}$ where $\hat{r}$ is the direction of
-the line of sight.  
+the line of sight.   There are also the parameters $p$ that linearly affect the mean of the Gaussian, a primary example being the combination
+$M+5\log{H_0}$ where $M$ is the SN absolute magnitude and $H_0$ is the Hubble constant.
+
+The Gaussian likelihood as a function of $\theta$ is
+\[
+\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}\right\}.
+\]
+and
+\begin{equation}
+-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y} + \text{const}.
+\label{log:eqn}
+\end{equation}
+
 Assuming linear dependence of the parameters,
@@
 \right)
-+ \sum_\theta \theta
++ \sum_\alpha \theta_\alpha
 \begin{pmatrix}
-\bar{N}^2\xi_{gg,\theta} & \bar{N}\xi_{g v,\theta} \\
-\bar{N}\xi_{v g,\theta} & \xi_{vv,\theta}
+\bar{N}^2\xi_{gg,{\theta_\alpha}} & \bar{N}\xi_{g v,{\theta_\alpha}} \\
+\bar{N}\xi_{v g,{\theta_\alpha}} & \xi_{vv,{\theta_\alpha}}
 \end{pmatrix}\\
 &=
-\left(
-\begin{array}{cc}
-\bar{N}\textbf{I} & 0 \nonumber \\
-0 & \bm{\sigma}_v^2\textbf{I}
-\end{array}
-\right)
-+ \sum_\alpha \theta_\alpha \textbf{S}_\alpha .
+\textbf{C}_0
++ \sum_\alpha \theta_\alpha \textbf{S}_\alpha,
 \end{align} 
+where the comma represents the partial derivative.
 
+%\begin{comment}
+%The inverse can be expanded as
+%\begin{align}
+%\textbf{C}^{-1}
+%&=
+%\textbf{C}_0^{-1} \left( 1 +  \sum_\alpha \theta_\alpha \textbf{S}_\alpha \textbf{C}_0^{-1} \right)^{-1}\\
+%&=
+%\textbf{C}_0^{-1} \sum_{i=0}^{\infty} (-1)^i \left(\sum_\alpha \theta_\alpha \textbf{S}_\alpha \textbf{C}_0^{-1} \right)^i.
+%\end{align} 
+%
+%In our case
+%\begin{align}
+% \textbf{S}_\alpha \textbf{C}_0^{-1} & = 
+% \begin{pmatrix}
+% \bar{N}\xi_{gg,\theta_\alpha} & \frac{\bar{N}}{ \bm{\sigma}_v^2} \xi_{gv,\theta_\alpha}\\
+% \xi_{vg,\theta_\alpha} & \frac{1}{ \bm{\sigma}_v^2} \xi_{vv,\theta_\alpha}\\
+% \end{pmatrix}
+% \end{align}
+%
+%We will eventually take the limit where $\bar{N} \rightarrow 0$.
+%Then
+%\begin{align}
+%\left(\sum_\alpha \theta_\alpha \textbf{S}_\alpha \textbf{C}_0^{-1} \right)^i &  \approx 
+%\sum_{\alpha}  \ldots \sum_{\delta}  \theta_{\alpha} \ldots \theta_\delta
+%  \begin{pmatrix}
+% \frac{\bar{N}}{ \bm{\sigma}_v^{2(i-1)}} \xi_{gv,\theta_\alpha}  \xi_{vg,\theta_\beta} \ldots  \xi_{vg,\theta_\gamma} & \frac{\bar{N}}{ \bm{\sigma}_v^{2i}} \xi_{gv,\theta_\alpha} \xi_{vv,\theta_\beta} \ldots  \xi_{vv,\theta_\gamma} \\
+% \frac{1}{ \bm{\sigma}_v^{2(i-1)}} \xi_{vv,\theta_\alpha} \xi_{vg,\theta_\beta}\ldots  \xi_{vg,\theta_\gamma}   & \frac{1}{ \bm{\sigma}_v^{2i}} \xi_{vv,\theta_\alpha} \xi_{vv,\theta_\beta}\ldots  \xi_{vv,\theta_\gamma}\\
+% \end{pmatrix}.
+%\end{align}
+%\end{comment}
 %\left(
@@
 
-The Gaussian likelihood as a function of $\theta$
-\[
-\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}\right\}.
-\]\par
-To make the expression cleaner, we take the log of the likelihood and focus on the $A$-dependent part:
-\begin{equation}
--2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y} + \text{const}.
-\label{log:eqn}
-\end{equation}
-
-The estimator and Fisher matrix for this Gaussian likelihood are
+The estimator and Fisher matrix for this Gaussian likelihood are well known (and derived in  \S\ref{appendix:sec}).
 \begin{align}
-\hat{\theta}_\alpha & = \frac{1}{2} F^{(\theta \theta)\,-1}_{\alpha \beta} [(\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)]\\
+\hat{\theta}_\alpha & = \frac{1}{2} F^{(\theta \theta)\,-1}_{\alpha \beta} [(\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)] \nonumber \\
 \hat{p}_\alpha  &=  F^{(pp)\, -1}_{\alpha \beta}  (\textbf{y}^T  \textbf{C}^{-1} x_\beta)
+\label{estimators:eqn}
 \end{align}
@@
 
-We focus on the parameters $\theta$ related to the covariance matrix .
-The term
-\begin{align}
-\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \Delta \textbf{y}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \Delta \textbf{y}\\
-&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
-\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
-\end{align}
-Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
-for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i,j\in N_{cell}} = \int n dV$, where $n$ is the galaxy number density,
+%Let us consider the limit where there are an infinite number of cells and $\bar{N} \rightarrow 0$ and $\bar{N}\sum_{i,j\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density,
+Denoting $\textbf{M}_\beta = \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1}$ and splitting the calculation into quadrants  corresponding to the galaxy and velocity indices,
+the complicated terms are
 \begin{align*}
-\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \sum_{i,j\in N_{gal}} \xi_{\delta\delta,\theta_\beta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
-& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
-& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
-& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
-&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
+\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}  
+ %& = \frac{1}{\bar{N}^2}\int \int n(x_1) n(x_2) (N(x_1)-\bar{N}) M_{gg}(x_1,x_2) (N(x_2)-\bar{N})  d^3x_1  d^3x_2\\
+%& +   \frac{2}{\bar{N}}\int \sum_{i=1}^{N} n(x_1) (N(x_1)-\bar{N}) M_{gv}(x_1,x_i) \Delta v_i  d^3x_1 \\
+%& + \sum_{i=1}^{N}  \sum_{j=1}^{N}  \Delta v_i M_{vv}(x_i,x_j) \Delta v_j\\
+&= \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}} N_i M_{\beta,gg}(x_i,x_j) N_j 
+ - 2\bar{N}   \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}}  M_{\beta,gg}(x_i,x_j) N_j \\
+& +\bar{N}^2 \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}}  M_{\beta,gg}(x_i,x_j) 
+ +   2 \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  N_i M_{\beta,gv}(x_i,x_j) \Delta v_j  \\
+& -   2 \bar{N}  \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}}  M_{\beta,gv}(x_i,x_j) \Delta v_j 
++ \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  \Delta v_i M_{\beta,vv}(x_i,x_j) \Delta v_j ,
+\end{align*}
+\begin{align*}
+\tr (\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{S}_\alpha) & 
+=  \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}} M_{\beta,gg}(x_i,x_j)\textbf{S}_{\alpha,gg}(x_j,x_i)  +  2 \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}} M_{\beta,gv}(x_i,x_j)\textbf{S}_{\alpha,vg} (x_j,x_i)\\
+& + \sum_{i=1}^{N_{\text gal}} \sum_{j=1}^{N_{\text gal}} M_{\beta,vv}(x_i,x_j) \textbf{S}_{\alpha,vv}(x_j,x_i),
+\end{align*}
+\begin{align*}
+\tr (\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & 
+=  \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}} M_{\beta,gg}(x_i,x_j)\textbf{C}_{0,gg}(x_j,x_i)  + \sum_{i=1}^{N_{\text gal}} \sum_{j=1}^{N_{\text gal}} M_{\beta,vv}(x_i,x_j) \textbf{C}_{0,vv}(x_j,x_i)\\
+ & 
+= \bar{N} \sum_{i=1}^{N_{\text cell}} M_{\beta,gg}(x_i,x_i)  + \sum_{i=1}^{N_{\text gal}}  M_{\beta,vv}(x_i,x_i) \sigma^{2}_{vi} .
 \end{align*}
 
-The trace term depends on only the zero-lag correlation
-\begin{align}
- \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\beta) \\
- & = \bar{N} \tr(\xi_{\delta\delta,\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
- & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i).
-\end{align}
+An interesting limit is when there are infinitely small cell sizes.  Then $N_{gal}$ (the number of cells with galaxies) approaches the number
+of galaxies, those cells have $N_i=1$,  and $\bar{N}\sum_{i\in N_{cell}} = n \int dV$, where $n$ is the galaxy number density.
+The calculation then is then `binless', with sums over sample objects and integrals to account for the volume where there are no objects.
+In this limit, a Poisson rather than Gaussian model distribution better describes expectations.
 
-The term in the Fisher matrix, to the leading term in $\bar{N}$ is
-\begin{align}
- \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1} \textbf{S}_\beta) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\alpha \textbf{C}_0^{-1} \textbf{S}_\beta)  \\
- & = \bar{N}^2 \tr(\xi_{\delta \delta,\alpha}\xi_{\delta \delta,\beta})+
- \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
- & = \int n(x) n(y) \xi_{\delta \delta,\alpha}(x,y)\xi_{\delta \delta,\beta}(y,x) d^3x d^3y \\
- & + \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
- & +  \sum_{i,j\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\alpha}(x_i,x_j)  \frac{1}{\sigma^2_{v,j}}  \xi_{vv,\beta}(x_j,x_i).
-\end{align}
+Finding the roots of Eq.~\ref{estimators:eqn} to get the estimators is non-trivial.  In particular the $\hat{\theta}$ appears in $\mathbf{C}^{-1}$, a large matrix that
+isn't obviously easy to invert.  Using the renormalization group, \cite{2019PhRvD..99d3538M} reparameterizes the likelihood such that much of the parameter-dependence
+gets shifted to other terms leaving a smaller
+band-diagonal matrix that remains to be inverted.  Applying his methodology to our problem is an interesting direction to pursue.
+
+For our purposes, having a precise answer is probably unnecessary.
+Solving the right hand side of Eq.~\ref{estimators:eqn} to lowest (zeroth) order in $\theta$ gives a useful approximate answer.  Replacing
+$\textbf{C}^{-1}$ with the diagonal $\textbf{C}_0^{-1}$ gives
+\begin{align*}
+\textbf{M}_{\beta,gg} & \approx \xi_{gg,\theta_\beta}\\
+\textbf{M}_{\beta,gv} & \approx \xi_{gv,\theta_\beta} \bm{\sigma}^{-2}_v\\
+\textbf{M}_{\beta,vv} & \approx \bm{\sigma}^{-2}_v \xi_{vv,\theta_\beta} \bm{\sigma}^{-2}_v .
+\end{align*}
+so that
+\begin{align*}
+\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}  
+&\approx \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}} N_i \xi_{gg,\theta_\beta}(x_i,x_j) N_j 
+ - 2\bar{N}   \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}}  \xi_{gg,\theta_\beta}(x_i,x_j)N_j \\
+& +\bar{N}^2 \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}} \xi_{gg,\theta_\beta}(x_i,x_j)
+ +   2 \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  N_i  \xi_{gv,\theta_\beta}(x_i,x_j)\frac{\Delta v_j}{\sigma^2_{v,j}} \\
+& -   2 \bar{N}  \sum_{i=1}^{N_{\text cell}} \sum_{i=1}^{N_{\text gal}}   \xi_{gv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma^2_{v,j}} 
++ \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}   \frac{\Delta v_i}{\sigma^2_{v,i}}  \xi_{vv,\theta_\beta}(x_i,x_j)  \frac{\Delta v_j}{\sigma^2_{v,j}},\\
+%\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \sum_{i,j\in N_{gal}} N_i N_j \xi_{\delta\delta,\theta_\beta}(x_i,x_j) -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
+%& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
+%& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
+%& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
+%&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
+%&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
+\end{align*}
+\begin{align*}
+\tr (\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{S}_\alpha) & 
+= \bar{N}^2  \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}} \xi_{gg,\theta_\beta}(x_i,x_j)\xi_{gg,\theta_\alpha}(x_j,x_i) 
++2 \bar{N}   \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}}\frac{1}{\sigma^{2}_{vj}}   \xi_{gv,\theta_\beta} (x_i,x_j)\xi_{\alpha,vg} (x_j,x_i)\\
+& + \sum_{i=1}^{N_{\text gal}} \sum_{j=1}^{N_{\text gal}} \frac{1}{\sigma^{2}_{vi} \sigma^{2}_{vj}} \xi_{vv,\theta_\beta}(x_i,x_j)\xi_{vv,\theta_\alpha}(x_j,x_i) 
+\end{align*}
+\begin{align*}
+\tr (\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & 
+=   \bar{N} \sum_{i=1}^{N_{\text cell}}  \xi_{gg,\theta_\beta}(x_i,x_i)  + \sum_{i=1}^{N_{\text gal}} \frac{1}{ \sigma^{2}_{vi}}\xi_{vv,\theta_\alpha}(x_i,x_i)
+\end{align*}
+
+The SN~Ia rate is $2.69 \times 10^{-5} \text{Mpc}^{-3} \text{yr}^{-1}$, meaning that after 10 years there
+will be around 1 supernova every 3700~$\text{Mpc}^{3}$.
+The Universe is $2.4 \times 10^9$~Mpc$^3$ out to $z=0.2$.  A rough bound on the number of necessary cells is $6\times 10^{5}$.
+When taking the linear approximation the calculation scales as $N_{cell}^2$, so this is possible to do on a laptop.
+The exact solution requires matrix factorizations that scale as $N_{cell}^3$, which is not so possible on a laptop.
+
+%%\begin{comment}
+%%We focus on the parameters $\theta$ related to the covariance matrix.
+%%The term
+%%\begin{align}
+%%\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \Delta \textbf{y}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \Delta \textbf{y}\\
+%%&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
+%%\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+%%\end{align}
+%%Noting that $N_i=1$ when the cell $i$ contains a galaxy and zero otherwise and that only cells with a peculiar velocity measurement contribute ($\sigma_v =\infty$
+%%for cells without such a measurement), and that in the limit of infinitely small cells, $\bar{N}\sum_{i\in N_{cell}} = \int n dV$, where $n$ is the galaxy number density,
+%%\end{comment}
+%
+%
+%
+%The trace term depends on only the zero-lag correlation
+%\begin{align}
+% \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\beta) \\
+% & = \bar{N} \tr(\xi_{\delta\delta,\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
+%% & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i).
+%\end{align}
+%
+%The term in the Fisher matrix, to the leading term in $\bar{N}$ is
+%\begin{align}
+% \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1} \textbf{S}_\beta) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\alpha \textbf{C}_0^{-1} \textbf{S}_\beta)  \\
+% & = \bar{N}^2 \tr(\xi_{\delta \delta,\alpha}\xi_{\delta \delta,\beta})+
+% \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
+% & = \int n(x) n(y) \xi_{\delta \delta,\alpha}(x,y)\xi_{\delta \delta,\beta}(y,x) d^3x d^3y \\
+% & + \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
+% & +  \sum_{i,j\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\alpha}(x_i,x_j)  \frac{1}{\sigma^2_{v,j}}  \xi_{vv,\beta}(x_j,x_i).
+%\end{align}
 
@@
 \section{Estimator and Fisher Matrix}
+\label{appendix:sec}
 Suppose we have a likelihood that has parameters $\theta$ and $p$ such that
@@
 \begin{align}
-\frac{df}{d\theta_\alpha} &= \frac{1}{2}\left(\tr(C^{-1}S_\alpha)   + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right)\\ 
+\frac{df}{d\theta_\alpha} &= \frac{1}{2}\left(\tr(C^{-1}S_\alpha)   - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right)\\ 
 \frac{df}{dp_{\alpha}} &= -(y-y_0)^T C^{-1} x_\alpha\\ 
-\frac{df^2}{d\theta_{\alpha} d\theta_\beta} &= \frac{1}{2}  \tr (C^{-1} S_\beta C^{-1}  S_\alpha) +  (y-y_0)^T (
-C^{-1} S_{\beta } C^{-1} S_\alpha C^{-1}
+\frac{df^2}{d\theta_{\alpha} d\theta_\beta} &=- \frac{1}{2}  \tr (C^{-1} S_\alpha C^{-1}  S_\beta) +  (y-y_0)^T (
+C^{-1} S_{\alpha} C^{-1} S_\beta C^{-1}
 )
 (y-y_0) \\ 
-\frac{df^2}{dp_{\alpha} dp_\beta} &= x^T_\beta C^{-1} x_\alpha\\ 
-\frac{df^2}{dp_\alpha \,d\theta_\beta} &=-  (y-y_0)^T C^{-1}  S _\beta C^{-1} x_\alpha\\
-F_{\alpha \beta}^{(\theta \theta)} & \equiv \left \langle  \frac{df^2}{d\theta_{\alpha} d\theta_\beta} \right \rangle = \frac{1}{2} \tr (C^{-1} S_\beta C^{-1}  S_\alpha) \\ 
-F_{\alpha \beta}^{(p p)}  & \equiv \left \langle  \frac{df^2}{dp_{\alpha} dp_\beta} \right \rangle= x^T_\beta C^{-1} x_\alpha\\ 
-F_{\alpha \beta}^{(\theta p)}  & \equiv \left \langle  \frac{df^2}{dp_\alpha \,d\theta_\beta} \right \rangle =0.
+\frac{df^2}{dp_{\alpha} dp_\beta} &= x^T_\alpha C^{-1} x_\beta\\ 
+\frac{df^2}{d\theta_\alpha\,dp_\beta} &= (y-y_0)^T C^{-1}  S _\alpha C^{-1} x_\beta.
 \end{align}
+For a Gaussian distribution
+\begin{align}
+F_{\alpha \beta}^{(\theta \theta)} & = \left \langle  \frac{df^2}{d\theta_{\alpha} d\theta_\beta} \right \rangle =  -\frac{1}{2} \tr (C^{-1} S_\alpha C^{-1}  S_\beta)    + \tr (C^{-1} S_\alpha C^{-1}  S_\beta) \\
+& = \frac{1}{2} \tr (C^{-1} S_\alpha C^{-1}  S_\beta) \\
+F_{\alpha \beta}^{(p p)}  & = \left \langle  \frac{df^2}{dp_{\alpha} dp_\beta} \right \rangle=  x^T_\alpha C^{-1} x_\beta\\ 
+F_{\alpha \beta}^{(\theta p)}  & = \left \langle  \frac{df^2}{d\theta_\alpha \,dp_\beta} \right \rangle =0,
+\end{align}
+where the ensemble averages in the Fisher matrices satisfy  $\langle y-y_0 \rangle = 0$, $\langle (y-y_0) (y-y_0)^T\rangle = C$.
 
@@
 \begin{align}
-0 & = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C)  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
-& = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} (C_0 + \hat{\theta}_\beta S_\beta) )  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
-& =\hat{ \theta}_\beta F^{(\theta \theta)}_{\alpha \beta} +  \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C_0)  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
-\hat{ \theta}_\beta  & = - \frac{1}{2} F^{(\theta \theta)\, -1}_{\beta \alpha} \left(\tr(C^{-1}S_\alpha C^{-1} C_0)  + (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+0 & = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C)  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+& = \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} (C_0 + \hat{\theta}_\beta S_\beta) )  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+& =\hat{ \theta}_\beta F^{(\theta \theta)}_{\alpha \beta} +  \frac{1}{2}\left(\tr(C^{-1}S_\alpha C^{-1} C_0)  - (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0)\right) \\
+\hat{ \theta}_\beta  & =  \frac{1}{2} F^{(\theta \theta)\, -1}_{\beta \alpha} \left( (y-y_0)^T  C^{-1} S_{\alpha} C^{-1}  (y-y_0) - \tr(C^{-1}S_\alpha C^{-1} C_0) \right) \\
 0 &=  (y-\hat{p}_\beta x_\beta )^T C^{-1} x_\alpha \\
@@
 C. Adams, C. Blake. 
-Improving constraints on the growth rate of structure by modelling the density-velocity cross-correlation in the 6dF Galaxy Survey. arXiv:1706.05205 [astro-ph.CO].
+Improving constraints on the growth rate of structure by modelling the density-velocity cross-correlation in the 6dF Galaxy Survey. arXiv:1706.05205 [astro-ph.CO]
+\bibitem[McDonald(2019)]{2019PhRvD..99d3538M} McDonald, P.\ 2019, PRD, 99, 43538
 \end{thebibliography}

### 2019-06-11 doc/src/partials.py
bug in OmegaM

@@ -38,7 +38,7 @@ def Cinverse_partial(m00,m11,m01,dm00,dm11,dm01):
         - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
 
 def OmegaM(a,OmegaM0=OmegaM0):
-    return OmegaM0/(OmegaM0 + (1-OmegaM0)*a**3)
+    return OmegaM0*a**3/(OmegaM0 + (1-OmegaM0)*a**3)
 
 def integrand_dlnfs(a,*args):
     return numpy.log(OmegaM(a,*args))* OmegaM(a,*args)**0.55 / a

### 2019-08-07 doc/point.tex
More generalized version

@@
 For any 2 cells, we can compute their correlation, a function of the distance $r$ between the cells only.
-In our application, we consider $\xi_{\delta\delta}$ (the overdensity-overdensity correlation), $\xi_{\delta v}$ (overdensity-velocity correlation), $\xi_{v\delta}$ (velocity-overdensity correlation), and $\xi_{vv}$ (velocity-velocity correlation).\par
+In our application, we consider $\xi_{\hat{\delta}\hat{\delta}}$ (the overdensity-overdensity correlation), $\xi_{\hat{\delta} v}$ (overdensity-velocity correlation), $\xi_{v\hat{\delta}}$ (velocity-overdensity correlation), and $\xi_{vv}$ (velocity-velocity correlation).\par
 
@@
 
-%The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \Delta v_1, \hdots, \Delta v_{N_{gal}}]^T$, where $\Delta v$ is the deviation of the measured velocity from the expected,
+%The data  is $\textbf{x}= [N_1-\bar{N}, N_2-\bar{N}, \hdots, N_{N_{cell}}-\bar{N}, \hat{\delta} v_1, \hdots, \hat{\delta} v_{N_{gal}}]^T$, where $\hat{\delta} v$ is the deviation of the measured velocity from the expected,
 The data in the cells are denoted as $\textbf{y} = [N_1, N_2, \hdots, N_{N_{cell}}, v_1, \hdots, v_{N_{gal}}]$
@@
  $\textbf{y}_0 = p_\alpha \textbf{x}_\alpha$.
-Their difference is  denoted as $\Delta \textbf{y}= \textbf{y} - \textbf{y}_0 $ and $\Delta v= v-v_0$.  (Although the note is written in terms of velocities
+For example, there could be  a $p_\alpha = H_0$, $x_\alpha=0$ for the counts elements, and $x_\alpha=d_L$ for the velocity elements. 
+Their difference is  denoted as $\hat{\delta} \textbf{y}= \textbf{y} - \textbf{y}_0 $ and $\hat{\delta} v= v-v_0$.  (Although the note is written in terms of velocities
 in practice we expect to work with magnitudes.)
@@
 \[
-\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}\right\}.
+\mathcal{L}(\theta) = \frac{1}{\sqrt{\left(2\pi\right)^{N_{cell}+N_{gal}} \det \textbf{C}}}\exp\left\{-\frac{1}{2}\hat{\delta} \textbf{y}^T\textbf{C}^{-1}\hat{\delta} \textbf{y}\right\}.
 \]
@@
 \begin{equation}
--2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y} + \text{const}.
+-2\ln\mathcal{L}(\theta) = \ln\det\textbf{C} + \hat{\delta} \textbf{y}^T\textbf{C}^{-1}\hat{\delta} \textbf{y} + \text{const}.
 \label{log:eqn}
@@
 %\left(\sum_\alpha \theta_\alpha \textbf{S}_\alpha \textbf{C}_0^{-1} \right)^i &  \approx 
-%\sum_{\alpha}  \ldots \sum_{\delta}  \theta_{\alpha} \ldots \theta_\delta
+%\sum_{\alpha}  \ldots \sum_{\hat{\delta}}  \theta_{\alpha} \ldots \theta_\hat{\delta}
 %  \begin{pmatrix}
@@
 %\begin{array}{cc}
-%\bar{N}^2\xi_{\delta\delta} & \bar{N}\xi_{\delta v} \\
-%\bar{N}\xi_{\delta v} & \xi_{vv}
+%\bar{N}^2\xi_{\hat{\delta}\hat{\delta}} & \bar{N}\xi_{\hat{\delta} v} \\
+%\bar{N}\xi_{\hat{\delta} v} & \xi_{vv}
 %\end{array}
@@
 \begin{align}
-\hat{\theta}_\alpha & = \frac{1}{2} F^{(\theta \theta)\,-1}_{\alpha \beta} [(\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)] \nonumber \\
+\hat{\theta}_\alpha & = \frac{1}{2} F^{(\theta \theta)\,-1}_{\alpha \beta} [(\hat{\delta} \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \hat{\delta} \textbf{y}) - \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0)] \nonumber \\
 \hat{p}_\alpha  &=  F^{(pp)\, -1}_{\alpha \beta}  (\textbf{y}^T  \textbf{C}^{-1} x_\beta)
@@
 \begin{align*}
-\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}  
+\hat{\delta} \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \hat{\delta} \textbf{y}  
  %& = \frac{1}{\bar{N}^2}\int \int n(x_1) n(x_2) (N(x_1)-\bar{N}) M_{gg}(x_1,x_2) (N(x_2)-\bar{N})  d^3x_1  d^3x_2\\
-%& +   \frac{2}{\bar{N}}\int \sum_{i=1}^{N} n(x_1) (N(x_1)-\bar{N}) M_{gv}(x_1,x_i) \Delta v_i  d^3x_1 \\
-%& + \sum_{i=1}^{N}  \sum_{j=1}^{N}  \Delta v_i M_{vv}(x_i,x_j) \Delta v_j\\
+%& +   \frac{2}{\bar{N}}\int \sum_{i=1}^{N} n(x_1) (N(x_1)-\bar{N}) M_{gv}(x_1,x_i) \hat{\delta} v_i  d^3x_1 \\
+%& + \sum_{i=1}^{N}  \sum_{j=1}^{N}  \hat{\delta} v_i M_{vv}(x_i,x_j) \hat{\delta} v_j\\
 &= \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}} N_i M_{\beta,gg}(x_i,x_j) N_j 
@@
 & +\bar{N}^2 \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}}  M_{\beta,gg}(x_i,x_j) 
- +   2 \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  N_i M_{\beta,gv}(x_i,x_j) \Delta v_j  \\
-& -   2 \bar{N}  \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}}  M_{\beta,gv}(x_i,x_j) \Delta v_j 
-+ \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  \Delta v_i M_{\beta,vv}(x_i,x_j) \Delta v_j ,
+ +   2 \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  N_i M_{\beta,gv}(x_i,x_j) \hat{\delta} v_j  \\
+& -   2 \bar{N}  \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text gal}}  M_{\beta,gv}(x_i,x_j) \hat{\delta} v_j 
++ \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  \hat{\delta} v_i M_{\beta,vv}(x_i,x_j) \hat{\delta} v_j ,
 \end{align*}
@@
 \begin{align*}
-\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y}  
+\hat{\delta} \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \hat{\delta} \textbf{y}  
 &\approx \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}} N_i \xi_{gg,\theta_\beta}(x_i,x_j) N_j 
@@
 & +\bar{N}^2 \sum_{i=1}^{N_{\text cell}} \sum_{j=1}^{N_{\text cell}} \xi_{gg,\theta_\beta}(x_i,x_j)
- +   2 \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  N_i  \xi_{gv,\theta_\beta}(x_i,x_j)\frac{\Delta v_j}{\sigma^2_{v,j}} \\
-& -   2 \bar{N}  \sum_{i=1}^{N_{\text cell}} \sum_{i=1}^{N_{\text gal}}   \xi_{gv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma^2_{v,j}} 
-+ \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}   \frac{\Delta v_i}{\sigma^2_{v,i}}  \xi_{vv,\theta_\beta}(x_i,x_j)  \frac{\Delta v_j}{\sigma^2_{v,j}},\\
-%\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \sum_{i,j\in N_{gal}} N_i N_j \xi_{\delta\delta,\theta_\beta}(x_i,x_j) -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,\theta_\beta,ij}\\
-%& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij} \\
-%& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
-%& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,\theta_\beta}(x_i,x_j) + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,\theta_\beta}(x_i,x_j) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-%&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\delta\delta,\theta_\beta}(x_i,y) + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,\theta_\beta}(x_i,y) \right) d^3y\\
-%&  + \int n(x) n(y) \xi_{\delta\delta,\theta_\beta}(x,y)d^3xd^3y.
+ +   2 \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}  N_i  \xi_{gv,\theta_\beta}(x_i,x_j)\frac{\hat{\delta} v_j}{\sigma^2_{v,j}} \\
+& -   2 \bar{N}  \sum_{i=1}^{N_{\text cell}} \sum_{i=1}^{N_{\text gal}}   \xi_{gv,\theta_\beta}(x_i,x_j) \frac{\hat{\delta} v_j}{\sigma^2_{v,j}} 
++ \sum_{i=1}^{N_{\text gal}}  \sum_{j=1}^{N_{\text gal}}   \frac{\hat{\delta} v_i}{\sigma^2_{v,i}}  \xi_{vv,\theta_\beta}(x_i,x_j)  \frac{\hat{\delta} v_j}{\sigma^2_{v,j}},\\
+%\hat{\delta} \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \hat{\delta} \textbf{y} & \approx \sum_{i,j\in N_{gal}} N_i N_j \xi_{\hat{\delta}\hat{\delta},\theta_\beta}(x_i,x_j) -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\hat{\delta}\hat{\delta},\theta_\beta,ij}\\
+%& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\hat{\delta}\hat{\delta},\theta_\beta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,\theta_\beta,ij} \\
+%& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,\theta_\beta,ij}  + \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\\
+%& = \sum_{i,j\in N_{gal}} \left( \xi_{\hat{\delta}\hat{\delta},\theta_\beta}(x_i,x_j) + 2  \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,\theta_\beta}(x_i,x_j) + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,\theta_\beta}(x_i,x_j) \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\right)\\
+%&-2  \sum_{i\in N_{gal}}\int n(x_i) \left(\xi_{\hat{\delta}\hat{\delta},\theta_\beta}(x_i,y) + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{v \hat{\delta},\theta_\beta}(x_i,y) \right) d^3y\\
+%&  + \int n(x) n(y) \xi_{\hat{\delta}\hat{\delta},\theta_\beta}(x,y)d^3xd^3y.
 \end{align*}
@@
 %%\begin{align}
-%%\Delta \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \Delta \textbf{y} & \approx \Delta \textbf{y}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \Delta \textbf{y}\\
-%%&=(\textbf{N}-\bar{N}) \xi_{\delta\delta,\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+ 
-%%\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+%%\hat{\delta} \textbf{y}^T \textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \hat{\delta} \textbf{y} & \approx \hat{\delta} \textbf{y}^T \textbf{C}_0^{-1} \textbf{S}_\beta \textbf{C}_0^{-1} \hat{\delta} \textbf{y}\\
+%%&=(\textbf{N}-\bar{N}) \xi_{\hat{\delta}\hat{\delta},\theta_\beta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\hat{\delta} v,\theta_\beta}\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2}+ 
+%%\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2} \xi_{vv,\theta_\beta}\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2}.
 %%\end{align}
@@
 % \tr(\textbf{C}^{-1} \textbf{S}_\beta \textbf{C}^{-1} \textbf{C}_0) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\beta) \\
-% & = \bar{N} \tr(\xi_{\delta\delta,\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
-%% & = \int n(x) \xi_{\delta \delta,\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i).
+% & = \bar{N} \tr(\xi_{\hat{\delta}\hat{\delta},\beta}) + \tr(\frac{1}{\sigma^2_v} \xi_{vv,\beta})\\
+%% & = \int n(x) \xi_{\hat{\delta} \hat{\delta},\beta}(x,x) d^3x +  \sum_{i\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\beta}(x_i,x_i).
 %\end{align}
@@
 % \tr(\textbf{C}^{-1} \textbf{S}_\alpha \textbf{C}^{-1} \textbf{S}_\beta) & \approx  \tr(\textbf{C}_0^{-1} \textbf{S}_\alpha \textbf{C}_0^{-1} \textbf{S}_\beta)  \\
-% & = \bar{N}^2 \tr(\xi_{\delta \delta,\alpha}\xi_{\delta \delta,\beta})+
-% \bar{N} \tr(\xi_{\delta v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\delta,\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
-% & = \int n(x) n(y) \xi_{\delta \delta,\alpha}(x,y)\xi_{\delta \delta,\beta}(y,x) d^3x d^3y \\
-% & + \int n(x) \sum_{i\in N_{gal}} \xi_{\delta v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\delta,\beta} (x_i,x) d^3x \\
+% & = \bar{N}^2 \tr(\xi_{\hat{\delta} \hat{\delta},\alpha}\xi_{\hat{\delta} \hat{\delta},\beta})+
+% \bar{N} \tr(\xi_{\hat{\delta} v,\alpha} \frac{1}{\bm{\sigma}^2_v} \xi_{v\hat{\delta},\beta} ) + \tr(\frac{1}{\bm{\sigma}^2_v} \xi_{vv,\alpha}  \frac{1}{\bm{\sigma}^2_v}  \xi_{vv,\beta})\\
+% & = \int n(x) n(y) \xi_{\hat{\delta} \hat{\delta},\alpha}(x,y)\xi_{\hat{\delta} \hat{\delta},\beta}(y,x) d^3x d^3y \\
+% & + \int n(x) \sum_{i\in N_{gal}} \xi_{\hat{\delta} v,\alpha}(x,x_i) \frac{1}{\sigma^2_{v,i}} \xi_{v\hat{\delta},\beta} (x_i,x) d^3x \\
 % & +  \sum_{i,j\in N_{gal}} \frac{1}{\sigma^2_{v,i}} \xi_{vv,\alpha}(x_i,x_j)  \frac{1}{\sigma^2_{v,j}}  \xi_{vv,\beta}(x_j,x_i).
@@
 
-First, we focus on the  term $\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
-by   introducing the matrices $\textbf{M}$ and $\Delta \textbf{y}$,
-where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \Delta \textbf{y} \textbf{M}$ such that $\Delta \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
-Now, our matrix $\textbf{C}$ can be re-written in terms of $\Delta \textbf{y}$ and $\textbf{M}$:
+First, we focus on the  term $\hat{\delta} \textbf{y}^T\textbf{C}^{-1}\hat{\delta} \textbf{y}$. We rewrite $\textbf{C}$ by normalizing out its uncorrelated $\textbf{C}_0$ term
+by   introducing the matrices $\textbf{M}$ and $\hat{\delta} \textbf{y}$,
+where $\textbf{M}=\sqrt{\textbf{C}_0}$ and $\textbf{S} = \textbf{M} \hat{\delta} \textbf{y} \textbf{M}$ such that $\hat{\delta} \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
+Now, our matrix $\textbf{C}$ can be re-written in terms of $\hat{\delta} \textbf{y}$ and $\textbf{M}$:
 \begin{equation}
-\textbf{C} = \textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M}.
+\textbf{C} = \textbf{M}(\textbf{I}+A\hat{\delta} \textbf{y})\textbf{M}.
 \end{equation}
@@
 \begin{align}
-\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\Delta \textbf{y}+A^2 \Delta \textbf{y}^2 + \ldots) \textbf{M}^{-1} \\
+\textbf{C}^{-1} & = \textbf{M}^{-1}(1-A\hat{\delta} \textbf{y}+A^2 \hat{\delta} \textbf{y}^2 + \ldots) \textbf{M}^{-1} \\
 & = \textbf{C}_0^{-1}   -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +\frac{A^2}{2} \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} +O(A^3) + \ldots 
@@
 \begin{align}
-\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
-\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \hat{\delta} \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\hat{\delta}\hat{\delta}}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\hat{\delta} v}\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2}+
+\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2}.
 \end{align}
@@
 \begin{align*}
-\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
-& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
-& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
-& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
-& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
+\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \hat{\delta} \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\hat{\delta}\hat{\delta},ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\hat{\delta}\hat{\delta},ij}\\
+& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\hat{\delta}\hat{\delta},ij} + 2 \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij} \\
+& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij}  + \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\\
+& = \sum_{i,j\in N_{gal}} \left( \xi_{\hat{\delta}\hat{\delta},ij} + 2  \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\right)\\
+&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\hat{\delta}\hat{\delta},iV} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{v \hat{\delta},iV}\right) dV\\
+& +n^2  \int \xi_{\hat{\delta}\hat{\delta},V_1V_2}dV_1dV_2.
 \end{align*}
 
-We evaluate $\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y}$ by splitting up the counts and velocities.  The data vector $\Delta \textbf{y}$ is composed of the counts
- $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\Delta v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\delta\delta}$, $\bar{N}\xi_{\delta v}$,
+We evaluate $\hat{\delta} \textbf{y}^T \textbf{C}^{-1}\hat{\delta} \textbf{y}$ by splitting up the counts and velocities.  The data vector $\hat{\delta} \textbf{y}$ is composed of the counts
+ $\textbf{N}-\bar{N}$ and peculiar velocities $\mathbf{\hat{\delta} v}$, whereas $\textbf{S}$ is comprised of $\bar{N}^2\xi_{\hat{\delta}\hat{\delta}}$, $\bar{N}\xi_{\hat{\delta} v}$,
 and $\xi_{vv}$.  In this manner
 \begin{align}
-\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\delta\delta}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\delta v}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}+
-\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\Delta v}}{\bm{\sigma}_v^2}.
+\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \hat{\delta} \textbf{y}& = (\textbf{N}-\bar{N}) \xi_{\hat{\delta}\hat{\delta}}  (\textbf{N}-\bar{N})  + 2  (\textbf{N}-\bar{N}) \xi_{\hat{\delta} v}\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2}+
+\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2} \xi_{vv}\frac{\mathbf{\hat{\delta} v}}{\bm{\sigma}_v^2}.
 \end{align}
@@
 \begin{align*}
-\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\delta\delta,ij}\\
-& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\delta\delta,ij} + 2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} \\
-& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}  + \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\\
-& = \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV\\
-& +n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2.
+\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \hat{\delta} \textbf{y}& = \sum_{i,j\in N_{gal}} \xi_{\hat{\delta}\hat{\delta},ij} -2 \bar{N} \sum_{i\in N_{gal},j\in N_{cell}} \xi_{\hat{\delta}\hat{\delta},ij}\\
+& +\bar{N}^2  \sum_{i,j\in N_{cell}}  \xi_{\hat{\delta}\hat{\delta},ij} + 2 \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij} \\
+& -2\bar{N}\sum_{i\in N_{cell}, j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij}  + \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\\
+& = \sum_{i,j\in N_{gal}} \left( \xi_{\hat{\delta}\hat{\delta},ij} + 2  \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\right)\\
+&-2 n \sum_{i\in N_{gal}}\int \left(\xi_{\hat{\delta}\hat{\delta},iV} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{v \hat{\delta},iV}\right) dV\\
+& +n^2  \int \xi_{\hat{\delta}\hat{\delta},V_1V_2}dV_1dV_2.
 \end{align*}
@@
 \begin{pmatrix}
- \bar{N}^2  \xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}  & \bar{N} \xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
- \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
+ \bar{N}^2  \xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}  & \bar{N} \xi_{\hat{\delta} v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \\
+ \bar{N} \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \hat{\delta}}   &  \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}
 \end{pmatrix}
@@
 \begin{align}
-\Delta \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \Delta \textbf{y}
-& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
- + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV\\
-& +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2.
+\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} \hat{\delta} \textbf{y}
+& = \sum_{i,j\in N_{gal}} \left( \left(\xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}\right)_{ij}
+ + 2  \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \left(\xi_{\hat{\delta} v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\right)\\
+&-2 n \sum_{i\in N_{gal}}\int \left( \left(\xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}\right)_{iV} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \hat{\delta}}  \right)_{iV} \right) dV\\
+& +n^2  \int  \left(\xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}\right)_{V_1V_2}dV_1dV_2.
 \end{align}
 
-We have assembled the contributions of $\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y}$ to second order in $A$.
+We have assembled the contributions of $\hat{\delta} \textbf{y}^T \textbf{C}^{-1}\hat{\delta} \textbf{y}$ to second order in $A$.
 
@@
 \begin{align*}
-\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M})) \\
-&= \ln\det(\textbf{I}+A\Delta \textbf{y}) +\text{const} \\
-&= A \tr \Delta \textbf{y} -\frac{A^2}{2}\tr \Delta \textbf{y}^2+ O(A^3) + \ldots\\
+\ln\det(\textbf{C}) &=\ln\det(\textbf{M}(\textbf{I}+A\hat{\delta} \textbf{y})\textbf{M})) \\
+&= \ln\det(\textbf{I}+A\hat{\delta} \textbf{y}) +\text{const} \\
+&= A \tr \hat{\delta} \textbf{y} -\frac{A^2}{2}\tr \hat{\delta} \textbf{y}^2+ O(A^3) + \ldots\\
 \end{align*}
@@
 \begin{equation}
-\Delta \textbf{y} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
+\hat{\delta} \textbf{y} = \textbf{M}^{-1}\textbf{S}\textbf{M}^{-1} = 
 \begin{pmatrix}
-\bar{N}\xi_{\delta\delta} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\delta v} \\
-\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\delta} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
+\bar{N}\xi_{\hat{\delta}\hat{\delta}} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\hat{\delta} v} \\
+\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\hat{\delta}} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
 \end{pmatrix}.
@@
 \begin{equation}
-\bar{N} \tr \xi_{\delta\delta} = n \int \xi_{\delta\delta,VV} dV
+\bar{N} \tr \xi_{\hat{\delta}\hat{\delta}} = n \int \xi_{\hat{\delta}\hat{\delta},VV} dV
 \end{equation}
@@
 
-The trace of $\Delta \textbf{y}^2$ has terms, to leading order in $\bar{N}$, has terms
+The trace of $\hat{\delta} \textbf{y}^2$ has terms, to leading order in $\bar{N}$, has terms
 \begin{equation}
-\bar{N}  \tr  \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} = n \int \left( \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} \right)_{VV} dV
+\bar{N}  \tr  \xi_{\hat{\delta} v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \hat{\delta}} = n \int \left( \xi_{\hat{\delta} v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \hat{\delta}} \right)_{VV} dV
 \end{equation}
@@
 \begin{align}
-c_1 &=- \sum_{i,j\in N_{gal}} \left( \xi_{\delta\delta,ij} + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-& +\sum_{i \in N_{gal}} \left( \frac{1}{\sigma_{v,i}^2} \xi_{vv,ii}   + 2n \int \left(\xi_{\delta\delta,iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v \delta,iV}\right) dV  \right)\\
-& +n \int \xi_{\delta\delta,VV} dV  -n^2  \int \xi_{\delta\delta,V_1V_2}dV_1dV_2
+c_1 &=- \sum_{i,j\in N_{gal}} \left( \xi_{\hat{\delta}\hat{\delta},ij} + 2  \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\right)\\
+& +\sum_{i \in N_{gal}} \left( \frac{1}{\sigma_{v,i}^2} \xi_{vv,ii}   + 2n \int \left(\xi_{\hat{\delta}\hat{\delta},iV} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{v \hat{\delta},iV}\right) dV  \right)\\
+& +n \int \xi_{\hat{\delta}\hat{\delta},VV} dV  -n^2  \int \xi_{\hat{\delta}\hat{\delta},V_1V_2}dV_1dV_2
 \end{align}
@@
 c_2 & = \frac{1}{2} \Biggl( 
- \sum_{i,j\in N_{gal}} \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{ij}
- + 2  \frac{\Delta v_j}{\sigma_{v,j}^2} \left(\xi_{\delta v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\Delta v_j}{\sigma_{v,j}^2}\right)\\
-&- \sum_{i \in N_{gal}}\left( \left(  \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} \right)_{ii}  +2n \int \left( \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{iV} + \frac{\Delta v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \delta}  \right)_{iV} \right) dV  \right)\\
-&-n  \int \left( \xi_{\delta v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \delta} \right)_{VV} dV  +n^2  \int  \left(\xi_{\delta v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \delta}\right)_{V_1V_2}dV_1dV_2 \Biggr).
+ \sum_{i,j\in N_{gal}} \left( \left(\xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}\right)_{ij}
+ + 2  \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \left(\xi_{\hat{\delta} v} \frac{1} {\bm{\sigma}_v^2} \xi_{vv}  \right)_{ij} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \left( \xi_{vv} \frac{1}{\bm{\sigma}_v^2} \xi_{vv}\right)_{ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}\right)\\
+&- \sum_{i \in N_{gal}}\left( \left(  \xi_{vv} \frac{1}{\sigma_{v}^2}  \frac{1}{\sigma_{v}^2} \xi_{vv} \right)_{ii}  +2n \int \left( \left(\xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}\right)_{iV} + \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \left( \xi_{v v} \frac{1} {\bm{\sigma}_v^2} \xi_{v \hat{\delta}}  \right)_{iV} \right) dV  \right)\\
+&-n  \int \left( \xi_{\hat{\delta} v} \frac{1}{\sigma_{v}}  \frac{1}{\sigma_{v}}  \xi_{v \hat{\delta}} \right)_{VV} dV  +n^2  \int  \left(\xi_{\hat{\delta} v}\frac{1}{\bm{\sigma}_v^2}  \xi_{v \hat{\delta}}\right)_{V_1V_2}dV_1dV_2 \Biggr).
 \end{align}
@@
 Given the enormous size of $\textbf{C}$, computing the log likelihood directly is not possible. In the following, we seek an approximation method accurate to first order in $A$. \par
-First, we focus on the second term, $\Delta \textbf{y}^T\textbf{C}^{-1}\Delta \textbf{y}$.
+First, we focus on the second term, $\hat{\delta} \textbf{y}^T\textbf{C}^{-1}\hat{\delta} \textbf{y}$.
 Looking at the definition of $\textbf{C}$ in equation (1), we want to factor $\textbf{C}_0$ into an identity matrix.
-To do so, we introduce matrix $\textbf{M}$ and $\Delta \textbf{y}$.
-We let $\textbf{M}=\sqrt{\textbf{C}_0}$, and $\textbf{S} = \textbf{M} \Delta \textbf{y} \textbf{M}$.
-Then $\Delta \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
-Now, our matrix $\textbf{C}$ can be re-written in terms of $\Delta \textbf{y}$ and $\textbf{M}$:
+To do so, we introduce matrix $\textbf{M}$ and $\hat{\delta} \textbf{y}$.
+We let $\textbf{M}=\sqrt{\textbf{C}_0}$, and $\textbf{S} = \textbf{M} \hat{\delta} \textbf{y} \textbf{M}$.
+Then $\hat{\delta} \textbf{y}=\textbf{M}^{-1}\textbf{S}\textbf{M}^{-1}$.
+Now, our matrix $\textbf{C}$ can be re-written in terms of $\hat{\delta} \textbf{y}$ and $\textbf{M}$:
 \begin{equation}
-\textbf{C} = \textbf{M}(\textbf{I}+A\Delta \textbf{y})\textbf{M},
+\textbf{C} = \textbf{M}(\textbf{I}+A\hat{\delta} \textbf{y})\textbf{M},
 \end{equation}
@@
 \[
-\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\Delta \textbf{y}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
+\textbf{C}^{-1} = \textbf{M}^{-1}(1-A\hat{\delta} \textbf{y}) \textbf{M}^{-1} = -A \textbf{C}_0^{-1}\textbf{S}\textbf{C}_0^{-1} + \text{const}.
 \] \par
@@
 \begin{align*}
-s &=\Delta \textbf{y}^T \textbf{C}^{-1}\Delta \textbf{y} \\
-&= -A\left(\Delta \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\Delta \textbf{y}\right).
+s &=\hat{\delta} \textbf{y}^T \textbf{C}^{-1}\hat{\delta} \textbf{y} \\
+&= -A\left(\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\hat{\delta} \textbf{y}\right).
 \end{align*} \par
@@
 \[
--\left(\Delta \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\Delta \textbf{y}\right) = s_1 +s_2+\hdots +s_6
+-\left(\hat{\delta} \textbf{y}^T \textbf{C}_0^{-1}\right)\textbf{S}\left(\textbf{C}_0^{-1}\hat{\delta} \textbf{y}\right) = s_1 +s_2+\hdots +s_6
 \]
 \begin{align*}
-s_1&= -\sum_{i,j\in N_{gal}} \xi_{\delta\delta,ij},  \\
-s_2 &= 2 \sum_{i\in N_{gal},j\in N_{cell}} \bar{N} \xi_{\delta\delta,ij}, \\
+s_1&= -\sum_{i,j\in N_{gal}} \xi_{\hat{\delta}\hat{\delta},ij},  \\
+s_2 &= 2 \sum_{i\in N_{gal},j\in N_{cell}} \bar{N} \xi_{\hat{\delta}\hat{\delta},ij}, \\
 &= 2\bar{N} \frac{b^2}{2\pi} \sum_{i\in N_{gal}} \int r^2drd\Omega  \int dk P(k) k^2 j_0 (k |\textbf{r}-\textbf{r}_i|), \\
-s_3 &= -\sum_{i,j\in N_{cell}} \bar{N}^2  \xi_{\delta\delta,ij}, \\ 
+s_3 &= -\sum_{i,j\in N_{cell}} \bar{N}^2  \xi_{\hat{\delta}\hat{\delta},ij}, \\ 
 &=-\bar{N}^2 \frac{b^2}{2\pi} \int r_1^2dr_1d\Omega_1 \int r_2^2dr_2 d\Omega_2 \int dk P(k) k^2 j_0 (k |\textbf{r}_1-\textbf{r}_2|), \\
-s_4 &=- \sum_{i,j\in N_{gal}} \frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\Delta v_j}{\sigma_{v,j}^2}, \\
-s_5 &= -2 \sum_{i,j\in N_{gal}} \frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij}, \\
-s_6 &= 2\sum_{i\in N_{cell}, j\in N_{gal}} \bar{N}\frac{\Delta v_j}{\sigma_{v,j}^2} \xi_{\delta v,ij},  \\
-&= 2\bar{N}\sum_{j\in N_{gal}} \frac{aHfb\Delta v_j}{2\pi^2 \sigma^2_{v,j}}\int r^2 dr d\Omega \int dk P(k) k ((\hat{\textbf{r}}-\hat{\textbf{r}_j})\cdot\hat{\textbf{r}_j}) j_1(k|\textbf{r}-\textbf{r}_j|).
-%s_7 &= A  \sum_{j\in N_{cell}, i\in N_{gal}} \bar{N}\frac{\Delta v_i}{\sigma_{v,i}^2} \xi_{v\delta,ij}  \\
-%&= A\bar{N} \sum_{i\in N_{gal}} \frac{+aHfb\Delta v_i}{2\pi^2 \sigma^2_{v,i}}\int r^2 dr d\Omega \int dk P(k) k (\textbf{r}\cdot\textbf{r}_i) j_1(k|\textbf{r}-\textbf{r}_i|)
+s_4 &=- \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{vv,ij} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2}, \\
+s_5 &= -2 \sum_{i,j\in N_{gal}} \frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij}, \\
+s_6 &= 2\sum_{i\in N_{cell}, j\in N_{gal}} \bar{N}\frac{\hat{\delta} v_j}{\sigma_{v,j}^2} \xi_{\hat{\delta} v,ij},  \\
+&= 2\bar{N}\sum_{j\in N_{gal}} \frac{aHfb\hat{\delta} v_j}{2\pi^2 \sigma^2_{v,j}}\int r^2 dr d\Omega \int dk P(k) k ((\hat{\textbf{r}}-\hat{\textbf{r}_j})\cdot\hat{\textbf{r}_j}) j_1(k|\textbf{r}-\textbf{r}_j|).
+%s_7 &= A  \sum_{j\in N_{cell}, i\in N_{gal}} \bar{N}\frac{\hat{\delta} v_i}{\sigma_{v,i}^2} \xi_{v\hat{\delta},ij}  \\
+%&= A\bar{N} \sum_{i\in N_{gal}} \frac{+aHfb\hat{\delta} v_i}{2\pi^2 \sigma^2_{v,i}}\int r^2 dr d\Omega \int dk P(k) k (\textbf{r}\cdot\textbf{r}_i) j_1(k|\textbf{r}-\textbf{r}_i|)
 \end{align*} \par
@@
 \begin{array}{cc}
-\bar{N}\xi_{\delta\delta} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\delta v} \\
-\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\delta} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
+\bar{N}\xi_{\hat{\delta}\hat{\delta}} & \frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{\hat{\delta} v} \\
+\frac{\sqrt{\bar{N}}}{\sigma_{v}} \xi_{v\hat{\delta}} & \frac{1}{\sigma_{v}^2} \xi_{vv} 
 \end{array}
@@
 \begin{array}{cc}
-\bar{N}\xi_{\delta\delta} & \xi_{\delta v} \\
-\bar{N}\bm{\sigma}^{-1} \xi_{v\delta} & \bm{\sigma}^{-1} \xi_{vv}
+\bar{N}\xi_{\hat{\delta}\hat{\delta}} & \xi_{\hat{\delta} v} \\
+\bar{N}\bm{\sigma}^{-1} \xi_{v\hat{\delta}} & \bm{\sigma}^{-1} \xi_{vv}
 \end{array}
@@
 \begin{array}{cc}
-\bar{N}^2 \xi_{\delta\delta}^2 + \bar{N}\xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\xi_{\delta\delta}\xi_{\delta v}+\xi_{\delta v}\bm{\sigma}^{-1}\xi_{vv} \\
-\bar{N}^2\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta\delta}+\bar{N}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta v} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}
+\bar{N}^2 \xi_{\hat{\delta}\hat{\delta}}^2 + \bar{N}\xi_{\hat{\delta} v}\bm{\sigma}^{-1} \xi_{v\hat{\delta}} & \bar{N}\xi_{\hat{\delta}\hat{\delta}}\xi_{\hat{\delta} v}+\xi_{\hat{\delta} v}\bm{\sigma}^{-1}\xi_{vv} \\
+\bar{N}^2\bm{\sigma}^{-1} \xi_{v\hat{\delta}}\xi_{\hat{\delta}\hat{\delta}}+\bar{N}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\hat{\delta}} & \bar{N}\bm{\sigma}^{-1} \xi_{v\hat{\delta}}\xi_{\hat{\delta} v} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}
 \end{array}
@@
 \begin{array}{cc}
-\bar{N} \xi_{\delta\delta}^2 + \xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\xi_{\delta\delta}\xi_{\delta v}\bm{\sigma}^{-1}+\xi_{\delta v}\bm{\sigma}^{-1}\xi_{vv}\bm{\sigma}^{-1} \\
-\bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta\delta}+\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\delta} & \bar{N}\bm{\sigma}^{-1} \xi_{v\delta}\xi_{\delta v}\bm{\sigma}^{-1} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1}
+\bar{N} \xi_{\hat{\delta}\hat{\delta}}^2 + \xi_{\hat{\delta} v}\bm{\sigma}^{-1} \xi_{v\hat{\delta}} & \bar{N}\xi_{\hat{\delta}\hat{\delta}}\xi_{\hat{\delta} v}\bm{\sigma}^{-1}+\xi_{\hat{\delta} v}\bm{\sigma}^{-1}\xi_{vv}\bm{\sigma}^{-1} \\
+\bar{N}\bm{\sigma}^{-1} \xi_{v\hat{\delta}}\xi_{\hat{\delta}\hat{\delta}}+\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{v\hat{\delta}} & \bar{N}\bm{\sigma}^{-1} \xi_{v\hat{\delta}}\xi_{\hat{\delta} v}\bm{\sigma}^{-1} + \bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1} \xi_{vv}\bm{\sigma}^{-1}
 \end{array}
@@
 \begin{align*}
-\textbf{A}_{ij} &=(\bar{N} \xi_{\delta\delta}^2 + \xi_{\delta v}\bm{\sigma}^{-1} \xi_{v\delta})_{ij} \\
-&= \frac{\bar{N}b^4}{4\pi^4} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|) +\sum_{m}\sigma_{m}^{-2}\xi_{\delta v,im}\xi_{v\delta,mj} \\
+\textbf{A}_{ij} &=(\bar{N} \xi_{\hat{\delta}\hat{\delta}}^2 + \xi_{\hat{\delta} v}\bm{\sigma}^{-1} \xi_{v\hat{\delta}})_{ij} \\
+&= \frac{\bar{N}b^4}{4\pi^4} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|) +\sum_{m}\sigma_{m}^{-2}\xi_{\hat{\delta} v,im}\xi_{v\hat{\delta},mj} \\
 &=\frac{\bar{N}b^4}{4\pi^4} \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2) k_1^2k_2^2j_0(k_1|\textbf{r}-\textbf{r}_i|)j_0(k_2|\textbf{r}-\textbf{r}_j|)\\
@@
 \textbf{B}_{ij} &= \frac{\bar{N}aHfb^3\sigma_j^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|)\\
-&+ \sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{\delta v,im}\xi_{vv,mj} \\
+&+ \sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{\hat{\delta} v,im}\xi_{vv,mj} \\
 &=\frac{\bar{N}aHfb^3\sigma_j^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) \\
@@
 \textbf{C}_{ij} &=\frac{\bar{N}aHfb^3\sigma_i^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_ij_0(k_2|\textbf{r}-\textbf{r}_j|)j_1(k_1|\textbf{r}-\textbf{r}_i|)\\
-&+ \sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{vv,im}\xi_{v\delta,mj} \\
+&+ \sigma_j^{-2}\sum_m \sigma_m^{-2}\xi_{vv,im}\xi_{v\hat{\delta},mj} \\
 &=\frac{\bar{N}aHfb^3\sigma_i^{-2}}{4\pi^4}\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_ij_0(k_2|\textbf{r}-\textbf{r}_j|)j_1(k_1|\textbf{r}-\textbf{r}_i|)\\
@@
 &\times (\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m(\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m \\
-S_4 &= \sum_{i,j} \textbf{D}_{ij}\Delta v_i \Delta v_j \\
-&= \frac{\bar{N}a^2H^2f^2b^2}{4\pi^4}\sum_{i,j}\sigma_i^{-2}\sigma_j^{-2}\Delta v_i \Delta v_j \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_i(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_j\times \\
-&\times j_1(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) +\sum_{m,i,j} \sigma_i^{-2}\sigma_j^{-2}\sigma_m^{-2}\Delta v_i \Delta v_j\xi_{vv,im}\xi_{vv,mj}\\
-S_5 &= 2\sum_{i,j}\textbf{B}_{ij}\Delta v_j \\
-&=\frac{\bar{N}aHfb^3}{2\pi^4}\sum_{i,j}\sigma_j^{-2}\Delta v_j\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) \\
-&+\frac{\sigma_j^{-2}aHfb}{\pi^2}\sum_{m,i,j}\Delta v_j \sigma_m^{-2} \xi_{vv,mj}\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m\\
-S_6 &=-2\bar{N} \sum_{i\in N_{cell},j\in N_{gal}}\textbf{B}_{ij} \Delta v_j\\
-&=\frac{-\bar{N}^2aHfb^3}{2\pi^4}\sum_{j}\sigma_j^{-2}\Delta v_j\int r^2r_i^2drdr_id\Omega d\Omega_i \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)\times \\
-&\times j_1(k_2|\textbf{r}-\textbf{r}_j|)-\frac{\bar{N}aHfb}{\pi^2}\sum_{m,j}\sigma_j^{-2} \Delta v_j \sigma_m^{-2} \xi_{vv,mj}\int r_i^2 dr_id\Omega_i\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m
+S_4 &= \sum_{i,j} \textbf{D}_{ij}\hat{\delta} v_i \hat{\delta} v_j \\
+&= \frac{\bar{N}a^2H^2f^2b^2}{4\pi^4}\sum_{i,j}\sigma_i^{-2}\sigma_j^{-2}\hat{\delta} v_i \hat{\delta} v_j \int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2^2(\hat{\textbf{r}_i}-\hat{\textbf{r}})\cdot\hat{\textbf{r}}_i(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_j\times \\
+&\times j_1(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) +\sum_{m,i,j} \sigma_i^{-2}\sigma_j^{-2}\sigma_m^{-2}\hat{\delta} v_i \hat{\delta} v_j\xi_{vv,im}\xi_{vv,mj}\\
+S_5 &= 2\sum_{i,j}\textbf{B}_{ij}\hat{\delta} v_j \\
+&=\frac{\bar{N}aHfb^3}{2\pi^4}\sum_{i,j}\sigma_j^{-2}\hat{\delta} v_j\int r^2drd\Omega \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)j_1(k_2|\textbf{r}-\textbf{r}_j|) \\
+&+\frac{\sigma_j^{-2}aHfb}{\pi^2}\sum_{m,i,j}\hat{\delta} v_j \sigma_m^{-2} \xi_{vv,mj}\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m\\
+S_6 &=-2\bar{N} \sum_{i\in N_{cell},j\in N_{gal}}\textbf{B}_{ij} \hat{\delta} v_j\\
+&=\frac{-\bar{N}^2aHfb^3}{2\pi^4}\sum_{j}\sigma_j^{-2}\hat{\delta} v_j\int r^2r_i^2drdr_id\Omega d\Omega_i \iint dk_1dk_2 P(k_1)P(k_2)k_1^2k_2(\hat{\textbf{r}}-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_jj_0(k_1|\textbf{r}-\textbf{r}_i|)\times \\
+&\times j_1(k_2|\textbf{r}-\textbf{r}_j|)-\frac{\bar{N}aHfb}{\pi^2}\sum_{m,j}\sigma_j^{-2} \hat{\delta} v_j \sigma_m^{-2} \xi_{vv,mj}\int r_i^2 dr_id\Omega_i\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m
 \end{align*}\par
@@
 \begin{align*}
-D_1 + D_2 + D_3 + D_4 &\equiv -\frac{\bar{N}^2}{2} \sum_{i,j\in N_{cell}} \xi_{\delta\delta,ij}\xi_{\delta\delta,ji} -\frac{\bar{N}}{2}\sum_{i \in N_{cell},j\in N_{gal}} \sigma_j^{-2}\xi_{\delta v,ij} \xi_{v\delta,ji} \\
-&-\frac{\bar{N}}{2} \sum_{i\in N_{gal},j\in N_{cell}} \sigma_i^{-2} \xi_{v\delta,ij}\xi_{\delta v,ji} - \frac{1}{2} \sum_{i,j} \sigma_i^{-2}\sigma_j^{-2} \xi_{vv,ij}\xi_{vv,ji}
+D_1 + D_2 + D_3 + D_4 &\equiv -\frac{\bar{N}^2}{2} \sum_{i,j\in N_{cell}} \xi_{\hat{\delta}\hat{\delta},ij}\xi_{\hat{\delta}\hat{\delta},ji} -\frac{\bar{N}}{2}\sum_{i \in N_{cell},j\in N_{gal}} \sigma_j^{-2}\xi_{\hat{\delta} v,ij} \xi_{v\hat{\delta},ji} \\
+&-\frac{\bar{N}}{2} \sum_{i\in N_{gal},j\in N_{cell}} \sigma_i^{-2} \xi_{v\hat{\delta},ij}\xi_{\hat{\delta} v,ji} - \frac{1}{2} \sum_{i,j} \sigma_i^{-2}\sigma_j^{-2} \xi_{vv,ij}\xi_{vv,ji}
 \end{align*}\par
@@
 &\times (\hat{\textbf{r}}_m-\hat{\textbf{r}}_j)\cdot\hat{\textbf{r}}_m \\
-S_6' &=-\frac{\bar{N}aHfb}{\pi^2}\sum_{m,j}\sigma_j^{-2} \Delta v_j \sigma_m^{-2} \xi_{vv,mj}\int r_i^2 dr_id\Omega_i\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m
+S_6' &=-\frac{\bar{N}aHfb}{\pi^2}\sum_{m,j}\sigma_j^{-2} \hat{\delta} v_j \sigma_m^{-2} \xi_{vv,mj}\int r_i^2 dr_id\Omega_i\int dk P(k)k j_1(k|\textbf{r}_i-\textbf{r}_m|)(\hat{\textbf{r}}_i-\hat{\textbf{r}}_m)\cdot\hat{\textbf{r}}_m
 \end{align*}
@@
 \end{align}
+\newpage
+\section{Model}
 
+Suppose we have two sets of data, the number of galaxies $N$ in a volume
+bin and the radial velocities of galaxies $v$.  These may or may not be the same galaxies.
+The likelihood is then
+\begin{align}
+L(\theta; N, v) & =p(N, v| \theta) \\
+&= \int  p(N | \delta)  p(v | u)  p(\delta,v| \theta) d\delta\, du\\ 
+& = \int  \prod_{i=1}^{N_{bin}}   p(N_i | \delta_i)p(v| u) p(\delta, u | \theta)d\delta\, du\\ 
+& = \int  \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty  p(N_i | \tilde{N}_i) p(\tilde{N}_i | \delta) p(v| u) p(\delta, u | \theta) d\delta\, du \\
+& = \int  \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))p(v| u) p(\delta, u | \theta) d\delta\, du \\
+&= \int  \exp \left[ \sum_{i=1}^{N_{bin}} \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)} \right]  p(v| u) p(\delta, u | \theta) d\delta\, du
+\end{align}
+Here $\delta$ and $i$ are  the mass-density  and momentum fields at the bin/galaxy positions, $\tilde{N}_i$ is the latent parameter
+for the  total number of galaxies in bin $i$, $N_{0i}(\hat{\delta})$ is the expected
+average number of galaxies in bin $i$, and $p_i$ is the probability of for a galaxy in cell $i$ to enter the sample.
+$\mathcal{B}$ is the binomial distribution and $\mathcal{P}$ is the Poisson distribution.
+
+The summation over bins an be split up into two terms assuming there are no cells with greater than two galaxies.
+The first is a term as if there were zero galaxies in all bins, $N_i=0\ \forall i$.
+\begin{align}
+ \sum_{i=1}^{N_{bin}} \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(0; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)} & =
+\sum_{i=1}^{N_{bin}} \ln{  \left( \sum_{\tilde{N}_i =0}^\infty   (1-p_i)^{ \tilde{N}_i} \frac{N_{0i}^{\tilde{N}_i} e^{-N_{0i}}}{\tilde{N}_i!} \right)}\\ 
+& = \sum_{i=1}^{N_{bin}} (1-p_i)N_{0i} - N_{0i}\\
+& = - \sum_{i=1}^{N_{bin}} p_i N_{0i}  \\
+& = - \int  p(x) n_{0}(x) d^3x
+\end{align}
+where in the limit of  infinitesimally small bins $N_{0}$ approaches zero and
+the integral replaces the summation.
+The second term is over those cells that have one galaxy:
+\begin{align}
+\sum_{i \in \{N_i \ne 0\}}
+\ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)} -
+ \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(0; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)}\\
+ =\sum_{i \in \{N_i \ne 0\}} \ln{  \left( \sum_{\tilde{N}_i =1}^\infty   \tilde{N}_i p_i (1-p_i)^{ \tilde{N}_i-1} \frac{N_{0i}^{\tilde{N}_i} e^{-N_{0i}}}{\tilde{N}_i!} \right)}
+- \ln{  \left( \sum_{\tilde{N}_i =0}^\infty   (1-p_i)^{ \tilde{N}_i} \frac{N_{0i}^{\tilde{N}_i} e^{-N_{0i}}}{\tilde{N}_i!} \right)} \\
+ = \sum_{i=1}^{N_{\text{galaxies}}} \ln{  \left(  p_i  {N_{0i} } \right)} \\
+  = \sum_{i=1}^{N_{\text{galaxies}}} \ln{  \left(  p_i  n_{0}(x_i) \hat{\delta} V \right)}
+\end{align}
+
+Then the likelihood is
+\begin{align}
+L(\theta; N, v) 
+&\propto \int  \exp \left(- \int  p(x) n_{0}(x) d^3x\right)  \prod_{i=1}^{N_{\text{galaxies}}} p_i n_0(x_i) p(v_i| u_i) p(\delta, u | \theta) d\delta\, du .
+\end{align}
+The exponential term depends on $\delta$ at every differential volume.  The
+remaining terms of the integrand depend only on the $\delta_i$ and $u_i$ at galaxy coordinates.
+
+When the universe is modeled as a periodic cube much larger than the survey. the the density field can be describe by discrete points. 
+Further assuming a cutoff in power at $k_{max}$, the field can be described by a finite grid.  Then the parameters $\delta$ and $u$ will be
+the Fourier density and momentum at grid points, having a covariance matrix composed of four diagonal submatrices. 
+The density and momentum at any point in real space can be constructed from the parameters.
+
+\newpage
+\section{How to analyze}
+% Requires the booktabs if the memoir class is not being used
+\begin{table}[htbp]
+   \centering
+   %\topcaption{Table captions are better up top} % requires the topcapt package
+   \begin{tabular}{|c|c|} % Column formatting, @{} suppresses leading/trailing space
+   	\hline
+	SN Ia rest rate & $0.65 \times 2.69 \times 10^{-5} \left(\frac{h}{0.7}\right)^3 \text{Mpc}^{-3} y^{-1}$\\
+	rate in $k=0.1 h \text{Mpc}^{-1}$ &$ 1.75 \times 10^{-2} \text{y}^{-1}$\\
+	Full sky volume $z_{\text{max}}=0.2$ & $2.30 \times 10^{9} \left(\frac{0.7}{h}\right)^3 \text{Mpc}^3$\\
+	number of cells $k=0.1 h \text{Mpc}^{-1}$   & $7.87 \times 10^5$\\
+	number of SNe in full sky& $1.38 \times 10^4$ \\
+	peak of matter power spectrum $P$ & $3 \times 10^4 (h^{-1} \text{Mpc})^3$\\
+	$nP$ &$1.53 \mbox{y}^{-1}$ \\
+	\hline
+   \end{tabular}
+   \caption{Useful numbers.}
+   \label{tab:booktabs}
+\end{table}
+
+There are fewer point objects than cells:
+\begin{itemize}
+\item  After 10 years there are still fewer SNe than cells for $k=0.1 h \text{Mpc}^{-1}$. 
+\item After 1 year of 100\% coverage the $gg$ survey is sample-noise dominated.  The higher number density from a
+galaxy survey, which would have more points than cells, is unnecessary.
+\end{itemize}
+
+For the velocity field is makes sense to work with points.
+
+For the density field this is not obvious.  While
+
+The density field probed by points requires integrals.  The numerical integration could be done by using cells though Monte Carlo integration seems to be the standard and 
+is presumably more efficient.
+
+For velocity-velocity using points is straightforward.
+
+For density-density
+
+\begin{table}[htbp]
+   \centering
+   %\topcaption{Table captions are better up top} % requires the topcapt package
+   \begin{tabular}{|c|c|c|} % Column formatting, @{} suppresses leading/trailing space
+   	\hline
+	Type & Pro & Con \\ \hline
+	Point & \\
+	Grid & \\
+	Conditional Correlations Likelihood & \\
+	Correlation Function Estimator & \\
+	Fourier Transform & \\
+	\hline
+   \end{tabular}
+   \caption{Methods.}
+   \label{tab:booktabs}
+\end{table}
+\newpage
 \begin{thebibliography}{9}
@@
 
-\end{document}
+\end{document}~/pro

### 2019-08-31 doc/src/partials.py
fast version for fisher matrix calcs

@@ -3,6 +3,7 @@
 import numpy
 import matplotlib.pyplot as plt
 import matplotlib
+from matplotlib import ticker, cm
 import scipy.integrate as integrate
 from astropy.cosmology import FlatLambdaCDM
 matplotlib.rcParams['font.size'] = 14
@@ -295,6 +296,27 @@ def traces(z,mu,ng,duration,sigm,restrate):
         numpy.array(lO_ind),numpy.array(bO_ind),numpy.array(OO_ind),numpy.array(lOsigM),numpy.array(bOsigM),numpy.array(OOsigM), \
         numpy.array(lO_vonly), numpy.array(OO_vonly)
 
+def traces_fast(z,mu,ng,duration,sigm,restrate):
+
+    cmatrices = Cmatrices(z,mu,ng,duration,sigm,restrate)
+
+    ll=[]
+    lb=[]
+    bb=[]
+    lO=[]
+    bO=[]
+    OO=[]
+
+    for C, Cinv, C_l,C_b,Cinvn,CinvsigM, C_Om in zip(cmatrices[0],cmatrices[1],cmatrices[2],cmatrices[3],cmatrices[4],cmatrices[5],cmatrices[6]):
+        ll.append(numpy.trace(Cinv@C_l@Cinv@C_l))
+        bb.append(numpy.trace(Cinv@C_b@Cinv@C_b))
+        lb.append(numpy.trace(Cinv@C_l@Cinv@C_b))
+        lO.append(numpy.trace(Cinv@C_l@Cinv@C_Om))
+        bO.append(numpy.trace(Cinv@C_b@Cinv@C_Om))
+        OO.append(numpy.trace(Cinv@C_Om@Cinv@C_Om))
+
+    return numpy.array(ll),numpy.array(bb),numpy.array(lb),numpy.array(lO),numpy.array(bO),numpy.array(OO)
+
 def muintegral(z,ng,duration,sigm,restrate):
     mus=numpy.arange(0,1.001,0.05)
 
@@ -363,6 +385,29 @@ def muintegral(z,ng,duration,sigm,restrate):
     return 2*ll,2*bb,2*bl, 2*lls,2*bbs,2*bls,2*ll_ind,2*bb_ind,2*bl_ind, 2*llsigM,2*bbsigM,2*blsigM,2*ll_vonly, 2*lO, 2*bO, 2*OO, \
[…]

### 2019-08-31 doc/point.tex
snapshot

@@
 bin and the radial velocities of galaxies $v$.  These may or may not be the same galaxies.
+We also assume that the universe is periodic with a minimum length scale, meaning that fields can be
+described by a finite number of Fourier modes.  Given a model of gravity, we can deterministically calculate the
+velocity field given the density field.
 The likelihood is then
@@
 L(\theta; N, v) & =p(N, v| \theta) \\
-&= \int  p(N | \delta)  p(v | u)  p(\delta,v| \theta) d\delta\, du\\ 
-& = \int  \prod_{i=1}^{N_{bin}}   p(N_i | \delta_i)p(v| u) p(\delta, u | \theta)d\delta\, du\\ 
-& = \int  \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty  p(N_i | \tilde{N}_i) p(\tilde{N}_i | \delta) p(v| u) p(\delta, u | \theta) d\delta\, du \\
-& = \int  \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))p(v| u) p(\delta, u | \theta) d\delta\, du \\
-&= \int  \exp \left[ \sum_{i=1}^{N_{bin}} \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)} \right]  p(v| u) p(\delta, u | \theta) d\delta\, du
+& = \int p(N,v,\delta | \theta) d\delta \\\
+& = p(v|\theta)  \int p(N | \delta) p(\delta| v, \theta) d\delta \\
+& = p(v|\theta)  \int \prod_{i=1}^{N_{bin}}  p(N_i | \delta) p(\delta| v, \theta) d\delta \\
+& = p(v|\theta)  \int \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty    p(N_i | \tilde{N}_i) p(\tilde{N}_i | \delta)  p(\delta| v, \theta) d\delta \\
+& = p(v|\theta)  \int \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty    \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))  p(\delta| v, \theta) d\delta \\
+& =  p(v|\theta)  \int  \exp \left[ \sum_{i=1}^{N_{bin}} \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)} \right]  p(\delta| v, \theta) d\delta
+%&= \int  p(N | \delta)  p(v | u)  p(\delta,v| \theta) d\delta\, du\\ 
+%& = \int  \prod_{i=1}^{N_{bin}}   p(N_i | \delta_i)p(v| u) p(\delta, u | \theta)d\delta\, du\\ 
+%& = \int  \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty  p(N_i | \tilde{N}_i) p(\tilde{N}_i | \delta) p(v| u) p(\delta, u | \theta) d\delta\, du \\
+%& = \int  \prod_{i=1}^{N_{bin}}  \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))p(v| u) p(\delta, u | \theta) d\delta\, du \\
+%&= \int  \exp \left[ \sum_{i=1}^{N_{bin}} \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)} \right]  p(v| u) p(\delta, u | \theta) d\delta\, du
 \end{align}
@@
 the integral replaces the summation.
+
+Consider a cell with $N_i$ galaxies.  It contributes a 
+\begin{align}
+\ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty  \mathcal{B}(N_i; \tilde{N}_i, p_i) \mathcal{P}(\tilde{N}_i ; N_{0i}(\delta))\right)}\\
+= \ln{  \left( \sum_{\tilde{N}_i =N_i}^\infty \frac{\tilde{N}_i!}{N_i!(\tilde{N}_i-N_i)!}   p_i^{N_i} (1-p_i)^{ \tilde{N}_i-N_i} \frac{N_{0i}^{\tilde{N}_i} e^{-N_{0i}}}{\tilde{N}_i!} \right)}
+ \\
+ = \ln{  \left( \frac{(p_iN_{0i})^N_ie^{-p_iN_{0i}}}{N_i!}  \right)} \\
+ =  N_i \ln{\left(  p_i  N_{0i} \right) } -p_i N_{0i} - \ln{N_i!}.
+\end{align}
+For those cells that have a galaxy subtract out the probability for zero objects.  Each of these cells thus contribute
+\begin{equation}
+ N_i \ln{\left(  p_i  N_{0i} \right) } - \ln{N_i!}.
+\end{equation}
+
+The exponent term is proportional to (we don't care about multiplicative terms independent of the parameters)
+\begin{align}
+\exp\left(- \sum_{i=1}^{N_{bin}} p_i N_{0i}  \right)
+ \prod_{i=1}^{N_{bin}}  (p_i N_{0i})^{N_i} 
+\end{align}
+
 The second term is over those cells that have one galaxy:
@@
  = \sum_{i=1}^{N_{\text{galaxies}}} \ln{  \left(  p_i  {N_{0i} } \right)} \\
-  = \sum_{i=1}^{N_{\text{galaxies}}} \ln{  \left(  p_i  n_{0}(x_i) \hat{\delta} V \right)}
+  = \sum_{i=1}^{N_{\text{galaxies}}} \ln{  \left(  p_i  n_{0}(x_i) \delta V \right)}
 \end{align}
@@
 L(\theta; N, v) 
-&\propto \int  \exp \left(- \int  p(x) n_{0}(x) d^3x\right)  \prod_{i=1}^{N_{\text{galaxies}}} p_i n_0(x_i) p(v_i| u_i) p(\delta, u | \theta) d\delta\, du .
+&\propto p(v|\theta)  \int \exp \left(- \int  p(x) n_{0}(x) d^3x\right)  \prod_{i=1}^{N_{\text{galaxies}}} p(x_i) n_0(x_i)  p(\delta| v, \theta) d\delta \\
+& \approx p(v|\theta)  \int \prod_{j=1}^{N_\text{grid}} \exp \left( - \frac{p(x_j) n_{0}(x_j)}{N_\text{grid}}\right) \prod_{i=1}^{N_{\text{galaxies}}} p(x_i) n_0(x_i)  p(\delta| v, \theta) d\delta 
+%&\propto \int  \exp \left(- \int  p(x) n_{0}(x) d^3x\right)  \prod_{i=1}^{N_{\text{galaxies}}} p(x_i) n_0(x_i) p(v_i| u_i) p(\delta, u | \theta) d\delta\, du \\
+%& \approx  \int  \prod_{j=1}^{N_\text{grid}} \exp \left( - \frac{p(x_j) n_{0}(x_j)}{N_\text{grid}}\right) \prod_{i=1}^{N_{\text{galaxies}}} p(x_i) n_0(x_i) p(v_i| u_i) p(\delta, u | \theta) d\delta\, du.
 \end{align}
@@
 
-When the universe is modeled as a periodic cube much larger than the survey. the the density field can be describe by discrete points. 
-Further assuming a cutoff in power at $k_{max}$, the field can be described by a finite grid.  Then the parameters $\delta$ and $u$ will be
-the Fourier density and momentum at grid points, having a covariance matrix composed of four diagonal submatrices. 
-The density and momentum at any point in real space can be constructed from the parameters.
+When the universe is modeled as periodic, the fields therein can be expanded as a Fourier series.  Truncating the expansion at some $k_{max}$
+means that the a field can be described by a finite set of numbers.  The parameters $\delta$ and $u$ will be
+the Fourier density and momentum at grid points.  Realizations of these parameters can be easily generated because their 
+ covariance matrix is composed of four diagonal submatrices. 
+The density and momentum at any point in real space can be constructed from the parameters using the Fourier transform.
 

### 2019-10-28 doc/src/partials.py
fix D normalization

@@ -63,7 +63,8 @@ def integrand_D(a,*args):
 
 # normalized to a=1
 def D(a,*args):
-    return numpy.exp(-integrate.quad(integrand_D, a,1,args=args)[0])
+    return 1./1000*numpy.exp(integrate.quad(integrand_D, 1./1000,a,args=(gamma,OmegaM0))[0])
+    # return numpy.exp(-integrate.quad(integrand_D, a,1,args=args)[0])
 
 def dOmdOm0(a,OmegaM0=OmegaM0):
     den= (OmegaM0 + (1-OmegaM0)*a**3) 

### 2019-10-28 doc/src/partials_fD.py

@@ -1,0 +1,465 @@
+#!/usr/bin/env python
+
+import numpy
+import matplotlib.pyplot as plt
+import matplotlib
+from matplotlib import ticker, cm
+import scipy.integrate as integrate
+from astropy.cosmology import FlatLambdaCDM
+import astropy
+matplotlib.rcParams['font.size'] = 14
+matplotlib.rcParams['lines.linewidth'] = 2.0
+
+# cosmology
+OmegaM0 = 0.28
+cosmo = FlatLambdaCDM(H0=100, Om0=OmegaM0)
+
+gamma=0.55
+
+# tho code is hard wired to change these numbers requires surgery
+zmax=0.2
+nbins = 2
+binwidth = zmax/nbins
+bincenters = binwidth/2 + binwidth*numpy.arange(nbins)
+
+# power spectrum from CAMB
+matter = numpy.loadtxt("/Users/akim/project/PeculiarVelocity/pv/dragan/matterpower.dat")
+
+# vv
+# f=OmegaM0**0.55
+
+# SN properties
+restrate_Ia = 0.65*  2.69e-5*(1/0.70)**3 # h^3/Mpc^3 yr^{-1}
+sigm_Ia = 0.08
+
+# galaxy density (TAIPAN Howlett et al.)
+ng = 1e-3 # not used in current code
+
+def Cinverse(m00,m11,m01):
+    return numpy.array([[m11,-m01],[-m01,m00]])/(m00*m11 - m01**2)
+
+def Cinverse_partial(m00,m11,m01,dm00,dm11,dm01):
+    den = (m00*m11 - m01**2)
+    return numpy.array([[dm11,2*m01*dm01],[2*m01*dm01,dm00]])/den \
+        - numpy.array([[m11,-m01],[-m01,m00]])/den**2 * (dm00*m11 + m00*dm11 - 2*m01*dm01)
+
+def OmegaM(a,OmegaM0=OmegaM0):
+    return OmegaM0/(OmegaM0 + (1-OmegaM0)*a**3)
+
+def integrand_dlnfs(a,*args):
+    return numpy.log(OmegaM(a,*args))* OmegaM(a,*args)**0.55 / a
+
+# def dlnfs8dg(a,*args):
[…]

### 2019-10-28 doc/point.tex
new version with fD as parameter in two bins Bug in normalization of D

@@
 \end{align}
+
 \newpage
+\section{More Complete Model}
+
+The model consists of top-level parameters
+\begin{itemize}
+\item $\gamma$: Growth index;
+\item $\Omega_{M0}$: Mass density;
+\item $b$: Host-galaxy bias;
+\item $\sigma^2_{v,\text{NL}}$: Contribution of velocities due to overdensities of scales $k>k_{v,\text{max}}$;
+\item $\sigma^2_{\rho,\text{NL}}$: Overdensities of scales $k>k_{\rho,\text{max}}$;
+\item $\mathcal{M} = M -5\log{H_0}$: The zeropoint of the SN Hubble diagram;
+\item $\sigma_M$: SN~Ia intrinsic magnitude dispersion.
+\end{itemize}
+For shorthand these are collectively labeled with $\theta$.
+
+The model latent parameters are
+\begin{itemize}
+\item $\bar{\delta} | \gamma, \Omega_{M0}$: Fourier modes that describe the linear density field through $P_{\delta \delta}(f\sigma_8(\gamma, \Omega_{M0}))$;
+\item $\delta | \bar{\delta}, \sigma^2_{\rho,\text{NL}}$: The density field including non-linear density effects;
+\item $\delta_g | \delta, b$: The galaxy density field;
+\item $v | \delta$: The velocity field excluding non-linear velocity effects;
+\item $\{u_\alpha, \mu_\alpha + 5\log{H_0} \}| \delta_g $: Comoving coordinates for the  SNe in the universe;
+\item $\Delta M_\alpha | \sigma_M$: Absolute magnitude difference from the average for SN $\alpha$;
+\item $\Delta v_\alpha| \sigma^2_{v,\text{NL}}$:  Non-linear peculiar velocity for SN $\alpha$;
+\end{itemize}
+From these parameters we have the expected values of the SN observables
+\begin{align*}
+m_\alpha &=  (\mu_\alpha + 5\log{H_0})+\mathcal{M} + \Delta M_\alpha\\
+z_\alpha &= f(\mu_\alpha,v_\alpha+\Delta v_\alpha).
+%n_\alpha(u,m)& = n_0 \sum_\alpha \delta(m_\alpha,z_\alpha).
+\end{align*}
+%There is another set of latent parameters for a set of ``virtual'' SNe.
+%\begin{itemize}
+%\item $\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g$: True comoving coordinates for the set of SN;;
+%\item $\Delta M_\alpha | \sigma_M$: Absolute magnitude difference from the average for SN $\alpha$;
+%\end{itemize}
+For shorthand these are collectively called $\phi$.
+
+The data is a set of supernovae.  The data for supernova $i$ are 
+\begin{itemize}
+\item $\hat{u}_i $: Angular coordinates RA and Dec;
+\item $\hat{m}_i $: peak magnitude (proxy for distance);
+\item $\hat{z}_i $: redshift;
+\end{itemize}
+
+I have a model where the probability of the data is
+\begin{align}
+p(\bar{\delta}|\gamma,\Omega_{M0})  p(\delta| \bar{\delta},\sigma^2_{\rho,\text{NL}})  p(\delta_g|\delta,b) p(v|\delta)\\
+p(\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g) \\
+p(\{\Delta M_\alpha\} | \sigma_M) p(\{m_\alpha\} | \{\mu_\alpha + 5\log{H_0}\},\{\Delta M_\alpha\},\mathcal{M})\\
+    p(\{\Delta v_\alpha\} | \sigma^2_v)
+     p(\{z_\alpha\}| \{\mu_\alpha\}, v, \{\Delta v_\alpha\})\\
+ %    p(\{\hat{u}_\alpha\}| \{u_\alpha\}) p(\{\hat{m}_\alpha\}| \{m_\alpha\})  \\
+     p( \{\hat{u}_i\}, \{\hat{m}_i\}  | \{u_\alpha\}, \{m_\alpha\})       p(\{\hat{z}_i\}| \{z_\alpha\}).
+\end{align}
+The first line describes the fields, the second the SNe, the third their apparent magnitudes, the fourth apparent redshifts, and the last
+connects the model observables with their experimental realizations.
+The $p( \{\hat{u}_i\}, \{\hat{m}_i\}  | \{u_\alpha\}, \{m_\alpha\})$ does not depend on ordering.
+
+The term $p(\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g)$ is the probability of the appearance of objects at certain coordinates and non-appearance elsewhere.
+It is convenient to rewrite this term as
+\begin{align}
+p(\{u_\alpha, \mu_\alpha + 5\log{H_0}\} | \delta_g) & = p(0| \delta_g) \prod_\alpha \frac{ p(u_\alpha, \mu_\alpha + 5\log{H_0} | \delta_g)}{p(0_\alpha| \delta_g)}  \\
+& = e^{-{\bar{N}}}  \prod_\alpha \frac{\bar{n}_\alpha^{n_\alpha}}{n_\alpha!}
+\end{align}
+where $\bar{N}$ is the expected number of discovered SN, $\bar{n}$ is the expected number of SN discovered at position $\alpha$, and
+$n_\alpha$ is the number of SNe found at the same $\alpha$ position.  We expect $\bar{n} \propto \delta_g$.
+
+The survey has a selection function that is dependent on the angular position and observed magnitude $S(u,m)$.
+It is thus straightforward to include the selection when calculating $p( \{\hat{u}_i\}, \{\hat{m}_i\}  | \{u_\alpha\}, \{m_\alpha\})$.
+
+\newpage
+
+The average magnitude distance modulus of the bin is $\mu_\beta$ and the bin has linear radial peculiar velocity $v_\beta$.
+Each SN~Ia in that space is parameterized by $\Delta M_i$ and $\Delta v_i$.  We assume RA and Dec are measured
+perfectly, e.g. no lensing.
 \section{Model}

### 2019-10-28 doc/src/partials.py

@@ -58,8 +58,11 @@ def integrand_dDdg(a,*args):
 def dDdg(a,*args):
     return D(a,*args) * integrate.quad(integrand_dDdg, 1e-8, a)[0]
 
-def integrand_D(a,*args):
-    return  OmegaM(a,*args)**0.55 / a
+# def integrand_D(a,*args):
+#     return  OmegaM(a,*args)**0.55 / a
+
+def integrand_D(a,gamma=gamma, OmegaM0=OmegaM0,*args):
+    return  OmegaM(a,OmegaM0,*args)**gamma / a    
 
 # normalized to a=1
 def D(a,*args):

### 2019-10-28 doc/src/partials_fD.py

@@ -118,41 +118,41 @@ def finvp(f00,f11,f01,f00p,f11p,f01p):
     den = f00*f11 -f01**2
     return f11p /den - f11*(f00p*f11 + f00*f11p - 2*f01*f01p)/den**2
 
-def finvp3d(f00,f01,f02,f11,f12,f22,f00p,f01p,f02p,f11p,f12p,f22p):
-    den = f00*f11*f22 - f00*f12**2-f11*f02**2-f22*f01**2 + 2*f01*f02*f12
-    denp = f00p*f11*f22 + f00*f11p*f22+f00*f11*f22p  \
-        - f00p*f12**2 - 2*f00*f12*f12p \
-        - f11p*f02**2 - 2*f11*f02*f02p \
-        - f22p*f01**2 - 2* f22*f01*f01p \
-        + 2*(f01p*f02*f12 + f01*f02p*f12 + f01*f02*f12p)
+# def finvp3d(f00,f01,f02,f11,f12,f22,f00p,f01p,f02p,f11p,f12p,f22p):
+#     den = f00*f11*f22 - f00*f12**2-f11*f02**2-f22*f01**2 + 2*f01*f02*f12
+#     denp = f00p*f11*f22 + f00*f11p*f22+f00*f11*f22p  \
+#         - f00p*f12**2 - 2*f00*f12*f12p \
+#         - f11p*f02**2 - 2*f11*f02*f02p \
+#         - f22p*f01**2 - 2* f22*f01*f01p \
+#         + 2*(f01p*f02*f12 + f01*f02p*f12 + f01*f02*f12p)
 
-    # full matrix
-    # mat = numpy.array([ \
-    #     [f11*f22-f12**2, f02*f12-f01*f22, f01*f12-f00*f12], \
-    #     [0, f00*f22-f02**2, f02*f01-f02*f11], \
-    #     [0, 0, f00*f11-f01**2] \
-    #     ])
-    #00 element
-    mat = f11*f22-f12**2
+#     # full matrix
+#     # mat = numpy.array([ \
+#     #     [f11*f22-f12**2, f02*f12-f01*f22, f01*f12-f00*f12], \
+#     #     [0, f00*f22-f02**2, f02*f01-f02*f11], \
+#     #     [0, 0, f00*f11-f01**2] \
+#     #     ])
+#     #00 element
+#     mat = f11*f22-f12**2
 
-    # mat[1,0]= mat[0,1]
[…]
